### 2015-08-14 paper.tex
CREATE

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference]{IEEEtran}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.
\usepackage{pifont}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{mybeamer}
\usepackage{draftwatermark}

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Flow-Directed Trace Analysis for Post-Silicon Validation}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Hao Zheng, Yuting Cao}
\IEEEauthorblockA{Computer Science and Engineering\\
University of South Florida\\
Tampa, Florida 33620}
\and
\IEEEauthorblockN{Sandip Ray, Jin Yang}
\IEEEauthorblockA{Strategic CAD Labs\\
Intel Corporation\\
Portland, Oregon
}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
Validation of multicore systems-on-chip (SoC) is very challenging, and main reasons include the limited internal operation observability and non-determinism due to multiple clock domains.   Very often, the observed signal traces offer limited values for debuggers to understand system internal behavior.  This paper presents a trace analysis approach for post-silicon validation.  This approach takes system flow specifications describing system behavior at a high level and an observed signal trace, and interprets the trace with respect to the system flow specification.  The result from the trace interpretation is a set of flow instances such that their execution during a test explains the observed signal trace.   This approach transforms partially observed and non-deterministic signal traces to a representation that is more understandable, and offers more helpful information to locate design defects and a measure to evaluate validation coverage.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
%
%Post silicon debug is getting more challenging for current complex system-on-chip design.
%
%Using system level flows in silicon trace analysis helps to reduce the volume of data to be stored and analyzed, difficulty caused by inherent non-determinism.  
%
%Also, abstracts signal traces to system use cases, which helps validation engineering better understand the behavior displayed by the signal traces.

Systems-on-chips (SoCs) have become more complex as more functionalities are integrated on chips.  To reduce design time, more external IPs are integrated into current SoC designs. At the high level, a SoC can be viewed as a network of IP blocks communicating by a large number of protocols.  A typical SoC design consists of multiple programmable processor cores where a number of software threads can be executed simultaneously.  The resulting designs usually operate in multiple clock domains, and show a high degree of concurrency.  Due to the sophisticated communication mechanisms employed in  current SoC designs to support complex functional behaviors, traditional processor-centric validation has to be enhanced to focus inter-core communications.

To make validation even more challenging, the high degree of integration in current SoC designs causes traditional chip observation points to become inaccessible.  A SoC design usually contains millions of signals.  In post-silicon validation, thousands of those signals are tapped for observation.  Due to the limited availability of the chip pins, only about a hundred of the tapped signals can be traced.   This severely limits observability hinders post-silicon debuggers' capability to root cause an observed failure.  To make the situation even worse, the observed traces often display non-deterministic behavior due to many concurrent operations occurred in multiple clock domains in a design.  The above factors make debugging highly difficult in failure root-causing, and need to be addressed.

In a typical design flow, a design starts with some high-level specification, particularly for communication protocols.   The abstract protocol specification usually captures information exchanged among IP blocks and how it is exchanged to support functional behavior. 
Graphical representations of protocols such as Message Sequence Charts~\cite{Harel03} or Live Sequence Charts~\cite{Damm99} describe protocols intuitively.  However, its success in hardware industry is limited.  Recently, message flows in BPMN~\cite{Krstic14HOST} have been being advocated as a formalism for specifying architecture requirements
 or system-level use cases such as power management, security, etc.  In general, a flow describes a set of message sequences across multiple IP blocks following a system communication protocol.  Specification of system flows BPMN is intuitive and easier for human to understand and to reason about system behavior.  An example of flow specification in BPMN is shown in Figure~\ref{flow-spec-ex}.

\begin{figure}[tb]
\begin{center}
\includegraphics[height=2.5in]{figures/bpmn-flow-ex}
\caption{A graphical representation of a SoC firmware load protocol in BPMN~\cite{Krstic14HOST}.}
\label{flow-spec-ex}
\end{center}
\end{figure}

As indicated above, interpreting and understanding the raw signal traces observed from executing the chip is extremely difficult due to the limited observability and non-determinism.  This paper proposes a trace analysis method which analyzes and interprets the observed signal traces at the level of system flow specifications.  During the step of analysis and interpretation, this method tries to find out possible executions of system flows such that the observed signal trace is a result of such executions.  The resulting scenarios of flow executions often give debuggers more clear and better structured system-level views of the internal operations of the system under debug.  With a better understanding of system operations in a test environment, therefore, it would make easier for debuggers to pinpoint and root cause system failures.  In addition to debugging, another important benefit is better coverage evaluation as a scenario of flow executions captures the information about the types and number of instances of flows executed by the chip in a test environment.  That information can help to determine the quality and coverage of validation.

%---
\section{Labeled Petri-Nets}

Although system flows depicted in BPMN as in Fig.~\ref{flow-spec-ex} are intuitive for system architects to understand communications across multiple components, they are not suitable for algorithmic analysis.  In this section, we present an alternative formalism, \emph{labeled Petri Nets}, to represent communication aspects of system flows with the intuitive descriptions of internal operations for IP blocks abstracted away whenever possible.  Labeled Petri nets are capable of describing sequencing, concurrency, and choices.   This formalism has well defined operational semantics, efficient analysis analysis algorithms and and has been used widely in modeling and analyzing communication protocols, concurrent programs, etc.

%--- LPN definition with levels
%Specifically, a labeled Petri-net (LPN) is a tuple $(P, T, M_0, V, E, L)$ where
%\begin{itemize}
%\item $P$ is a finite set of places,
%\item $T$ is a finite set of transitions,
%\item $M_0$ is the set of initially marked places, also referred to as the initial marking.
%\item $V$ is a finite set of auxiliary variables of integer type,
%\item $E$ is a finite set of events,
%\item $L$ is a labeling function that maps each transition $t \in T$ to a triple $(g, e, A)$ where
%	\begin{itemize}
%	\item $g$ is a predicate over $V$,
%	\item $e \in E$ is an event,
%	\item $A$ is a set of assignments to some variables in $V$.  
%	\end{itemize}
%\end{itemize}
%
%For each transition $t \in T$, its preset, denoted as $\bullet{t} \subseteq P$, is the set of places connected to $t$, and its postset, denoted as $t\bullet \subseteq P$, is the set of places that $t$ is connected to.  A marking of a LPN is a set of places marked with tokens, and a state of a LPN is $(M, \alpha)$ where $M$ is a marking and $\alpha$ is an assignment to all auxiliary variables in $V$.  
%
%The operational semantics of a LPN is defined by transition executions.  A transition can be executed after it is {\em enabled}.  A transition $t = (g, e, A)$ is enabled in a state $(M, \alpha)$ if every place in its preset is included in the marking, i.e. $\bullet{t} \subseteq M$, and the values of the auxiliary variables make the predicate $g$ be true, i.e. $\alpha \models g$.  Execution of $t$ results in a new state $(M', \alpha')$ such that 
%\[
%M' = (M - \bullet{t}) \cup t\bullet,
%\]
%and $\alpha'$ is a new assignment where the values of the auxiliary variables are updated with respect to the assignments $A$.  Furthermore, when $t$ is executed, the event $e$ is emitted.  SoC

Specifically, a labeled Petri-net (LPN) is a tuple $(P, T, s_0, E, L)$ where
\begin{itemize}
\item $P$ is a finite set of places,
\item $T$ is a finite set of transitions,
\item $\mathit{init}$ is the set of initially marked places, also referred to as the initial marking.
\item $E$ is a finite set of events,
\item $L: T \rightarrow E$ is a labeling function that maps each transition $t \in T$ to an event $e \in E$.
\end{itemize}

For each transition $t \in T$, its preset, denoted as $\bullet{t} \subseteq P$, is the set of places connected to $t$, and its postset, denoted as $t\bullet \subseteq P$, is the set of places that $t$ is connected to.  A marking $s \subseteq P$ of a LPN is a subset of places marked with tokens, and it is also referred to as a state of a LPN.  The initial marking $\mathit{init}$ is also the initial state of the LPN.

The operational semantics of a LPN is defined by transition executions.  A transition can be executed after it is {\em enabled}.  A transition $t \in T$ is enabled in a state $s$ if every place in its preset is included in the marking, i.e. $\bullet{t} \subseteq s$.  Execution of $t$ results in a new state $s'$ such that 
\[
s' = (s - \bullet{t}) \cup t\bullet,
\]
and the emission of event $e$ labeled for $t$.

A simple communication protocol specification between two components represented by a LPN is shown in~Figure~\ref{fig-flow-ex-1}.  In this and the following figures for LPNs, the labeled circles denote places, and the labeled boxes denote transitions.  Each transition is labeled with its name and the associated event.  The solid black places without outgoing edges are {\em terminals}, which indicate termination of protocols represented by the LPNs.  In this LPN, the initial state $\mathit{init} = \{p_1\}$.  In this protocol, component~1 sends two consecutive messages $\mathit{msg}_1$ to component~2, which responds with either $\mathit{msg}_2$ or $\mathit{msg}_3$ non-deterministically.  If $\mathit{msg}_2$ is responded, component~1 repeats the protocol by sending two consecutive $\mathit{msg}_1$ again.  Otherwise, component~1 stops the protocol upon receiving $\mathit{msg}_3$.  This LPN represents two possible communication sequences as shown below.
\[
\begin{array}{l}
\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_1\;\ldots \\
\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\ldots, \mathit{msg}_3\;
\end{array}
\]

\begin{figure}[tb]
\begin{center}
\resizebox{2.5in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(t1)	at	($(p1) + (0,-1.1)$) {$t_1:msg_1$};
	\node[circ] 	(p2) 		at ($(t1) + (0,-1.1)$) {$p_2$};
	\node[rectbox]	(t2)	at	($(p2) + (0,-1.1)$) {$t_2:msg_2$};
	\node[circ] 	(p3) 		at ($(t2) + (0,-1.1)$) {$p_3$};
	\node[rectbox]	(t3)	at	($(p3) + (3,0)$) {$t_3:msg_3$};
	\node[rectbox]	(t4)	at	($(p3) + (-3,0)$) {$t_4:msg_4$};
	\node[circ,fill=black] 	(term) 		at ($(t4) + (0,-1.1)$) {$~~~$};

	% Edges
	\draw[->,weight3]	(p1) -- (t1);
	\draw[->,weight3]	(t1) -- (p2);
	\draw[->,weight3]	(p2) -- (t2);
	\draw[->,weight3]	(t2) -- (p3);
	\draw[->,weight3]	(p3) -- (t3);
	\draw[->,weight3]	(p3) -- (t4);
	\draw[->,weight3]	(t4) -- (term);	
	\draw[->,weight3]	(t3) -- ($(p1)+(3,0)$) -- (p1);

\end{tikzpicture}
}
\caption{An example of a LPN representing a system flow.}
\label{fig-flow-ex-1}
\end{center}
\end{figure}

% @COMMENT This example is for LPNs with guards.
%The above protocol specification can be made more precise by describing how component~2 would respond without using non-determinism.  For example, the system architect may wish to specify that component~2  responds with $\mathit{msg_2}$ to every sequence of two $\mathit{msg_1}$ for five such sequences in a row.  On the sixth such sequence, component~2 responds with $\mathit{msg_3}$.  The LPN modeling such protocol is shown in Figure~\ref{fig-flow-ex-2}.  An alternative without using auxiliary variables in the transition predicates is by unrolling the loops for the specified number of times; however this would result in a larger and more complex LPN.  
%
%\begin{figure}[htbp]
%\begin{center}
%\includegraphics[height=1.6in]{figures/flow-ex-2}
%\caption{Another example of a LPN representing the system flow in Fig.~\ref{fig-flow-ex-1} with a fixed number of message exchanges.}
%\label{fig-flow-ex-2}
%\end{center}
%\end{figure}

%=====
\section{Flow-Directed Trace Analysis}

In a typical validation setting, the system under debug (SUD) is executed in a test environment until it is terminated by the test environment or the system crashes due to a failure.  During the execution, a trace on a small number of observable signals is streamed off the chip for debugging.  Due to the limited observability and inherent non-determinism in today's SoC designs, the observed signal trace is difficult to understand, thus providing limited values for debugging.  In this section, we describe a trace analysis method where the observed signal traces are interpreted at the level of system flows.  In general, the trace analysis can offer debuggers a structured view of communications among the IP blocks during the SUD execution by deriving the types and numbers of system flows activated during SUD executions from the observed signal traces.

The trace analysis method consists of two steps: \emph{trace abstraction} and \emph{interpretation}.  The trace abstraction takes a signal trace observed during the execution of the SUD, and abstracts it with respect to information captured in the system flows.  This requires recognizing signal events and mapping the signal events or sequences of the signal events to flow events appeared in system flow specifications.  For example, a data write message may be a single event in a system flow, however, such a flow event maybe implemented in hardware as a sequence of events including a header, a number of data word transfers and a tail.  This abstraction requires a mapping relation from flow events to sequences of signal events, and it is reasonable to assume that this relation is available.

The trace interpretation takes a finite trace of flow events resulting from the trace abstraction and a set of system flows in LPNs $\vec{F}$, and generates a set of possible system flow execution scenarios.  A \emph{flow execution scenario} is defined as $\{(F_{i,j}, s_{i,j})\}$ where in each element $(F_{i,j}, s_{i,j})$, $F_{i,j}$ is the $j$th activated instance of flow $F_i \in \vec{F}$, and $s_{i,j}$ is a state of $F_{i,j}$.  A flow execution scenario indicates that at a certain point of SUD execution, what types of flows and the number of instances of a particular flow are activated and their corresponding current states.

An observed trace of flow events $\rho = e_1e_2\ldots e_n$ is a result of SUD executing the flow instances of some flow execution scenario.  The goal of the trace interpretation is to derive such scenario.  Let $\mathit{accept(F_{i,j}, s_{i,j}, e)}$ be a function that determines if event $e$ can be emitted by $F_{i,j}$ in state $s_{i,j}$.  This function is used during the trace interpretation for checking whether an event of an observed trace is a result of executing some flow instance.  Formally,  $\mathit{accept(F_{i,j}, s_{i,j}, e)}$ returns $(F_{i,j}, s'_{i,j})$ if there exists a transition $t$ in $F_i$ such that $L(t) = e$ and $\bullet t \subseteq s_{i,j}$.  In this case, $s'_{i,j} = (s_{i,j} - \bullet t) \cup t\bullet$.  It returns $\emptyset$ if no such $t$ exists in $F_i$.

Given a trace of flow events $\rho = e_1e_2\ldots e_n$, the trace interpretation algorithm starts with an empty set of of flow execution scenario $\mathit{Scen} = \emptyset$.  Then, for each $e_h$ where $1 \leq h \leq n$ starting $h=1$, and for each $\mathit{scen} \in \mathit{Scen}$, the following two steps are performed. 
\begin{description}
\item[Step 1]~~For each $(F_{i,j}, s_{i,j}) \in \mathit{scen}$, if $\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = (F_{i,j}, s'_{i,j})$, create a new scenario $\mathit{scen}' = (\mathit{scen} - (F_{i,j}, s_{i,j})) \cup (F_{i,j}, s'_{i,j})$, which is added into $\mathit{Scen}'$.

\item[Step 2]~~For each $F_i \in \vec{F}$, create a new instance $F_{i, j+1}$.  If $\mathit{accept}(F_{i,j+1}, \mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$,  create a new scenario $\mathit{scen}' = \mathit{scen} \cup (F_{i,j+1}, s'_{i,j+1})$, which is added into $\mathit{Scen}'$.
\end{description}
After $e_h$ is processed, $Scen = Scen'$, and the above two steps repeat for the next event $e_{h+1}$.

If every events in $\rho$ is successfully mapped to some flow instance, this algorithm returns a set of flow execution scenarios such that every flow instance is in its terminal state.  On the other hand, inconsistent events can also be encountered.  An event $e$ is \emph{inconsistent} if for each flow execution scenario $\mathit{scen} \in \mathit{Scen}$, the following two conditions hold.
\begin{enumerate}
\item For each $(F_{i,j}, s_{i,j}) \in \mathit{scen}$, $\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = \emptyset$,
\item For each $F_i \in \vec{F}$, $\mathit{accept}(F_{i}, \mathit{init}_{i}, e_h) = \emptyset$.
\end{enumerate}
An inconsistent event is the one produced by SUD execution but cannot be mapped to any flow instances no matter how the trace prior to event $e$ is interpreted.  Inconsistent events indicates possible causes of system failures.

Based on the above discussion, the trace interpretation algorithm returns two pieces of information: 1) a set $G$ of flow execution scenarios where every flow instance in every scenario is in its terminal state, 2) a set $B$ of pairs, each of which includes a set of flow execution scenarios and an inconsistent event.  The set $B$ provides valuable information for debuggers to root cause system failures.  One of these two sets can empty.   With the full observability, the set $G$ includes a single flow execution scenario derived for a trace.  In reality, it is always the case that the SUD is only partially observable.  Therefore, due to the lack of information for precise interpretation, a set of flow execution scenarios is typically derived for a given trace as the result of the trace analysis.

% @Note  \red{\bf Another assumption:  the observed signal trace captures the values on the observed signals in every cycle from the beginning.  Too strong an assumption??}

To illustrate the basic idea of the trace analysis method, consider the system flow shown in Figure~\ref{fig-flow-ex-1}.  Let $F_1$ denote such flow.  Suppose that a hardware system implements flow $F_1$, and the following trace of flow events is abstracted from an observed signal trace as a result of executing such system.     
\[
\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_3\;\ldots
\]  
This trace is interpreted from the first event to the last in order to derive all possible flow execution scenarios.  At the beginning, the first event $\mathit{msg}_1$ is processed.  According to the flow specification $F_1$, we know that one instance of such flow $F_1$, $F_{1,1}$,  is activated by the SUD as $\mathit{accept}(F_{1,1}, init_1, \mathit{msg}_1) = (F_{1,1}, \{p_2\})$ where $\mathit{init} = \{p_1\}$ is the initial state of $F_1$.  As the result, the flow execution scenario after interpreting the first event $\mathit{msg_1}$ is $\{(F_{1,1}, \{p_2\})\}$.

Next, the second $\mathit{msg}_1$ is interpreted on both scenarios   This event could be the result of two possible cases.  In the first case, this event is the result of the continuing execution of $F_{1,1}$ as $\mathit{accept}(F_{1,1}, \{p_2\}, \mathit{msg}_1) = (F_{1,1}, \{p_3\})$.  In the second case, the system may activate another instance of $F_1$, $F_{1,2}$ such that the second event $\mathit{msg_1}$ is a result of executing this new instance.  Therefore, the interpretation of the first two events $\mathit{msg}_1$ leads to two flow execution scenarios as shown below.
\[
\label{tr-analy-s1}\tag{1}
\begin{array}{cl}
1. & \{(F_{1,1}, \{p_3\})\},\\
2. & \{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_2\})\} 
\end{array}
\]

Now, consider the third $\mathit{msg}_1$ for each of the two scenario derived in~(\ref{tr-analy-s1}).    For the execution scenario~1, $F_{1,1}$ is not able to accept $\mathit{msg}_1$ as it is in state $\{p_3\}$.  On the other hand, this event could be the result of activation of a new flow instance.  Therefore, this execution scenario can be revised accordingly as 
\[
\label{tr-analy-s2}\tag{2}
\{(F_{1,1}, \{p_3\}),(F_{1,2},\{p_2\})\}.
\]   
For the execution scenario~2, event $\mathit{msg}_1$ could be the result of continuing execution of $F_{1,1}$ or $F_{1,2}$, or it could be a result of activation of a new flow instance.  Therefore,  three new execution scenarios can be derived as shown below for this event.
\[
\label{tr-analy-s3}\tag{3}
\begin{array}{l}
\{(F_{1,1}, \{p_3\}), (F_{1,2}, \{p_2\})\},\\
\{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_3\})\}, \\
\{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_2\}), (F_{1,3}, \{p_2\})\}
\end{array}
\]
Since the flow execution scenario in~(\ref{tr-analy-s2}) already exists in~(\ref{tr-analy-s3}), the three flow execution scenarios shown in~(\ref{tr-analy-s3}) is the result from the interpretation of the first three events $\mathit{msg}_1$.

The next flow event in the trace $\mathit{msg_2}$ is analyzed for the flow execution scenarios as shown in~(\ref{tr-analy-s3}).   For the first scenario $\{(F_{1,1}, \{p_3\}), (F_{1,2}, \{p_2\})\}$, event $\mathit{msg_2}$ can only be the result from executing $F_{1,1}$ as $\mathit{accept}(F_{1,1}, \{p_3\}, \mathit{msg_2}) = (F_{1,1}, \{p_1\})$.   Based on the same reasoning, this event can only be the result from executing $F_{1,2}$ in the second scenario, and it moves $F_{1,2}$ to a new state $\{p_3\}$ too.  The interesting case is the third scenario where none of the flow instances can allow $\mathit{msg_2}$ to happen.  This is due to the fact that the flow must be in $\{p_3\}$ for $\mathit{msg_2}$ to happen.  This indicates the third system execution scenario is impossible for the prefix of the flow event trace upto $\mathit{msg_2}$, therefore this scenario is ignored from further analysis.  After analyzing event $\mathit{msg_2}$, the updated system execution scenarios are shown below.
\[
\begin{array}{l}
\{(F_{1,1}, \{p_1\}), (F_{1,2}, \{p_2\})\},\ \{(F_{1,1}, \{p2\}), (F_{1,2}, \{p_1\})\}.
\end{array}
\]

The last flow event in the trace is $\mathit{msg}_3$.  Considering the above two possible system executions, neither can allow this event to be produced as none of the flow instances in both system executions is in state $\{p_3\}$.  What this means is that the system does {\em not} implement the flow specification correctly as it produces something not allowed by the specification.  By examining the system executions right before the ``buggy'' event, debuggers may gain more information on when and where the problem might be.  The trace interpretation algorithm adds these two scenarios along with the fifth event $\mathit{msg_3}$ into $B$, and returns it for debuggers to analyze.

%---
\subsection{Trace Analysis with Partial Observability}

In hardware that implements a given system flow specification, a flow event is defined as an event or a sequence of events on a set of signals.  Due to the limited number of pins on the boundary of chips available for observation, only a small fraction of system signals can be observed during debug.  In this section, we discuss how the trace analysis method presented above can be adapted to deal with signal traces of partial observability.

In general, a signal trace of partial observability corresponding a set of traces of flow events due to the ambiguous interpretation of signal events.  In the following, we discuss two cases for trace abstraction on partial observability: mapping a single signal event to a flow event or mapping a sequence of signal events to a flow event.  A signal event is defined as a state on or an assignment to a set of signals.

Hereafter, the term {\em flow traces} is used to refer to traces of flow events. 
Consider the following example for the first case.  Suppose that there are three flow events: $e_1$, $e_2$, and $e_3$, which are implemented in hardware by the signal events shown in the list below.  We use Boolean expressions to represent signal events for the discussion.
\[
\begin{array}{cl}
e_1: & abc\\
e_2: & \bar{a}bc\\
e_3: & a\bar{b}c
\end{array}
\] 
Now suppose that only signals $b$ and $c$ are observable, and a signal trace of this partial observability is obtained below.
\[
bc\ bc \ \bar{b}c
\]
During the trace abstraction step, the first two signal events $bc$ can be mapped to $\{e_1, e_2\}$ since $a$ is not observable, and the last one $b'c$ is mapped to $\{e_3\}$.  Therefore, this signal trace is abstracted to four flow traces, $\{e_1, e_2\} \times \{e_1, e_2\} \times \{e_3\}$.

Next, we consider the case where a flow event is mapped from a sequence of signal events.  Now suppose that two other flow events are implemented by sequences of signal events as defined in the list below.
\[
\begin{array}{cl}
e_4: & abc\ \bar{a}bc\\
e_5: & abc\ abc\ abc\ \bar{a}bc\\
\end{array}
\] 
Given an observed trace of the same observability shown below
\[
bc\ bc\ bc \ bc,
\]
it is abstracted to the following flow traces.
\[
e_4e_4,\; \textvisiblespace e_4 \textvisiblespace,\;  e_4 \textvisiblespace \textvisiblespace,\;  \textvisiblespace \textvisiblespace e_4, \; e_5
\] 
where $\textvisiblespace$ denotes signal events that are not mapped to any flow events.  Note that the above abstraction leads to three distinct flow traces as the middle three correspond to the same trace of flow events.

From the above discussions, it can be seen that a signal trace of partial observability is generally mapped to a set of flow traces.  As the the number of signals available for observation becomes smaller, the number of flow event traces corresponding to a signal trace as a result of the abstraction can be enormous.  This phenomenon can increase the complexity of the trace interpretation as it can cause the number of possible flow execution scenarios generated during the analysis to explode.   Possible solutions to address this issue include better trace signal selection for observation and assistance from debuggers' insights.  Trace signal selection itself is an important and difficult subject, and a detailed discussion of it is out of scope of this paper.   Next, we briefly describe how the debuggers' insights of a system's architecture can help to address the complexity issue in the trace analysis.

\subsection{Inputs from Validators}

During the trace interpretation, the number of intermediate flow execution scenarios may become too large due to ambiguous interpretation from signal events to flow events or from flow events to flow execution scenarios.  The explosion of the intermediate results can significantly slow down the performance of the trace analysis.  To address this problem, the debuggers can use their insights and understanding of the SUD to trim the total number of possible system executions derived from the trace analysis.  When a SUD is validated, a debugger is assumed to have deep knowledge about the system's architecture and microarchitecture, and how the test environment affects the system executions.  For instance, he/she may have knowledge that in a test environment, the maximal number of instances of a flow can be activated by the SUD, or a flow can be activated after certain other flows have terminated, etc.  These debugging insights can be encoded as constraints into the trace analysis method, which can then be used to eliminate a large number of flow execution scenarios that violate these constraints during the trace interpretation step.

This approach can be flexible in that it allows a debugger to analyze the observed traces in a trail-and-error manner if the precise knowledge of the system (micro-)architecture is hard to come by.  For instance, the debugger might initially make a very restricted assumption on how the SUD executes a flow specification, and these assumptions can potentially lead to an empty set of flow execution scenarios.  Depending on which of these assumptions triggered during the trace interpretation step, the debugger can study these assumptions more carefully, and relax some or all of them for the next run of analysis.  This iteration can be repeated as many times as necessary until some results deemed meaningful are produced.

Alternatively, if all derived execution scenarios seem to be plausible, the implication that a debugger may draw from this result is that the failure may be independent of the flows being observed.  Therefore, the testing environment can be adjusted in order for a different part or different behavior of the SUD to be observed.  This idea, closely related to trace signal selection, is critical for post-silicon validation, and a detailed discussion can only be presented in a separate paper.

%=====
\section{Case Study}

The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a simple SoC design with two CPU cores, memory, and a peripheral device connected by an interconnect as shown in Fig.~\ref{fig-soc-ex}.  Since the analysis method presented in this paper is communication centric, the detailed computations of these blocks are not modeled.  Instead, the modeling is focused on how they participate in flows for different system level use cases.

\begin{figure}[tb]
\begin{center}
\includegraphics[width=2in]{figures/simple-ex}
\caption{A simple SoC example.}
\label{fig-soc-ex}
\end{center}
\end{figure}

\subsection{Flow Specification}

In this case study, four system flows are implemented in the simple SoC model. They include cache coherent memory access operations and a memory-mapped peripheral read operation initiated from the CPUs, a message signaled interrupt operation initiated from the peripheral device.  These flow specifications capture how messages are exchanged for different use cases.  In this model, a message is defined with the following format, $(\mbox{Src}, \mbox{Dest}, \mathit{Cmd}, \mathit{Addr})$, where Src and Dest refer to the source and destination components of messages, $\mathit{Cmd}$ refers to the operations that the destination component should perform, and $\mathit{Addr}$ refers to memory addresses where $\mathit{Cmd}$ applies.  $\mathit{Cmd}$ can be memory accesses or not.  If it is not, then the $\mathit{Addr}$ field of messages is ignored.  In this case study, memory mapped IO mechanism is used.  Furthermore, detailed memory addresses are not modeled. Instead, the address space is partitioned to main memory addresses and peripheral addresses.  In messages, the $\mathit{Addr}$ field is replaced with either $M$ representing a memory address or $P$ representing an address to a peripheral device.

The LPN as shown in Fig.~\ref{fig-cpu-write} specifies a system flow where {CPU2} initiates a memory write operation.   In this flow, CPU2 initiates a memory write request followed by a data valid message to the interconnect.  The data valid messages are used to model availability or validity of data for transfer.  Concurrently, the interconnect inquiries CPU1 if it holds a more updated version of the data by sending a memory write message $\mathit{msg_2}$.   CPU1 generates one of two possible responses.  If CPU1 holds the more updated data in its cache in the same memory space that CPU2 intends to write, a cache hit message $\mathit{msg_4}$ followed by $\mathit{msg_6}$ are sent to the interconnect.  Otherwise, CPU1 sends a cache miss message $\mathit{msg_5}$.  After getting the response from CPU1, Interconnect sends a write request to the memory unit.  This flow is symmetric for CPU1.

\begin{figure}[tb]
\begin{center}
\resizebox{2.8in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(msg1)	at	($(p1) + (0,-1.1)$) {$msg_1$};
	\node[circ] 	(p2) 		at ($(msg1) + (-3,0)$) {$p_2$};
	\node[circ] 	(p3) 		at ($(msg1) + (3,0)$) {$p_3$};
	\node[rectbox]	(msg2)	at	($(p2) + (0,-1.1)$) {$msg_2$};
	\node[rectbox]	(msg3)	at	($(p3) + (0,-1.5)$) {$msg_3$};
	\node[circ] 	(p4) 		at ($(msg2) + (0,-1.1)$) {$p_4$};
	\node[rectbox]	(msg4)	at	($(p4) + (0,-1.1)$) {$msg_4$};
	\node[rectbox]	(msg5)	at	($(p4) + (3,0)$) {$msg_5$};
	\node[circ] 	(p6) 		at ($(msg4) + (0,-1.1)$) {$p_6$};
	\node[circ] 	(p5) 		at ($(msg3) + (0,-1.4)$) {$p_5$};
	\node[rectbox,rotate=90]	(msg6)	at	($(p6) + (1.5,0)$) {$msg_6$};
	\node[circ] 	(p7) 		at ($(msg6) + (1.5,0)$) {$p_7$};
	\node[rectbox]	(msg7)	at	($(p5) + (0,-1.5)$) {$msg_7$};
	\node[terminal]	(p8)	at	($(msg7) + (2,0)$) {$p_8$};
	% Edges
	\draw[->,weight3]	(p1) -- (msg1);
	\draw[->,weight3]	(msg1) -- (p2);
	\draw[->,weight3]	(msg1) -- (p3);
	\draw[->,weight3]	(p2) -- (msg2);
	\draw[->,weight3]	(msg2) -- (p4);
	\draw[->,weight3]	(p4) -- (msg4);
	\draw[->,weight3]	(p4) -- (msg5);
	\draw[->,weight3]	(msg4) -- (p6);
	\draw[->,weight3]	(p6) -- (msg6);
	\draw[->,weight3]	(msg6) -- (p7);
	\draw[->,weight3]	(p7) -- (msg7);
	\draw[->,weight3]	(msg7) -- (p8);
	\draw[->,weight3]	(msg5) -- (p7);
	\draw[->,weight3]	(p3) -- (msg3);
	\draw[->,weight3]	(msg3) -- (p5);
	\draw[->,weight3]	(p5) -- (msg7);

	\node[]	(msg-def)	at	($(p1) + (0,-8.5)$) {
	\begin{minipage}{3in}
	Definition of the messages:
\[
\begin{array}{clllll}
msg_1: & (\mbox{CPU2},	& \mbox{Icon},		& \mathit{Wr}, 	& M) \\
msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& \mathit{Wr},			& M) \\
msg_3: & (\mbox{CPU2},	& \mbox{Icon}, 	& \mathit{DVal},	& -) \\
msg_4: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Hit},	& -) \\
msg_5: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Miss},	& -) \\
msg_6: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{DVal},	& -) \\
msg_7: & (\mbox{Icon},	& \mbox{Mem}, 		& \mathit{Wr},			& M) \\
\end{array}
\]
\end{minipage}
		};
\end{tikzpicture}
}
\caption{Flow specification ($F_1$) of a cache coherent write operation initiated from \texttt{CPU2}.}
\label{fig-cpu-write}
\end{center}
\end{figure}

The LPN specification as shown in Fig.~\ref{fig-cpu-read} captures the system flow where CPU2 initiates a memory read operation.  Basically, CPU2 sends a memory read message to Interconnect, which then generates two concurrent threads, one checks if CPU1 has the more updated data in the memory space for the read operation, and the other thread to get data from the memory.  Once Interconnect gets both responses from CPU1 and memory, it synchronizes the responses, and writes the correct data to the CPU2's cache and memory in parallel.  Again, this specification is symmetric for CPU1.

\begin{figure}[htbp]
\begin{center}
\resizebox{3.2in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(msg1)	at	($(p1) + (0,-1.1)$) {$msg_1$};
	\node[circ] 	(p2) 		at ($(msg1) + (-3,0)$) {$p_2$};
	\node[circ] 	(p3) 		at ($(msg1) + (2,0)$) {$p_3$};
	\node[rectbox]	(msg2)	at	($(p2) + (0,-1.1)$) {$msg_2$};
	\node[rectbox]	(msg3)	at	($(p3) + (2,0)$) {$msg_3$};
	\node[circ] 	(p4) 		at ($(msg2) + (0,-1.1)$) {$p_4$};
	\node[rectbox]	(msg4)	at	($(p4) + (0,-1.1)$) {$msg_4$};
	\node[rectbox]	(msg5)	at	($(p4) + (3,0)$) {$msg_5$};
	\node[circ] 	(p6) 		at ($(msg4) + (0,-1.1)$) {$p_6$};
	\node[circ] 	(p5) 		at ($(msg3) + (0,-1.1)$) {$p_5$};
	\node[rectbox,rotate=90]	(msg6)	at	($(p6) + (1.5,0)$) {$msg_6$};
	\node[circ] 	(p7) 		at ($(msg6) + (1.5,0)$) {$p_7$};
	\node[rectbox]	(msg7)	at	($(p5) + (0,-1.1)$) {$msg_7$};
	\node[circ]	(p8)	at	($(msg7) + (0,-1.1)$) {$p_8$};
	\node[rectbox]	(msg8)	at	($(p8) + (0,-1.1)$) {$msg_8$};
	\node[circ]	(p9)	at	($(msg8) + (-1.5,-1.1)$) {$p_9$};
	\node[circ]	(p10)	at	($(msg8) + (1.5,-1.1)$) {$p_{10}$};
	\node[rectbox]	(msg9)	at	($(p9) + (0,-1.1)$) {$msg_9$};
	\node[rectbox]	(msg10)	at	($(p10) + (0,-1.1)$) {$msg_{10}$};
	\node[terminal]	(p11)	at	($(msg9) + (0,-1.1)$) {$p_{11}$};
	\node[terminal]	(p12)	at	($(msg10) + (0,-1.1)$) {$p_{12}$};

	% Edges
	\draw[->,weight3]	(p1) -- (msg1);
	\draw[->,weight3]	(msg1) -- (p2);
	\draw[->,weight3]	(msg1) -- (p3);
	\draw[->,weight3]	(p2) -- (msg2);
	\draw[->,weight3]	(msg2) -- (p4);
	\draw[->,weight3]	(p4) -- (msg4);
	\draw[->,weight3]	(p4) -- (msg5);
	\draw[->,weight3]	(msg4) -- (p6);
	\draw[->,weight3]	(p6) -- (msg6);
	\draw[->,weight3]	(msg6) -- (p7);
	\draw[->,weight3]	(p7) -- (msg8);
	\draw[->,weight3]	(msg7) -- (p8);
	\draw[->,weight3]	(msg5) -- (p7);
	\draw[->,weight3]	(p3) -- (msg3);
	\draw[->,weight3]	(msg3) -- (p5);
	\draw[->,weight3]	(p5) -- (msg7);
	\draw[->,weight3]	(p8) -- (msg8);
	\draw[->,weight3]	(msg8) -- (p9);
	\draw[->,weight3]	(msg8) -- (p10);
	\draw[->,weight3]	(p9)  -- (msg9);
	\draw[->,weight3]	(p10)  -- (msg10);
	\draw[->,weight3]	(msg9) -- (p11);
	\draw[->,weight3]	(msg10) -- (p12);

	\node[]	(msg-def)	at	($(p1) + (-2.7,-10)$) {
	\begin{minipage}{2in}
	Definition of the messages:
\[
\begin{array}{clllll}
msg_1: & (\mbox{CPU2},	& \mbox{Icon},		& \mathit{Rd}, 		& M) \\
msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& \mathit{Rd},	& M) \\
msg_3: & (\mbox{Icon},	& \mbox{Mem}, 		& \mathit{Rd},			& M) \\
msg_4: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Hit},	& -) \\
msg_5: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Miss},& -) \\
msg_6: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{DVal},	& -) \\
msg_7: & (\mbox{Mem},		& \mbox{Icon}, & \mathit{DVal},			& -) \\
msg_8: & (-,		& -, 		& -,				& -) \\
msg_9: & (\mbox{Icon},	& \mbox{Mem}, 		& \mathit{Wr},			& M) \\
msg_{10}: & (\mbox{Icon},	& \mbox{CPU2}, 		& \mathit{Wr},			& M) \\
\end{array}
\]
\end{minipage}
		};

\end{tikzpicture}
}
\caption{Flow specification ($F_2$) of a cache coherent read operation from CPU2.}
\label{fig-cpu-read}
\end{center}
\end{figure}

The LPN specification as shown in Fig.~\ref{fig-peri-read} captures a system flow for CPU initiated memory mapped peripheral read operations.  When CPU2 tries to read the peripheral device (device hereafter), a read message with address $P$ is sent to Interconnect, which then sends this message to CPU1 and the device simultaneously.  When CPU1 sees this message, it responds with a cache miss message as the address $P$ points to the device.  At the meantime, the device responds with a data valid message indicating the availability of the requested data.  Finally, Interconnect synchronizes the both responses, and sends a data valid message back to CPU2.   This specification is also symmetric for CPU1.

\begin{figure}[tb]
\begin{center}
\resizebox{3.2in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(msg1)	at	($(p1) + (0,-1.1)$) {$msg_1$};
	\node[circ] 	(p2) 		at ($(msg1) + (-1.5,0)$) {$p_2$};
	\node[circ] 	(p3) 		at ($(msg1) + (1.5,0)$) {$p_3$};
	\node[rectbox]	(msg2)	at	($(p2) + (-1.5,0)$) {$msg_2$};
	\node[rectbox]	(msg4)	at	($(p3) + (1.5,0)$) {$msg_4$};
	\node[circ] 	(p4) 		at ($(msg2) + (-1.5,0)$) {$p_4$};
	\node[rectbox]	(msg3)	at	($(p4) + (0,-1.2)$) {$msg_3$};
	\node[circ] 	(p6) 		at ($(msg3) + (0,-1.2)$) {$p_6$};
	\node[circ] 	(p5) 		at ($(msg4) + (1.5,0)$) {$p_5$};
	\node[rectbox]	(msg5)	at	($(p5) + (0,-1.2)$) {$msg_5$};
	\node[circ]	(p7)	at	($(msg5) + (0,-1.2)$) {$p_7$};
	\node[rectbox]	(msg6)	at	($(p6) + (4.5,0)$) {$msg_6$};
	\node[terminal]	(p8)	at	($(msg6) + (0,-1.2)$) {$p_8$};

	% Edges
	\draw[->,weight3]	(p1) -- (msg1);
	\draw[->,weight3]	(msg1) -- (p2);
	\draw[->,weight3]	(msg1) -- (p3);
	\draw[->,weight3]	(p2) -- (msg2);
	\draw[->,weight3]	(msg2) -- (p4);
	\draw[->,weight3]	(p4) -- (msg3);
	\draw[->,weight3]	(msg3) -- (p6);
	\draw[->,weight3]	(p6) -- (msg6);
	\draw[->,weight3]	(msg6) -- (p8);
	\draw[->,weight3]	(p3) -- (msg4);
	\draw[->,weight3]	(msg4) -- (p5);
	\draw[->,weight3]	(p5) -- (msg5);
	\draw[->,weight3]	(msg5) -- (p7);
	\draw[->,weight3]	(p7) -- (msg6);
	\draw[->,weight3]	(msg6) -- (p8);

		\node[]	(msg-def)	at	($(p8) + (0,-2.5)$) {
	\begin{minipage}{3in}
	Definition of the messages:
\[
\begin{array}{clllll}
msg_1: & (\mbox{CPU2},	& \mbox{Icon},		& Rd, 	& P) \\
msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& Rd,		& P) \\
msg_3: & (\mbox{CPU1},	& \mbox{Icon}, 	& Miss,			& -) \\
msg_4: & (\mbox{Icon},	& \mbox{Dev}, 	& Rd,	& P) \\
msg_5: & (\mbox{Dev},	& \mbox{Icon}, 	& Dval,	& -) \\
msg_6: & (\mbox{Icon},	& \mbox{CPU2}, 	& Dval,	& -) \\\end{array}
\]
\end{minipage}
		};
\end{tikzpicture}
}
\caption{Flow specification ($F_3$) of a read access to peripheral device from CPU2.}
\label{fig-peri-read}
\end{center}
\end{figure}

The last LPN specification as shown in Fig.~\ref{fig-msi} captures how interrupts from the device are handled.  In this case study, all interrupts are directed to CPU1.  When the device triggers an interrupts, it sends a message with $\mathit{Intr}$ in the command field.  Then, Interconnect notifies CPU1 by sending a message with $\mathit{MSI}$ and $I$ in the command and address fields, respectively, where $I$ is the symbol referring to the entry points to interrupt service routines.   CPU1 responds with a cache miss message as the receipt of the interrupt.

\begin{figure}[htbp]
\begin{center}
\resizebox{3.5in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(msg1)	at	($(p1) + (1.5, 0)$) {$msg_1$};
	\node[circ] 	(p2) 		at ($(msg1) + (1.5, 0)$) {$p_2$};
	\node[rectbox]	(msg2)	at	($(p2) + (1.5,0)$) {$msg_2$};
	\node[circ] 	(p3) 		at ($(msg2) + (1.5, 0)$) {$p_3$};
	\node[rectbox]	(msg3)	at	($(p3) + (1.5, 0)$) {$msg_3$};
	\node[terminal] 	(p4) 		at ($(msg3) + (1.5,0)$) {$p_4$};

	% Edges
	\draw[->,weight3]	(p1) -- (msg1);
	\draw[->,weight3]	(msg1) -- (p2);
	\draw[->,weight3]	(p2) -- (msg2);
	\draw[->,weight3]	(msg2) -- (p3);
	\draw[->,weight3]	(p3) -- (msg3);
	\draw[->,weight3]	(msg3) -- (p4);

\node[]	(msg-def)	at	($(msg2) + (0,-2)$) {
	\begin{minipage}{3in}
	Definition of the messages:
\[
\begin{array}{clllll}
msg_1: & (\mbox{Dev},	& \mbox{Icon},		& \mathit{Intr}, 	& -) \\
msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& \mathit{MSI},	& I) \\
msg_3: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Miss},	& -) \\
\end{array}
\]
\end{minipage}
		};
\end{tikzpicture}
}
\caption{Flow specification ($F_4$) of handling interrupts from the peripheral device.}
\label{fig-msi}
\end{center}
\end{figure}

%
%
\subsection{Results and Discussions}
%
%
The transaction level model of the simple system implementing the four flow specifications shown in the previous section is described in SystemC.  Each component is described as a SystemC module, which may include a number of threads to model concurrency.  The entire model is concurrent and operates in asynchronous mode.

When testing this model, both CPUs and the peripheral device are set up as flow instance generators.  They randomly generate the first message in a corresponding flow specification to start a flow instance, and react to incoming messages by generating new messages as defined in the flow specification.  In the model, monitors are embedded to observe the messages generated by each component.  When a message is observed, it is written to an output trace file for analysis.

Even though this example is conceptually simple, getting the model to correctly implement the flow specifications is not straightforward.  On the other hand, results from the trace analysis greatly help the debugging process by providing information to locate problems quickly.  For example, in an early version of the model, the trace shown in Table~\ref{table-trace} is observed.  
\begin{table*}
\caption{An observed trace of messages for trace analysis.}
\label{table-trace}
\begin{center}
\begin{tabular}{llll}
1~~(CPU2, Icon, Rd, M) & 2~~(Icon, CPU1, Rd, M) & 3~~(Icon, Mem, Rd, M) & 4~~(Mem,  Icon, DVal, -) \\
5~~(CPU1, Icon, Hit, -) & 6~~(CPU1, Icon, DVal, -) & 7~~(Icon, CPU2, Wr, M) & 8~~(Icon, Mem, Wr, M) \\
9~~(CPU2, Icon, Rd, M) & 10~(Icon, CPU1, Rd, M) & 11~(Icon, Mem, Rd, M) & 12~(CPU2, Icon, Rd, P) \\
13~(Icon, CPU1, Rd, P) & 14~(CPU1, Icon, Miss, -) & 15~(Icon, Dev, Rd, P) & 16~(CPU1, Icon, DVal, -) \\
\end{tabular}
\end{center}
\end{table*}
The trace analysis finds out that messages~$1-8$ in the trace are the results of execution of an instance of flow $F_2$ as shown in Fig.~\ref{fig-cpu-read}, and the flow execution scenario is $\{(F_{2,1}, \{p_{11}, p_{12}\})\}$.  Messages~$9-11$ are analyzed as the results of executing another instance of flow $F_2$, message~$12-13$ as the results of executing an instance of flow $F_3$ as shown in Fig.~\ref{fig-peri-read}.  The flow execution scenario after the first thirteen messages in the trace is 
\[
\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_4, p_5\}), (F_{3,1}, \{p_4, p_3\})\}.
\]  
Message~$14$ can be  the result from executing $F_{2,2}$ or $F_{3,1}$, therefore it leads to two following flow execution scenarios:
\begin{equation}
\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_7, p_5\}), (F_{3,1}, \{p_4, p_3\})\}
\end{equation}
\begin{equation}
\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_4, p_5\}), (F_{3,1}, \{p_6, p_3\})\}
\end{equation}
In either scenario, after mapping message~$15$ to flow $F_{3,1}$, message~$16$ cannot be mapped to any existing flow instance or to a new flow instance.  From this inconsistent message, we know that it is generated by CPU1.  In scenario~(1), CPU1 is in the state after generating the message reporting a cache miss.  In this case, the DataValid message should not generated.  In scenario~(2), CPU1 is in the state before generating the message reporting either a cache hit or miss, and again the DataValid message should not generated.   This inconsistent message helps to locate a bug in the CPU1 model where a DataValid message is generated after either a cache hit or miss message is generated.  After fixing this bug,  in a few more iterations of analysis and debugging, the trace analysis can eventually extract all initiated flow instances in the model, all in their terminal states, thus showing that the model implements the four flows correctly.

In the second experiment, partial observability is taken into account during the trace analysis with the assumption that the command $\mathit{Wr}$ and $\mathit{Rd}$ and addresses $M$ and $P$ are indistinguishable due to the lack of observability.  This partial observability is simulated with a modification to the monitors such that in each observed message, command $\mathit{Wr}$ or $\mathit{Rd}$ is replaced with $(\mathit{Wr},\mathit{Rd})$ and address $M$ or $P$ is replaced with $(M,P)$.  The version of the model generating the trace shown in Table~\ref{table-trace} is reused with the modified the monitors, and the generated trace with the simulated partial observability is shown in Table~\ref{table-trace-2}.  Similarly, only the first sixteen messages are shown. 
\begin{table*}
\caption{A trace of messages with partial observability for trace analysis.}
\label{table-trace-2}
\begin{center}
\resizebox{7.3in}{!}{
\begin{tabular}{llll}
1~~(CPU2, Icon, (Wr, Rd), (M, P)) & 2~~(Icon, CPU1, (Wr, Rd), (M, P)) & 3~~(Icon, Mem, (Wr, Rd), (M, P)) & 4~~(Mem,  Icon, DVal, -) \\
5~~(CPU1, Icon, (Hit, Miss), -) & 6~~(CPU1, Icon, DVal, -) & 7~~(Icon, CPU2, (Wr, Rd), (M, P)) & 8~~(Icon, Mem, (Wr, Rd), (M, P)) \\
9~~(CPU2, Icon, (Wr, Rd), (M, P)) & 10~(Icon, CPU1, (Wr, Rd), (M, P)) & 11~(Icon, Mem, (Wr, Rd), (M, P)) & 12~(CPU2, Icon, (Wr, Rd), (M, P)) \\
13~(Icon, CPU1, (Wr, Rd), (M, P)) & 14~(CPU1, Icon, (Hit, Miss), -) & 15~(Icon, Dev, (Wr, Rd), (M, P)) & 16~(CPU1, Icon, DVal, -) \\
\end{tabular}
}
\end{center}
\end{table*}

Each message in the trace with partial observability is referred to as an {\em super} message to distinguish it from the messages of full observability.  The traces of super messages are referred to as {\em super} traces.  For example, the first super message in the trace from Table~\ref{table-trace-2}, $\mbox{(CPU2, Icon, (Wr, Rd), (M, P))}$, corresponds to four distinct messages:
$\mbox{(CPU2, Icon, Wr, M)}$, $\mbox{(CPU2, Icon, Wr, P)}$, $\mbox{(CPU2, Icon, Rd, M)}$, and $\mbox{(CPU2, Icon, Rd, P)}$.  Some of these messages do not exist in the flow specification, and are ignored during the trace analysis.   In the above example, message $\mbox{(CPU2, Icon, Wr, P)}$ is ignored.

%Each super trace represents a set of traces, each of which is interpreted to derive a set of flow execution scenarios.  These flow execution scenarios can show different inconsistent messages.  For example, the particular trace of the super trace shown in Table~\ref{table-trace-2} below
%\[
%\mbox{1 (CPU2, Icon, Wr, M), 2~(Icon, CPU1, Rd, M)}, \ldots
%\]
%leads to the following flow execution scenario $\{(F_{1,1}, \langle p_2,p_3\rangle)\}$ showing that the second message is inconsistent.  In addition, this super trace also represents a trace as shown in Table~\ref{table-trace} that leads to a real bug in the model as discussed above.  It is also possible in some situation that a super trace may represent a trace with no inconsistent messages.  Therefore, the trace analysis on a super trace can result in a number of flow execution scenarios revealing different problems or no problem.  This ambiguity is caused by the partial observability.

Each super trace represents a set of traces, each of which is interpreted to derive a set of flow execution scenarios.
Due to the partial observability, the number of traces represented by a super trace can become very large. For example, the super trace shown in Table~\ref{table-trace-2} represents about twelve thousand possible traces for that short sequence of messages.  The trace analysis algorithm returns the set of flow execution scenarios for each trace for examination, and a very large number of possible flow execution scenarios can be generated.  The large number of possible flow execution scenarios not only produces too much information that can overwhelm system validators, but also degrades the performance of the trace analysis algorithm by consuming too much memory.  As indicated above,  the validators' insights on the SUD can be utilized to trim the possibilities.  For example, if the validator knows that no flow $F_1$ in Fig.~\ref{fig-cpu-write} is activated in the testing environment, this insight helps to eliminate all flow execution scenarios that include instances of $F_1$ by interpreting message~$\#1$ as either $\mbox{(CPU2, Icon, Rd, M)}$ or $\mbox{(CPU2, Icon, Rd, P)}$.  Consider another insight such that a new instance of flow $F_2$ as in Fig.~\ref{fig-cpu-read} can be initiated only after the completion of the previous instance of $F_2$.  If an instant of $F_2$ is assumed to be initiated by the super message $\#9~\mbox{(CPU2, Icon, (Wr, Rd), (M, P))}$ by interpreting it to $\mbox{(CPU2, Icon, Rd, M)}$ during the trace analysis, the super message~$\#12$ can only be interpreted to $\mbox{(CPU2, Icon, Wr, M)}$ or $\mbox{(CPU2, Icon, Rd, P)}$ as the instance of $F_2$ initiated by the super message $\#9$ has not been completed yet at this point.   According to the above discussion, the validators' insights help restrict how super messages are interpreted, thus reducing the number of flow execution scenarios that can be generated.
At the end of the analysis, all possible flow execution scenarios are returned to system validators for examination.

\section{Case Study II}

The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a SoC design from GEM5 with two ARM CPU cores, and one memory connected by an interconnect as shown in Fig 1. Since the analysis method presented in this paper is communication centric, the detailed computations of these blocks are not modeled. Instead, the modeling is focused on how they participate in flows for different system level use cases.

\begin{figure} 
\centerline{
\includegraphics[width=4in]{figures/Fig4.png}}
\caption{SoC platform structure.}
\label{fig_example1.ps}
\end{figure}

\subsection{Flow Specification}

In this case study, two system flows are implemented in the SoC model. They include a cache coherent read request from CPU and a cache coherent write request from CPU. In the following , a brief overview for the implemented instances of these flows is given. These flow specifications capture how messages are exchanged for different use cases. In this model, a message is defined with the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform.

\begin{figure} 
\centerline{
\includegraphics[width=3in]{figures/Fig5.png}}
\caption{Flow specification ($F_1$) of a cache coherent write operation initiated from CPU1}
\label{fig_example1.ps}
\end{figure}

The LPN as shown in Fig 2 specifies a system flow where CPU1 initiates a memory write operation. In this flow, CPU1 initiates a memory write request to L1 Cache. Next, depends on whether the required data is included in cache1, the cache1 will generate three possible responses. First, the cache1 will send read exclusive request message msg2 to interconnect if data not present in cache; Or if the data is shared by CPU0, it will send upgrade message msg16 to interconnect to disable CPU0's ownership of this block of data, else if CPU1 has exclusive right of required data, cache1 will perform the write operation and sent an response message msg15 to CPU1. Afterward, interconnect will sent request to all connected component (data cache and instruction cache of each CPU and memory). Once the interconnect get the response from either CPU0 or memory, it will generate a response message to cache1 and cache1 will generate a write response message msg14 (or msg 21 ) to CPU1. The flow is symmetric for CPU2.

The LPN specification as shown in Fig 3 captures the system flow where CPU1 initiate a memory read operation. Similar to memory write, CPU1 will first initiate a memory read request message msg1 to cache1. Cache1 can generate three possible responses. If the requested data is not in cache1, cache1 will generate an storeCondReq message msg2 to interconnect asking for data. Else if the requested data is inside of cache1 but it's shared, cache1 will generate a upgrade request message msg22 to make sure it has the newest data; Else if cache1 has and data and it's not shared, it will generate a read response message msg15 to CPU1. Afterward, interconnect will send corresponding request to all its connected component and when it received the response, generate response message to cache1 and cache1 will generate the read response message msg14 or msg34 to CPU1. This specification is also symmetric for CPU2.

\begin{figure} 
\centerline{
\includegraphics[width=3in]{figures/Fih6.png}}

\caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1}
\label{fig_example1.ps}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Results and Discussions}

The transaction level model of the simple system implementing the two flow specifications shown in the previous section is implemented by Gem5. We build a simple platform in Gem5 with components described in Fig 1. By running two simple executables on each CPU simultaneously, we produce a result file including all the communication between each components. The executable assigned to CPU1 is read one file three times for one letter and writing one letter for three times to the same file. CPU2's executable will do the same functionality to the same file, only difference is that CPU2 will write first and then read.

Even though this example is conceptually simple, getting the model to correctly implement the flow specifications is not straightforward. On the other hand, results from the trace analysis greatly help the debugging process by providing information to locate problems quickly. For example, the following trace is observed in Fig. 4.

\begin{figure} 
\centerline{
\includegraphics[width=4in]{figures/Fig.png}}
\caption{Example of trace result}
\end{figure}
\
The trace analysis finds out that messages 1-18 maps to two CPU0 read as
\vspace{2 mm}

$\{(F2^1,<p29>),  (F2^2, <p13>)\}$

\vspace{2 mm}

message 19 can be result of either flow instances, therefore it leads to two floowling flow execution scenarios:
\vspace{2 mm}

$\{(F2^1,<p30>),  (F2^2, <p13>)\}$

$\{(F2^1,<p29>),  (F2^2, <p14>)\}$
\vspace{2 mm}

In either scenario, message 20 can't be mapped to any existing flow instance or to a new flow instance. From this example, we can see that the error is originated by CPU0, and the problem is cache0 send the read response to CPU0 before it received any response from memory bus.%% In a few more iterations of analysis and debugging, the trace analysis can eventually extract all initiated flow instances in the model, all in their terminal states, thus showing that the model implements the two flows correctly.

%=====
\section{Related Work}

Our work is closely related to communication-centric and transaction based debug.  An early pioneering work is described in~\cite{Goossens2007NOCS}, which advocates the focus on observing activities on the interconnect network among IP blocks, and mapping these activities to transactions for better correlation between computations and communications.  Therefore, the communication transactions, as a result of software execution, provide an interface between computation and communication, and facilitate  system-level debug.  This work is extended in~\cite{Vermeulen2009VLSI-DAT,Goossens2009DATE}.  However, this line of work is focused on the network-on-chip (NoC) architecture for interconnect using the run/stop debug control method.

A similar transaction-based debug approach is presented in~\cite{Gharehbaghi2012ISQED}.  Furthermore, it proposes an automated extraction of state machines at transaction level from  high level design models.  From an observed failure trace, it performs backtracking on this transaction level state machine to derive a set of transaction traces that lead to the observed failure state.  In the subsequent step, bounded model checking with the constraints on the internal variables is used to refine the set of transaction traces to remove the infeasible traces.  This approach requires user inputs to identify impossible transaction sequences, and may not find the states causing the failure if the transaction traces leading to the observed failure state is long.  Backtracking from the observed failure state requires pre-image computation, which can be computationally expensive.  A transaction-based online debug approach is proposed in~\cite{Dehbash2014} to address these issues.  This approach utilizes a transaction debug pattern specification language~\cite{Gharehbaghi2009ICCD} to define properties that transactions should meet.  These transaction properties are checked at runtime by programming debug units in the on-chip debug infrastructure, and the system can be stopped shortly after a violation is detected for any one of those properties.  In this sense, it can be viewed as the hardware assertion approaches in~\cite{Boule2007ISQED} elevated to the transaction level.

In~\cite{Singerman2011DAC}, a coherent workflow is described where the result from the pre-silicon validation stage can be carried over to the post-silicon stage to improve efficiency and productivity of post-silicon debug.  This workflow is centered on a repository of system events and simple transactions defined by architects and IP designers.  It spans across a wide spectrum of the post-silicon validation including DFx instrumentation, test generation, coverage, and debug.  The DFx instruments are automatically inserted into the design RTL code driven by the defined transactions.  This instrumentation is optimized for making a large set of events and transactions observable.   Test generation is also optimized to generate only the necessary but sufficient tests to allow all defined transactions to be exercised.   Moreover, coverage for post-silicon validation is now defined at the abstract level of events and transactions rather than the raw signals, and thus can be evaluated more efficiently.  In~\cite{Abarbanel2014DAC}, a model at an even higher-level of abstraction, {\em flows}, is proposed.  Flows are used to specify more sophisticated cross-IP transactions such as power management, security, etc, and to facilitate reuse of the efforts of the architectural analysis to check HW/SW implementations.

%
%--
\section{Conclusion}

This paper presents a trace analysis based method for post-silicon validation by interpreting observed raw signal traces at the level of system flow specifications.  The derived flow execution scenarios provide more structured information on system operations, which is more understandable to system validators.   This information can help to locate design defects more easily, and also provides a measurement of validation coverage.

Due to partial observability, this approach may derive a large number of different flow execution scenarios for a given signal trace.  Insights from system validators can help to eliminate some false scenarios due to the partial observability.  An interesting future direction is formalization of the validators' insights using temporal logic on flows so that the validators can express their intents more precisely and concisely.

The trace analysis approach presented in this paper needs to be iterated with different observations selected in different iterations in order to eliminate the false scenarios and to root cause system failures as quickly as possible.  The observation selection and stitching signal traces of different observations together for the above goal will also be pursued in the future.

% conference papers do not normally have an appendix

% use section* for acknowledgement
%\section*{Acknowledgment}
%
%
%The authors would like to thank...
%

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{soc}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
%
%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
%\end{thebibliography}

% that's all folks
\end{document}

### 2015-08-16 paper.tex
graph changed, added more txt

@@
 \end{table*}
-The trace analysis finds out that messages~$1-8$ in the trace are the results of execution of an instance of flow $F_2$ as shown in Fig.~\ref{fig-cpu-read}, and the flow execution scenario is $\{(F_{2,1}, \{p_{11}, p_{12}\})\}$.
+The trace analysis finds out that messages~$1-8$ in the trace are the results of execution of an instance of flow $F_2$ as shown, and the flow execution scenario is $\{(F_{2,1}, \{p_{11}, p_{12}\})\}$.
 Messages~$9-11$ are analyzed as the results of executing another instance of flow $F_2$, message~$12-13$ as the results of executing an instance of flow $F_3$ as shown in Fig.~\ref{fig-peri-read}.
@@
 
-The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a SoC design from GEM5 with two ARM CPU cores, and one memory connected by an interconnect as shown in Fig 1.
+\begin{figure} 
+\centerline{
+\includegraphics[width=3.8in]{figures/Fig4.png}}
+\caption{SoC platform structure.}
+\label{SOC}
+\end{figure}
+
+The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a SoC design from GEM5 with two ARM CPU cores, each with 16KB data cache and 16KB instruction cache.
+One simple memory with 1GB capacity is connected with cache by an interconnect bus as shown in Fig~\ref{SOC}.
 Since the analysis method presented in this paper is communication centric, the detailed computations of these blocks are not modeled.
 Instead, the modeling is focused on how they participate in flows for different system level use cases.
-
-\begin{figure} 
-\centerline{
-\includegraphics[width=4in]{figures/Fig4.png}}
-\caption{SoC platform structure.}
-\label{fig_example1.ps}
-\end{figure}
+Originally all the component is connected by ports in GEM5, in order to trace the communication inside the system, we added an communicate monitor between each components connection.
+As showed in Fig~\ref{SOC}, take CPU0 and instruction cache for example, the communicate monitor between is called icache0.
+When CPU0 is about to send a command (read request as example), it will send to communicate monitor first, where it will store the request information in a file, and forward the request packet to instruction cache.
+
+By combining information from all 9 of the communicate monitors, we will have all the communication traces we need.
+As an virtualized SOC platform, GEM5 has two types of request, timing and atomic.
+Timing is what will be used in our system as it simulate delays generated by
+During our experiment, we took only timing request from the raw traces to eliminate initiation traces that are useless for our research.
 
@@
 In this model, a message is defined with the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform.
-
 \begin{figure} 
 \centerline{
-\includegraphics[width=3in]{figures/Fig5.png}}
+\includegraphics[width=3.7In]{figures/Fig5.png}}
+
+{\footnotesize
+\[
+\begin{array}{llll}
+msg_0: (&\mbox{ CPU1},&\mbox{writeReq},&\mbox{cache1   })\\       
+msg_1: (&\mbox{ cache1},&\mbox{ readExreq },&\mbox{Bus     })\\        
+msg_2: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2 })\\  
+msg_3: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2         })\\   
+msg_4: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2           })\\  
+msg_5: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2 })\\  
+msg_6: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce1       })\\     
+msg_7: (&\mbox{ cache1},&\mbox{readExreq},&\mbox{cpu1           })\\  
+msg_8: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ Memory })\\  
+msg_9: (&\mbox{ true                                          })\\  
+msg_{10}: (&\mbox{ Memory},&\mbox{ readExres},&\mbox{ Bus        })\\  
+msg_{11}: (&\mbox{ cache2},&\mbox{ readExres},&\mbox{ Bus })\\  
+msg_{12}: (&\mbox{ Bus},&\mbox{ readExres},&\mbox{ cache1        })\\  
+msg_{13}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1         })\\  
+msg_{14}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1 })\\  
+msg_{15}: (&\mbox{ cache1},&\mbox{UpgradeReq             })\\  
+msg_{16}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ cahce2      })\\   
+msg_{17}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ Memory })\\  
+msg_{18}: (&\mbox{ cache2},&\mbox{UpgradeRes},&\mbox{ Bus     })\\  
+msg_{19}: (&\mbox{ Bus},&\mbox{ UpgradeRes},&\mbox{ cache1      })\\  
+msg_{20}: (&\mbox{ cache1},&\mbox{ WriteRes},&\mbox{ CPU1 })\\  
+\end{array}
+\]}
 \caption{Flow specification ($F_1$) of a cache coherent write operation initiated from CPU1}
-\label{fig_example1.ps}
+\label{write-flow}
 \end{figure}
 
-The LPN as shown in Fig 2 specifies a system flow where CPU1 initiates a memory write operation.
+The LPN as shown in Fig~\ref{write-flow} specifies a system flow where CPU1 initiates a memory write operation.
 In this flow, CPU1 initiates a memory write request to L1 Cache.
@@
 
-The LPN specification as shown in Fig 3 captures the system flow where CPU1 initiate a memory read operation.
+\begin{figure} 
+\centerline{
+\includegraphics[width=4in]{figures/Fih6.png}}
+
+{\footnotesize
+\[
+\begin{array}{llll}
+t0: (&\mbox{ CPU1},&\mbox{ReadReq},&\mbox{cache1  })\\                   
+t1: (&\mbox{ cache1},&\mbox{ StoreCondreq },&\mbox{Bus })\\           
+t2: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2 })\\
+t3: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2       })\\      
+t4: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2           })\\ 
+t5: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2 })\\
+t6: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce1     })\\       
+t7: (&\mbox{ cache1},&\mbox{StoreCondreq},&\mbox{cpu1           })\\ 
+t8: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ Memory })\\
+t9: (&\mbox{ true                                        })\\
+t10: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus            })\\    
+t11: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus })\\
+t12: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1      })\\            
+t13: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1          })\\  
+t14: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
+t21: (&\mbox{ cache1},&\mbox{ LoadLockedreq },&\mbox{Bus })\\     
+t22: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\
+t23: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2 })\\
+t24: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\ 
+t25: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2     })\\
+t26: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce1 })\\
+t27: (&\mbox{ cache1},&\mbox{LoadLockedreq},&\mbox{cpu1       })\\
+t28: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ Memory     })\\
+t29: (&\mbox{ true })\\
+t30: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus       })\\     
+t31: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus    })\\
+t32: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1 })\\
+t33: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
+\end{array}
+\]}
+\caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1}
+\label{read-flow}
+
+\end{figure}
+The LPN specification as shown in Fig~\ref{read-flow} captures the system flow where CPU1 initiate a memory read operation.
 Similar to memory write, CPU1 will first initiate a memory read request message msg1 to cache1.
@@
 
-\begin{figure} 
-\centerline{
-\includegraphics[width=3in]{figures/Fih6.png}}
-
-\caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1}
-\label{fig_example1.ps}
-\end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The transaction level model of the simple system implementing the two flow specifications shown in the previous section is implemented by Gem5.
-We build a simple platform in Gem5 with components described in Fig 1.
+We build a simple platform in Gem5 with components described in Fig~\ref{SOC}.
 By running two simple executables on each CPU simultaneously, we produce a result file including all the communication between each components.
@@
 
+\begin{table*}
+\caption{{Example of trace result}}
+\label{example-trace}
+\begin{center}
+{\footnotesize
+
+\begin{tabular}{llll}
+1~~ (cpu0,cache0,readreq)				&2~~ (cache0,membus,LoadLockedreq)		&3~~ (membus,mem,LoadLockedreq) 		&4~~ (membus,cache0,LoadLockedreq)\\
+
+5~~(cache0,cpu0,LoadLockedreq)		&6~~(cpu0,cache0,readreq)		&7~~(cache0,membus,StoreCondreq)		&8~~(membus,mem,StoreCondreq)\\
+
+9~~(membus,cache0,StoreCondreq)		&10~~(cache0,cpu0,StoreCondreq)		&11~~(membus,cache1,StoreCondreq)		&12~~(cache1,cpu1,StoreCondreq)\\
+
+13~~(membus,cache1,StoreCondreq)		&14~~(cache1,cpu1,StoreCondreq)		&15~~(membus,cache1,LoadLockedreq)	&16~~(cache1,cpu1,LoadLockedreq)\\
+
+17~~(membus,cache1,LoadLockedreq)	&18~~(cache1,cpu1,LoadLockedreq)		&19~~(mem,membus,readres)			&20~~(cache0,cpu0,readres)\\
+
+20~~(membus,cache0,readres)			&22~~(mem,membus,readres)			&23~~(membus,cache0,readres)				&24~~(cache0,cpu0,readres)\\
+\end{tabular}}
+\end{center}
+\end{table*}
+
 Even though this example is conceptually simple, getting the model to correctly implement the flow specifications is not straightforward.
 On the other hand, results from the trace analysis greatly help the debugging process by providing information to locate problems quickly.
-For example, the following trace is observed in Fig.
-4.
-
-\begin{figure} 
-\centerline{
-\includegraphics[width=4in]{figures/Fig.png}}
-\caption{Example of trace result}
-\end{figure}
+For example, the following trace as shown in Table~\ref{example-trace} is observed.
 \

### 2015-08-17 paper.tex
minor

@@
 By combining information from all 9 of the communicate monitors, we will have all the communication traces we need.
-As an virtualized SOC platform, GEM5 has two types of request, timing and atomic.
-Timing is what will be used in our system as it simulate delays generated by
-During our experiment, we took only timing request from the raw traces to eliminate initiation traces that are useless for our research.
+As an virtualized SOC platform, GEM5 has three types of request: timing, atomic and functional.
+Timing request are the most detailed request.
+This type of request include the modeling of queuing delay and resource contention.
+A successfully sent timing will receive an response at some point in the future.
+Atomic accesses are a faster than detailed access.
+They are used for fast forwarding and warming up caches and return an approximate time to complete the request without any resource contention or queuing delay.
+Functional accesses are used for things such as loading binaries, examining/changing variables in the simulated system, and allowing a remote debugger to be attached to the simulator.
+In our case, we specified all our request to be timing.
+However, some of the system initiation is still atomic and it's useless for our research, so we only took the messages that are timing request from our collected data.
+(From GEM5 online documentaries)
 
@@
 They include a cache coherent read request from CPU and a cache coherent write request from CPU.
+To avoid messages that are not included in this two flow, like write back, cache flushing, we only use data that is related to these two flow.
+
 In the following , a brief overview for the implemented instances of these flows is given.
@@
 \subsection{Results and Discussions}
-
-The transaction level model of the simple system implementing the two flow specifications shown in the previous section is implemented by Gem5.
-We build a simple platform in Gem5 with components described in Fig~\ref{SOC}.
-By running two simple executables on each CPU simultaneously, we produce a result file including all the communication between each components.
+We produce a result file including all the communication between each components by running two simple programs on each CPU simultaneously, .
 The executable assigned to CPU1 is read one file three times for one letter and writing one letter for three times to the same file.
 CPU2's executable will do the same functionality to the same file, only difference is that CPU2 will write first and then read.
+One thing we should know about GEM5 is that even if we specify these two program two be ran concurrently, and GEM5 try to produce an concurrent result.
+The data we collected shows that it's not the real nondeterministic concurrency.
+What GEM5 did is letting the two CPU execute one instruction in turn, so the order is determined, no matter how many time we run the program, it will always produce the same result.
+We tried other virtual SOC platform softwares, and this is the best nondeterministic concurrency we can get.
+Even if this is slightly off with what the real chip should work, it still suit our purpose of testing the correctness of our algorithm.
+
+As we mentioned before, we will take away unrelated messages for easier testing purpose.
+After taking out non-timing request, the number of communication messages reduced from 343581 to 133676, and after removing the message that's not related to tested flow specification, number of messages became 120949.
 

### 2015-08-17 paper.tex
Grammar mistake

--- 2015-08-16 paper.tex
@@
 %% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
+%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compSoC.tex
 %%*************************************************************************
@@
 \documentclass[conference]{IEEEtran}
-% Add the compsoc option for Computer Society conferences.
+% Add the compSoC option for Computer SoCiety conferences.
 %
@@
 %\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
+% A popular package from the American Mathematical SoCiety that provides
 % many useful and powerful commands for dealing with mathematics. If using
@@
 In this and the following figures for LPNs, the labeled circles denote places, and the labeled boxes denote transitions.
-Each transition is labeled with its name and the associated event.
+Each transition is labeled with its name and the asSoCiated event.
 The solid black places without outgoing edges are {\em terminals}, which indicate termination of protocols represented by the LPNs.
@@
 
-The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a simple SoC design with two CPU cores, memory, and a peripheral device connected by an interconnect as shown in Fig.~\ref{fig-soc-ex}.
+The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a simple SoC design with two CPU cores, memory, and a peripheral device connected by an interconnect as shown in Fig.~\ref{fig-SoC-ex}.
 Since the analysis method presented in this paper is communication centric, the detailed computations of these blocks are not modeled.
@@
 \caption{A simple SoC example.}
-\label{fig-soc-ex}
+\label{fig-SoC-ex}
 \end{center}
@@
 \caption{SoC platform structure.}
-\label{SOC}
+\label{SoC}
 \end{figure}
@@
 The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a SoC design from GEM5 with two ARM CPU cores, each with 16KB data cache and 16KB instruction cache.
-One simple memory with 1GB capacity is connected with cache by an interconnect bus as shown in Fig~\ref{SOC}.
+A simple memory controller with 1GB capacity is connected with CPUs and caches by an interconnect bus as shown in Fig.~\ref{SoC}.
 Since the analysis method presented in this paper is communication centric, the detailed computations of these blocks are not modeled.
-Instead, the modeling is focused on how they participate in flows for different system level use cases.
-Originally all the component is connected by ports in GEM5, in order to trace the communication inside the system, we added an communicate monitor between each components connection.
-As showed in Fig~\ref{SOC}, take CPU0 and instruction cache for example, the communicate monitor between is called icache0.
-When CPU0 is about to send a command (read request as example), it will send to communicate monitor first, where it will store the request information in a file, and forward the request packet to instruction cache.
-
-By combining information from all 9 of the communicate monitors, we will have all the communication traces we need.
-As an virtualized SOC platform, GEM5 has two types of request, timing and atomic.
-Timing is what will be used in our system as it simulate delays generated by
-During our experiment, we took only timing request from the raw traces to eliminate initiation traces that are useless for our research.
+Instead, the modeling focuses on how they participate in flows for different system level use cases.
+Originally all of the components are connected by ports in GEM5.
+Ports exist in peers, and by connecting them, one component can send request or response through ports.
+In order to trace the communication inside the system, we added an communication monitor between each components connection.
+The communication monitor is a MemObject that monitors the communication  between two ports in the system.
+
+By combining the informations from all of the nine communication monitors, required communication traces are obtained.
+As a virtualized SoC platform, GEM5 has three types of request: timing, atomic and functional.
+Timing request include the modeling of queuing delay and resource contention.
+Atomic request is a faster than timing request with no delay.
+Functional request is used for loading binaries, examining/changing variables in the simulated system, and so on.
+In our case, we specify all our request to be timed as it's the most detailed one.However, some of the system initiation is still atomic and it's not related to our research, so we only took the messages that are timing request from our collected data.
 
@@
 They include a cache coherent read request from CPU and a cache coherent write request from CPU.
-In the following , a brief overview for the implemented instances of these flows is given.
+To avoid messages that are not included in these two flows, like write back and cache flushing, we only use data that are related to these two flows.
+
+In the following section, a brief overview for the implemented instances of these flows are given.
 These flow specifications capture how messages are exchanged for different use cases.
@@
 
-The LPN as shown in Fig~\ref{write-flow} specifies a system flow where CPU1 initiates a memory write operation.
+The LPN as shown in Fig.~\ref{write-flow} specifies a system flow where CPU1 initiates a memory write operation.
 In this flow, CPU1 initiates a memory write request to L1 Cache.
 Next, depends on whether the required data is included in cache1, the cache1 will generate three possible responses.
-First, the cache1 will send read exclusive request message msg2 to interconnect if data not present in cache; Or if the data is shared by CPU0, it will send upgrade message msg16 to interconnect to disable CPU0's ownership of this block of data, else if CPU1 has exclusive right of required data, cache1 will perform the write operation and sent an response message msg15 to CPU1.
-Afterward, interconnect will sent request to all connected component (data cache and instruction cache of each CPU and memory).
-Once the interconnect get the response from either CPU0 or memory, it will generate a response message to cache1 and cache1 will generate a write response message msg14 (or msg 21 ) to CPU1.
+First, the cache1 will send read exclusive request message msg2 to interconnect if data is not present in cache.; Or if the data is shared by CPU0, it will send upgrade message msg16 to interconnect to disable CPU0's ownership of this block of data, else if CPU1 has exclusive right of required data, cache1 will perform the write operation and sent an response message msg15 to CPU1.
+Afterwards, interconnect will sent request to all connected component (data cache and instruction cache of each CPU and memory).
+Once the interconnect obtains the response from either CPU0 or memory, it will generate a response message to cache1, then cache1 will generate a write response message msg14 (or msg 21 ) to CPU1.
 The flow is symmetric for CPU2.
@@
 \end{figure}
-The LPN specification as shown in Fig~\ref{read-flow} captures the system flow where CPU1 initiate a memory read operation.
+The LPN specification as shown in Fig.~\ref{read-flow} captures the system flow where CPU1 initiate a memory read operation.
 Similar to memory write, CPU1 will first initiate a memory read request message msg1 to cache1.
 Cache1 can generate three possible responses.
-If the requested data is not in cache1, cache1 will generate an storeCondReq message msg2 to interconnect asking for data.
-Else if the requested data is inside of cache1 but it's shared, cache1 will generate a upgrade request message msg22 to make sure it has the newest data; Else if cache1 has and data and it's not shared, it will generate a read response message msg15 to CPU1.
-Afterward, interconnect will send corresponding request to all its connected component and when it received the response, generate response message to cache1 and cache1 will generate the read response message msg14 or msg34 to CPU1.
+First, if the requested data is not in cache1, cache1 will generate an storeCondReq message msg2 to interconnect asking for data.
+Second, if the requested data is inside of cache1 but it's shared, cache1 will generate a upgrade request message msg22 to make sure it has the newest data; Last,  if cache1 has and data and it's not shared, it will generate a read response message msg15 to CPU1.
+Afterwards, interconnect will send corresponding request to all of its connected components.
+When interconnect receives the response, it will generate response message to cache1.
+Then cache1 will generate the read response message msg14 or msg34 to CPU1.
 This specification is also symmetric for CPU2.
@@
 \subsection{Results and Discussions}
-
-The transaction level model of the simple system implementing the two flow specifications shown in the previous section is implemented by Gem5.
-We build a simple platform in Gem5 with components described in Fig~\ref{SOC}.
-By running two simple executables on each CPU simultaneously, we produce a result file including all the communication between each components.
-The executable assigned to CPU1 is read one file three times for one letter and writing one letter for three times to the same file.
-CPU2's executable will do the same functionality to the same file, only difference is that CPU2 will write first and then read.
+We produce a result file including all of the communication messages between every components from two simple programs running simultaneously on each CPU.
+The program assigned to CPU1 read one file three times for one letter and writing one letter for three times to the same file.
+CPU2's program will do the same read and write functionalities to the same file, only difference is that CPU2 will write first and then read.
+One thing we should know about GEM5 is that even when we run these two programs concurrently, it will attempt to produce a concurrent result.
+The data we collected shows that it's not the real nondeterministic concurrency.
+What GEM5 did was it allowed two CPUs to execute its own instruction in turn, therefore the order was deterministic.
+Therefore, no matter how many times we ran the program, it produced the same result.
+We tried other virtual SoC platform softwares, and this was the best nondeterministic concurrency we can get.
+Even if this is slightly off with what the real chip should work, it still surve our purpose of testing the correctness of our algorithm.
+
+As we mentioned before, we will take away unrelated messages for easier testing purpose.
+After taking out non-timing request, the number of communication messages reduced from 343581 to 133676, and after removing the message that's not related to tested flow specification, number of messages became 120949.
+
+The total time we used for processing the filtered data is 3 seconds, with 12MB data usage.
 
@@
 % argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{soc}
+\bibliography{SoC}
 %

### 2015-08-17 paper.tex
graph changed, some redundant deleted

@@
 \centerline{
-\includegraphics[width=3.8in]{figures/Fig4.png}}
+\includegraphics[width=3.7in]{figures/Fig4.png}}
 \caption{SoC platform structure.}
@@
 
-The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a SoC design from GEM5 with two ARM CPU cores, each with 16KB data cache and 16KB instruction cache.
+The proposed trace analysis method is demonstrated on a transaction level model of a SoC design from GEM5.
+The SoC platform built for this case study has two ARM CPU cores, each with 16KB data cache and 16KB instruction cache.
 A simple memory controller with 1GB capacity is connected with CPUs and caches by an interconnect bus as shown in Fig.~\ref{SoC}.
-Since the analysis method presented in this paper is communication centric, the detailed computations of these blocks are not modeled.
-Instead, the modeling focuses on how they participate in flows for different system level use cases.
+This modeling focuses on how they participate in flows for different system level use cases.
 Originally all of the components are connected by ports in GEM5.
@@
 In order to trace the communication inside the system, we added an communication monitor between each components connection.
-The communication monitor is a MemObject that monitors the communication  between two ports in the system.
+The communication monitor is an Object that monitors the communication  between two ports in the system.
+The communication monitor will not affect any functionality of the system, it only record the communication activities into an output file.
 
@@
 As we mentioned before, we will take away unrelated messages for easier testing purpose.
-After taking out non-timing request, the number of communication messages reduced from 343581 to 133676, and after removing the message that's not related to tested flow specification, number of messages became 120949.
+After removing the message that's not related to tested flow specification, the number of communication messages reduced from 343581 to 120949.
 
 The total time we used for processing the filtered data is 3 seconds, with 12MB data usage.
-
-\begin{table*}
-\caption{{Example of trace result}}
-\label{example-trace}
-\begin{center}
-{\footnotesize
-
-\begin{tabular}{llll}
-1~~ (cpu0,cache0,readreq)				&2~~ (cache0,membus,LoadLockedreq)		&3~~ (membus,mem,LoadLockedreq) 		&4~~ (membus,cache0,LoadLockedreq)\\
-
-5~~(cache0,cpu0,LoadLockedreq)		&6~~(cpu0,cache0,readreq)		&7~~(cache0,membus,StoreCondreq)		&8~~(membus,mem,StoreCondreq)\\
-
-9~~(membus,cache0,StoreCondreq)		&10~~(cache0,cpu0,StoreCondreq)		&11~~(membus,cache1,StoreCondreq)		&12~~(cache1,cpu1,StoreCondreq)\\
-
-13~~(membus,cache1,StoreCondreq)		&14~~(cache1,cpu1,StoreCondreq)		&15~~(membus,cache1,LoadLockedreq)	&16~~(cache1,cpu1,LoadLockedreq)\\
-
-17~~(membus,cache1,LoadLockedreq)	&18~~(cache1,cpu1,LoadLockedreq)		&19~~(mem,membus,readres)			&20~~(cache0,cpu0,readres)\\
-
-20~~(membus,cache0,readres)			&22~~(mem,membus,readres)			&23~~(membus,cache0,readres)				&24~~(cache0,cpu0,readres)\\
-\end{tabular}}
-\end{center}
-\end{table*}
-
-Even though this example is conceptually simple, getting the model to correctly implement the flow specifications is not straightforward.
-On the other hand, results from the trace analysis greatly help the debugging process by providing information to locate problems quickly.
-For example, the following trace as shown in Table~\ref{example-trace} is observed.
-\
-The trace analysis finds out that messages 1-18 maps to two CPU0 read as
-\vspace{2 mm}
-
-$\{(F2^1,<p29>),  (F2^2, <p13>)\}$
-
-\vspace{2 mm}
-
-message 19 can be result of either flow instances, therefore it leads to two floowling flow execution scenarios:
-\vspace{2 mm}
-
-$\{(F2^1,<p30>),  (F2^2, <p13>)\}$
-
-$\{(F2^1,<p29>),  (F2^2, <p14>)\}$
-\vspace{2 mm}
-
-In either scenario, message 20 can't be mapped to any existing flow instance or to a new flow instance.
-From this example, we can see that the error is originated by CPU0, and the problem is cache0 send the read response to CPU0 before it received any response from memory bus.%% In a few more iterations of analysis and debugging, the trace analysis can eventually extract all initiated flow instances in the model, all in their terminal states, thus showing that the model implements the two flows correctly.
+The 120949 messages matche to 3370 CPU write flows generated from CPU1 and 3308 CPU write flows from CPU2.
+There are also 21584 CPU read flows from CPU1 and 21341 CPU read flows generated from CPU2.
+The number of read and write generated from each CPU is very similar as both of them are doing the same times of read and write functions.
 

### 2015-08-19 paper.tex
Revert 5fad844..ce4a0a9
This rolls back to commit 5fad844267858dc27d42494b92fe89793a098172.

@@
 
+%
+%--
+%
 \section{Case Study II}
@@
 \centerline{
-\includegraphics[width=3.7in]{figures/Fig4.png}}
+\includegraphics[width=3.3in]{figures/Fig4.png}}
 \caption{SoC platform structure.}
@@
 
-The proposed trace analysis method is demonstrated on a transaction level model of a SoC design from GEM5.
-The SoC platform built for this case study has two ARM CPU cores, each with 16KB data cache and 16KB instruction cache.
-A simple memory controller with 1GB capacity is connected with CPUs and caches by an interconnect bus as shown in Fig.~\ref{SoC}.
-This modeling focuses on how they participate in flows for different system level use cases.
-Originally all of the components are connected by ports in GEM5.
-Ports exist in peers, and by connecting them, one component can send request or response through ports.
-In order to trace the communication inside the system, we added an communication monitor between each components connection.
-The communication monitor is an Object that monitors the communication  between two ports in the system.
-The communication monitor will not affect any functionality of the system, it only record the communication activities into an output file.
-
-By combining the informations from all of the nine communication monitors, required communication traces are obtained.
+In order to find out the efficiency of the trace analysis method for more realistic examples, in this case study, a more detailed transaction level model of a SoC is constructed within the GEM5 environment~\cite{Binkert2011}.
+This SoC model consists of two ARM Cortex-A9 cores, each of which contains two separate 16KB data and instruction caches.
+The caches are connected to a 1GB memory through a memory bus model.
+The architecture of this SoC model is shown in Fig.~\ref{SoC}.
+
+In this model, components communicate with each other by sending and receiving various request and response messages.
+In order to observe and trace communications occurring inside this model during execution, monitors are attached to links connecting the components.
+These monitors record the messages flowing through the links they are attached to, and store them into output trace files.
+
+%By combining the informations from all of the nine communication monitors, required communication traces are obtained.
 As a virtualized SoC platform, GEM5 has three types of request: timing, atomic and functional.
@@
 
-\subsection{Flow Specification}
-
-In this case study, two system flows are implemented in the SoC model.
-They include a cache coherent read request from CPU and a cache coherent write request from CPU.
-To avoid messages that are not included in these two flows, like write back and cache flushing, we only use data that are related to these two flows.
-
-In the following section, a brief overview for the implemented instances of these flows are given.
+%\subsection{Flow Specification}
+
+For this model, we consider the flow specification describing the cache coherence protocol supported in GEM5 that is used to build the model in Fig~\ref{SoC}.
+This specification consists of two flows, one for cache coherent read requests and one for  cache coherent write requests, from CPUs.
+Two such flows describe the cache coherent protocol for each CPU.
+As there are two CPUs in this model, there are actually four such flows considered for this model.
+More detailed information about the flow specification can be found at \red{github???}
+
+%In the following section, a brief overview for the implemented instances of these flows are given.
 These flow specifications capture how messages are exchanged for different use cases.
 In this model, a message is defined with the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform.
-\begin{figure} 
-\centerline{
-\includegraphics[width=3.7In]{figures/Fig5.png}}
-
-{\footnotesize
-\[
-\begin{array}{llll}
-msg_0: (&\mbox{ CPU1},&\mbox{writeReq},&\mbox{cache1   })\\       
-msg_1: (&\mbox{ cache1},&\mbox{ readExreq },&\mbox{Bus     })\\        
-msg_2: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2 })\\  
-msg_3: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2         })\\   
-msg_4: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2           })\\  
-msg_5: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2 })\\  
-msg_6: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce1       })\\     
-msg_7: (&\mbox{ cache1},&\mbox{readExreq},&\mbox{cpu1           })\\  
-msg_8: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ Memory })\\  
-msg_9: (&\mbox{ true                                          })\\  
-msg_{10}: (&\mbox{ Memory},&\mbox{ readExres},&\mbox{ Bus        })\\  
-msg_{11}: (&\mbox{ cache2},&\mbox{ readExres},&\mbox{ Bus })\\  
-msg_{12}: (&\mbox{ Bus},&\mbox{ readExres},&\mbox{ cache1        })\\  
-msg_{13}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1         })\\  
-msg_{14}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1 })\\  
-msg_{15}: (&\mbox{ cache1},&\mbox{UpgradeReq             })\\  
-msg_{16}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ cahce2      })\\   
-msg_{17}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ Memory })\\  
-msg_{18}: (&\mbox{ cache2},&\mbox{UpgradeRes},&\mbox{ Bus     })\\  
-msg_{19}: (&\mbox{ Bus},&\mbox{ UpgradeRes},&\mbox{ cache1      })\\  
-msg_{20}: (&\mbox{ cache1},&\mbox{ WriteRes},&\mbox{ CPU1 })\\  
-\end{array}
-\]}
-\caption{Flow specification ($F_1$) of a cache coherent write operation initiated from CPU1}
-\label{write-flow}
-\end{figure}
-
-The LPN as shown in Fig.~\ref{write-flow} specifies a system flow where CPU1 initiates a memory write operation.
+%\begin{figure} 
+%\centerline{
+%\includegraphics[width=3.7In]{figures/Fig5.png}}
+%
+%{\footnotesize
+%\[
+%\begin{array}{llll}
+%msg_0: (&\mbox{ CPU1},&\mbox{writeReq},&\mbox{cache1   })\\       
+%msg_1: (&\mbox{ cache1},&\mbox{ readExreq },&\mbox{Bus     })\\        
+%msg_2: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2 })\\  
+%msg_3: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2         })\\   
+%msg_4: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2           })\\  
+%msg_5: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2 })\\  
+%msg_6: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce1       })\\     
+%msg_7: (&\mbox{ cache1},&\mbox{readExreq},&\mbox{cpu1           })\\  
+%msg_8: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ Memory })\\  
+%msg_9: (&\mbox{ true                                          })\\  
+%msg_{10}: (&\mbox{ Memory},&\mbox{ readExres},&\mbox{ Bus        })\\  
+%msg_{11}: (&\mbox{ cache2},&\mbox{ readExres},&\mbox{ Bus })\\  
+%msg_{12}: (&\mbox{ Bus},&\mbox{ readExres},&\mbox{ cache1        })\\  
+%msg_{13}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1         })\\  
+%msg_{14}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1 })\\  
+%msg_{15}: (&\mbox{ cache1},&\mbox{UpgradeReq             })\\  
+%msg_{16}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ cahce2      })\\   
+%msg_{17}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ Memory })\\  
+%msg_{18}: (&\mbox{ cache2},&\mbox{UpgradeRes},&\mbox{ Bus     })\\  
+%msg_{19}: (&\mbox{ Bus},&\mbox{ UpgradeRes},&\mbox{ cache1      })\\  
+%msg_{20}: (&\mbox{ cache1},&\mbox{ WriteRes},&\mbox{ CPU1 })\\  
+%\end{array}
+%\]}
+%\caption{Flow specification ($F_1$) of a cache coherent write operation initiated from CPU1}
+%\label{write-flow}
+%\end{figure}
+%
+%The LPN as shown in Fig.~\ref{write-flow} specifies a system flow where CPU1 initiates a memory write operation.
 In this flow, CPU1 initiates a memory write request to L1 Cache.
@@
 The flow is symmetric for CPU2.
-
-\begin{figure} 
-\centerline{
-\includegraphics[width=4in]{figures/Fih6.png}}
-
-{\footnotesize
-\[
-\begin{array}{llll}
-t0: (&\mbox{ CPU1},&\mbox{ReadReq},&\mbox{cache1  })\\                   
-t1: (&\mbox{ cache1},&\mbox{ StoreCondreq },&\mbox{Bus })\\           
-t2: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2 })\\
-t3: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2       })\\      
-t4: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2           })\\ 
-t5: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2 })\\
-t6: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce1     })\\       
-t7: (&\mbox{ cache1},&\mbox{StoreCondreq},&\mbox{cpu1           })\\ 
-t8: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ Memory })\\
-t9: (&\mbox{ true                                        })\\
-t10: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus            })\\    
-t11: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus })\\
-t12: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1      })\\            
-t13: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1          })\\  
-t14: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
-t21: (&\mbox{ cache1},&\mbox{ LoadLockedreq },&\mbox{Bus })\\     
-t22: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\
-t23: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2 })\\
-t24: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\ 
-t25: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2     })\\
-t26: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce1 })\\
-t27: (&\mbox{ cache1},&\mbox{LoadLockedreq},&\mbox{cpu1       })\\
-t28: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ Memory     })\\
-t29: (&\mbox{ true })\\
-t30: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus       })\\     
-t31: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus    })\\
-t32: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1 })\\
-t33: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
-\end{array}
-\]}
-\caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1}
-\label{read-flow}
-
-\end{figure}
-The LPN specification as shown in Fig.~\ref{read-flow} captures the system flow where CPU1 initiate a memory read operation.
+%
+%
+%\begin{figure} 
+%\centerline{
+%\includegraphics[width=4in]{figures/Fih6.png}}
+%
+%{\footnotesize
+%\[
+%\begin{array}{llll}
+%t0: (&\mbox{ CPU1},&\mbox{ReadReq},&\mbox{cache1  })\\                   
+%t1: (&\mbox{ cache1},&\mbox{ StoreCondreq },&\mbox{Bus })\\           
+%t2: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2 })\\
+%t3: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2       })\\      
+%t4: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2           })\\ 
+%t5: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2 })\\
+%t6: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce1     })\\       
+%t7: (&\mbox{ cache1},&\mbox{StoreCondreq},&\mbox{cpu1           })\\ 
+%t8: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ Memory })\\
+%t9: (&\mbox{ true                                        })\\
+%t10: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus            })\\    
+%t11: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus })\\
+%t12: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1      })\\            
+%t13: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1          })\\  
+%t14: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
+%t21: (&\mbox{ cache1},&\mbox{ LoadLockedreq },&\mbox{Bus })\\     
+%t22: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\
+%t23: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2 })\\
+%t24: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\ 
+%t25: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2     })\\
+%t26: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce1 })\\
+%t27: (&\mbox{ cache1},&\mbox{LoadLockedreq},&\mbox{cpu1       })\\
+%t28: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ Memory     })\\
+%t29: (&\mbox{ true })\\
+%t30: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus       })\\     
+%t31: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus    })\\
+%t32: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1 })\\
+%t33: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
+%\end{array}
+%\]}
+%\caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1}
+%\label{read-flow}
+%
+%\end{figure}
+%The LPN specification as shown in Fig.~\ref{read-flow} captures the system flow where CPU1 initiate a memory read operation.
 Similar to memory write, CPU1 will first initiate a memory read request message msg1 to cache1.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Results and Discussions}
-We produce a result file including all of the communication messages between every components from two simple programs running simultaneously on each CPU.
+%\subsection{Results and Discussions}
+
+In order to trigger cache coherence protocols, a simple two-threaded program including one producer and one consumer is created and executed on the CPUs, each of which executes one thread.
+\red{How is shared memory is simulated???}  GEM5 does not support true concurrency.
+When there are two programs running on the CPUs, GEM5 alternates the executions between the two  CPUs.
+
+%We produce a result file including all of the communication messages between every components from two simple programs running simultaneously on each CPU.
 The program assigned to CPU1 read one file three times for one letter and writing one letter for three times to the same file.
@@
 
-As we mentioned before, we will take away unrelated messages for easier testing purpose.
-After removing the message that's not related to tested flow specification, the number of communication messages reduced from 343581 to 120949.
-
-The total time we used for processing the filtered data is 3 seconds, with 12MB data usage.
-The 120949 messages matche to 3370 CPU write flows generated from CPU1 and 3308 CPU write flows from CPU2.
-There are also 21584 CPU read flows from CPU1 and 21341 CPU read flows generated from CPU2.
-The number of read and write generated from each CPU is very similar as both of them are doing the same times of read and write functions.
+After the SoC model finishes executing the program, there are totally $343581$ messages collected in the trace file.
+Not all of the messages are relevant to the flow specification as many are used by GEM5 to initialize its simulation environment.
+After removing those irrelevant messages, the number of messages in the trace file is to reduced to $120949$.
+
+The time taken to remove the irrelevant messages from the trace is negligible.
+The total runtime and the peak memory usage for the trace analysis algorithm to finish the reduced trace are 3 seconds and 12MB, respectively.
+From the trace, Table~\ref{table-case-2} shows the number of instances extracted for the four flows describing cache coherent read/write operations.
+\begin{table}[td]
+\caption{The number of flow instances derived by the trace analysis}
+\begin{center}
+\begin{tabular}{|c|c|}
+\hline
+Flows & $\#$Instances \\
+\hline
+\hline
+CPU1 Read		&  $21584$\\
+CPU1 Write	&  $3370$\\
+CPU2 Read		&  $21341$\\
+CPU2 Write	&  $3308$\\
+\hline
+\end{tabular}
+\end{center}
+\label{table-case-2}
+\end{table}%
+
+\red{Also need to consider the case where some monitors are removed?}
 

### 2015-08-19 paper.tex
changed graph

@@
 As there are two CPUs in this model, there are actually four such flows considered for this model.
-More detailed information about the flow specification can be found at \red{github???}
+%More detailed information about the flow specification is explained in the following section.
+
+%found at \red{github???}
 
@@
 In order to trigger cache coherence protocols, a simple two-threaded program including one producer and one consumer is created and executed on the CPUs, each of which executes one thread.
-\red{How is shared memory is simulated???}  GEM5 does not support true concurrency.
+\red{How is shared memory is simulated???} .
+The producer will write to a specified memory address, and the consumer will read from the same address.
+GEM5 does not support true concurrency.
 When there are two programs running on the CPUs, GEM5 alternates the executions between the two  CPUs.

### 2015-08-21 Flow Specification.tex
added documentation
flow specification

\documentclass[conference]{IEEEtran}

\usepackage{pifont}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{mybeamer}
\usepackage{draftwatermark}

\begin{document}

\title{Flow Specification in GEM5}

\maketitle

\begin{abstract}
This paper will go through a brief overview for the implemented instances of these flows are given. These flow specifications capture how messages are exchanged for different use cases. In this model, a message is defined with the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform. 
\end{abstract}

\section{write flow specification}

 \begin{figure} 
 \centerline{
 \includegraphics[width=3.4In]{figures/Fig5.png}}

 {\footnotesize
 \[
 \begin{array}{llll}
 msg_0: (&\mbox{ CPU1},&\mbox{writeReq},&\mbox{icache1   })\\       
 msg_1: (&\mbox{ dcache1},&\mbox{ readExreq },&\mbox{Bus     })\\        
 msg_2: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ dcahce2 })\\  
 msg_3: (&\mbox{ dcache2},&\mbox{readExreq},&\mbox{cpu2         })\\   
 msg_4: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ icahce2           })\\  
 msg_5: (&\mbox{ icache2},&\mbox{readExreq},&\mbox{cpu2 })\\  
 msg_6: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ icahce1       })\\     
 msg_7: (&\mbox{ dcache1},&\mbox{readExreq},&\mbox{cpu1           })\\  
 msg_8: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ Memory })\\  
 msg_9: (&\mbox{ true                                          })\\  
 msg_{10}: (&\mbox{ Memory},&\mbox{ readExres},&\mbox{ Bus        })\\  
 msg_{11}: (&\mbox{ icache2},&\mbox{ readExres},&\mbox{ Bus })\\  
 msg_{12}: (&\mbox{ Bus},&\mbox{ readExres},&\mbox{ dcache1        })\\  
 msg_{13}: (&\mbox{ icache1},&\mbox{ writeRes},&\mbox{ CPU1         })\\  
 msg_{14}: (&\mbox{ icache1},&\mbox{ writeRes},&\mbox{ CPU1 })\\  
 msg_{15}: (&\mbox{ dcache1},&\mbox{UpgradeReq             })\\  
 msg_{16}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ icahce2      })\\   
 msg_{17}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ Memory })\\  
 msg_{18}: (&\mbox{ icache2},&\mbox{UpgradeRes},&\mbox{ Bus     })\\  
 msg_{19}: (&\mbox{ Bus},&\mbox{ UpgradeRes},&\mbox{ icache1      })\\  
 msg_{20}: (&\mbox{ icache1},&\mbox{ WriteRes},&\mbox{ CPU1 })\\  
 \end{array}
 \]}
 \caption{Flow specification ($F_1$) of a cache coherent write operation initiated from CPU1}
 \label{write-flow}
 \end{figure}

 The LPN as shown in Fig.~\ref{write-flow} specifies a system flow where CPU1 initiates a memory write operation. In this flow, CPU1 initiates a memory write request to L1 Cache. Next, depends on whether the required data is included in cache1, the cache1 will generate three possible responses. First, the cache1 will send read exclusive request message msg2 to interconnect if data is not present in cache; Or if the data is shared by CPU0, it will send upgrade message msg16 to interconnect to disable CPU0's ownership of this block of data, else if CPU1 has exclusive right of required data, cache1 will perform the write operation and sent an response message msg15 to CPU1. Afterwards, interconnect will sent request to all connected component (data cache and instruction cache of each CPU and memory). Once the interconnect obtains the response from either CPU0 or memory, it will generate a response message to cache1, then cache1 will generate a write response message msg14 (or msg 21 ) to CPU1. The flow is symmetric for CPU2.

 \section{read flow specification}

 \begin{figure} 
 \centerline{
 \includegraphics[width=3.33in]{figures/Fih6.png}}

 {\footnotesize
 \[
 \begin{array}{llll}
 msg0: (&\mbox{ CPU1},&\mbox{ReadReq},&\mbox{icache1  })\\                   
 msg1: (&\mbox{ dcache1},&\mbox{ StoreCondreq },&\mbox{Bus })\\           
 msg2: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ icahce2 })\\
 msg3: (&\mbox{ icache2},&\mbox{StoreCondreq},&\mbox{cpu2       })\\      
 msg4: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ dcahce2           })\\ 
 msg5: (&\mbox{ dcache2},&\mbox{StoreCondreq},&\mbox{cpu2 })\\
 msg6: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ dcahce1     })\\       
 msg7: (&\mbox{ icache1},&\mbox{StoreCondreq},&\mbox{cpu1           })\\ 
 msg8: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ Memory })\\
 msg9: (&\mbox{ true                                        })\\
 msg10: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus            })\\    
 msg11: (&\mbox{ icache2},&\mbox{ ReadRes},&\mbox{ Bus })\\
 msg12: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ dcache1      })\\            
 msg13: (&\mbox{ icache1},&\mbox{ ReadRes},&\mbox{ CPU1          })\\  
 msg14: (&\mbox{ icache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
 \end{array}
 \]}
 \caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1 to instruction cache}
 \label{read-flow}

 \end{figure}

 Read operation has two flow specification as different coherence protocols are implemented depend on which cache the read operation is initialized to.

 The LPN specification as shown in Fig.~\ref{read-flow} captures the system flow where CPU1 initiate a memory read operation to its instruction cache. When CPU1 will first initiate a memory read request message msg1 to icache1, ICache1 can generate two possible responses. First, if the requested data is not in DCache1, ICache1 will generate an storeCondReq message msg2 to interconnect asking for data. Second, if ICache1 has the exclusive right of the requested data, it will generate a read response message msg14 to CPU1.

 \begin{figure} 
 \centerline{
 \includegraphics[width=3.1in]{figures/readDcache.png}}

 {\footnotesize
 \[
 \begin{array}{llll}
 msg0: (&\mbox{ CPU1},&\mbox{ReadReq},&\mbox{dcache1  })\\                   
 msg1: (&\mbox{ dcache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
 msg2: (&\mbox{ icache1},&\mbox{ LoadLockedreq },&\mbox{Bus })\\     
 msg3: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ dcahce2     })\\
 msg4: (&\mbox{ dcache2},&\mbox{LoadLockedreq},&\mbox{cpu2 })\\
 msg5: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ icahce2     })\\ 
 msg6: (&\mbox{ icache2},&\mbox{LoadLockedreq},&\mbox{cpu2     })\\
 msg7: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ dcahce1 })\\
 msg8: (&\mbox{ icache1},&\mbox{LoadLockedreq},&\mbox{cpu1       })\\
 msg9: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ Memory     })\\
 msg10: (&\mbox{ true })\\
 msg11: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus       })\\     
 msg12: (&\mbox{ icache2},&\mbox{ ReadRes},&\mbox{ Bus    })\\
 msg13: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ icache1 })\\
 msg14: (&\mbox{ dcache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
 \end{array}
 \]}
 \caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1 to  data cache}
 \label{read-dcache}

 \end{figure}

The LPN specification shown in Fig.~\ref{read-dcahce} describes the system flow when CPU1 initiate a memory read operation to its data cache. DCache1 can also generate two possible responses to the read request. First, if the requested data is included in DCache1 but it's shared, cache1 will generate a load locked data request message msg21 to make sure it has the newest data. Second possibility is that when DCache1 has the requested, it will directly issue a read response message msg 33 to CPU1.

This two specifications are also symmetric for CPU2.
\end{document}

### 2015-08-23 Flow Specification.tex
k

@@
 
-This two specifications are also symmetric for CPU2.
+\begin{figure} 
+ \centerline{
+ \includegraphics[width=1.5in]{figures/storeFail.png}}
+
+ {\footnotesize
+ \[
+ \begin{array}{llll}
+ msg0: (&\mbox{ CPU1},&\mbox{StoreFailedReq},&\mbox{ICache1})\\                   
+ msg1: (&\mbox{ ICache1},&\mbox{ StoreFailedRes},&\mbox{ CPU1 })\\
+ \end{array}
+ \]}
+ \caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1 to  data cache}
+ \label{storefail}
+
+ \end{figure}
+Another flow implemented in GEM5 is when CPU initiate a store failed request to instruction cache, with a simple store failed response from instruction cache.
+Detailed flow specification is specified in Fig.~\ref{storefail}
+
+This three specifications are also symmetric for CPU2.
 \end{document}

### 2015-08-23 Flow Specification.tex
d

@@
 
-\title{Flow Specification in GEM5}
+\title{Flow Specification of Cache Coherence Protocol in GEM5}
 
@@
 \begin{abstract}
-This paper will go through a brief overview for the implemented instances of these flows are given.
-These flow specifications capture how messages are exchanged for different use cases.
-In this model, a message is defined with the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform.
+This paper will go through a brief overview for the implemented instances of given GEM5 flow specifications. These flow specifications capture how messages are exchanged for different use cases. 
 \end{abstract}
+
+The flow specifications are implemented in a SoC model using GEM5.
+The architecture of this SoC model is shown in Fig.~\ref{SoC}.
+This SoC model consists of two ARM Cortex-A9 cores, each of which contains two separate 16KB data and instruction caches.
+The caches are connected to a 1GB memory through a memory bus model.
+.
+
+In this model, components communicate with each other by sending and receiving various request and response messages.
+In order to observe and trace communications occurring inside this model during execution, monitors are attached to links connecting the components.
+These monitors record the messages flowing through the links they are attached to, and store them into output trace files.
+The recorded mssage is defined in the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform.
+
+\begin{figure} 
+\centerline{
+\includegraphics[width=3.3in]{figures/Fig4.png}}
+\caption{SoC platform structure.}
+\label{SoC}
+\end{figure}
 

### 2015-08-23 Flow Specification.tex
added sequence chart

@@
 These monitors record the messages flowing through the links they are attached to, and store them into output trace files.
-The recorded mssage is defined in the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform.
+The recorded message is defined in the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform.
 
@@
 \section{write flow specification}
+
+In this section, we will take CPU1 as an example to explain how write flow specification works in our SoC system.
+
+Write flow specification will be triggered by CPU1 initiating a memory write request to its data cache.
+Next, depends on whether the required data is included in cache1, and the status of requested data,  ICache1 can generate three possible responses.
+
+First, the cache1 will send read exclusive request message msg2 to interconnect if data is not present in cache, as shown in Fig.~\ref{write3};
+
+ \begin{figure} 
+ \centerline{
+ \includegraphics[width=3.9In]{figures/write3.png}}
+ \caption{Flow sequence chart of write operation when requested data is not included in Dcache.
+ReadExRes can also be sent from Memory if Dcache2 doesn't have requested data.
+This sequence chart is symmetric for CPU2.
+}
+ \label{write3}
+ \end{figure}
+
+ \begin{figure} 
+ \centerline{
+ \includegraphics[width=3.9In]{figures/write2.png}}
+ \caption{Flow sequence chart of write operation when requested data is shared by another component.
+UpgradeRes can also be sent from Memory if Dcache2 doesn't have requested data.
+This sequence chart is symmetric for CPU2.
+}
+ \label{write2}
+ \end{figure}
+
+ \begin{figure} 
+ \includegraphics[width=2In]{figures/write1.png}
+ \caption{Flow sequence chart of write operation when XCache has the exclusive right of requested data. XCache can be instruction cache or data cache. This sequence chart is symmetric for CPU2. }
+ \label{write1}
+ \end{figure}
+
+Second, if the data is shared by CPU0, it will send upgrade request message msg16 to interconnect to disable CPU0's ownership of this block of data, as shown in Fig.!\ref{write2};
+
+Last, if CPU1 has exclusive right of required data, ICache1 will perform the write operation and sent an response message msg15 to CPU1, as shown in Fig.~\ref{write1}.
 
@@
  msg_9: (&\mbox{ true                                          })\\  
- msg_{10}: (&\mbox{ Memory},&\mbox{ readExres},&\mbox{ Bus        })\\  
+ msg_{10}: (&\mbox{ Memory},&\mbox{ readExres},&\mbox{ Bus})\\  
  msg_{11}: (&\mbox{ icache2},&\mbox{ readExres},&\mbox{ Bus })\\  
- msg_{12}: (&\mbox{ Bus},&\mbox{ readExres},&\mbox{ dcache1        })\\  
+ msg_{12}: (&\mbox{ Bus},&\mbox{ readExres},&\mbox{ dcache1})\\  
  msg_{13}: (&\mbox{ icache1},&\mbox{ writeRes},&\mbox{ CPU1         })\\  
  msg_{14}: (&\mbox{ icache1},&\mbox{ writeRes},&\mbox{ CPU1 })\\  
- msg_{15}: (&\mbox{ dcache1},&\mbox{UpgradeReq             })\\  
+ msg_{15}: (&\mbox{ dcache1},&\mbox{UpgradeReq},&\mbox{Bus})\\  
  msg_{16}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ icahce2      })\\   
@@
  \]}
- \caption{Flow specification ($F_1$) of a cache coherent write operation initiated from CPU1}
+ \caption{Flow specification of a cache coherent write operation initiated from CPU1. This flow is symmetric for CPU2. }
  \label{write-flow}
@@
 
- The LPN as shown in Fig.~\ref{write-flow} specifies a system flow where CPU1 initiates a memory write operation.
-In this flow, CPU1 initiates a memory write request to L1 Cache.
-Next, depends on whether the required data is included in cache1, the cache1 will generate three possible responses.
-First, the cache1 will send read exclusive request message msg2 to interconnect if data is not present in cache; Or if the data is shared by CPU0, it will send upgrade message msg16 to interconnect to disable CPU0's ownership of this block of data, else if CPU1 has exclusive right of required data, cache1 will perform the write operation and sent an response message msg15 to CPU1.
-Afterwards, interconnect will sent request to all connected component (data cache and instruction cache of each CPU and memory).
-Once the interconnect obtains the response from either CPU0 or memory, it will generate a response message to cache1, then cache1 will generate a write response message msg14 (or msg 21 ) to CPU1.
-The flow is symmetric for CPU2.
+ Afterwards, interconnect will sent received request to all connected component (data cache and instruction cache of each CPU and memory).
+Once the interconnect obtains the response from either CPU2 or memory, it will generate a response message to cache1, ICache1 will generate a write response message msg14 (or msg 21 ) to CPU1.
+These three flow sequence charts are implemented as on flow specification in GEM5, as specified in Fig.~\ref{write-flow}.
 
  \section{read flow specification}
+ In this section, we will take CPU1 as an example to explain how read flow specification works in our SoC system.
+
+ \begin{figure} 
+ \centerline{
+ \includegraphics[width=3.9In]{figures/read3.png}}
+ \caption{Flow sequence chart of write operation when requested data is not included in Dcache.
+ReadExRes can also be sent from Memory if Dcache2 doesn't have requested data.
+This sequence chart is symmetric for CPU2.
+}
+ \label{read3}
+ \end{figure}
+
+ \begin{figure} 
+ \centerline{
+ \includegraphics[width=3.9In]{figures/read2.png}}
+ \caption{Flow sequence chart of write operation when requested data is shared by another component.
+UpgradeRes can also be sent from Memory if Dcache2 doesn't have requested data.
+This sequence chart is symmetric for CPU2.
+}
+ \label{read2}
+ \end{figure}
+
+ \begin{figure} 
+ \includegraphics[width=2In]{figures/read1.png}
+ \caption{Flow sequence chart of write operation when XCache has the exclusive right of requested data. XCache can be instruction cache or data cache. This sequence chart is symmetric for CPU2. }
+ \label{read1}
+ \end{figure}
+
+Read coherence protocol will be triggered by CPU1 initiating a memory read request to its cache.
+Depending on whether the required data is included in its cache, the status of requested data,  and which cache the request is sent to, cache can generate three possible responses.
+
+First, for both instruction cache and data cache, if they have the exclusive right of requested data, they will generate a read response carrying the data back to CPU.
+Detailed flow sequence chart is displayed in Fig.~\ref{read1}.
+
+Second, when ICache1 decide that the requested data is not present, it will issue an StoreCondReq to memory bus asking for data.
+Finally, if DCache1 has the data, but it's shared, it will issue an LoadLockedReq to memory bus.
+Flow sequence charts for these two possibility are shown in Fig.\~ref{read1} and Fig.\~ref{read}.
+
+Since instruction cache and data cache have different coherence protocol, there are two flow specifications for read operation generated by CPU1.
+The LPN specification as shown in Fig.~\ref{read-flow} captures the system flow where CPU1 initiate a memory read to instruction cache.
+And Fig.~\ref{read-dcache} specifies read operation to data cache from CPU1.
 
@@
  \]}
- \caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1 to instruction cache}
+ \caption{Flow specification of a cache coherent read operation initiated from CPU1 to instruction cache.This flow is symmetric for CPU2. }
  \label{read-flow}
@@
  \end{figure}
-
- Read operation has two flow specification as different coherence protocols are implemented depend on which cache the read operation is initialized to.
-
- The LPN specification as shown in Fig.~\ref{read-flow} captures the system flow where CPU1 initiate a memory read operation to its instruction cache.
-When CPU1 will first initiate a memory read request message msg1 to icache1, ICache1 can generate two possible responses.
-First, if the requested data is not in DCache1, ICache1 will generate an storeCondReq message msg2 to interconnect asking for data.
-Second, if ICache1 has the exclusive right of the requested data, it will generate a read response message msg14 to CPU1.
 
@@
  \]}
- \caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1 to  data cache}
+ \caption{Flow specification of a cache coherent read operation initiated from CPU1 to  data cache. This flow is symmetric for CPU2. }
  \label{read-dcache}
@@
 
-The LPN specification shown in Fig.~\ref{read-dcahce} describes the system flow when CPU1 initiate a memory read operation to its data cache.
-DCache1 can also generate two possible responses to the read request.
-First, if the requested data is included in DCache1 but it's shared, cache1 will generate a load locked data request message msg21 to make sure it has the newest data.
-Second possibility is that when DCache1 has the requested, it will directly issue a read response message msg 33 to CPU1.
+\section{failed read specification}
 
@@
  \begin{array}{llll}
- msg0: (&\mbox{ CPU1},&\mbox{StoreFailedReq},&\mbox{ICache1})\\                   
- msg1: (&\mbox{ ICache1},&\mbox{ StoreFailedRes},&\mbox{ CPU1 })\\
+ msg0: (&\mbox{ CPU1},&\mbox{StoreCondFailedReq},&\mbox{ICache1})\\                   
+ msg1: (&\mbox{ ICache1},&\mbox{ StoreCondFailedRes},&\mbox{ CPU1 })\\
  \end{array}
  \]}
- \caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1 to  data cache}
+ \caption{Flow specification of failed read. This flow is symmetric for CPU2. }
  \label{storefail}

### 2015-08-23 paper.tex
k
number changed

@@
 Not all of the messages are relevant to the flow specification as many are used by GEM5 to initialize its simulation environment.
-After removing those irrelevant messages, the number of messages in the trace file is to reduced to $120949$.
+After removing those irrelevant messages, the number of messages in the trace file is to reduced to $121138$.
 
@@
 \hline
-CPU1 Read		&  $21584$\\
+CPU1 Read		&  $4002$\\
 CPU1 Write	&  $3370$\\
-CPU2 Read		&  $21341$\\
+CPU1 Failed Read  &$95$\\
+CPU2 Read		&  $3955$\\
 CPU2 Write	&  $3308$\\
+CPU2 Failed Read  &$93$\\
 \hline

### 2015-08-24 Flow Specification.tex
layout changes

@@
 
-Write flow specification will be triggered by CPU1 initiating a memory write request to its data cache.
-Next, depends on whether the required data is included in cache1, and the status of requested data,  ICache1 can generate three possible responses.
+Coherence protocol for write operation can be triggered by CPU1 initiating a memory write request to its data cache.
+Next, depends on whether the required data is included in cache1, and the status of requested data,  cache1 can generate three possible responses.
 
-First, the cache1 will send read exclusive request message msg2 to interconnect if data is not present in cache, as shown in Fig.~\ref{write3};
+First, the cache1 will send read exclusive request message msg2 to interconnect if data is not present in cache, as shown in Fig.~\ref{write3}.
 
@@
  \centerline{
- \includegraphics[width=3.9In]{figures/write3.png}}
+ \includegraphics[width=4In]{figures/write3.png}}
  \caption{Flow sequence chart of write operation when requested data is not included in Dcache.
@@
  \centerline{
- \includegraphics[width=3.9In]{figures/write2.png}}
+ \includegraphics[width=4In]{figures/write2.png}}
  \caption{Flow sequence chart of write operation when requested data is shared by another component.
@@
  \begin{figure} 
- \includegraphics[width=2In]{figures/write1.png}
+ \includegraphics[width=1.9In]{figures/write1.png}
  \caption{Flow sequence chart of write operation when XCache has the exclusive right of requested data. XCache can be instruction cache or data cache. This sequence chart is symmetric for CPU2. }
@@
 
-Second, if the data is shared by CPU0, it will send upgrade request message msg16 to interconnect to disable CPU0's ownership of this block of data, as shown in Fig.!\ref{write2};
+Second, if the data is shared by CPU2, it will send upgrade request message msg16 to interconnect request disabling  CPU2's ownership of the data, as shown in Fig.~\ref{write2}.
 
@@
  \centerline{
- \includegraphics[width=3.4In]{figures/Fig5.png}}
+ \includegraphics[width=3.7In]{figures/Fig5.png}}
 
@@
  Afterwards, interconnect will sent received request to all connected component (data cache and instruction cache of each CPU and memory).
-Once the interconnect obtains the response from either CPU2 or memory, it will generate a response message to cache1, ICache1 will generate a write response message msg14 (or msg 21 ) to CPU1.
-These three flow sequence charts are implemented as on flow specification in GEM5, as specified in Fig.~\ref{write-flow}.
+Once the interconnect obtains the response from either CPU2 or memory, it will generate a response message to cache1, cache1 will generate a write response message msg14 (or msg 21 ) to CPU1.
+These three flow sequence charts are implemented together in one flow specification in GEM5, as specified in Fig.~\ref{write-flow}.
+
+ \section{failed read specification}
+
+\begin{figure} 
+ \includegraphics[width=1.5in]{figures/storeFail.png}
+
+ {\footnotesize
+ \[
+ \begin{array}{llll}
+ msg0: (&\mbox{ CPU1},&\mbox{StoreCondFailedReq},&\mbox{ICache1})\\                   
+ msg1: (&\mbox{ ICache1},&\mbox{ StoreCondFailedRes},&\mbox{ CPU1 })\\
+ \end{array}
+ \]}
+ \caption{Flow specification of failed read. This flow is symmetric for CPU2. }
+ \label{storefail}
+
+ \end{figure}
+Another flow implemented in GEM5 is when CPU initiate a store failed request to instruction cache, with a simple store failed response from instruction cache.
+Detailed flow specification is specified in Fig.~\ref{storefail}
+
+This three specifications are also symmetric for CPU2.
 
  \section{read flow specification}
- In this section, we will take CPU1 as an example to explain how read flow specification works in our SoC system.
+ In this section, we will also take CPU1 as an example to explain how read flow specification works in our SoC system.
 
@@
  \includegraphics[width=3.9In]{figures/read3.png}}
- \caption{Flow sequence chart of write operation when requested data is not included in Dcache.
-ReadExRes can also be sent from Memory if Dcache2 doesn't have requested data.
+ \caption{Flow sequence chart of read operation when requested data is shared by another component.
+LoadLockedRes can also be sent from Memory if Dcache2 doesn't have requested data.
 This sequence chart is symmetric for CPU2.
@@
  \includegraphics[width=3.9In]{figures/read2.png}}
- \caption{Flow sequence chart of write operation when requested data is shared by another component.
-UpgradeRes can also be sent from Memory if Dcache2 doesn't have requested data.
+ \caption{Flow sequence chart of read operation when requested data is not present.
+StoreCondRes can also be sent from Memory if Dcache2 doesn't have requested data.
 This sequence chart is symmetric for CPU2.
@@
  \includegraphics[width=2In]{figures/read1.png}
- \caption{Flow sequence chart of write operation when XCache has the exclusive right of requested data. XCache can be instruction cache or data cache. This sequence chart is symmetric for CPU2. }
+ \caption{Flow sequence chart of read operation when XCache has the exclusive right of requested data. XCache can be instruction cache or data cache. This sequence chart is symmetric for CPU2. }
  \label{read1}
@@
 
-Read coherence protocol will be triggered by CPU1 initiating a memory read request to its cache.
+Read coherence protocol can be triggered by CPU1 initiating a memory read request to its cache.
 Depending on whether the required data is included in its cache, the status of requested data,  and which cache the request is sent to, cache can generate three possible responses.
 
-First, for both instruction cache and data cache, if they have the exclusive right of requested data, they will generate a read response carrying the data back to CPU.
+First, for both instruction cache and data cache, if they have the exclusive right of requested data, they can generate a read response carrying the data back to CPU.
 Detailed flow sequence chart is displayed in Fig.~\ref{read1}.
@@
 Second, when ICache1 decide that the requested data is not present, it will issue an StoreCondReq to memory bus asking for data.
-Finally, if DCache1 has the data, but it's shared, it will issue an LoadLockedReq to memory bus.
+
+Finally, if DCache1 has the data, but it's shared, it will issue an LoadLockedReq to memory bus requesting exclusive right of the data.
 Flow sequence charts for these two possibility are shown in Fig.\~ref{read1} and Fig.\~ref{read}.
@@
  \centerline{
- \includegraphics[width=3.33in]{figures/Fih6.png}}
+ \includegraphics[width=3in]{figures/Fih6.png}}
 
@@
  \centerline{
- \includegraphics[width=3.1in]{figures/readDcache.png}}
+ \includegraphics[width=3in]{figures/readDcache.png}}
 
@@
 
-\section{failed read specification}
-
-\begin{figure} 
- \centerline{
- \includegraphics[width=1.5in]{figures/storeFail.png}}
-
- {\footnotesize
- \[
- \begin{array}{llll}
- msg0: (&\mbox{ CPU1},&\mbox{StoreCondFailedReq},&\mbox{ICache1})\\                   
- msg1: (&\mbox{ ICache1},&\mbox{ StoreCondFailedRes},&\mbox{ CPU1 })\\
- \end{array}
- \]}
- \caption{Flow specification of failed read. This flow is symmetric for CPU2. }
- \label{storefail}
-
- \end{figure}
-Another flow implemented in GEM5 is when CPU initiate a store failed request to instruction cache, with a simple store failed response from instruction cache.
-Detailed flow specification is specified in Fig.~\ref{storefail}
-
-This three specifications are also symmetric for CPU2.
 \end{document}

### 2015-09-05 paper.tex
hyperref
Test commit
Changed to ISQED format and also changed title

@@
 \usepackage{draftwatermark}
+\usepackage{hyperref}
+\hypersetup{
+    colorlinks=true,
+    linkcolor=blue,
+    filecolor=magenta,      
+    urlcolor=cyan,
+}
 
@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Flow-Directed Trace Analysis for Post-Silicon Validation}
-
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{\IEEEauthorblockN{Hao Zheng, Yuting Cao}
-\IEEEauthorblockA{Computer Science and Engineering\\
-University of South Florida\\
-Tampa, Florida 33620}
-\and
-\IEEEauthorblockN{Sandip Ray, Jin Yang}
-\IEEEauthorblockA{Strategic CAD Labs\\
-Intel Corporation\\
-Portland, Oregon
-}}
+\title{\Large {\bf Flow-guided Analysis of System-level Post-silicon
+  Traces Under Limited Observability}}
+
+\author{\large Hao Zheng$^1$, Yuting Cao$^1$, Sandip Ray$^2$, Jin Yang$^2$ \\
+$^1$Dept. of Computer Science and Engg., University of South Florida, Tampa, Fl 33620. USA. \\
+$^2$Strategic CAD Labs, Intel Corporation, Hillsboro, OR 97124.  USA.}
+\maketitle
+
+%% % author names and affiliations
+%% % use a multiple column layout for up to three different
+%% % affiliations
+%% \author{\IEEEauthorblockN{Hao Zheng, Yuting Cao}
+%% \IEEEauthorblockA{Computer Science and Engineering\\
+%% University of South Florida\\
+%% Tampa, Florida 33620}
+%% \and
+%% \IEEEauthorblockN{Sandip Ray, Jin Yang}
+%% \IEEEauthorblockA{Strategic CAD Labs\\
+%% Intel Corporation\\
+%% Portland, Oregon
+%% }}
 
@@
 % make the title area
-\maketitle
+%\maketitle
 
@@
 % creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
+% \IEEEpeermaketitle
 
@@
 As there are two CPUs in this model, there are actually four such flows considered for this model.
-%More detailed information about the flow specification is explained in the following section.
-
-%found at \red{github???}
+More detailed information about the flow specification can be found at \url{https://github.com/cao2/paper/blob/master/Flow\%20Specification.pdf}
 

### 2015-09-07 Flow Specification.tex
Revert "layout changes"
This reverts commit b80d8dccae5b1ac338fd11738c728fdc38514fcf.
Merge remote-tracking branch 'origin/master'
Conflicts:
	Flow Specification.pdf
	Flow Specification.synctex.gz

@@
  \centerline{
- \includegraphics[width=3.7In]{figures/Fig5.png}}
+ \includegraphics[width=3.4In]{figures/Fig5.png}}
 
@@
  msg_{18}: (&\mbox{ icache2},&\mbox{UpgradeRes},&\mbox{ Bus     })\\  
- msg_{19}: (&\mbox{ Bus},&\mbox{ UpgradeRes},&\mbox{ icache1      })\\  
- msg_{20}: (&\mbox{ icache1},&\mbox{ WriteRes},&\mbox{ CPU1 })\\  
+ msg_{19}: (&\mbox{ Bus},&\mbox{ UpgradeRes},&\mbox{ dcache1      })\\  
+ msg_{20}: (&\mbox{ icache1},&\mbox{ writeRes},&\mbox{ CPU1 })\\  
  \end{array}
@@
 These three flow sequence charts are implemented together in one flow specification in GEM5, as specified in Fig.~\ref{write-flow}.
-
- \section{failed read specification}
-
-\begin{figure} 
- \includegraphics[width=1.5in]{figures/storeFail.png}
-
- {\footnotesize
- \[
- \begin{array}{llll}
- msg0: (&\mbox{ CPU1},&\mbox{StoreCondFailedReq},&\mbox{ICache1})\\                   
- msg1: (&\mbox{ ICache1},&\mbox{ StoreCondFailedRes},&\mbox{ CPU1 })\\
- \end{array}
- \]}
- \caption{Flow specification of failed read. This flow is symmetric for CPU2. }
- \label{storefail}
-
- \end{figure}
-Another flow implemented in GEM5 is when CPU initiate a store failed request to instruction cache, with a simple store failed response from instruction cache.
-Detailed flow specification is specified in Fig.~\ref{storefail}
-
-This three specifications are also symmetric for CPU2.
 
@@
 
+\section{failed read specification}
+
+\begin{figure} 
+ \includegraphics[width=1.5in]{figures/storeFail.png}
+
+ {\footnotesize
+ \[
+ \begin{array}{llll}
+ msg0: (&\mbox{ CPU1},&\mbox{StoreCondFailedReq},&\mbox{ICache1})\\                   
+ msg1: (&\mbox{ ICache1},&\mbox{ StoreCondFailedRes},&\mbox{ CPU1 })\\
+ \end{array}
+ \]}
+ \caption{Flow specification of failed read. This flow is symmetric for CPU2. }
+ \label{storefail}
+
+ \end{figure}
+Another flow implemented in GEM5 is when CPU initiate a store failed request to instruction cache, with a simple store failed response from instruction cache.
+Detailed flow specification is specified in Fig.~\ref{storefail}
+
+This three specifications are also symmetric for CPU2.
 \end{document}

### 2015-09-07 paper.tex

@@
 %Also, abstracts signal traces to system use cases, which helps validation engineering better understand the behavior displayed by the signal traces.
+
+Post-silicon validation makes use of pre-production silicon
+integrated circuit (IC) to ensure that the fabricated system
+works as desired under actual operating conditions with real
+software.  Since the silicon executes at target clock speed,
+post-silicon executions are billions of times faster than
+RTL simulations, and even provide speed-up of several orders
+of magnitude over other pre-silicon platforms (\eg, FPGA,
+system-level emulation, etc.).  This makes it possible to
+explore deep design states which cannot be exercised in
+pre-silicon, and identify errors missed during pre-silicon
+validation and debug.  Post-silicon validation is a critical
+component of the design validation life-cycle for modern
+microprocessors and SoC designs.  Unfortunately, it is also
+a highly complex component, performed under aggressive
+schedules and accounting for more than $50\%$ of the overall
+design validation cost.  Consequently, it is crucial to
+develop techniques for streamlining and automating
+post-silicon validation activities.
+
+A central component of post-silicon validation is to
+correlate trace from silicon execution with the intended
+behavior of the sysem.  As an example, consider a silicon
+execution of a smartphone executing a post-silicon use case
+where the end-user browses the Web while listening to music
+and sending and receiving occasional text messages.  Such a
+an execution trace would involve activities from the CPU,
+audio controller, display controller, wireless radio
+antenna, etc., reflecting the interleaved execution of
+several communication protocols.  On the other hand, due to
+observability limitations, only a small small number of
+participating signals can be actually traced during silicon
+execution.  Furthermore, due to electrical perturbations,
+silicon data can be noisy, lossy, and ambiguous.
+Consequently, it is non-trivial to identify all
+participating protocols and pinpoint the ``right''
+interleaving that results in an observed trace.
+
+In this paper, we consider the problem of reconstructing
+protocol-level behavior from silicon traces in SoC designs.
 

### 2015-09-07 paper.tex
Added some stuff to intro
added a sentence to intro
fixing some git issues

@@
 
+\newcommand{\eg}{\mbox{{\em e.g.}}}
+
 % *** GRAPHICS RELATED PACKAGES ***
@@
 
-A central component of post-silicon validation is to
-correlate trace from silicon execution with the intended
-behavior of the sysem.  As an example, consider a silicon
-execution of a smartphone executing a post-silicon use case
-where the end-user browses the Web while listening to music
-and sending and receiving occasional text messages.  Such a
-an execution trace would involve activities from the CPU,
-audio controller, display controller, wireless radio
-antenna, etc., reflecting the interleaved execution of
-several communication protocols.  On the other hand, due to
-observability limitations, only a small small number of
-participating signals can be actually traced during silicon
-execution.  Furthermore, due to electrical perturbations,
-silicon data can be noisy, lossy, and ambiguous.
-Consequently, it is non-trivial to identify all
-participating protocols and pinpoint the ``right''
+A central component of post-silicon validation of SoC
+designs is to correlate trace from silicon execution with
+the intended system-level transactions.  An SoC design is
+typically composed of a large number of pre-designed
+hardware or software blocks (often referred to as
+``intellectual properties'' or ``IPs'') that coordinate
+through complex protocols to implement system-level
+behavior.  Any execution trace of the system involves a
+large number of interleaved instances of these protocols.
+For example, consider a smartphone executing a usage
+scenerio where the end-user browses the Web while listening
+to music and sending and receiving occasional text messages.
+Typical post-silicon validation use-case involve exercising
+such scenarios.  An execution trace for this scenario would
+involve activities from the CPU, audio controller, display
+controller, wireless radio antenna, etc., reflecting the
+interleaved execution of several communication protocols.
+On the other hand, due to observability limitations, only a
+small small number of participating signals can be actually
+traced during silicon execution.  Furthermore, due to
+electrical perturbations, silicon data can be noisy, lossy,
+and ambiguous.  Consequently, it is non-trivial to identify
+all participating protocols and pinpoint the ``right''
 interleaving that results in an observed trace.
@@
 protocol-level behavior from silicon traces in SoC designs.
+More specifically, given a collection of system-level
+communication protocols and a trace of (partially observed)
+hardware signals, our approach infers, with a certain
+measure of confidence, the protocol instances (and their
+interleavings) being exercised by the trace.  We show how to use symbolic simulation and
 

### 2015-09-08 paper.tex
Update section V with latest results
Only section V is updated with the latest results from Hao.

@@
 \usepackage{mybeamer}
-\usepackage{draftwatermark}
+%\usepackage{draftwatermark}
 \usepackage{hyperref}
@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{\Large {\bf Flow-guided Analysis of System-level Post-silicon
+\title{\Large {\bf Flow-Guided Analysis of System-level Post-silicon
   Traces Under Limited Observability}}
@@
 \author{\large Hao Zheng$^1$, Yuting Cao$^1$, Sandip Ray$^2$, Jin Yang$^2$ \\
-$^1$Dept. of Computer Science and Engg., University of South Florida, Tampa, Fl 33620. USA. \\
+$^1$Dept. of Computer Science and Eng., University of South Florida, Tampa, Fl 33620. USA. \\
 $^2$Strategic CAD Labs, Intel Corporation, Hillsboro, OR 97124.  USA.}
@@
 
-For this model, we consider the flow specification describing the cache coherence protocol supported in GEM5 that is used to build the model in Fig~\ref{SoC}.
-This specification consists of two flows, one for cache coherent read requests and one for  cache coherent write requests, from CPUs.
-Two such flows describe the cache coherent protocol for each CPU.
-As there are two CPUs in this model, there are actually four such flows considered for this model.
-More detailed information about the flow specification can be found at \url{https://github.com/cao2/paper/blob/master/Flow\%20Specification.pdf}
+For this model, we consider the flow specifications describing the cache coherence protocols supported in GEM5 that is used to build the model in Fig~\ref{SoC}.
+These flow specifications describe data/instruction read operations and data write operations initiated from CPUs.
+Three such flows describe the cache coherent protocols for each CPU.
+As there are two CPUs in this model, there are actually six such flows considered for this model.
+More detailed information about these protocols and the flow specifications can be found online\footnote{\url{https://github.com/cao2/paper/blob/master/Flow\%20Specification.pdf}} to meet the page limit.
 
@@
 
-In order to trigger cache coherence protocols, a simple two-threaded program including one producer and one consumer is created and executed on the CPUs, each of which executes one thread.
-\red{How is shared memory is simulated???} .
-The producer will write to a specified memory address, and the consumer will read from the same address.
-GEM5 does not support true concurrency.
+Two simple programs are written, one for each CPU.
+These simple programs read numbers from a file, perform some operations on these numbers, and store the results back to the file.
+How GEM5 supports shared memory multi-threaded program execution is unclear.
+Therefore, there is no data shared in both caches in this test.
+Furthermore, GEM5 does not support true concurrency.
 When there are two programs running on the CPUs, GEM5 alternates the executions between the two CPUs.
+To simulate asynchronous concurrency with the interleaving semantics, those two simple programs are instrumented with pseudo-blocking commands, one placed before each statement.
+A pseudo blocking command includes a random number generator that returns either $0$ or $1$ and a loop that only exits when the returned random number is $0$.
 
@@
 The total runtime and the peak memory usage for the trace analysis algorithm to finish the reduced trace are 3 seconds and 12MB, respectively.
-From the trace, Table~\ref{table-case-2} shows the number of instances extracted for the four flows describing cache coherent read/write operations.
-\begin{table}[td]
-\caption{The number of flow instances derived by the trace analysis}
+From the trace, Table~\ref{table-case-2} shows the number of instances extracted for the six flows describing cache coherent read/write operations initiated from both CPUs.
+\begin{table}[tb]
+\caption{The number of flow instances derived by the trace analysis with the full observability.}
 \begin{center}
-\begin{tabular}{|c|c|}
+\begin{tabular}{|l|c|}
 \hline
@@
 \hline
-CPU1 Read		&  $4002$\\
+CPU1 Data Read			&  $17582$\\
+CPU1 Instruction Read		&  $4002$\\
 CPU1 Write				&  $3370$\\
-CPU1 Failed Read  &$95$\\
-CPU2 Read		&  $3955$\\
+\hline
+CPU2 Data Read			&  $17386$\\
+CPU2 Instruction Read		&  $3955$\\
 CPU2 Write				&  $3308$\\
-CPU2 Failed Read  &$93$\\
 \hline
@@
 
-\red{Also need to consider the case where some monitors are removed?}
+To take the partial observability into account, the four monitors attached to the links between two CPUs and their caches are disabled.
+Then, the trace is generated by the remaining five monitors from the SoC model executing the same program.
+After removing all non-observable messages, the trace only contains 15089 messages.
+The numbers of the flow instances extracted by the trace analysis method are shown in Table~\ref{table-par-obs}.
+From these results, the numbers of the flow instances are dropped significantly compared to the results extracted from the trace with the full observability as shown in Table~\ref{table-case-2}.
+This difference is due to that some communications occurred in the system when executing the program involve the CPUs and their corresponding caches only, and the traffic on the links between the CPUs and their corresponding caches is not observable.
+Therefore, the instances of the flow specifications characterizing these communications do not exist in the trace.
+In other words, all extracted flow instances in Table~\ref{table-par-obs} characterize the communications that pass through the memory bus in the system model.
+The runtime and memory usage is similar to that for analyzing the trace of the full observability as shown above.
+
+\begin{table}[tb]
+\caption{The number of flow instances derived by the trace analysis with certain monitors disabled.}
+\begin{center}
+\begin{tabular}{|l|c|}
+\hline
+Flows & $\#$Instances \\
+\hline
+\hline
+CPU1 Data Read			&  $829$\\
+CPU1 Instruction Read		&  $169$\\
+CPU1 Write				&  $82$\\
+\hline
+CPU2 Data Read			&  $803$\\
+CPU2 Instruction Read		&  $190$\\
+CPU2 Write				&  $83$\\
+\hline
+\end{tabular}
+\end{center}
+\label{table-par-obs}
+\end{table}%
+
+In the third experiment, further partial observability is taken into consideration.
+In this experiment, only the five links involving the memory bus are still considered.
+However, an assumption is made that all messages passing the same link are not distinguishable due to the limitation of the observability.
+More specifically, the monitors are modified such that whenever a message is captured on one of the links, it dumps a set of messages passing through the same link into the trace file.
+Therefore, each line of the trace file corresponds to a set of messages.
+After applying the trace analysis to this trace,  a total of 13944 flow execution scenarios are extracted.
+This large number, compared to the number of extracted execution scenarios shown in the Table~\ref{table-case-2} and \ref{table-par-obs}, is due to the ambiguous interpretation of the messages with limited observability.
+
+The whole experiment takes about $15$ minutes and $420$~MB to finish, significantly higher than the numbers for analyzing traces where there is no ambiguity in the observed messages.
+This is due to the fact that a trace of ambiguous events is in fact a set of traces of original messages, which lead to large numbers of execution scenarios either during or at the end of the analysis.
+In this experiment, the peak number of executions during the analysis process is $70384$.
+Compared to the final number, many of the intermediate scenarios are invalid, and removed eventually.
+However, controlling the number of intermediate scenarios during the trace analysis is critical in order for the analysis to be tractable.
+Here, insights from validators can help.
 

### 2015-09-10 paper.tex
Committing changes to intro
Done editing intro

@@
 
+\begin{abstract}
+We consider the problem of reconstructing system-level
+behavior of an SoC design from a partially observed signal
+trace.  Solving this problem is a critical activity in
+post-silicon validation, and currently depends primarily on
+human creativity and insight.  We provide algorithms to
+automatically infer system-level transactions from
+incomplete, ambiguous, and noisy trace data, together with a
+measure of confidence.  We demonstrate the approach on
+illustrative system-level protocols for SoC models developed
+in SystemC as well as in an FPGA environment.
+\end{abstract}
+
 % make the title area
@@
 
-\begin{abstract}
-Validation of multicore systems-on-chip (SoC) is very challenging, and main reasons include the limited internal operation observability and non-determinism due to multiple clock domains.
+%% \begin{abstract}
+%% Validation of multicore systems-on-chip (SoC) is very challenging, and main reasons include the limited internal operation observability and non-determinism due to multiple clock domains.
 Very often, the observed signal traces offer limited values for debuggers to understand system internal behavior.
@@
 This approach transforms partially observed and non-deterministic signal traces to a representation that is more understandable, and offers more helpful information to locate design defects and a measure to evaluate validation coverage.
-\end{abstract}
+%% \end{abstract}
 % IEEEtran.cls defaults to using nonbold math in the Abstract.
@@
 measure of confidence, the protocol instances (and their
-interleavings) being exercised by the trace.  We show how to use symbolic simulation and
-
-Systems-on-chips (SoCs) have become more complex as more functionalities are integrated on chips.
+interleavings) being exercised by the trace.  Our approach
+is based on a formalization of system-level transactions via
+labeled Petri-Nets~\cite{petrinets}, which are capable of
+describing sequencing, concurrency, and choices of events.
+Given this formalization, we develop algorithms to infer
+system-level transactions from traces with missing, noisy,
+and ambiguous signal values, together with an estimate of
+confidence on the inference.  We demonstrate our approach on
+two SoC models: a SystemC prototype and a realistic
+implementation constructed within the GEM5
+environment~\cite{Binkert2011}.
+
+%% Systems-on-chips (SoCs) have become more complex as more functionalities are integrated on chips.
 To reduce design time, more external IPs are integrated into current SoC designs.
@@
 
-To make validation even more challenging, the high degree of integration in current SoC designs causes traditional chip observation points to become inaccessible.
+%% To make validation even more challenging, the high degree of integration in current SoC designs causes traditional chip observation points to become inaccessible.
 A SoC design usually contains millions of signals.
@@
 
-In a typical design flow, a design starts with some high-level specification, particularly for communication protocols.
+%% In a typical design flow, a design starts with some high-level specification, particularly for communication protocols.
 The abstract protocol specification usually captures information exchanged among IP blocks and how it is exchanged to support functional behavior.
-Graphical representations of protocols such as Message Sequence Charts~\cite{Harel03} or Live Sequence Charts~\cite{Damm99} describe protocols intuitively.
+%% Graphical representations of protocols such as Message Sequence Charts~\cite{Harel03} or Live Sequence Charts~\cite{Damm99} describe protocols intuitively.
 However, its success in hardware industry is limited.
 Recently, message flows in BPMN~\cite{Krstic14HOST} have been being advocated as a formalism for specifying architecture requirements
- or system-level use cases such as power management, security, etc.
+%%  or system-level use cases such as power management, security, etc.
 In general, a flow describes a set of message sequences across multiple IP blocks following a system communication protocol.
@@
 
-\begin{figure}[tb]
-\begin{center}
-\includegraphics[height=2.5in]{figures/bpmn-flow-ex}
-\caption{A graphical representation of a SoC firmware load protocol in BPMN~\cite{Krstic14HOST}.}
-\label{flow-spec-ex}
-\end{center}
-\end{figure}
-
-As indicated above, interpreting and understanding the raw signal traces observed from executing the chip is extremely difficult due to the limited observability and non-determinism.
+%% \begin{figure}[tb]
+%% \begin{center}
+%% \includegraphics[height=2.5in]{figures/bpmn-flow-ex}
+%% \caption{A graphical representation of a SoC firmware load protocol in BPMN~\cite{Krstic14HOST}.}
+%% \label{flow-spec-ex}
+%% \end{center}
+%% \end{figure}
+
+%% As indicated above, interpreting and understanding the raw signal traces observed from executing the chip is extremely difficult due to the limited observability and non-determinism.
 This paper proposes a trace analysis method which analyzes and interprets the observed signal traces at the level of system flow specifications.
@@
 %---
-\section{Labeled Petri-Nets}
+
+\section{Background}
+
+\subsection{System-level Transactions in SoC Designs}
+
+\subsection{Labeled Petri-Nets}
 

### 2015-09-11 paper.tex
Done editing intro

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{\Large {\bf Flow-Guided Analysis of System-level Post-silicon
+\title{\Large {\bf Protocol-Guided Analysis of Post-silicon
   Traces Under Limited Observability}}
@@
 \subsection{System-level Transactions in SoC Designs}
+
+An SoC design is typically composed of a large number of
+pre-designed hardware or software blocks (often referred to
+as ``intellectual properties'' or ``IPs'') that coordinate
+through complex protocols to implement system-level
+behavior.  Fig.~\ref{fig:flow-spec-ex} shows one such
+protocol, involving th
+\begin{figure}[tb]
+\begin{center}
+\includegraphics[height=2.5in]{figures/bpmn-flow-ex}
+\caption{A graphical representation of a SoC firmware load protocol in BPMN~\cite{Krstic14HOST}.}
+\label{flow-spec-ex}
+\end{center}
+\end{figure}
 

### 2015-09-12 paper.tex
revised section II
Figure 1 and 2 are merged such that a LPN is created for the BPMN in
figure 1.  The last paragraph of section II is also revised accordingly.

@@
 is based on a formalization of system-level transactions via
-labeled Petri-Nets~\cite{petrinets}, which are capable of
+labeled Petri-Nets, which are capable of
 describing sequencing, concurrency, and choices of events.
@@
 confidence on the inference.  We demonstrate our approach on
-two SoC models: a SystemC prototype and a realistic
+two SoC models: a SystemC prototype and a more realistic
 implementation constructed within the GEM5
@@
 
+\begin{figure*}[tb]
+\begin{center}
+\begin{tabular}{cc}
+\begin{minipage}{3in}
+\includegraphics[height=2.5in]{figures/bpmn-flow-ex}
+\end{minipage}
+&
+\begin{minipage}{3in}
+\resizebox{2.8in}{!}{
+\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
+\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
+\tikzset{circ/.style={circle, draw,thick}}
+\tikzset{terminal/.style={circle, draw,line width=.8mm}}
+\tikzset{weight3/.style={line width=.3mm}}
+	% Nodes:
+	\node[circ]		(p1) 		at (0,0) {$p_1$};
+	\node[rectbox]	(t1)	at	($(p1) + (0,-1.1)$) {$t_1:\langle {\tt Driver:Device:Load} \_fw\rangle$};
+	\node[circ] 	(p2) 		at ($(t1) + (0,-1.1)$) {$p_2$};
+	\node[rectbox]	(t2)	at	($(p2) + (0,-1.1)$) {$t_2:\langle {\tt Device:CE:Auth\_req} \rangle$};
+	\node[circ] 	(p3) 		at ($(t2) + (0,-1.1)$) {$p_3$};
+	\node[rectbox]	(t3)	at	($(p3) + (0,-1.1)$) {$t_3:\langle {\tt CE:Device:Auth\_resp}\rangle$};
+	\node[circ] 	(p4) 		at ($(t3) + (0,-1.1)$) {$p_4$};
+	\node[rectbox]	(t4)	at	($(p4) + (-3,-1.1)$) {$t_4:\langle {\tt Device:Driver:Report} \rangle$};
+	\node[rectbox]	(t5)	at	($(p4) + (3,-1.1)$) {$t_5:\langle {\tt Device:CE:Ack} \rangle$};
+	\node[circ,fill=black] 	(term1) 		at ($(t4) + (0,-1.1)$) {$~~~$};
+	\node[circ,fill=black] 	(term2) 		at ($(t5) + (0,-1.1)$) {$~~~$};
+
+	% Edges
+	\draw[->,weight3]	(p1) -- (t1);
+	\draw[->,weight3]	(t1) -- (p2);
+	\draw[->,weight3]	(p2) -- (t2);
+	\draw[->,weight3]	(t2) -- (p3);
+	\draw[->,weight3]	(p3) -- (t3);
+	\draw[->,weight3]	(t3) -- (p4);
+	\draw[->,weight3]	(p4) -- ($(p4)+(-3,0)$) -- (t4);
+	\draw[->,weight3]	(p4) -- ($(p4)+(3,0)$) -- (t5);
+	\draw[->,weight3]	(t4) -- (term1);	
+	\draw[->,weight3]	(t5) -- (term2);
+
+\end{tikzpicture}
+}
+\end{minipage}
+\\
+(a) & (b)
+\end{tabular}
+\caption{(a) A graphical representation of a SoC firmware load protocol in BPMN~\cite{Krstic14HOST}, and (b) the LPN capturing the communication protocol in (a).}
+\label{flow-spec-ex}
+\end{center}
+\end{figure*}
+
 \section{Background}
@@
 protocol, involving th
-\begin{figure}[tb]
-\begin{center}
-\includegraphics[height=2.5in]{figures/bpmn-flow-ex}
-\caption{A graphical representation of a SoC firmware load protocol in BPMN~\cite{Krstic14HOST}.}
-\label{flow-spec-ex}
-\end{center}
-\end{figure}
 
@@
 In this section, we present an alternative formalism, \emph{labeled Petri Nets}, to represent communication aspects of system flows with the intuitive descriptions of internal operations for IP blocks abstracted away whenever possible.
-Labeled Petri nets are capable of describing sequencing, concurrency, and choices.
+Labeled Petri nets, which are a subclass of traditional Petri-nets~\cite{Murata89} where Petri-net transitions are labeled with user defined information,  are capable of describing sequencing, concurrency, and choices.
 This formalism has well defined operational semantics, efficient analysis analysis algorithms and and has been used widely in modeling and analyzing communication protocols, concurrent programs, etc.
@@
 
-A simple communication protocol specification between two components represented by a LPN is shown in~Figure~\ref{fig-flow-ex-1}.
+The communication protocol in the BMPN shown in~Fig.~\ref{flow-spec-ex}(a) is represented by the LPN shown in~Figure~\ref{flow-spec-ex}(b).
 In this and the following figures for LPNs, the labeled circles denote places, and the labeled boxes denote transitions.
-Each transition is labeled with its name and the asSoCiated event.
+Each transition is labeled with its name and the associated event.
+Each event has a form of $\langle {\tt src, dest, cmd} \rangle$ where ${\tt cmd}$ is a command sent from a source component ${\tt src}$ to a destination component ${\tt dest}$.
 The solid black places without outgoing edges are {\em terminals}, which indicate termination of protocols represented by the LPNs.
-In this LPN, the initial state $\mathit{init} = \{p_1\}$.
-In this protocol, component~1 sends two consecutive messages $\mathit{msg}_1$ to component~2, which responds with either $\mathit{msg}_2$ or $\mathit{msg}_3$ non-deterministically.
-If $\mathit{msg}_2$ is responded, component~1 repeats the protocol by sending two consecutive $\mathit{msg}_1$ again.
-Otherwise, component~1 stops the protocol upon receiving $\mathit{msg}_3$.
-This LPN represents two possible communication sequences as shown below.
-\[
-\begin{array}{l}
-\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_1\;\ldots \\
-\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\ldots, \mathit{msg}_3\;
-\end{array}
-\]
-
-\begin{figure}[tb]
-\begin{center}
-\resizebox{2.5in}{!}{
-\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
-\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
-\tikzset{circ/.style={circle, draw,thick}}
-\tikzset{terminal/.style={circle, draw,line width=.8mm}}
-\tikzset{weight3/.style={line width=.3mm}}
-	% Nodes:
-	\node[circ]		(p1) 		at (0,0) {$p_1$};
-	\node[rectbox]	(t1)	at	($(p1) + (0,-1.1)$) {$t_1:msg_1$};
-	\node[circ] 	(p2) 		at ($(t1) + (0,-1.1)$) {$p_2$};
-	\node[rectbox]	(t2)	at	($(p2) + (0,-1.1)$) {$t_2:msg_2$};
-	\node[circ] 	(p3) 		at ($(t2) + (0,-1.1)$) {$p_3$};
-	\node[rectbox]	(t3)	at	($(p3) + (3,0)$) {$t_3:msg_3$};
-	\node[rectbox]	(t4)	at	($(p3) + (-3,0)$) {$t_4:msg_4$};
-	\node[circ,fill=black] 	(term) 		at ($(t4) + (0,-1.1)$) {$~~~$};
-
-	% Edges
-	\draw[->,weight3]	(p1) -- (t1);
-	\draw[->,weight3]	(t1) -- (p2);
-	\draw[->,weight3]	(p2) -- (t2);
-	\draw[->,weight3]	(t2) -- (p3);
-	\draw[->,weight3]	(p3) -- (t3);
-	\draw[->,weight3]	(p3) -- (t4);
-	\draw[->,weight3]	(t4) -- (term);	
-	\draw[->,weight3]	(t3) -- ($(p1)+(3,0)$) -- (p1);
-
-\end{tikzpicture}
-}
-\caption{An example of a LPN representing a system flow.}
-\label{fig-flow-ex-1}
-\end{center}
-\end{figure}
+The initial marking is $\mathit{init} = \{p_1\}$.
+In this LPN model, only the communication portion of the BPMN specification is represented while the computation portion is ignored.
 
@@
 
-% @Note  \red{\bf Another assumption:  the observed signal trace captures the values on the observed signals in every cycle from the beginning.  Too strong an assumption??}
+\red{\bf The following illustration may not be needed (from here to the end of this subsection) if seciton IV is kept.  Or we keep the illustration below but remove section IV which uses 3 pages.}
 

### 2015-09-16 paper.tex
Starting with LPN

@@
 
-\author{\large Hao Zheng$^1$, Yuting Cao$^1$, Sandip Ray$^2$, Jin Yang$^2$ \\
-$^1$Dept. of Computer Science and Eng., University of South Florida, Tampa, Fl 33620. USA. \\
-$^2$Strategic CAD Labs, Intel Corporation, Hillsboro, OR 97124.  USA.}
+%% \author{\large Hao Zheng$^1$, Yuting Cao$^1$, Sandip Ray$^2$, Jin Yang$^2$ \\
+%% $^1$Dept. of Computer Science and Eng., University of South Florida, Tampa, Fl 33620. USA. \\
+%% $^2$Strategic CAD Labs, Intel Corporation, Hillsboro, OR 97124.  USA.}
 \maketitle
@@
 is based on a formalization of system-level transactions via
-labeled Petri-Nets, which are capable of
-describing sequencing, concurrency, and choices of events.
-Given this formalization, we develop algorithms to infer
-system-level transactions from traces with missing, noisy,
-and ambiguous signal values, together with an estimate of
-confidence on the inference.  We demonstrate our approach on
-two SoC models: a SystemC prototype and a more realistic
+labeled Petri-Nets, which are capable of describing
+sequencing, concurrency, and choices of events.  Given this
+formalization, we develop algorithms to infer system-level
+transactions from traces with missing, noisy, and ambiguous
+signal values, together with an estimate of confidence on
+the inference.  We demonstrate our approach on two SoC case
+studies: a SystemC prototype and a more realistic
 implementation constructed within the GEM5
@@
 \begin{minipage}{3in}
-\includegraphics[height=2.5in]{figures/bpmn-flow-ex}
+\includegraphics[height=2in]{figures/bpmn-flow-ex}
 \end{minipage}
@@
 \end{tabular}
-\caption{(a) A graphical representation of a SoC firmware load protocol in BPMN~\cite{Krstic14HOST}, and (b) the LPN capturing the communication protocol in (a).}
+\caption{(a) A graphical representation of a SoC firmware
+  load protocol~\cite{Krstic14HOST}.  (b) LPN formalization.
+  Each event has a form of $\langle {\tt src, dest, cmd}
+  \rangle$ where ${\tt cmd}$ is a command sent from a source
+  component ${\tt src}$ to a destination component ${\tt
+    dest}$. The solid black places without outgoing edges
+  are {\em terminals}, which indicate termination of
+  protocols represented by the LPNs.}
 \label{flow-spec-ex}
@@
 
-\section{Background}
-
-\subsection{System-level Transactions in SoC Designs}
-
-An SoC design is typically composed of a large number of
-pre-designed hardware or software blocks (often referred to
-as ``intellectual properties'' or ``IPs'') that coordinate
-through complex protocols to implement system-level
-behavior.  Fig.~\ref{fig:flow-spec-ex} shows one such
-protocol, involving th
-
-\subsection{Labeled Petri-Nets}
-
-Although system flows depicted in BPMN as in Fig.~\ref{flow-spec-ex} are intuitive for system architects to understand communications across multiple components, they are not suitable for algorithmic analysis.
-In this section, we present an alternative formalism, \emph{labeled Petri Nets}, to represent communication aspects of system flows with the intuitive descriptions of internal operations for IP blocks abstracted away whenever possible.
-Labeled Petri nets, which are a subclass of traditional Petri-nets~\cite{Murata89} where Petri-net transitions are labeled with user defined information,  are capable of describing sequencing, concurrency, and choices.
-This formalism has well defined operational semantics, efficient analysis analysis algorithms and and has been used widely in modeling and analyzing communication protocols, concurrent programs, etc.
+\section{Labeled Petri-Nets}
+
+Labeled Petri-nets (LPN) is a formalization of state
+transition systems that is capable of describing sequencing,
+concurrency, and choices.  Fig.~\ref{flow-spec-ex} shows the
+formalization of a simple SoC protocol with Petri-Nets.
+Formally, an LPN is a tuple $(P, T, s_0, E, L)$ where $P$ is
+a finite set of {\em places}, $T$ is a finite set of {\em
+  transitions}, $\mathit{init}$ is the set of initially
+marked places, also referred to as the {\em initial
+  marking}, $E$ is a finite set of {\em events}, and $L: T
+\rightarrow E$ is a labeling function that maps each
+transition $t \in T$ to an event $e \in E$.  For each
+transition $t \in T$, its preset, denoted as $\bullet{t}
+\subseteq P$, is the set of places connected to $t$, and its
+postset, denoted as $t\bullet \subseteq P$, is the set of
+places that $t$ is connected to.  A marking $s \subseteq P$
+of a LPN is a subset of places marked with tokens, and it is
+also referred to as a state of a LPN.  The initial marking
+$\mathit{init}$ is also the initial state of the LPN.
+
+%% This formalism has well defined
+%% operational semantics, efficient analysis analysis
+%% algorithms and and has been used widely in modeling and
+%% analyzing communication protocols, concurrent programs, etc.
 
@@
 
-Specifically, a labeled Petri-net (LPN) is a tuple $(P, T, s_0, E, L)$ where
-\begin{itemize}
-\item $P$ is a finite set of places,
-\item $T$ is a finite set of transitions,
-\item $\mathit{init}$ is the set of initially marked places, also referred to as the initial marking.
-\item $E$ is a finite set of events,
-\item $L: T \rightarrow E$ is a labeling function that maps each transition $t \in T$ to an event $e \in E$.
-\end{itemize}
-
-For each transition $t \in T$, its preset, denoted as $\bullet{t} \subseteq P$, is the set of places connected to $t$, and its postset, denoted as $t\bullet \subseteq P$, is the set of places that $t$ is connected to.
-A marking $s \subseteq P$ of a LPN is a subset of places marked with tokens, and it is also referred to as a state of a LPN.
-The initial marking $\mathit{init}$ is also the initial state of the LPN.
-
-The operational semantics of a LPN is defined by transition executions.
+%% Formally, an LPN is a tuple $(P, T, s_0, E, L)$ where $P$ is
+%% a finite set of {\em places}, $T$ is a finite set of {\em
+%%   transitions}, $\mathit{init}$ is the set of initially
+%% marked places, also referred to as the {\em initial
+%%   marking}, $E$ is a finite set of {\em events}, and $L: T
+%% \rightarrow E$ is a labeling function that maps each
+%% transition $t \in T$ to an event $e \in E$.  For each
+%% transition $t \in T$, its preset, denoted as $\bullet{t}
+%% \subseteq P$, is the set of places connected to $t$, and its
+%% postset, denoted as $t\bullet \subseteq P$, is the set of
+%% places that $t$ is connected to.  A marking $s \subseteq P$
+%% of a LPN is a subset of places marked with tokens, and it is
+%% also referred to as a state of a LPN.  The initial marking
+%% $\mathit{init}$ is also the initial state of the LPN.
+
+%% The operational semantics of a LPN is defined by transition executions.
 A transition can be executed after it is {\em enabled}.
@@
 Execution of $t$ results in a new state $s'$ such that
-\[
-s' = (s - \bullet{t}) \cup t\bullet,
-\]
-and the emission of event $e$ labeled for $t$.
-
-The communication protocol in the BMPN shown in~Fig.~\ref{flow-spec-ex}(a) is represented by the LPN shown in~Figure~\ref{flow-spec-ex}(b).
-In this and the following figures for LPNs, the labeled circles denote places, and the labeled boxes denote transitions.
-Each transition is labeled with its name and the associated event.
-Each event has a form of $\langle {\tt src, dest, cmd} \rangle$ where ${\tt cmd}$ is a command sent from a source component ${\tt src}$ to a destination component ${\tt dest}$.
-The solid black places without outgoing edges are {\em terminals}, which indicate termination of protocols represented by the LPNs.
-The initial marking is $\mathit{init} = \{p_1\}$.
-In this LPN model, only the communication portion of the BPMN specification is represented while the computation portion is ignored.
+%% \[
+%% s' = (s - \bullet{t}) \cup t\bullet,
+%% \]
+%% and the emission of event $e$ labeled for $t$.
+
+%% The communication protocol in the BMPN shown
+%% in~Fig.~\ref{flow-spec-ex}(a) is represented by the LPN
+%% shown in~Figure~\ref{flow-spec-ex}(b).  In this and the
+%% following figures for LPNs, the labeled circles denote
+%% places, and the labeled boxes denote transitions.  Each
+%% transition is labeled with its name and the associated
+%% event.  Each event has a form of $\langle {\tt src, dest,
+%%   cmd} \rangle$ where ${\tt cmd}$ is a command sent from a
+%% source component ${\tt src}$ to a destination component
+%% ${\tt dest}$. The solid black places without outgoing edges
+%% are {\em terminals}, which indicate termination of protocols
+%% represented by the LPNs.  The initial marking is
+%% $\mathit{init} = \{p_1\}$.  In this LPN model, only the
+%% communication portion of the BPMN specification is
+%% represented while the computation portion is ignored.
 

### 2015-09-25 paper.tex
Added Sections 1-3

@@
-%% bare_conf.tex
-%% V1.3
-%% 2007/01/11
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
-
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compSoC.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print - the typesetting of the document will not typically be
-% affected with changes in paper size (but the bottom and side margins will).
-% Use the testflow package mentioned above to verify correct handling of
-% both paper sizes by the user's LaTeX system.
-%
-% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
-% should be used if it is desired that the figures are to be displayed in
-% draft mode.
-%
 \documentclass[conference]{IEEEtran}
-% Add the compSoC option for Computer SoCiety conferences.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
 \usepackage{pifont}
@@
 
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-
 \newcommand{\eg}{\mbox{{\em e.g.}}}
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at: 
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical SoCiety that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\usepackage[tight,footnotesize]{subfigure}
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\usepackage[caption=false]{caption}
-%\usepackage[font=footnotesize]{subfig}
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\usepackage[caption=false,font=footnotesize]{subfig}
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
+\newcommand{\ie}{\mbox{{\em i.e.}}}
+\newcommand{\viz}{\mbox{{\em viz.}}}
+
 \title{\Large {\bf Protocol-Guided Analysis of Post-silicon
@@
 %% $^2$Strategic CAD Labs, Intel Corporation, Hillsboro, OR 97124.  USA.}
+
+\begin{document}
+
 \maketitle
-
-%% % author names and affiliations
-%% % use a multiple column layout for up to three different
-%% % affiliations
-%% \author{\IEEEauthorblockN{Hao Zheng, Yuting Cao}
-%% \IEEEauthorblockA{Computer Science and Engineering\\
-%% University of South Florida\\
-%% Tampa, Florida 33620}
-%% \and
-%% \IEEEauthorblockN{Sandip Ray, Jin Yang}
-%% \IEEEauthorblockA{Strategic CAD Labs\\
-%% Intel Corporation\\
-%% Portland, Oregon
-%% }}
-
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-% 
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3}, 
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
 
@@
 
-% make the title area
-%\maketitle
-
-%% \begin{abstract}
-%% Validation of multicore systems-on-chip (SoC) is very challenging, and main reasons include the limited internal operation observability and non-determinism due to multiple clock domains.
-Very often, the observed signal traces offer limited values for debuggers to understand system internal behavior.
-This paper presents a trace analysis approach for post-silicon validation.
-This approach takes system flow specifications describing system behavior at a high level and an observed signal trace, and interprets the trace with respect to the system flow specification.
-The result from the trace interpretation is a set of flow instances such that their execution during a test explains the observed signal trace.
-This approach transforms partially observed and non-deterministic signal traces to a representation that is more understandable, and offers more helpful information to locate design defects and a measure to evaluate validation coverage.
-%% \end{abstract}
-% IEEEtran.cls defaults to using nonbold math in the Abstract.
-% This preserves the distinction between vectors and scalars. However,
-% if the conference you are submitting to favors bold math in the abstract,
-% then you can use LaTeX's standard command \boldmath at the very start
-% of the abstract to achieve this. Many IEEE journals/conferences frown on
-% math in the abstract anyway.
-
-% no keywords
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-% \IEEEpeermaketitle
-
 \section{Introduction}
-% no \IEEEPARstart
-%
-%Post silicon debug is getting more challenging for current complex system-on-chip design.
-%
-%Using system level flows in silicon trace analysis helps to reduce the volume of data to be stored and analyzed, difficulty caused by inherent non-determinism.  
-%
-%Also, abstracts signal traces to system use cases, which helps validation engineering better understand the behavior displayed by the signal traces.
 
@@
 
-A central component of post-silicon validation of SoC
-designs is to correlate trace from silicon execution with
-the intended system-level transactions.  An SoC design is
-typically composed of a large number of pre-designed
-hardware or software blocks (often referred to as
-``intellectual properties'' or ``IPs'') that coordinate
-through complex protocols to implement system-level
-behavior.  Any execution trace of the system involves a
-large number of interleaved instances of these protocols.
-For example, consider a smartphone executing a usage
-scenerio where the end-user browses the Web while listening
-to music and sending and receiving occasional text messages.
-Typical post-silicon validation use-case involve exercising
-such scenarios.  An execution trace for this scenario would
-involve activities from the CPU, audio controller, display
-controller, wireless radio antenna, etc., reflecting the
-interleaved execution of several communication protocols.
-On the other hand, due to observability limitations, only a
-small small number of participating signals can be actually
-traced during silicon execution.  Furthermore, due to
-electrical perturbations, silicon data can be noisy, lossy,
-and ambiguous.  Consequently, it is non-trivial to identify
-all participating protocols and pinpoint the ``right''
-interleaving that results in an observed trace.
+A key component of post-silicon validation of SoC designs is
+to correlate trace from silicon execution with the intended
+system-level transactions.  An SoC design is typically
+composed of a large number of pre-designed hardware or
+software blocks (often referred to as ``intellectual
+properties'' or ``IPs'') that coordinate through complex
+protocols to implement the system-level behavior.  Any
+execution trace of the system involves a large number of
+interleaved instances of these protocols.  For example,
+consider a smartphone executing a usage scenerio where the
+end-user browses the Web while listening to music and
+sending and receiving occasional text messages.  Typical
+post-silicon validation use-case involve exercising such
+scenarios.  An execution trace would involve activities from
+the CPU, audio controller, display controller, wireless
+radio antenna, etc., reflecting the interleaved execution of
+several communication protocols.  On the other hand, due to
+observability limitations, only a small small number of
+participating signals can be actually traced during silicon
+execution.  Furthermore, due to electrical perturbations,
+silicon data can be noisy, lossy, and ambiguous.
+Consequently, it is non-trivial to identify all
+participating protocols and pinpoint the interleaving that
+results in an observed trace.
 
@@
 protocol-level behavior from silicon traces in SoC designs.
-More specifically, given a collection of system-level
-communication protocols and a trace of (partially observed)
-hardware signals, our approach infers, with a certain
-measure of confidence, the protocol instances (and their
-interleavings) being exercised by the trace.  Our approach
-is based on a formalization of system-level transactions via
-labeled Petri-Nets, which are capable of describing
-sequencing, concurrency, and choices of events.  Given this
-formalization, we develop algorithms to infer system-level
-transactions from traces with missing, noisy, and ambiguous
-signal values, together with an estimate of confidence on
-the inference.  We demonstrate our approach on two SoC case
-studies: a SystemC prototype and a more realistic
-implementation constructed within the GEM5
+Given a collection of system-level communication protocols
+and a trace of (partially observed) hardware signals, our
+approach infers, with a certain measure of confidence, the
+protocol instances (and their interleavings) being exercised
+by the trace.  The approach is based on a formalization of
+system-level transactions via labeled Petri-Nets, which are
+capable of describing sequencing, concurrency, and choices
+over system events.  We develop algorithms to infer
+system-level transactions from traces with missing, noisy,
+and ambiguous signal values, together with an estimate of
+confidence on the inference.  We demonstrate our approach on
+two SoC case studies: a pedagogical SystemC prototype and a
+more realistic implementation constructed within the GEM5
 environment~\cite{Binkert2011}.
-
-%% Systems-on-chips (SoCs) have become more complex as more functionalities are integrated on chips.
-To reduce design time, more external IPs are integrated into current SoC designs.
-At the high level, a SoC can be viewed as a network of IP blocks communicating by a large number of protocols.
-A typical SoC design consists of multiple programmable processor cores where a number of software threads can be executed simultaneously.
-The resulting designs usually operate in multiple clock domains, and show a high degree of concurrency.
-Due to the sophisticated communication mechanisms employed in  current SoC designs to support complex functional behaviors, traditional processor-centric validation has to be enhanced to focus inter-core communications.
-
-%% To make validation even more challenging, the high degree of integration in current SoC designs causes traditional chip observation points to become inaccessible.
-A SoC design usually contains millions of signals.
-In post-silicon validation, thousands of those signals are tapped for observation.
-Due to the limited availability of the chip pins, only about a hundred of the tapped signals can be traced.
-This severely limits observability hinders post-silicon debuggers' capability to root cause an observed failure.
-To make the situation even worse, the observed traces often display non-deterministic behavior due to many concurrent operations occurred in multiple clock domains in a design.
-The above factors make debugging highly difficult in failure root-causing, and need to be addressed.
-
-%% In a typical design flow, a design starts with some high-level specification, particularly for communication protocols.
-The abstract protocol specification usually captures information exchanged among IP blocks and how it is exchanged to support functional behavior.
-%% Graphical representations of protocols such as Message Sequence Charts~\cite{Harel03} or Live Sequence Charts~\cite{Damm99} describe protocols intuitively.
-However, its success in hardware industry is limited.
-Recently, message flows in BPMN~\cite{Krstic14HOST} have been being advocated as a formalism for specifying architecture requirements
-%%  or system-level use cases such as power management, security, etc.
-In general, a flow describes a set of message sequences across multiple IP blocks following a system communication protocol.
-Specification of system flows BPMN is intuitive and easier for human to understand and to reason about system behavior.
-An example of flow specification in BPMN is shown in Figure~\ref{flow-spec-ex}.
-
-%% \begin{figure}[tb]
-%% \begin{center}
-%% \includegraphics[height=2.5in]{figures/bpmn-flow-ex}
-%% \caption{A graphical representation of a SoC firmware load protocol in BPMN~\cite{Krstic14HOST}.}
-%% \label{flow-spec-ex}
-%% \end{center}
-%% \end{figure}
-
-%% As indicated above, interpreting and understanding the raw signal traces observed from executing the chip is extremely difficult due to the limited observability and non-determinism.
-This paper proposes a trace analysis method which analyzes and interprets the observed signal traces at the level of system flow specifications.
-During the step of analysis and interpretation, this method tries to find out possible executions of system flows such that the observed signal trace is a result of such executions.
-The resulting scenarios of flow executions often give debuggers more clear and better structured system-level views of the internal operations of the system under debug.
-With a better understanding of system operations in a test environment, therefore, it would make easier for debuggers to pinpoint and root cause system failures.
-In addition to debugging, another important benefit is better coverage evaluation as a scenario of flow executions captures the information about the types and number of instances of flows executed by the chip in a test environment.
-That information can help to determine the quality and coverage of validation.
-
-%---
 
@@
 
-\section{Labeled Petri-Nets}
+\section{Background}
+
+\subsection{SoC Protocols and Post-silicon Trace Analysis}
+
+An SoC design involves integration of a number of IPs that
+communicate through complex protocols.  Such system-level
+protocols are typically specified in architecture documents
+as message flow diagrams.  As illustration,
+Fig~\ref{flow-spec-ex}(a) shows one such diagram for a
+protcol to authenticate and load a firmware during system
+boot for firmware upgrade.  During validation, the system
+under debug (SUD) exercises some complex system-level
+use-case which involves interleaved execution of possibly a
+large number of such protocols.  A trace of a small number
+of hardware signals is then shipped off-chip analysis.  The
+off-chip analysis includes two broad phases: (1)~trace
+abstraction, and (2)~trace interpretation.  Trace
+abstraction maps the hardware trace into higher-level
+architectural constructs, \eg, messages, operations, etc.: a
+message such as {\tt Authorization request} may be
+implemented in hardware through a Boolean or temporal
+combination of specific hardware signals in the NoC fabric
+between {\tt Device} and {\tt CE}, \eg, as a sequence
+containing a header, a specific value of a sequence of data
+words, etc.  We will refer to such architectural constructs
+as {\em protocol events} or {\em flow events}.  Note that
+due to limited observability, it may not be possible to map
+a given set of (observed) hardware signals to uniquely to a
+flow event.  Finally, the trace may include several
+instances of the same protocol executing concurrently, \eg,
+a firmware authentication protocol may be invoked when
+another instance of the protocol has not completed.
+
+Trace interpretation entails mapping flow events created
+during trace abstraction to system-level protocols in order
+to identify the set of protocol instances (and
+interleavings) responsible for creating the observed
+behavior.  The trace may identify a problem in the protocols
+themselves, \eg an interleaving of some protocol executions
+may lead to an unexpected message being sent or cause the
+system to crash.  More commonly, one finds a bug in the {\em
+  implementation} of the protocol, \ie, a trace inconsistent
+with any possible interleaving of the protocol executions.
+Identifying these problems involves significant human
+expertise, and can often take days to weeks of effort.
+
+%% In a typical validation setting, the system under debug
+%% (SUD) is executed in a test environment until it is
+%% terminated by the test environment or the system crashes
+%% due to a failure.  During the execution, a trace on a
+%% small number of observable signals is streamed off the
+%% chip for debugging.
+
+\subsection{Labeled Petri-Nets}
 
@@
 transition systems that is capable of describing sequencing,
-concurrency, and choices.  Fig.~\ref{flow-spec-ex} shows the
-formalization of a simple SoC protocol with Petri-Nets.
+concurrency, and choices.  Fig.~\ref{flow-spec-ex}(b)
+illustrates how to use LPN to formalize protocols.
 Formally, an LPN is a tuple $(P, T, s_0, E, L)$ where $P$ is
@@
 $\mathit{init}$ is also the initial state of the LPN.
+
+%% with respect to information captured in the system
+%% flows.  This requires recognizing signal events and mapping
+%% the signal events or sequences of the signal events to flow
+%% events appeared in system flow specifications.  For example,
+%% a data write message may be a single event in a system flow,
+%% however, such a flow event maybe implemented in hardware as
+%% a sequence of events including a header, a number of data
+%% word transfers and a tail.  This abstraction requires a
+%% mapping relation from flow events to sequences of signal
+%% events, and it is reasonable to assume that this relation is
+%% available.
+
+%% The trace interpretation takes a finite trace of flow events
+%% resulting from the trace abstraction and a set of system
+%% flows in LPNs $\vec{F}$, and generates a set of possible
+%% system flow execution scenarios.  A \emph{flow execution
+%%   scenario} is defined as $\{(F_{i,j}, s_{i,j})\}$ where in
+%% each element $(F_{i,j}, s_{i,j})$, $F_{i,j}$ is the $j$th
+%% activated instance of flow $F_i \in \vec{F}$, and $s_{i,j}$
+%% is a state of $F_{i,j}$.  A flow execution scenario
+%% indicates that at a certain point of SUD execution, what
+%% types of flows and the number of instances of a particular
+%% flow are activated and their corresponding current states.
 
@@
 %=====
-\section{Flow-Directed Trace Analysis}
-
-In a typical validation setting, the system under debug (SUD) is executed in a test environment until it is terminated by the test environment or the system crashes due to a failure.
-During the execution, a trace on a small number of observable signals is streamed off the chip for debugging.
-Due to the limited observability and inherent non-determinism in today's SoC designs, the observed signal trace is difficult to understand, thus providing limited values for debugging.
-In this section, we describe a trace analysis method where the observed signal traces are interpreted at the level of system flows.
-In general, the trace analysis can offer debuggers a structured view of communications among the IP blocks during the SUD execution by deriving the types and numbers of system flows activated during SUD executions from the observed signal traces.
-
-The trace analysis method consists of two steps: \emph{trace abstraction} and \emph{interpretation}.
-The trace abstraction takes a signal trace observed during the execution of the SUD, and abstracts it with respect to information captured in the system flows.
-This requires recognizing signal events and mapping the signal events or sequences of the signal events to flow events appeared in system flow specifications.
-For example, a data write message may be a single event in a system flow, however, such a flow event maybe implemented in hardware as a sequence of events including a header, a number of data word transfers and a tail.
-This abstraction requires a mapping relation from flow events to sequences of signal events, and it is reasonable to assume that this relation is available.
-
-The trace interpretation takes a finite trace of flow events resulting from the trace abstraction and a set of system flows in LPNs $\vec{F}$, and generates a set of possible system flow execution scenarios.
-A \emph{flow execution scenario} is defined as $\{(F_{i,j}, s_{i,j})\}$ where in each element $(F_{i,j}, s_{i,j})$, $F_{i,j}$ is the $j$th activated instance of flow $F_i \in \vec{F}$, and $s_{i,j}$ is a state of $F_{i,j}$.
-A flow execution scenario indicates that at a certain point of SUD execution, what types of flows and the number of instances of a particular flow are activated and their corresponding current states.
-
-An observed trace of flow events $\rho = e_1e_2\ldots e_n$ is a result of SUD executing the flow instances of some flow execution scenario.
-The goal of the trace interpretation is to derive such scenario.
-Let $\mathit{accept(F_{i,j}, s_{i,j}, e)}$ be a function that determines if event $e$ can be emitted by $F_{i,j}$ in state $s_{i,j}$.
-This function is used during the trace interpretation for checking whether an event of an observed trace is a result of executing some flow instance.
-Formally,  $\mathit{accept(F_{i,j}, s_{i,j}, e)}$ returns $(F_{i,j}, s'_{i,j})$ if there exists a transition $t$ in $F_i$ such that $L(t) = e$ and $\bullet t \subseteq s_{i,j}$.
-In this case, $s'_{i,j} = (s_{i,j} - \bullet t) \cup t\bullet$.
-It returns $\emptyset$ if no such $t$ exists in $F_i$.
-
-Given a trace of flow events $\rho = e_1e_2\ldots e_n$, the trace interpretation algorithm starts with an empty set of of flow execution scenario $\mathit{Scen} = \emptyset$.
-Then, for each $e_h$ where $1 \leq h \leq n$ starting $h=1$, and for each $\mathit{scen} \in \mathit{Scen}$, the following two steps are performed.
-\begin{description}
-\item[Step 1]~~For each $(F_{i,j}, s_{i,j}) \in \mathit{scen}$, if $\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = (F_{i,j}, s'_{i,j})$, create a new scenario $\mathit{scen}' = (\mathit{scen} - (F_{i,j}, s_{i,j})) \cup (F_{i,j}, s'_{i,j})$, which is added into $\mathit{Scen}'$.
-
-\item[Step 2]~~For each $F_i \in \vec{F}$, create a new instance $F_{i, j+1}$.
-If $\mathit{accept}(F_{i,j+1}, \mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$,  create a new scenario $\mathit{scen}' = \mathit{scen} \cup (F_{i,j+1}, s'_{i,j+1})$, which is added into $\mathit{Scen}'$.
-\end{description}
-After $e_h$ is processed, $Scen = Scen'$, and the above two steps repeat for the next event $e_{h+1}$.
-
-If every events in $\rho$ is successfully mapped to some flow instance, this algorithm returns a set of flow execution scenarios such that every flow instance is in its terminal state.
-On the other hand, inconsistent events can also be encountered.
-An event $e$ is \emph{inconsistent} if for each flow execution scenario $\mathit{scen} \in \mathit{Scen}$, the following two conditions hold.
-\begin{enumerate}
-\item For each $(F_{i,j}, s_{i,j}) \in \mathit{scen}$, $\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = \emptyset$,
-\item For each $F_i \in \vec{F}$, $\mathit{accept}(F_{i}, \mathit{init}_{i}, e_h) = \emptyset$.
-\end{enumerate}
-An inconsistent event is the one produced by SUD execution but cannot be mapped to any flow instances no matter how the trace prior to event $e$ is interpreted.
-Inconsistent events indicates possible causes of system failures.
-
-Based on the above discussion, the trace interpretation algorithm returns two pieces of information: 1) a set $G$ of flow execution scenarios where every flow instance in every scenario is in its terminal state, 2) a set $B$ of pairs, each of which includes a set of flow execution scenarios and an inconsistent event.
-The set $B$ provides valuable information for debuggers to root cause system failures.
-One of these two sets can empty.
-With the full observability, the set $G$ includes a single flow execution scenario derived for a trace.
-In reality, it is always the case that the SUD is only partially observable.
-Therefore, due to the lack of information for precise interpretation, a set of flow execution scenarios is typically derived for a given trace as the result of the trace analysis.
-
-\red{\bf The following illustration may not be needed (from here to the end of this subsection) if seciton IV is kept.  Or we keep the illustration below but remove section IV which uses 3 pages.}
-
-To illustrate the basic idea of the trace analysis method, consider the system flow shown in Figure~\ref{fig-flow-ex-1}.
+
+\section{Flow-Directed Trace Interpretation}
+
+%% In a typical validation setting, the system under debug
+%% (SUD) is executed in a test environment until it is
+%% terminated by the test environment or the system crashes
+%% due to a failure.  During the execution, a trace on a
+%% small number of observable signals is streamed off the
+%% chip for debugging.  Due to the limited observability and
+%% inherent non-determinism in today's SoC designs, the
+%% observed signal trace is difficult to understand, thus
+%% providing limited values for debugging.  In this section,
+%% we describe a trace analysis method where the observed
+%% signal traces are interpreted at the level of system
+%% flows.  In general, the trace analysis can offer
+%% debuggers a structured view of communications among the
+%% IP blocks during the SUD execution by deriving the types
+%% and numbers of system flows activated during SUD
+%% executions from the observed signal traces.
+
+%% The trace analysis method consists of two steps:
+%% \emph{trace abstraction} and \emph{interpretation}.  The
+%% trace abstraction takes a signal trace observed during
+%% the execution of the SUD, and abstracts it with respect
+%% to information captured in the system flows.  This
+%% requires recognizing signal events and mapping the signal
+%% events or sequences of the signal events to flow events
+%% appeared in system flow specifications.  For example, a
+%% data write message may be a single event in a system
+%% flow, however, such a flow event maybe implemented in
+%% hardware as a sequence of events including a header, a
+%% number of data word transfers and a tail.  This
+%% abstraction requires a mapping relation from flow events
+%% to sequences of signal events, and it is reasonable to
+%% assume that this relation is available.
+
+In this section we formalize the trace interpretation
+problem in terms of labeled Petri-nets, and discuss our
+algorithms to address the problem.  For pedagogical reasons,
+here we assume full observability of all hardware signals
+involved in the flow events.  In the next section we will
+extend the approach to partial observability.
+
+\medskip
+
+\noindent 
+{\bf Notations and formalization.}  We assume that the set
+of system flows is provided as a collection ${\vec{F}}$ of
+LPNs.  We define a {\em flow execution scenario} to be the
+set of pairs $(F_{i,j}, s_{i,j})$ where $F_{i,j}$ is the
+$j$th instance of flow $F_i \in {\vec{F}}$, and $s_{i,j}$ is
+a state of $F_{i,j}$.  A flow execution scenario indicates
+the set of protocols and the number of instances of a
+particular protocol are activated and their corresponding
+current states.  Since we assume full observability, we view
+an {\em observed trace} $\rho = e_1e_2\ldots e_n$ is a
+sequence of events.  Given an observed trace $\rho$, the
+goal of of trace interpretation is to construct a set of
+candidate flow execution scenarios whose execution can
+create the sequence of events in $\rho$.  We call such
+execution scenarios {\em compliant} with $\rho$.  Let
+$\mathit{accept(F_{i,j}, s_{i,j}, e)}$ be a function that
+determines if event $e$ can be emitted by $F_{i,j}$ in state
+$s_{i,j}$.  Formally, $\mathit{accept(F_{i,j}, s_{i,j}, e)}$
+returns $(F_{i,j}, s'_{i,j})$ if there exists a transition
+$t$ in $F_i$ such that $L(t) = e$ and $\bullet t \subseteq
+s_{i,j}$.  In this case, $s'_{i,j} = (s_{i,j} - \bullet t)
+\cup t\bullet$.  It returns $\emptyset$ if no such $t$
+exists in $F_i$.
+
+\medskip
+
+Given an obsered trace $\rho$ and the set $\vec{F}$ of LPNs,
+Algorithm~\ref{algo:compliance} provides the basic steps for
+computing a set of compliant flow execution scenarios.
+\begin{algorithm}[ht]
+\DontPrintSemicolon
+$\mathit{Scen} \gets \emptyset$\;
+\ForEach {$h, \; 1 \leq h \leq n $} {
+$\mathit{found} \gets {\tt true}$ \;
+  \ForEach {$(F_{i,j}, s_{i,j}) \in \mathit{scen}$} {
+    \If {$\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = (F_{i,j}, s'_{i,j})$} {
+      $\mathit{scen} \gets (\mathit{scen} - (F_{i,j}, s_{i,j})) \cup (F_{i,j}, s'_{i,j})$\;
+      $\mathit{scen} \gets \mathit{scen} \cup \mathit{Scen}'$\;
+      $\mathit{found} \gets {\tt false}$ \;
+    }
+  }
+  \ForEach {$F_i \in \vec{F}$} {
+      create a new instance $F_{i, j+1}$ \;
+      \If {$\mathit{accept}(F_{i,j+1},\mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$} {
+    $\mathit{scen} \gets \mathit{scen} \cup (F_{i,j+1}, s'_{i,j+1})$ \;
+        $\mathit{found} \gets {\tt false}$ \;
+      }
+    }
+  \If {$\mathit{found} == {\tt true}$} {
+    \Return ${\tt Inconsistent}$\;
+  }
+}
+\Return $\mathit{scen}$ \;
+\caption{$\textsc{Check-Compliance}(\vec{F}, \, \rho)$}
+\label{algo:compliance}
+\end{algorithm}
+The algorithm operates by keeping track (in variable
+$\mathit{scen}$) of a set of candidate protocol instances
+compliant with each prefix of $\rho$.  At each iteration, we
+update this set by either extending some member of
+$\mathit{scen}$ or initiating a new protocol instance.  If
+neither is possible then we report that the trace is {\tt
+  inconsistent}, \ie, there is no possible interleaving of
+the protocol instances from $\vec{F}$ that is compliant with
+$\rho$.
+
+%% Given a trace of flow events $\rho = e_1e_2\ldots e_n$, the
+%% trace interpretation algorithm starts with an empty set of
+%% of flow execution scenario $\mathit{Scen} = \emptyset$.
+%% Then, for each $e_h$ where $1 \leq h \leq n$ starting $h=1$,
+%% and for each $\mathit{scen} \in \mathit{Scen}$, the
+%% following two steps are performed.
+%% \begin{description}
+%% \item[Step 1]~~For each $(F_{i,j}, s_{i,j}) \in
+%%   \mathit{scen}$, if $\mathit{accept}(F_{i,j}, s_{i,j}, e_h)
+%%   = (F_{i,j}, s'_{i,j})$, create a new scenario
+%%   $\mathit{scen}' = (\mathit{scen} - (F_{i,j}, s_{i,j}))
+%%   \cup (F_{i,j}, s'_{i,j})$, which is added into
+%%   $\mathit{Scen}'$.
+
+%% \item[Step 2]~~For each $F_i \in \vec{F}$, create a new
+%%   instance $F_{i, j+1}$.  If $\mathit{accept}(F_{i,j+1},
+%%   \mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$,
+%%   create a new scenario $\mathit{scen}' = \mathit{scen} \cup
+%%   (F_{i,j+1}, s'_{i,j+1})$, which is added into
+%%   $\mathit{Scen}'$.
+%% \end{description}
+
+%% After $e_h$ is processed, $Scen = Scen'$, and the above two
+%% steps repeat for the next event $e_{h+1}$.
+
+%% If every events in $\rho$ is successfully mapped to some
+%% flow instance, this algorithm returns a set of flow
+%% execution scenarios such that every flow instance is in its
+%% terminal state.  On the other hand, inconsistent events can
+%% also be encountered.  An event $e$ is \emph{inconsistent} if
+%% for each flow execution scenario $\mathit{scen} \in
+%% \mathit{Scen}$, the following two conditions hold.
+
+%% \begin{enumerate}
+%% \item For each $(F_{i,j}, s_{i,j}) \in \mathit{scen}$,
+%%   $\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = \emptyset$,
+%% \item For each $F_i \in \vec{F}$, $\mathit{accept}(F_{i},
+%%   \mathit{init}_{i}, e_h) = \emptyset$.
+%% \end{enumerate}
+
+%% An inconsistent event is the one produced by SUD execution
+%% but cannot be mapped to any flow instances no matter how the
+%% trace prior to event $e$ is interpreted.  Inconsistent
+%% events indicates possible causes of system failures.
+
+%% Based on the above discussion, the trace interpretation
+%% algorithm returns two pieces of information: 1) a set $G$ of
+%% flow execution scenarios where every flow instance in every
+%% scenario is in its terminal state, 2) a set $B$ of pairs,
+%% each of which includes a set of flow execution scenarios and
+%% an inconsistent event.  The set $B$ provides valuable
+%% information for debuggers to root cause system failures.
+%% One of these two sets can empty.  With the full
+%% observability, the set $G$ includes a single flow execution
+%% scenario derived for a trace.  In reality, it is always the
+%% case that the SUD is only partially observable.  Therefore,
+%% due to the lack of information for precise interpretation, a
+%% set of flow execution scenarios is typically derived for a
+%% given trace as the result of the trace analysis.
+
+%% \red{\bf The following illustration may not be needed (from here to the end of this subsection) if seciton IV is kept.  Or we keep the illustration below but remove section IV which uses 3 pages.}
+
+%% To illustrate the basic idea of the trace analysis method, consider the system flow shown in Figure~\ref{fig-flow-ex-1}.
 Let $F_1$ denote such flow.
 Suppose that a hardware system implements flow $F_1$, and the following trace of flow events is abstracted from an observed signal trace as a result of executing such system.
-\[
-\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_3\;\ldots
-\]  
-This trace is interpreted from the first event to the last in order to derive all possible flow execution scenarios.
+%% \[
+%% \mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_3\;\ldots
+%% \]  
+%% This trace is interpreted from the first event to the last in order to derive all possible flow execution scenarios.
 At the beginning, the first event $\mathit{msg}_1$ is processed.
@@
 
-Next, the second $\mathit{msg}_1$ is interpreted on both scenarios   This event could be the result of two possible cases.
+%% Next, the second $\mathit{msg}_1$ is interpreted on both scenarios   This event could be the result of two possible cases.
 In the first case, this event is the result of the continuing execution of $F_{1,1}$ as $\mathit{accept}(F_{1,1}, \{p_2\}, \mathit{msg}_1) = (F_{1,1}, \{p_3\})$.
@@
 Therefore, the interpretation of the first two events $\mathit{msg}_1$ leads to two flow execution scenarios as shown below.
-\[
-\label{tr-analy-s1}\tag{1}
-\begin{array}{cl}
-1. & \{(F_{1,1}, \{p_3\})\},\\
-2. & \{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_2\})\} 
-\end{array}
-\]
-
-Now, consider the third $\mathit{msg}_1$ for each of the two scenario derived in~(\ref{tr-analy-s1}).
+%% \[
+%% \label{tr-analy-s1}\tag{1}
+%% \begin{array}{cl}
+%% 1. & \{(F_{1,1}, \{p_3\})\},\\
+%% 2. & \{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_2\})\} 
+%% \end{array}
+%% \]
+
+%% Now, consider the third $\mathit{msg}_1$ for each of the two scenario derived in~(\ref{tr-analy-s1}).
 For the execution scenario~1, $F_{1,1}$ is not able to accept $\mathit{msg}_1$ as it is in state $\{p_3\}$.
@@
 Therefore, this execution scenario can be revised accordingly as
-\[
-\label{tr-analy-s2}\tag{2}
-\{(F_{1,1}, \{p_3\}),(F_{1,2},\{p_2\})\}.
-\]   
-For the execution scenario~2, event $\mathit{msg}_1$ could be the result of continuing execution of $F_{1,1}$ or $F_{1,2}$, or it could be a result of activation of a new flow instance.
+%% \[
+%% \label{tr-analy-s2}\tag{2}
+%% \{(F_{1,1}, \{p_3\}),(F_{1,2},\{p_2\})\}.
+%% \]   
+%% For the execution scenario~2, event $\mathit{msg}_1$ could be the result of continuing execution of $F_{1,1}$ or $F_{1,2}$, or it could be a result of activation of a new flow instance.
 Therefore,  three new execution scenarios can be derived as shown below for this event.
-\[
-\label{tr-analy-s3}\tag{3}
-\begin{array}{l}
-\{(F_{1,1}, \{p_3\}), (F_{1,2}, \{p_2\})\},\\
-\{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_3\})\}, \\
-\{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_2\}), (F_{1,3}, \{p_2\})\}
-\end{array}
-\]
-Since the flow execution scenario in~(\ref{tr-analy-s2}) already exists in~(\ref{tr-analy-s3}), the three flow execution scenarios shown in~(\ref{tr-analy-s3}) is the result from the interpretation of the first three events $\mathit{msg}_1$.
-
-The next flow event in the trace $\mathit{msg_2}$ is analyzed for the flow execution scenarios as shown in~(\ref{tr-analy-s3}).
+%% \[
+%% \label{tr-analy-s3}\tag{3}
+%% \begin{array}{l}
+%% \{(F_{1,1}, \{p_3\}), (F_{1,2}, \{p_2\})\},\\
+%% \{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_3\})\}, \\
+%% \{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_2\}), (F_{1,3}, \{p_2\})\}
+%% \end{array}
+%% \]
+%% Since the flow execution scenario in~(\ref{tr-analy-s2}) already exists in~(\ref{tr-analy-s3}), the three flow execution scenarios shown in~(\ref{tr-analy-s3}) is the result from the interpretation of the first three events $\mathit{msg}_1$.
+
+%% The next flow event in the trace $\mathit{msg_2}$ is analyzed for the flow execution scenarios as shown in~(\ref{tr-analy-s3}).
 For the first scenario $\{(F_{1,1}, \{p_3\}), (F_{1,2}, \{p_2\})\}$, event $\mathit{msg_2}$ can only be the result from executing $F_{1,1}$ as $\mathit{accept}(F_{1,1}, \{p_3\}, \mathit{msg_2}) = (F_{1,1}, \{p_1\})$.
@@
 After analyzing event $\mathit{msg_2}$, the updated system execution scenarios are shown below.
-\[
-\begin{array}{l}
-\{(F_{1,1}, \{p_1\}), (F_{1,2}, \{p_2\})\},\ \{(F_{1,1}, \{p2\}), (F_{1,2}, \{p_1\})\}.
-\end{array}
-\]
-
-The last flow event in the trace is $\mathit{msg}_3$.
+%% \[
+%% \begin{array}{l}
+%% \{(F_{1,1}, \{p_1\}), (F_{1,2}, \{p_2\})\},\ \{(F_{1,1}, \{p2\}), (F_{1,2}, \{p_1\})\}.
+%% \end{array}
+%% \]
+
+%% The last flow event in the trace is $\mathit{msg}_3$.
 Considering the above two possible system executions, neither can allow this event to be produced as none of the flow instances in both system executions is in state $\{p_3\}$.
@@
 %---
-\subsection{Trace Analysis with Partial Observability}
-
-In hardware that implements a given system flow specification, a flow event is defined as an event or a sequence of events on a set of signals.
-Due to the limited number of pins on the boundary of chips available for observation, only a small fraction of system signals can be observed during debug.
-In this section, we discuss how the trace analysis method presented above can be adapted to deal with signal traces of partial observability.
+\section{Trace Analysis with Partial Observability}
+
+The trace interpretation discussion in the preceding section
+assumed full observability.  In this section, we discuss how
+to adapt the method to deal with silicon traces under
+limited (partial) obsrvability.
+
+%% In hardware that implements a given system flow
+%% specification, a flow event is defined as an event or a
+%% sequence of events on a set of signals.  Due to the limited
+%% number of pins on the boundary of chips available for
+%% observation, only a small fraction of system signals can be
+%% observed during debug.  In this section, we discuss how the
+%% trace analysis method presented above can be adapted to deal
+%% with signal traces of partial observability.
 

### 2015-09-28 case1.tex
added one more experiment into case study
Section III, updated algorithm 1, added a small example to illustrate
the trace analysis.

section V: results from one more experiment are added.

Case study I is removed to save space.

--- 2015-09-25 paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\usepackage{pifont}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{mybeamer}
-%\usepackage{draftwatermark}
-\usepackage{hyperref}
-\hypersetup{
-    colorlinks=true,
-    linkcolor=blue,
-    filecolor=magenta,      
-    urlcolor=cyan,
-}
-
-\newcommand{\eg}{\mbox{{\em e.g.}}}
-\newcommand{\ie}{\mbox{{\em i.e.}}}
-\newcommand{\viz}{\mbox{{\em viz.}}}
-
-\title{\Large {\bf Protocol-Guided Analysis of Post-silicon
-  Traces Under Limited Observability}}
-
-%% \author{\large Hao Zheng$^1$, Yuting Cao$^1$, Sandip Ray$^2$, Jin Yang$^2$ \\
-%% $^1$Dept. of Computer Science and Eng., University of South Florida, Tampa, Fl 33620. USA. \\
-%% $^2$Strategic CAD Labs, Intel Corporation, Hillsboro, OR 97124.  USA.}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-We consider the problem of reconstructing system-level
-behavior of an SoC design from a partially observed signal
-trace.  Solving this problem is a critical activity in
-post-silicon validation, and currently depends primarily on
-human creativity and insight.  We provide algorithms to
-automatically infer system-level transactions from
-incomplete, ambiguous, and noisy trace data, together with a
-measure of confidence.  We demonstrate the approach on
-illustrative system-level protocols for SoC models developed
-in SystemC as well as in an FPGA environment.
-\end{abstract}
-
-\section{Introduction}
-
-Post-silicon validation makes use of pre-production silicon
-integrated circuit (IC) to ensure that the fabricated system
-works as desired under actual operating conditions with real
-software.  Since the silicon executes at target clock speed,
-post-silicon executions are billions of times faster than
-RTL simulations, and even provide speed-up of several orders
-of magnitude over other pre-silicon platforms (\eg, FPGA,
-system-level emulation, etc.).  This makes it possible to
-explore deep design states which cannot be exercised in
-pre-silicon, and identify errors missed during pre-silicon
-validation and debug.  Post-silicon validation is a critical
-component of the design validation life-cycle for modern
-microprocessors and SoC designs.  Unfortunately, it is also
-a highly complex component, performed under aggressive
-schedules and accounting for more than $50\%$ of the overall
-design validation cost.  Consequently, it is crucial to
-develop techniques for streamlining and automating
-post-silicon validation activities.
-
-A key component of post-silicon validation of SoC designs is
-to correlate trace from silicon execution with the intended
-system-level transactions.  An SoC design is typically
-composed of a large number of pre-designed hardware or
-software blocks (often referred to as ``intellectual
-properties'' or ``IPs'') that coordinate through complex
-protocols to implement the system-level behavior.  Any
-execution trace of the system involves a large number of
-interleaved instances of these protocols.  For example,
-consider a smartphone executing a usage scenerio where the
-end-user browses the Web while listening to music and
-sending and receiving occasional text messages.  Typical
-post-silicon validation use-case involve exercising such
-scenarios.  An execution trace would involve activities from
-the CPU, audio controller, display controller, wireless
-radio antenna, etc., reflecting the interleaved execution of
-several communication protocols.  On the other hand, due to
-observability limitations, only a small small number of
-participating signals can be actually traced during silicon
-execution.  Furthermore, due to electrical perturbations,
-silicon data can be noisy, lossy, and ambiguous.
-Consequently, it is non-trivial to identify all
-participating protocols and pinpoint the interleaving that
-results in an observed trace.
-
-In this paper, we consider the problem of reconstructing
-protocol-level behavior from silicon traces in SoC designs.
-Given a collection of system-level communication protocols
-and a trace of (partially observed) hardware signals, our
-approach infers, with a certain measure of confidence, the
-protocol instances (and their interleavings) being exercised
-by the trace.  The approach is based on a formalization of
-system-level transactions via labeled Petri-Nets, which are
-capable of describing sequencing, concurrency, and choices
-over system events.  We develop algorithms to infer
-system-level transactions from traces with missing, noisy,
-and ambiguous signal values, together with an estimate of
-confidence on the inference.  We demonstrate our approach on
-two SoC case studies: a pedagogical SystemC prototype and a
-more realistic implementation constructed within the GEM5
-environment~\cite{Binkert2011}.
-
-\begin{figure*}[tb]
-\begin{center}
-\begin{tabular}{cc}
-\begin{minipage}{3in}
-\includegraphics[height=2in]{figures/bpmn-flow-ex}
-\end{minipage}
-&
-\begin{minipage}{3in}
-\resizebox{2.8in}{!}{
-\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
-\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
-\tikzset{circ/.style={circle, draw,thick}}
-\tikzset{terminal/.style={circle, draw,line width=.8mm}}
-\tikzset{weight3/.style={line width=.3mm}}
-	% Nodes:
-	\node[circ]		(p1) 		at (0,0) {$p_1$};
-	\node[rectbox]	(t1)	at	($(p1) + (0,-1.1)$) {$t_1:\langle {\tt Driver:Device:Load} \_fw\rangle$};
-	\node[circ] 	(p2) 		at ($(t1) + (0,-1.1)$) {$p_2$};
-	\node[rectbox]	(t2)	at	($(p2) + (0,-1.1)$) {$t_2:\langle {\tt Device:CE:Auth\_req} \rangle$};
-	\node[circ] 	(p3) 		at ($(t2) + (0,-1.1)$) {$p_3$};
-	\node[rectbox]	(t3)	at	($(p3) + (0,-1.1)$) {$t_3:\langle {\tt CE:Device:Auth\_resp}\rangle$};
-	\node[circ] 	(p4) 		at ($(t3) + (0,-1.1)$) {$p_4$};
-	\node[rectbox]	(t4)	at	($(p4) + (-3,-1.1)$) {$t_4:\langle {\tt Device:Driver:Report} \rangle$};
-	\node[rectbox]	(t5)	at	($(p4) + (3,-1.1)$) {$t_5:\langle {\tt Device:CE:Ack} \rangle$};
-	\node[circ,fill=black] 	(term1) 		at ($(t4) + (0,-1.1)$) {$~~~$};
-	\node[circ,fill=black] 	(term2) 		at ($(t5) + (0,-1.1)$) {$~~~$};
-
-	% Edges
-	\draw[->,weight3]	(p1) -- (t1);
-	\draw[->,weight3]	(t1) -- (p2);
-	\draw[->,weight3]	(p2) -- (t2);
-	\draw[->,weight3]	(t2) -- (p3);
-	\draw[->,weight3]	(p3) -- (t3);
-	\draw[->,weight3]	(t3) -- (p4);
-	\draw[->,weight3]	(p4) -- ($(p4)+(-3,0)$) -- (t4);
-	\draw[->,weight3]	(p4) -- ($(p4)+(3,0)$) -- (t5);
-	\draw[->,weight3]	(t4) -- (term1);	
-	\draw[->,weight3]	(t5) -- (term2);
-
-\end{tikzpicture}
-}
-\end{minipage}
-\\
-(a) & (b)
-\end{tabular}
-\caption{(a) A graphical representation of a SoC firmware
-  load protocol~\cite{Krstic14HOST}.  (b) LPN formalization.
-  Each event has a form of $\langle {\tt src, dest, cmd}
-  \rangle$ where ${\tt cmd}$ is a command sent from a source
-  component ${\tt src}$ to a destination component ${\tt
-    dest}$. The solid black places without outgoing edges
-  are {\em terminals}, which indicate termination of
-  protocols represented by the LPNs.}
-\label{flow-spec-ex}
-\end{center}
-\end{figure*}
-
-\section{Background}
-
-\subsection{SoC Protocols and Post-silicon Trace Analysis}
-
-An SoC design involves integration of a number of IPs that
-communicate through complex protocols.  Such system-level
-protocols are typically specified in architecture documents
-as message flow diagrams.  As illustration,
-Fig~\ref{flow-spec-ex}(a) shows one such diagram for a
-protcol to authenticate and load a firmware during system
-boot for firmware upgrade.  During validation, the system
-under debug (SUD) exercises some complex system-level
-use-case which involves interleaved execution of possibly a
-large number of such protocols.  A trace of a small number
-of hardware signals is then shipped off-chip analysis.  The
-off-chip analysis includes two broad phases: (1)~trace
-abstraction, and (2)~trace interpretation.  Trace
-abstraction maps the hardware trace into higher-level
-architectural constructs, \eg, messages, operations, etc.: a
-message such as {\tt Authorization request} may be
-implemented in hardware through a Boolean or temporal
-combination of specific hardware signals in the NoC fabric
-between {\tt Device} and {\tt CE}, \eg, as a sequence
-containing a header, a specific value of a sequence of data
-words, etc.  We will refer to such architectural constructs
-as {\em protocol events} or {\em flow events}.  Note that
-due to limited observability, it may not be possible to map
-a given set of (observed) hardware signals to uniquely to a
-flow event.  Finally, the trace may include several
-instances of the same protocol executing concurrently, \eg,
-a firmware authentication protocol may be invoked when
-another instance of the protocol has not completed.
-
-Trace interpretation entails mapping flow events created
-during trace abstraction to system-level protocols in order
-to identify the set of protocol instances (and
-interleavings) responsible for creating the observed
-behavior.  The trace may identify a problem in the protocols
-themselves, \eg an interleaving of some protocol executions
-may lead to an unexpected message being sent or cause the
-system to crash.  More commonly, one finds a bug in the {\em
-  implementation} of the protocol, \ie, a trace inconsistent
-with any possible interleaving of the protocol executions.
-Identifying these problems involves significant human
-expertise, and can often take days to weeks of effort.
-
-%% In a typical validation setting, the system under debug
-%% (SUD) is executed in a test environment until it is
-%% terminated by the test environment or the system crashes
-%% due to a failure.  During the execution, a trace on a
-%% small number of observable signals is streamed off the
-%% chip for debugging.
-
-\subsection{Labeled Petri-Nets}
-
-Labeled Petri-nets (LPN) is a formalization of state
-transition systems that is capable of describing sequencing,
-concurrency, and choices.  Fig.~\ref{flow-spec-ex}(b)
-illustrates how to use LPN to formalize protocols.
-Formally, an LPN is a tuple $(P, T, s_0, E, L)$ where $P$ is
-a finite set of {\em places}, $T$ is a finite set of {\em
-  transitions}, $\mathit{init}$ is the set of initially
-marked places, also referred to as the {\em initial
-  marking}, $E$ is a finite set of {\em events}, and $L: T
-\rightarrow E$ is a labeling function that maps each
-transition $t \in T$ to an event $e \in E$.  For each
-transition $t \in T$, its preset, denoted as $\bullet{t}
-\subseteq P$, is the set of places connected to $t$, and its
-postset, denoted as $t\bullet \subseteq P$, is the set of
-places that $t$ is connected to.  A marking $s \subseteq P$
-of a LPN is a subset of places marked with tokens, and it is
-also referred to as a state of a LPN.  The initial marking
-$\mathit{init}$ is also the initial state of the LPN.
-
-%% with respect to information captured in the system
-%% flows.  This requires recognizing signal events and mapping
-%% the signal events or sequences of the signal events to flow
-%% events appeared in system flow specifications.  For example,
-%% a data write message may be a single event in a system flow,
-%% however, such a flow event maybe implemented in hardware as
-%% a sequence of events including a header, a number of data
-%% word transfers and a tail.  This abstraction requires a
-%% mapping relation from flow events to sequences of signal
-%% events, and it is reasonable to assume that this relation is
-%% available.
-
-%% The trace interpretation takes a finite trace of flow events
-%% resulting from the trace abstraction and a set of system
-%% flows in LPNs $\vec{F}$, and generates a set of possible
-%% system flow execution scenarios.  A \emph{flow execution
-%%   scenario} is defined as $\{(F_{i,j}, s_{i,j})\}$ where in
-%% each element $(F_{i,j}, s_{i,j})$, $F_{i,j}$ is the $j$th
-%% activated instance of flow $F_i \in \vec{F}$, and $s_{i,j}$
-%% is a state of $F_{i,j}$.  A flow execution scenario
-%% indicates that at a certain point of SUD execution, what
-%% types of flows and the number of instances of a particular
-%% flow are activated and their corresponding current states.
-
-%% This formalism has well defined
-%% operational semantics, efficient analysis analysis
-%% algorithms and and has been used widely in modeling and
-%% analyzing communication protocols, concurrent programs, etc.
-
-%--- LPN definition with levels
-%Specifically, a labeled Petri-net (LPN) is a tuple $(P, T, M_0, V, E, L)$ where
-%\begin{itemize}
-%\item $P$ is a finite set of places,
-%\item $T$ is a finite set of transitions,
-%\item $M_0$ is the set of initially marked places, also referred to as the initial marking.
-%\item $V$ is a finite set of auxiliary variables of integer type,
-%\item $E$ is a finite set of events,
-%\item $L$ is a labeling function that maps each transition $t \in T$ to a triple $(g, e, A)$ where
-%	\begin{itemize}
-%	\item $g$ is a predicate over $V$,
-%	\item $e \in E$ is an event,
-%	\item $A$ is a set of assignments to some variables in $V$.  
-%	\end{itemize}
-%\end{itemize}
-%
-%For each transition $t \in T$, its preset, denoted as $\bullet{t} \subseteq P$, is the set of places connected to $t$, and its postset, denoted as $t\bullet \subseteq P$, is the set of places that $t$ is connected to.
-A marking of a LPN is a set of places marked with tokens, and a state of a LPN is $(M, \alpha)$ where $M$ is a marking and $\alpha$ is an assignment to all auxiliary variables in $V$.
-%
-%The operational semantics of a LPN is defined by transition executions.
-A transition can be executed after it is {\em enabled}.
-A transition $t = (g, e, A)$ is enabled in a state $(M, \alpha)$ if every place in its preset is included in the marking, i.e.
-$\bullet{t} \subseteq M$, and the values of the auxiliary variables make the predicate $g$ be true, i.e.
-$\alpha \models g$.
-Execution of $t$ results in a new state $(M', \alpha')$ such that
-%\[
-%M' = (M - \bullet{t}) \cup t\bullet,
-%\]
-%and $\alpha'$ is a new assignment where the values of the auxiliary variables are updated with respect to the assignments $A$.  Furthermore, when $t$ is executed, the event $e$ is emitted.  SoC
-
-%% Formally, an LPN is a tuple $(P, T, s_0, E, L)$ where $P$ is
-%% a finite set of {\em places}, $T$ is a finite set of {\em
-%%   transitions}, $\mathit{init}$ is the set of initially
-%% marked places, also referred to as the {\em initial
-%%   marking}, $E$ is a finite set of {\em events}, and $L: T
-%% \rightarrow E$ is a labeling function that maps each
-%% transition $t \in T$ to an event $e \in E$.  For each
-%% transition $t \in T$, its preset, denoted as $\bullet{t}
-%% \subseteq P$, is the set of places connected to $t$, and its
-%% postset, denoted as $t\bullet \subseteq P$, is the set of
-%% places that $t$ is connected to.  A marking $s \subseteq P$
-%% of a LPN is a subset of places marked with tokens, and it is
-%% also referred to as a state of a LPN.  The initial marking
-%% $\mathit{init}$ is also the initial state of the LPN.
-
-%% The operational semantics of a LPN is defined by transition executions.
-A transition can be executed after it is {\em enabled}.
-A transition $t \in T$ is enabled in a state $s$ if every place in its preset is included in the marking, i.e.
-$\bullet{t} \subseteq s$.
-Execution of $t$ results in a new state $s'$ such that
-%% \[
-%% s' = (s - \bullet{t}) \cup t\bullet,
-%% \]
-%% and the emission of event $e$ labeled for $t$.
-
-%% The communication protocol in the BMPN shown
-%% in~Fig.~\ref{flow-spec-ex}(a) is represented by the LPN
-%% shown in~Figure~\ref{flow-spec-ex}(b).  In this and the
-%% following figures for LPNs, the labeled circles denote
-%% places, and the labeled boxes denote transitions.  Each
-%% transition is labeled with its name and the associated
-%% event.  Each event has a form of $\langle {\tt src, dest,
-%%   cmd} \rangle$ where ${\tt cmd}$ is a command sent from a
-%% source component ${\tt src}$ to a destination component
-%% ${\tt dest}$. The solid black places without outgoing edges
-%% are {\em terminals}, which indicate termination of protocols
-%% represented by the LPNs.  The initial marking is
-%% $\mathit{init} = \{p_1\}$.  In this LPN model, only the
-%% communication portion of the BPMN specification is
-%% represented while the computation portion is ignored.
-
-% @COMMENT This example is for LPNs with guards.
-%The above protocol specification can be made more precise by describing how component~2 would respond without using non-determinism.
-For example, the system architect may wish to specify that component~2  responds with $\mathit{msg_2}$ to every sequence of two $\mathit{msg_1}$ for five such sequences in a row.
-On the sixth such sequence, component~2 responds with $\mathit{msg_3}$.
-The LPN modeling such protocol is shown in Figure~\ref{fig-flow-ex-2}.
-An alternative without using auxiliary variables in the transition predicates is by unrolling the loops for the specified number of times; however this would result in a larger and more complex LPN.
-%
-%\begin{figure}[htbp]
-%\begin{center}
-%\includegraphics[height=1.6in]{figures/flow-ex-2}
-%\caption{Another example of a LPN representing the system flow in Fig.~\ref{fig-flow-ex-1} with a fixed number of message exchanges.}
-%\label{fig-flow-ex-2}
-%\end{center}
-%\end{figure}
-
-%=====
-
-\section{Flow-Directed Trace Interpretation}
-
-%% In a typical validation setting, the system under debug
-%% (SUD) is executed in a test environment until it is
-%% terminated by the test environment or the system crashes
-%% due to a failure.  During the execution, a trace on a
-%% small number of observable signals is streamed off the
-%% chip for debugging.  Due to the limited observability and
-%% inherent non-determinism in today's SoC designs, the
-%% observed signal trace is difficult to understand, thus
-%% providing limited values for debugging.  In this section,
-%% we describe a trace analysis method where the observed
-%% signal traces are interpreted at the level of system
-%% flows.  In general, the trace analysis can offer
-%% debuggers a structured view of communications among the
-%% IP blocks during the SUD execution by deriving the types
-%% and numbers of system flows activated during SUD
-%% executions from the observed signal traces.
-
-%% The trace analysis method consists of two steps:
-%% \emph{trace abstraction} and \emph{interpretation}.  The
-%% trace abstraction takes a signal trace observed during
-%% the execution of the SUD, and abstracts it with respect
-%% to information captured in the system flows.  This
-%% requires recognizing signal events and mapping the signal
-%% events or sequences of the signal events to flow events
-%% appeared in system flow specifications.  For example, a
-%% data write message may be a single event in a system
-%% flow, however, such a flow event maybe implemented in
-%% hardware as a sequence of events including a header, a
-%% number of data word transfers and a tail.  This
-%% abstraction requires a mapping relation from flow events
-%% to sequences of signal events, and it is reasonable to
-%% assume that this relation is available.
-
-In this section we formalize the trace interpretation
-problem in terms of labeled Petri-nets, and discuss our
-algorithms to address the problem.  For pedagogical reasons,
-here we assume full observability of all hardware signals
-involved in the flow events.  In the next section we will
-extend the approach to partial observability.
-
-\medskip
-
-\noindent 
-{\bf Notations and formalization.}  We assume that the set
-of system flows is provided as a collection ${\vec{F}}$ of
-LPNs.  We define a {\em flow execution scenario} to be the
-set of pairs $(F_{i,j}, s_{i,j})$ where $F_{i,j}$ is the
-$j$th instance of flow $F_i \in {\vec{F}}$, and $s_{i,j}$ is
-a state of $F_{i,j}$.  A flow execution scenario indicates
-the set of protocols and the number of instances of a
-particular protocol are activated and their corresponding
-current states.  Since we assume full observability, we view
-an {\em observed trace} $\rho = e_1e_2\ldots e_n$ is a
-sequence of events.  Given an observed trace $\rho$, the
-goal of of trace interpretation is to construct a set of
-candidate flow execution scenarios whose execution can
-create the sequence of events in $\rho$.  We call such
-execution scenarios {\em compliant} with $\rho$.  Let
-$\mathit{accept(F_{i,j}, s_{i,j}, e)}$ be a function that
-determines if event $e$ can be emitted by $F_{i,j}$ in state
-$s_{i,j}$.  Formally, $\mathit{accept(F_{i,j}, s_{i,j}, e)}$
-returns $(F_{i,j}, s'_{i,j})$ if there exists a transition
-$t$ in $F_i$ such that $L(t) = e$ and $\bullet t \subseteq
-s_{i,j}$.  In this case, $s'_{i,j} = (s_{i,j} - \bullet t)
-\cup t\bullet$.  It returns $\emptyset$ if no such $t$
-exists in $F_i$.
-
-\medskip
-
-Given an obsered trace $\rho$ and the set $\vec{F}$ of LPNs,
-Algorithm~\ref{algo:compliance} provides the basic steps for
-computing a set of compliant flow execution scenarios.
-\begin{algorithm}[ht]
-\DontPrintSemicolon
-$\mathit{Scen} \gets \emptyset$\;
-\ForEach {$h, \; 1 \leq h \leq n $} {
-$\mathit{found} \gets {\tt true}$ \;
-  \ForEach {$(F_{i,j}, s_{i,j}) \in \mathit{scen}$} {
-    \If {$\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = (F_{i,j}, s'_{i,j})$} {
-      $\mathit{scen} \gets (\mathit{scen} - (F_{i,j}, s_{i,j})) \cup (F_{i,j}, s'_{i,j})$\;
-      $\mathit{scen} \gets \mathit{scen} \cup \mathit{Scen}'$\;
-      $\mathit{found} \gets {\tt false}$ \;
-    }
-  }
-  \ForEach {$F_i \in \vec{F}$} {
-      create a new instance $F_{i, j+1}$ \;
-      \If {$\mathit{accept}(F_{i,j+1},\mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$} {
-    $\mathit{scen} \gets \mathit{scen} \cup (F_{i,j+1}, s'_{i,j+1})$ \;
-        $\mathit{found} \gets {\tt false}$ \;
-      }
-    }
-  \If {$\mathit{found} == {\tt true}$} {
-    \Return ${\tt Inconsistent}$\;
-  }
-}
-\Return $\mathit{scen}$ \;
-\caption{$\textsc{Check-Compliance}(\vec{F}, \, \rho)$}
-\label{algo:compliance}
-\end{algorithm}
-The algorithm operates by keeping track (in variable
-$\mathit{scen}$) of a set of candidate protocol instances
-compliant with each prefix of $\rho$.  At each iteration, we
-update this set by either extending some member of
-$\mathit{scen}$ or initiating a new protocol instance.  If
-neither is possible then we report that the trace is {\tt
-  inconsistent}, \ie, there is no possible interleaving of
-the protocol instances from $\vec{F}$ that is compliant with
-$\rho$.
-
-%% Given a trace of flow events $\rho = e_1e_2\ldots e_n$, the
-%% trace interpretation algorithm starts with an empty set of
-%% of flow execution scenario $\mathit{Scen} = \emptyset$.
-%% Then, for each $e_h$ where $1 \leq h \leq n$ starting $h=1$,
-%% and for each $\mathit{scen} \in \mathit{Scen}$, the
-%% following two steps are performed.
-%% \begin{description}
-%% \item[Step 1]~~For each $(F_{i,j}, s_{i,j}) \in
-%%   \mathit{scen}$, if $\mathit{accept}(F_{i,j}, s_{i,j}, e_h)
-%%   = (F_{i,j}, s'_{i,j})$, create a new scenario
-%%   $\mathit{scen}' = (\mathit{scen} - (F_{i,j}, s_{i,j}))
-%%   \cup (F_{i,j}, s'_{i,j})$, which is added into
-%%   $\mathit{Scen}'$.
-
-%% \item[Step 2]~~For each $F_i \in \vec{F}$, create a new
-%%   instance $F_{i, j+1}$.  If $\mathit{accept}(F_{i,j+1},
-%%   \mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$,
-%%   create a new scenario $\mathit{scen}' = \mathit{scen} \cup
-%%   (F_{i,j+1}, s'_{i,j+1})$, which is added into
-%%   $\mathit{Scen}'$.
-%% \end{description}
-
-%% After $e_h$ is processed, $Scen = Scen'$, and the above two
-%% steps repeat for the next event $e_{h+1}$.
-
-%% If every events in $\rho$ is successfully mapped to some
-%% flow instance, this algorithm returns a set of flow
-%% execution scenarios such that every flow instance is in its
-%% terminal state.  On the other hand, inconsistent events can
-%% also be encountered.  An event $e$ is \emph{inconsistent} if
-%% for each flow execution scenario $\mathit{scen} \in
-%% \mathit{Scen}$, the following two conditions hold.
-
-%% \begin{enumerate}
-%% \item For each $(F_{i,j}, s_{i,j}) \in \mathit{scen}$,
-%%   $\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = \emptyset$,
-%% \item For each $F_i \in \vec{F}$, $\mathit{accept}(F_{i},
-%%   \mathit{init}_{i}, e_h) = \emptyset$.
-%% \end{enumerate}
-
-%% An inconsistent event is the one produced by SUD execution
-%% but cannot be mapped to any flow instances no matter how the
-%% trace prior to event $e$ is interpreted.  Inconsistent
-%% events indicates possible causes of system failures.
-
-%% Based on the above discussion, the trace interpretation
-%% algorithm returns two pieces of information: 1) a set $G$ of
-%% flow execution scenarios where every flow instance in every
-%% scenario is in its terminal state, 2) a set $B$ of pairs,
-%% each of which includes a set of flow execution scenarios and
-%% an inconsistent event.  The set $B$ provides valuable
-%% information for debuggers to root cause system failures.
-%% One of these two sets can empty.  With the full
-%% observability, the set $G$ includes a single flow execution
-%% scenario derived for a trace.  In reality, it is always the
-%% case that the SUD is only partially observable.  Therefore,
-%% due to the lack of information for precise interpretation, a
-%% set of flow execution scenarios is typically derived for a
-%% given trace as the result of the trace analysis.
-
-%% \red{\bf The following illustration may not be needed (from here to the end of this subsection) if seciton IV is kept.  Or we keep the illustration below but remove section IV which uses 3 pages.}
-
-%% To illustrate the basic idea of the trace analysis method, consider the system flow shown in Figure~\ref{fig-flow-ex-1}.
-Let $F_1$ denote such flow.
-Suppose that a hardware system implements flow $F_1$, and the following trace of flow events is abstracted from an observed signal trace as a result of executing such system.
-%% \[
-%% \mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_3\;\ldots
-%% \]  
-%% This trace is interpreted from the first event to the last in order to derive all possible flow execution scenarios.
-At the beginning, the first event $\mathit{msg}_1$ is processed.
-According to the flow specification $F_1$, we know that one instance of such flow $F_1$, $F_{1,1}$,  is activated by the SUD as $\mathit{accept}(F_{1,1}, init_1, \mathit{msg}_1) = (F_{1,1}, \{p_2\})$ where $\mathit{init} = \{p_1\}$ is the initial state of $F_1$.
-As the result, the flow execution scenario after interpreting the first event $\mathit{msg_1}$ is $\{(F_{1,1}, \{p_2\})\}$.
-
-%% Next, the second $\mathit{msg}_1$ is interpreted on both scenarios   This event could be the result of two possible cases.
-In the first case, this event is the result of the continuing execution of $F_{1,1}$ as $\mathit{accept}(F_{1,1}, \{p_2\}, \mathit{msg}_1) = (F_{1,1}, \{p_3\})$.
-In the second case, the system may activate another instance of $F_1$, $F_{1,2}$ such that the second event $\mathit{msg_1}$ is a result of executing this new instance.
-Therefore, the interpretation of the first two events $\mathit{msg}_1$ leads to two flow execution scenarios as shown below.
-%% \[
-%% \label{tr-analy-s1}\tag{1}
-%% \begin{array}{cl}
-%% 1. & \{(F_{1,1}, \{p_3\})\},\\
-%% 2. & \{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_2\})\} 
-%% \end{array}
-%% \]
-
-%% Now, consider the third $\mathit{msg}_1$ for each of the two scenario derived in~(\ref{tr-analy-s1}).
-For the execution scenario~1, $F_{1,1}$ is not able to accept $\mathit{msg}_1$ as it is in state $\{p_3\}$.
-On the other hand, this event could be the result of activation of a new flow instance.
-Therefore, this execution scenario can be revised accordingly as
-%% \[
-%% \label{tr-analy-s2}\tag{2}
-%% \{(F_{1,1}, \{p_3\}),(F_{1,2},\{p_2\})\}.
-%% \]   
-%% For the execution scenario~2, event $\mathit{msg}_1$ could be the result of continuing execution of $F_{1,1}$ or $F_{1,2}$, or it could be a result of activation of a new flow instance.
-Therefore,  three new execution scenarios can be derived as shown below for this event.
-%% \[
-%% \label{tr-analy-s3}\tag{3}
-%% \begin{array}{l}
-%% \{(F_{1,1}, \{p_3\}), (F_{1,2}, \{p_2\})\},\\
-%% \{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_3\})\}, \\
-%% \{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_2\}), (F_{1,3}, \{p_2\})\}
-%% \end{array}
-%% \]
-%% Since the flow execution scenario in~(\ref{tr-analy-s2}) already exists in~(\ref{tr-analy-s3}), the three flow execution scenarios shown in~(\ref{tr-analy-s3}) is the result from the interpretation of the first three events $\mathit{msg}_1$.
-
-%% The next flow event in the trace $\mathit{msg_2}$ is analyzed for the flow execution scenarios as shown in~(\ref{tr-analy-s3}).
-For the first scenario $\{(F_{1,1}, \{p_3\}), (F_{1,2}, \{p_2\})\}$, event $\mathit{msg_2}$ can only be the result from executing $F_{1,1}$ as $\mathit{accept}(F_{1,1}, \{p_3\}, \mathit{msg_2}) = (F_{1,1}, \{p_1\})$.
-Based on the same reasoning, this event can only be the result from executing $F_{1,2}$ in the second scenario, and it moves $F_{1,2}$ to a new state $\{p_3\}$ too.
-The interesting case is the third scenario where none of the flow instances can allow $\mathit{msg_2}$ to happen.
-This is due to the fact that the flow must be in $\{p_3\}$ for $\mathit{msg_2}$ to happen.
-This indicates the third system execution scenario is impossible for the prefix of the flow event trace upto $\mathit{msg_2}$, therefore this scenario is ignored from further analysis.
-After analyzing event $\mathit{msg_2}$, the updated system execution scenarios are shown below.
-%% \[
-%% \begin{array}{l}
-%% \{(F_{1,1}, \{p_1\}), (F_{1,2}, \{p_2\})\},\ \{(F_{1,1}, \{p2\}), (F_{1,2}, \{p_1\})\}.
-%% \end{array}
-%% \]
-
-%% The last flow event in the trace is $\mathit{msg}_3$.
-Considering the above two possible system executions, neither can allow this event to be produced as none of the flow instances in both system executions is in state $\{p_3\}$.
-What this means is that the system does {\em not} implement the flow specification correctly as it produces something not allowed by the specification.
-By examining the system executions right before the ``buggy'' event, debuggers may gain more information on when and where the problem might be.
-The trace interpretation algorithm adds these two scenarios along with the fifth event $\mathit{msg_3}$ into $B$, and returns it for debuggers to analyze.
-
-%---
-\section{Trace Analysis with Partial Observability}
-
-The trace interpretation discussion in the preceding section
-assumed full observability.  In this section, we discuss how
-to adapt the method to deal with silicon traces under
-limited (partial) obsrvability.
-
-%% In hardware that implements a given system flow
-%% specification, a flow event is defined as an event or a
-%% sequence of events on a set of signals.  Due to the limited
-%% number of pins on the boundary of chips available for
-%% observation, only a small fraction of system signals can be
-%% observed during debug.  In this section, we discuss how the
-%% trace analysis method presented above can be adapted to deal
-%% with signal traces of partial observability.
-
-In general, a signal trace of partial observability corresponding a set of traces of flow events due to the ambiguous interpretation of signal events.
-In the following, we discuss two cases for trace abstraction on partial observability: mapping a single signal event to a flow event or mapping a sequence of signal events to a flow event.
-A signal event is defined as a state on or an assignment to a set of signals.
-
-Hereafter, the term {\em flow traces} is used to refer to traces of flow events. 
-Consider the following example for the first case.
-Suppose that there are three flow events: $e_1$, $e_2$, and $e_3$, which are implemented in hardware by the signal events shown in the list below.
-We use Boolean expressions to represent signal events for the discussion.
-\[
-\begin{array}{cl}
-e_1: & abc\\
-e_2: & \bar{a}bc\\
-e_3: & a\bar{b}c
-\end{array}
-\] 
-Now suppose that only signals $b$ and $c$ are observable, and a signal trace of this partial observability is obtained below.
-\[
-bc\ bc \ \bar{b}c
-\]
-During the trace abstraction step, the first two signal events $bc$ can be mapped to $\{e_1, e_2\}$ since $a$ is not observable, and the last one $b'c$ is mapped to $\{e_3\}$.
-Therefore, this signal trace is abstracted to four flow traces, $\{e_1, e_2\} \times \{e_1, e_2\} \times \{e_3\}$.
-
-Next, we consider the case where a flow event is mapped from a sequence of signal events.
-Now suppose that two other flow events are implemented by sequences of signal events as defined in the list below.
-\[
-\begin{array}{cl}
-e_4: & abc\ \bar{a}bc\\
-e_5: & abc\ abc\ abc\ \bar{a}bc\\
-\end{array}
-\] 
-Given an observed trace of the same observability shown below
-\[
-bc\ bc\ bc \ bc,
-\]
-it is abstracted to the following flow traces.
-\[
-e_4e_4,\; \textvisiblespace e_4 \textvisiblespace,\;  e_4 \textvisiblespace \textvisiblespace,\;  \textvisiblespace \textvisiblespace e_4, \; e_5
-\] 
-where $\textvisiblespace$ denotes signal events that are not mapped to any flow events.
-Note that the above abstraction leads to three distinct flow traces as the middle three correspond to the same trace of flow events.
-
-From the above discussions, it can be seen that a signal trace of partial observability is generally mapped to a set of flow traces.
-As the the number of signals available for observation becomes smaller, the number of flow event traces corresponding to a signal trace as a result of the abstraction can be enormous.
-This phenomenon can increase the complexity of the trace interpretation as it can cause the number of possible flow execution scenarios generated during the analysis to explode.
-Possible solutions to address this issue include better trace signal selection for observation and assistance from debuggers' insights.
-Trace signal selection itself is an important and difficult subject, and a detailed discussion of it is out of scope of this paper.
-Next, we briefly describe how the debuggers' insights of a system's architecture can help to address the complexity issue in the trace analysis.
-
-\subsection{Inputs from Validators}
-
-During the trace interpretation, the number of intermediate flow execution scenarios may become too large due to ambiguous interpretation from signal events to flow events or from flow events to flow execution scenarios.
-The explosion of the intermediate results can significantly slow down the performance of the trace analysis.
-To address this problem, the debuggers can use their insights and understanding of the SUD to trim the total number of possible system executions derived from the trace analysis.
-When a SUD is validated, a debugger is assumed to have deep knowledge about the system's architecture and microarchitecture, and how the test environment affects the system executions.
-For instance, he/she may have knowledge that in a test environment, the maximal number of instances of a flow can be activated by the SUD, or a flow can be activated after certain other flows have terminated, etc.
-These debugging insights can be encoded as constraints into the trace analysis method, which can then be used to eliminate a large number of flow execution scenarios that violate these constraints during the trace interpretation step.
-
-This approach can be flexible in that it allows a debugger to analyze the observed traces in a trail-and-error manner if the precise knowledge of the system (micro-)architecture is hard to come by.
-For instance, the debugger might initially make a very restricted assumption on how the SUD executes a flow specification, and these assumptions can potentially lead to an empty set of flow execution scenarios.
-Depending on which of these assumptions triggered during the trace interpretation step, the debugger can study these assumptions more carefully, and relax some or all of them for the next run of analysis.
-This iteration can be repeated as many times as necessary until some results deemed meaningful are produced.
-
-Alternatively, if all derived execution scenarios seem to be plausible, the implication that a debugger may draw from this result is that the failure may be independent of the flows being observed.
-Therefore, the testing environment can be adjusted in order for a different part or different behavior of the SUD to be observed.
-This idea, closely related to trace signal selection, is critical for post-silicon validation, and a detailed discussion can only be presented in a separate paper.
-
 %=====
@@
 At the end of the analysis, all possible flow execution scenarios are returned to system validators for examination.
-
-%
-%--
-%
-\section{Case Study II}
-
-\begin{figure} 
-\centerline{
-\includegraphics[width=3.3in]{figures/Fig4.png}}
-\caption{SoC platform structure.}
-\label{SoC}
-\end{figure}
-
-In order to find out the efficiency of the trace analysis method for more realistic examples, in this case study, a more detailed transaction level model of a SoC is constructed within the GEM5 environment~\cite{Binkert2011}.
-This SoC model consists of two ARM Cortex-A9 cores, each of which contains two separate 16KB data and instruction caches.
-The caches are connected to a 1GB memory through a memory bus model.
-The architecture of this SoC model is shown in Fig.~\ref{SoC}.
-
-In this model, components communicate with each other by sending and receiving various request and response messages.
-In order to observe and trace communications occurring inside this model during execution, monitors are attached to links connecting the components.
-These monitors record the messages flowing through the links they are attached to, and store them into output trace files.
-
-%By combining the informations from all of the nine communication monitors, required communication traces are obtained.
-As a virtualized SoC platform, GEM5 has three types of request: timing, atomic and functional.
-Timing request include the modeling of queuing delay and resource contention.
-Atomic request is a faster than timing request with no delay.
-Functional request is used for loading binaries, examining/changing variables in the simulated system, and so on.
-In our case, we specify all our request to be timed as it's the most detailed one.However, some of the system initiation is still atomic and it's not related to our research, so we only took the messages that are timing request from our collected data.
-
-%\subsection{Flow Specification}
-
-For this model, we consider the flow specifications describing the cache coherence protocols supported in GEM5 that is used to build the model in Fig~\ref{SoC}.
-These flow specifications describe data/instruction read operations and data write operations initiated from CPUs.
-Three such flows describe the cache coherent protocols for each CPU.
-As there are two CPUs in this model, there are actually six such flows considered for this model.
-More detailed information about these protocols and the flow specifications can be found online\footnote{\url{https://github.com/cao2/paper/blob/master/Flow\%20Specification.pdf}} to meet the page limit.
-
-%In the following section, a brief overview for the implemented instances of these flows are given.
-These flow specifications capture how messages are exchanged for different use cases.
-In this model, a message is defined with the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform.
-%\begin{figure} 
-%\centerline{
-%\includegraphics[width=3.7In]{figures/Fig5.png}}
-%
-%{\footnotesize
-%\[
-%\begin{array}{llll}
-%msg_0: (&\mbox{ CPU1},&\mbox{writeReq},&\mbox{cache1   })\\       
-%msg_1: (&\mbox{ cache1},&\mbox{ readExreq },&\mbox{Bus     })\\        
-%msg_2: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2 })\\  
-%msg_3: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2         })\\   
-%msg_4: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2           })\\  
-%msg_5: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2 })\\  
-%msg_6: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce1       })\\     
-%msg_7: (&\mbox{ cache1},&\mbox{readExreq},&\mbox{cpu1           })\\  
-%msg_8: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ Memory })\\  
-%msg_9: (&\mbox{ true                                          })\\  
-%msg_{10}: (&\mbox{ Memory},&\mbox{ readExres},&\mbox{ Bus        })\\  
-%msg_{11}: (&\mbox{ cache2},&\mbox{ readExres},&\mbox{ Bus })\\  
-%msg_{12}: (&\mbox{ Bus},&\mbox{ readExres},&\mbox{ cache1        })\\  
-%msg_{13}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1         })\\  
-%msg_{14}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1 })\\  
-%msg_{15}: (&\mbox{ cache1},&\mbox{UpgradeReq             })\\  
-%msg_{16}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ cahce2      })\\   
-%msg_{17}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ Memory })\\  
-%msg_{18}: (&\mbox{ cache2},&\mbox{UpgradeRes},&\mbox{ Bus     })\\  
-%msg_{19}: (&\mbox{ Bus},&\mbox{ UpgradeRes},&\mbox{ cache1      })\\  
-%msg_{20}: (&\mbox{ cache1},&\mbox{ WriteRes},&\mbox{ CPU1 })\\  
-%\end{array}
-%\]}
-%\caption{Flow specification ($F_1$) of a cache coherent write operation initiated from CPU1}
-%\label{write-flow}
-%\end{figure}
-%
-%The LPN as shown in Fig.~\ref{write-flow} specifies a system flow where CPU1 initiates a memory write operation.
-In this flow, CPU1 initiates a memory write request to L1 Cache.
-Next, depends on whether the required data is included in cache1, the cache1 will generate three possible responses.
-First, the cache1 will send read exclusive request message msg2 to interconnect if data is not present in cache.; Or if the data is shared by CPU0, it will send upgrade message msg16 to interconnect to disable CPU0's ownership of this block of data, else if CPU1 has exclusive right of required data, cache1 will perform the write operation and sent an response message msg15 to CPU1.
-Afterwards, interconnect will sent request to all connected component (data cache and instruction cache of each CPU and memory).
-Once the interconnect obtains the response from either CPU0 or memory, it will generate a response message to cache1, then cache1 will generate a write response message msg14 (or msg 21 ) to CPU1.
-The flow is symmetric for CPU2.
-%
-%
-%\begin{figure} 
-%\centerline{
-%\includegraphics[width=4in]{figures/Fih6.png}}
-%
-%{\footnotesize
-%\[
-%\begin{array}{llll}
-%t0: (&\mbox{ CPU1},&\mbox{ReadReq},&\mbox{cache1  })\\                   
-%t1: (&\mbox{ cache1},&\mbox{ StoreCondreq },&\mbox{Bus })\\           
-%t2: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2 })\\
-%t3: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2       })\\      
-%t4: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2           })\\ 
-%t5: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2 })\\
-%t6: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce1     })\\       
-%t7: (&\mbox{ cache1},&\mbox{StoreCondreq},&\mbox{cpu1           })\\ 
-%t8: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ Memory })\\
-%t9: (&\mbox{ true                                        })\\
-%t10: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus            })\\    
-%t11: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus })\\
-%t12: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1      })\\            
-%t13: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1          })\\  
-%t14: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
-%t21: (&\mbox{ cache1},&\mbox{ LoadLockedreq },&\mbox{Bus })\\     
-%t22: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\
-%t23: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2 })\\
-%t24: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\ 
-%t25: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2     })\\
-%t26: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce1 })\\
-%t27: (&\mbox{ cache1},&\mbox{LoadLockedreq},&\mbox{cpu1       })\\
-%t28: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ Memory     })\\
-%t29: (&\mbox{ true })\\
-%t30: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus       })\\     
-%t31: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus    })\\
-%t32: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1 })\\
-%t33: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
-%\end{array}
-%\]}
-%\caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1}
-%\label{read-flow}
-%
-%\end{figure}
-%The LPN specification as shown in Fig.~\ref{read-flow} captures the system flow where CPU1 initiate a memory read operation.
-Similar to memory write, CPU1 will first initiate a memory read request message msg1 to cache1.
-Cache1 can generate three possible responses.
-First, if the requested data is not in cache1, cache1 will generate an storeCondReq message msg2 to interconnect asking for data.
-Second, if the requested data is inside of cache1 but it's shared, cache1 will generate a upgrade request message msg22 to make sure it has the newest data; Last,  if cache1 has and data and it's not shared, it will generate a read response message msg15 to CPU1.
-Afterwards, interconnect will send corresponding request to all of its connected components.
-When interconnect receives the response, it will generate response message to cache1.
-Then cache1 will generate the read response message msg14 or msg34 to CPU1.
-This specification is also symmetric for CPU2.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Results and Discussions}
-
-Two simple programs are written, one for each CPU.
-These simple programs read numbers from a file, perform some operations on these numbers, and store the results back to the file.
-How GEM5 supports shared memory multi-threaded program execution is unclear.
-Therefore, there is no data shared in both caches in this test.
-Furthermore, GEM5 does not support true concurrency.
-When there are two programs running on the CPUs, GEM5 alternates the executions between the two CPUs.
-To simulate asynchronous concurrency with the interleaving semantics, those two simple programs are instrumented with pseudo-blocking commands, one placed before each statement.
-A pseudo blocking command includes a random number generator that returns either $0$ or $1$ and a loop that only exits when the returned random number is $0$.
-
-%We produce a result file including all of the communication messages between every components from two simple programs running simultaneously on each CPU.
-The program assigned to CPU1 read one file three times for one letter and writing one letter for three times to the same file.
-CPU2's program will do the same read and write functionalities to the same file, only difference is that CPU2 will write first and then read.
-One thing we should know about GEM5 is that even when we run these two programs concurrently, it will attempt to produce a concurrent result.
-The data we collected shows that it's not the real nondeterministic concurrency.
-What GEM5 did was it allowed two CPUs to execute its own instruction in turn, therefore the order was deterministic.
-Therefore, no matter how many times we ran the program, it produced the same result.
-We tried other virtual SoC platform softwares, and this was the best nondeterministic concurrency we can get.
-Even if this is slightly off with what the real chip should work, it still surve our purpose of testing the correctness of our algorithm.
-
-After the SoC model finishes executing the program, there are totally $343581$ messages collected in the trace file.
-Not all of the messages are relevant to the flow specification as many are used by GEM5 to initialize its simulation environment.
-After removing those irrelevant messages, the number of messages in the trace file is to reduced to $121138$.
-
-The time taken to remove the irrelevant messages from the trace is negligible.
-The total runtime and the peak memory usage for the trace analysis algorithm to finish the reduced trace are 3 seconds and 12MB, respectively.
-From the trace, Table~\ref{table-case-2} shows the number of instances extracted for the six flows describing cache coherent read/write operations initiated from both CPUs.
-\begin{table}[tb]
-\caption{The number of flow instances derived by the trace analysis with the full observability.}
-\begin{center}
-\begin{tabular}{|l|c|}
-\hline
-Flows & $\#$Instances \\
-\hline
-\hline
-CPU1 Data Read			&  $17582$\\
-CPU1 Instruction Read		&  $4002$\\
-CPU1 Write				&  $3370$\\
-\hline
-CPU2 Data Read			&  $17386$\\
-CPU2 Instruction Read		&  $3955$\\
-CPU2 Write				&  $3308$\\
-\hline
-\end{tabular}
-\end{center}
-\label{table-case-2}
-\end{table}%
-
-To take the partial observability into account, the four monitors attached to the links between two CPUs and their caches are disabled.
-Then, the trace is generated by the remaining five monitors from the SoC model executing the same program.
-After removing all non-observable messages, the trace only contains 15089 messages.
-The numbers of the flow instances extracted by the trace analysis method are shown in Table~\ref{table-par-obs}.
-From these results, the numbers of the flow instances are dropped significantly compared to the results extracted from the trace with the full observability as shown in Table~\ref{table-case-2}.
-This difference is due to that some communications occurred in the system when executing the program involve the CPUs and their corresponding caches only, and the traffic on the links between the CPUs and their corresponding caches is not observable.
-Therefore, the instances of the flow specifications characterizing these communications do not exist in the trace.
-In other words, all extracted flow instances in Table~\ref{table-par-obs} characterize the communications that pass through the memory bus in the system model.
-The runtime and memory usage is similar to that for analyzing the trace of the full observability as shown above.
-
-\begin{table}[tb]
-\caption{The number of flow instances derived by the trace analysis with certain monitors disabled.}
-\begin{center}
-\begin{tabular}{|l|c|}
-\hline
-Flows & $\#$Instances \\
-\hline
-\hline
-CPU1 Data Read			&  $829$\\
-CPU1 Instruction Read		&  $169$\\
-CPU1 Write				&  $82$\\
-\hline
-CPU2 Data Read			&  $803$\\
-CPU2 Instruction Read		&  $190$\\
-CPU2 Write				&  $83$\\
-\hline
-\end{tabular}
-\end{center}
-\label{table-par-obs}
-\end{table}%
-
-In the third experiment, further partial observability is taken into consideration.
-In this experiment, only the five links involving the memory bus are still considered.
-However, an assumption is made that all messages passing the same link are not distinguishable due to the limitation of the observability.
-More specifically, the monitors are modified such that whenever a message is captured on one of the links, it dumps a set of messages passing through the same link into the trace file.
-Therefore, each line of the trace file corresponds to a set of messages.
-After applying the trace analysis to this trace,  a total of 13944 flow execution scenarios are extracted.
-This large number, compared to the number of extracted execution scenarios shown in the Table~\ref{table-case-2} and \ref{table-par-obs}, is due to the ambiguous interpretation of the messages with limited observability.
-
-The whole experiment takes about $15$ minutes and $420$~MB to finish, significantly higher than the numbers for analyzing traces where there is no ambiguity in the observed messages.
-This is due to the fact that a trace of ambiguous events is in fact a set of traces of original messages, which lead to large numbers of execution scenarios either during or at the end of the analysis.
-In this experiment, the peak number of executions during the analysis process is $70384$.
-Compared to the final number, many of the intermediate scenarios are invalid, and removed eventually.
-However, controlling the number of intermediate scenarios during the trace analysis is critical in order for the analysis to be tractable.
-Here, insights from validators can help.
-
-%=====
-\section{Related Work}
-
-Our work is closely related to communication-centric and transaction based debug.
-An early pioneering work is described in~\cite{Goossens2007NOCS}, which advocates the focus on observing activities on the interconnect network among IP blocks, and mapping these activities to transactions for better correlation between computations and communications.
-Therefore, the communication transactions, as a result of software execution, provide an interface between computation and communication, and facilitate  system-level debug.
-This work is extended in~\cite{Vermeulen2009VLSI-DAT,Goossens2009DATE}.
-However, this line of work is focused on the network-on-chip (NoC) architecture for interconnect using the run/stop debug control method.
-
-A similar transaction-based debug approach is presented in~\cite{Gharehbaghi2012ISQED}.
-Furthermore, it proposes an automated extraction of state machines at transaction level from  high level design models.
-From an observed failure trace, it performs backtracking on this transaction level state machine to derive a set of transaction traces that lead to the observed failure state.
-In the subsequent step, bounded model checking with the constraints on the internal variables is used to refine the set of transaction traces to remove the infeasible traces.
-This approach requires user inputs to identify impossible transaction sequences, and may not find the states causing the failure if the transaction traces leading to the observed failure state is long.
-Backtracking from the observed failure state requires pre-image computation, which can be computationally expensive.
-A transaction-based online debug approach is proposed in~\cite{Dehbash2014} to address these issues.
-This approach utilizes a transaction debug pattern specification language~\cite{Gharehbaghi2009ICCD} to define properties that transactions should meet.
-These transaction properties are checked at runtime by programming debug units in the on-chip debug infrastructure, and the system can be stopped shortly after a violation is detected for any one of those properties.
-In this sense, it can be viewed as the hardware assertion approaches in~\cite{Boule2007ISQED} elevated to the transaction level.
-
-In~\cite{Singerman2011DAC}, a coherent workflow is described where the result from the pre-silicon validation stage can be carried over to the post-silicon stage to improve efficiency and productivity of post-silicon debug.
-This workflow is centered on a repository of system events and simple transactions defined by architects and IP designers.
-It spans across a wide spectrum of the post-silicon validation including DFx instrumentation, test generation, coverage, and debug.
-The DFx instruments are automatically inserted into the design RTL code driven by the defined transactions.
-This instrumentation is optimized for making a large set of events and transactions observable.
-Test generation is also optimized to generate only the necessary but sufficient tests to allow all defined transactions to be exercised.
-Moreover, coverage for post-silicon validation is now defined at the abstract level of events and transactions rather than the raw signals, and thus can be evaluated more efficiently.
-In~\cite{Abarbanel2014DAC}, a model at an even higher-level of abstraction, {\em flows}, is proposed.
-Flows are used to specify more sophisticated cross-IP transactions such as power management, security, etc, and to facilitate reuse of the efforts of the architectural analysis to check HW/SW implementations.
-
-%
-%--
-\section{Conclusion}
-
-This paper presents a trace analysis based method for post-silicon validation by interpreting observed raw signal traces at the level of system flow specifications.
-The derived flow execution scenarios provide more structured information on system operations, which is more understandable to system validators.
-This information can help to locate design defects more easily, and also provides a measurement of validation coverage.
-
-Due to partial observability, this approach may derive a large number of different flow execution scenarios for a given signal trace.
-Insights from system validators can help to eliminate some false scenarios due to the partial observability.
-An interesting future direction is formalization of the validators' insights using temporal logic on flows so that the validators can express their intents more precisely and concisely.
-
-The trace analysis approach presented in this paper needs to be iterated with different observations selected in different iterations in order to eliminate the false scenarios and to root cause system failures as quickly as possible.
-The observation selection and stitching signal traces of different observations together for the above goal will also be pursued in the future.
-
-% conference papers do not normally have an appendix
-
-% use section* for acknowledgement
-%\section*{Acknowledgment}
-%
-%
-%The authors would like to thank...
-%
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{SoC}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-%\begin{thebibliography}{1}
-%
-%\bibitem{IEEEhowto:kopka}
-%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-%
-%\end{thebibliography}
-
-% that's all folks
-\end{document}

### 2015-09-28 paper.tex
added one more experiment into case study
Section III, updated algorithm 1, added a small example to illustrate
the trace analysis.

section V: results from one more experiment are added.

Case study I is removed to save space.

@@
 A key component of post-silicon validation of SoC designs is
-to correlate trace from silicon execution with the intended
+to correlate traces from silicon execution with the intended
 system-level transactions.  An SoC design is typically
@@
 interleaved instances of these protocols.  For example,
-consider a smartphone executing a usage scenerio where the
+consider a smartphone executing a usage scenario where the
 end-user browses the Web while listening to music and
 sending and receiving occasional text messages.  Typical
-post-silicon validation use-case involve exercising such
+post-silicon validation use-cases involve exercising such
 scenarios.  An execution trace would involve activities from
@@
 several communication protocols.  On the other hand, due to
-observability limitations, only a small small number of
+observability limitations, only a small number of
 participating signals can be actually traced during silicon
@@
 system-level transactions from traces with missing, noisy,
-and ambiguous signal values, together with an estimate of
-confidence on the inference.  We demonstrate our approach on
+and ambiguous signal values.
+% together with an estimate of confidence on the inference.  
+We demonstrate our approach on
 two SoC case studies: a pedagogical SystemC prototype and a
@@
 	\node[rectbox]	(t5)	at	($(p4) + (3,-1.1)$) {$t_5:\langle {\tt Device:CE:Ack} \rangle$};
-	\node[circ,fill=black] 	(term1) 		at ($(t4) + (0,-1.1)$) {$~~~$};
-	\node[circ,fill=black] 	(term2) 		at ($(t5) + (0,-1.1)$) {$~~~$};
+	\node[circ,fill=black] 	(term1) 		at ($(t4) + (0,-1.1)$) {\textcolor{white}{$p_5$}};
+	\node[circ,fill=black] 	(term2) 		at ($(t5) + (0,-1.1)$) {\textcolor{white}{$p_6$}};
 
@@
 due to limited observability, it may not be possible to map
-a given set of (observed) hardware signals to uniquely to a
-flow event.  Finally, the trace may include several
+a given set of (observed) hardware signals uniquely to a
+flow event.  Finally, the trace may a result from several
 instances of the same protocol executing concurrently, \eg,
@@
 behavior.  The trace may identify a problem in the protocols
-themselves, \eg an interleaving of some protocol executions
+themselves, \eg~an interleaving of some protocol executions
 may lead to an unexpected message being sent or cause the
@@
 of system flows is provided as a collection ${\vec{F}}$ of
-LPNs.  We define a {\em flow execution scenario} to be the
-set of pairs $(F_{i,j}, s_{i,j})$ where $F_{i,j}$ is the
+LPNs.  We define a {\em flow execution scenario} to be a
+set $\{(F_{i,j}, s_{i,j})\}$ where $F_{i,j}$ is the
 $j$th instance of flow $F_i \in {\vec{F}}$, and $s_{i,j}$ is
@@
 $s_{i,j}$.  Formally, $\mathit{accept(F_{i,j}, s_{i,j}, e)}$
-returns $(F_{i,j}, s'_{i,j})$ if there exists a transition
+returns $(F_{i,j}, s'_{i,j})$ where $s'_{i,j} = (s_{i,j} - \bullet t)
+\cup t\bullet$ if there exists a transition
 $t$ in $F_i$ such that $L(t) = e$ and $\bullet t \subseteq
-s_{i,j}$.  In this case, $s'_{i,j} = (s_{i,j} - \bullet t)
-\cup t\bullet$.  It returns $\emptyset$ if no such $t$
+s_{i,j}$.  It returns $\emptyset$ if no such $t$
 exists in $F_i$.
@@
 Given an obsered trace $\rho$ and the set $\vec{F}$ of LPNs,
-Algorithm~\ref{algo:compliance} provides the basic steps for
+Algorithm~\ref{algo:compliance} provides a basic procedure for
 computing a set of compliant flow execution scenarios.
@@
 \DontPrintSemicolon
-$\mathit{Scen} \gets \emptyset$\;
+Create an empty scenario $scen$\;
+$\mathit{Scen} = \{scen\}$\;
 \ForEach {$h, \; 1 \leq h \leq n $} {
 	$\mathit{found} \gets {\tt true}$ \;
-  \ForEach {$(F_{i,j}, s_{i,j}) \in \mathit{scen}$} {
+	$Scen' = \emptyset$\;
+	\ForEach {$scen \in Scen$} {
+  		\ForEach {$(F_{i,j}, s_{i,j}) \in \mathit{scen_1}$} {
     		\If {$\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = (F_{i,j}, s'_{i,j})$} {
-      $\mathit{scen} \gets (\mathit{scen} - (F_{i,j}, s_{i,j})) \cup (F_{i,j}, s'_{i,j})$\;
-      $\mathit{scen} \gets \mathit{scen} \cup \mathit{Scen}'$\;
+				Let $scen'$ be a copy of $scen$\;
+      		$\mathit{scen'} \gets (\mathit{scen'} - (F_{i,j}, s_{i,j})) \cup (F_{i,j}, s'_{i,j})$\;
+      		$\mathit{Scen'} \gets \mathit{scen'} \cup \mathit{Scen'}$\;
       		$\mathit{found} \gets {\tt false}$ \;
@@
       	\If {$\mathit{accept}(F_{i,j+1},\mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$} {
-    $\mathit{scen} \gets \mathit{scen} \cup (F_{i,j+1}, s'_{i,j+1})$ \;
+    			Let $scen'$ be a copy of $scen$\;
+				$\mathit{scen'} \gets \mathit{scen'} \cup (F_{i,j+1}, s'_{i,j+1})$ \;
+				$\mathit{Scen'} \gets \mathit{scen'} \cup \mathit{Scen'}$\;
         		$\mathit{found} \gets {\tt false}$ \;
+      	}
     	}
@@
   }
+  $Scen = Scen'$\;
 }
@@
 The algorithm operates by keeping track (in variable
-$\mathit{scen}$) of a set of candidate protocol instances
-compliant with each prefix of $\rho$.  At each iteration, we
-update this set by either extending some member of
-$\mathit{scen}$ or initiating a new protocol instance.  If
-neither is possible then we report that the trace is {\tt
-  inconsistent}, \ie, there is no possible interleaving of
+$\mathit{Scen}$) of a set of candidate flow execution scenarios
+compliant with each prefix of $\rho$.  At each iteration,
+for each event $e_h$ in the observed trace, we
+update $Scen$ by either extending a member of
+$\mathit{scen}$ or initiating a new protocol instance for 
+each $scen \in Scen$ with respect to $e_h$ in every possible way.
+If $e_h$ cannot be emitted by any existing or new flow instances, 
+then we report that the trace is {\tt  inconsistent}, \ie, 
+there is no possible interleaving of
 the protocol instances from $\vec{F}$ that is compliant with
@@
 
+To illustrate the basic idea of the trace analysis method, 
+consider the system flow shown in Figure~Fig.~\ref{flow-spec-ex}(b).
+Let $F_1$ denote such flow.  Suppose that a hardware system 
+implements flow $F_1$, and the following flow trace is abstracted 
+from an observed signal trace as a result of executing such system.
+\[
+	t_1\;t_2\;t_1\;t_2\;t_3\;t_3\;t_4\;t_5\;t_5\;t_4\ldots
+\]  
+where transition names in the LPN are used to represent the 
+flow events in the trace.  The first four events results in
+the following flow execution scenario
+\[
+	\{(F_{1,1}, \{p_3\}),~(F_{1,2}, \{p_3\})\}.
+\]
+For the first event $t_3$, it results in two execution scenarios 
+below depending on which flow instance emits $t_3$.
+\[
+\begin{array}{l}
+	\{(F_{1,1}, \{p_4\}),~(F_{1,2}, \{p_3\})\} \\
+	\{(F_{1,1}, \{p_3\}),~(F_{1,2}, \{p_4\})\}.
+\end{array}
+\]
+After handing the next event $t_3$, the above two execution scenarios
+are reduced to the one as shown below.
+\[
+	\{(F_{1,1}, \{p_4\}),~(F_{1,2}, \{p_4\})\}.
+\]
+By following Algorithm~\ref{algo:compliance} to handle the remaining 
+four events, the following execution scenario can be derived.
+\[
+	\{(F_{1,1}, \{p_5, p_6\}),~(F_{1,2}, \{p_5, p_6\})\}
+\]
 %% Given a trace of flow events $\rho = e_1e_2\ldots e_n$, the
@@
 
-%=====
+%%-- Case study I
+%\input{case1}
+
+%
+%--
+%
 \section{Case Study}
-
-The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a simple SoC design with two CPU cores, memory, and a peripheral device connected by an interconnect as shown in Fig.~\ref{fig-SoC-ex}.
-Since the analysis method presented in this paper is communication centric, the detailed computations of these blocks are not modeled.
-Instead, the modeling is focused on how they participate in flows for different system level use cases.
-
-\begin{figure}[tb]
-\begin{center}
-\includegraphics[width=2in]{figures/simple-ex}
-\caption{A simple SoC example.}
-\label{fig-SoC-ex}
-\end{center}
-\end{figure}
-
-\subsection{Flow Specification}
-
-In this case study, four system flows are implemented in the simple SoC model.
-They include cache coherent memory access operations and a memory-mapped peripheral read operation initiated from the CPUs, a message signaled interrupt operation initiated from the peripheral device.
-These flow specifications capture how messages are exchanged for different use cases.
-In this model, a message is defined with the following format, $(\mbox{Src}, \mbox{Dest}, \mathit{Cmd}, \mathit{Addr})$, where Src and Dest refer to the source and destination components of messages, $\mathit{Cmd}$ refers to the operations that the destination component should perform, and $\mathit{Addr}$ refers to memory addresses where $\mathit{Cmd}$ applies.
-$\mathit{Cmd}$ can be memory accesses or not.
-If it is not, then the $\mathit{Addr}$ field of messages is ignored.
-In this case study, memory mapped IO mechanism is used.
-Furthermore, detailed memory addresses are not modeled.
-Instead, the address space is partitioned to main memory addresses and peripheral addresses.
-In messages, the $\mathit{Addr}$ field is replaced with either $M$ representing a memory address or $P$ representing an address to a peripheral device.
-
-The LPN as shown in Fig.~\ref{fig-cpu-write} specifies a system flow where {CPU2} initiates a memory write operation.
-In this flow, CPU2 initiates a memory write request followed by a data valid message to the interconnect.
-The data valid messages are used to model availability or validity of data for transfer.
-Concurrently, the interconnect inquiries CPU1 if it holds a more updated version of the data by sending a memory write message $\mathit{msg_2}$.
-CPU1 generates one of two possible responses.
-If CPU1 holds the more updated data in its cache in the same memory space that CPU2 intends to write, a cache hit message $\mathit{msg_4}$ followed by $\mathit{msg_6}$ are sent to the interconnect.
-Otherwise, CPU1 sends a cache miss message $\mathit{msg_5}$.
-After getting the response from CPU1, Interconnect sends a write request to the memory unit.
-This flow is symmetric for CPU1.
-
-\begin{figure}[tb]
-\begin{center}
-\resizebox{2.8in}{!}{
-\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
-\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
-\tikzset{circ/.style={circle, draw,thick}}
-\tikzset{terminal/.style={circle, draw,line width=.8mm}}
-\tikzset{weight3/.style={line width=.3mm}}
-	% Nodes:
-	\node[circ]		(p1) 		at (0,0) {$p_1$};
-	\node[rectbox]	(msg1)	at	($(p1) + (0,-1.1)$) {$msg_1$};
-	\node[circ] 	(p2) 		at ($(msg1) + (-3,0)$) {$p_2$};
-	\node[circ] 	(p3) 		at ($(msg1) + (3,0)$) {$p_3$};
-	\node[rectbox]	(msg2)	at	($(p2) + (0,-1.1)$) {$msg_2$};
-	\node[rectbox]	(msg3)	at	($(p3) + (0,-1.5)$) {$msg_3$};
-	\node[circ] 	(p4) 		at ($(msg2) + (0,-1.1)$) {$p_4$};
-	\node[rectbox]	(msg4)	at	($(p4) + (0,-1.1)$) {$msg_4$};
-	\node[rectbox]	(msg5)	at	($(p4) + (3,0)$) {$msg_5$};
-	\node[circ] 	(p6) 		at ($(msg4) + (0,-1.1)$) {$p_6$};
-	\node[circ] 	(p5) 		at ($(msg3) + (0,-1.4)$) {$p_5$};
-	\node[rectbox,rotate=90]	(msg6)	at	($(p6) + (1.5,0)$) {$msg_6$};
-	\node[circ] 	(p7) 		at ($(msg6) + (1.5,0)$) {$p_7$};
-	\node[rectbox]	(msg7)	at	($(p5) + (0,-1.5)$) {$msg_7$};
-	\node[terminal]	(p8)	at	($(msg7) + (2,0)$) {$p_8$};
-	% Edges
-	\draw[->,weight3]	(p1) -- (msg1);
-	\draw[->,weight3]	(msg1) -- (p2);
-	\draw[->,weight3]	(msg1) -- (p3);
-	\draw[->,weight3]	(p2) -- (msg2);
-	\draw[->,weight3]	(msg2) -- (p4);
-	\draw[->,weight3]	(p4) -- (msg4);
-	\draw[->,weight3]	(p4) -- (msg5);
-	\draw[->,weight3]	(msg4) -- (p6);
-	\draw[->,weight3]	(p6) -- (msg6);
-	\draw[->,weight3]	(msg6) -- (p7);
-	\draw[->,weight3]	(p7) -- (msg7);
-	\draw[->,weight3]	(msg7) -- (p8);
-	\draw[->,weight3]	(msg5) -- (p7);
-	\draw[->,weight3]	(p3) -- (msg3);
-	\draw[->,weight3]	(msg3) -- (p5);
-	\draw[->,weight3]	(p5) -- (msg7);
-
-	\node[]	(msg-def)	at	($(p1) + (0,-8.5)$) {
-	\begin{minipage}{3in}
-	Definition of the messages:
-\[
-\begin{array}{clllll}
-msg_1: & (\mbox{CPU2},	& \mbox{Icon},		& \mathit{Wr}, 	& M) \\
-msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& \mathit{Wr},			& M) \\
-msg_3: & (\mbox{CPU2},	& \mbox{Icon}, 	& \mathit{DVal},	& -) \\
-msg_4: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Hit},	& -) \\
-msg_5: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Miss},	& -) \\
-msg_6: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{DVal},	& -) \\
-msg_7: & (\mbox{Icon},	& \mbox{Mem}, 		& \mathit{Wr},			& M) \\
-\end{array}
-\]
-\end{minipage}
-		};
-\end{tikzpicture}
-}
-\caption{Flow specification ($F_1$) of a cache coherent write operation initiated from \texttt{CPU2}.}
-\label{fig-cpu-write}
-\end{center}
-\end{figure}
-
-The LPN specification as shown in Fig.~\ref{fig-cpu-read} captures the system flow where CPU2 initiates a memory read operation.
-Basically, CPU2 sends a memory read message to Interconnect, which then generates two concurrent threads, one checks if CPU1 has the more updated data in the memory space for the read operation, and the other thread to get data from the memory.
-Once Interconnect gets both responses from CPU1 and memory, it synchronizes the responses, and writes the correct data to the CPU2's cache and memory in parallel.
-Again, this specification is symmetric for CPU1.
-
-\begin{figure}[htbp]
-\begin{center}
-\resizebox{3.2in}{!}{
-\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
-\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
-\tikzset{circ/.style={circle, draw,thick}}
-\tikzset{terminal/.style={circle, draw,line width=.8mm}}
-\tikzset{weight3/.style={line width=.3mm}}
-	% Nodes:
-	\node[circ]		(p1) 		at (0,0) {$p_1$};
-	\node[rectbox]	(msg1)	at	($(p1) + (0,-1.1)$) {$msg_1$};
-	\node[circ] 	(p2) 		at ($(msg1) + (-3,0)$) {$p_2$};
-	\node[circ] 	(p3) 		at ($(msg1) + (2,0)$) {$p_3$};
-	\node[rectbox]	(msg2)	at	($(p2) + (0,-1.1)$) {$msg_2$};
-	\node[rectbox]	(msg3)	at	($(p3) + (2,0)$) {$msg_3$};
-	\node[circ] 	(p4) 		at ($(msg2) + (0,-1.1)$) {$p_4$};
-	\node[rectbox]	(msg4)	at	($(p4) + (0,-1.1)$) {$msg_4$};
-	\node[rectbox]	(msg5)	at	($(p4) + (3,0)$) {$msg_5$};
-	\node[circ] 	(p6) 		at ($(msg4) + (0,-1.1)$) {$p_6$};
-	\node[circ] 	(p5) 		at ($(msg3) + (0,-1.1)$) {$p_5$};
-	\node[rectbox,rotate=90]	(msg6)	at	($(p6) + (1.5,0)$) {$msg_6$};
-	\node[circ] 	(p7) 		at ($(msg6) + (1.5,0)$) {$p_7$};
-	\node[rectbox]	(msg7)	at	($(p5) + (0,-1.1)$) {$msg_7$};
-	\node[circ]	(p8)	at	($(msg7) + (0,-1.1)$) {$p_8$};
-	\node[rectbox]	(msg8)	at	($(p8) + (0,-1.1)$) {$msg_8$};
-	\node[circ]	(p9)	at	($(msg8) + (-1.5,-1.1)$) {$p_9$};
-	\node[circ]	(p10)	at	($(msg8) + (1.5,-1.1)$) {$p_{10}$};
-	\node[rectbox]	(msg9)	at	($(p9) + (0,-1.1)$) {$msg_9$};
-	\node[rectbox]	(msg10)	at	($(p10) + (0,-1.1)$) {$msg_{10}$};
-	\node[terminal]	(p11)	at	($(msg9) + (0,-1.1)$) {$p_{11}$};
-	\node[terminal]	(p12)	at	($(msg10) + (0,-1.1)$) {$p_{12}$};
-
-	% Edges
-	\draw[->,weight3]	(p1) -- (msg1);
-	\draw[->,weight3]	(msg1) -- (p2);
-	\draw[->,weight3]	(msg1) -- (p3);
-	\draw[->,weight3]	(p2) -- (msg2);
-	\draw[->,weight3]	(msg2) -- (p4);
-	\draw[->,weight3]	(p4) -- (msg4);
-	\draw[->,weight3]	(p4) -- (msg5);
-	\draw[->,weight3]	(msg4) -- (p6);
-	\draw[->,weight3]	(p6) -- (msg6);
-	\draw[->,weight3]	(msg6) -- (p7);
-	\draw[->,weight3]	(p7) -- (msg8);
-	\draw[->,weight3]	(msg7) -- (p8);
-	\draw[->,weight3]	(msg5) -- (p7);
-	\draw[->,weight3]	(p3) -- (msg3);
-	\draw[->,weight3]	(msg3) -- (p5);
-	\draw[->,weight3]	(p5) -- (msg7);
-	\draw[->,weight3]	(p8) -- (msg8);
-	\draw[->,weight3]	(msg8) -- (p9);
-	\draw[->,weight3]	(msg8) -- (p10);
-	\draw[->,weight3]	(p9)  -- (msg9);
-	\draw[->,weight3]	(p10)  -- (msg10);
-	\draw[->,weight3]	(msg9) -- (p11);
-	\draw[->,weight3]	(msg10) -- (p12);
-
-	\node[]	(msg-def)	at	($(p1) + (-2.7,-10)$) {
-	\begin{minipage}{2in}
-	Definition of the messages:
-\[
-\begin{array}{clllll}
-msg_1: & (\mbox{CPU2},	& \mbox{Icon},		& \mathit{Rd}, 		& M) \\
-msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& \mathit{Rd},	& M) \\
-msg_3: & (\mbox{Icon},	& \mbox{Mem}, 		& \mathit{Rd},			& M) \\
-msg_4: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Hit},	& -) \\
-msg_5: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Miss},& -) \\
-msg_6: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{DVal},	& -) \\
-msg_7: & (\mbox{Mem},		& \mbox{Icon}, & \mathit{DVal},			& -) \\
-msg_8: & (-,		& -, 		& -,				& -) \\
-msg_9: & (\mbox{Icon},	& \mbox{Mem}, 		& \mathit{Wr},			& M) \\
-msg_{10}: & (\mbox{Icon},	& \mbox{CPU2}, 		& \mathit{Wr},			& M) \\
-\end{array}
-\]
-\end{minipage}
-		};
-
-\end{tikzpicture}
-}
-\caption{Flow specification ($F_2$) of a cache coherent read operation from CPU2.}
-\label{fig-cpu-read}
-\end{center}
-\end{figure}
-
-The LPN specification as shown in Fig.~\ref{fig-peri-read} captures a system flow for CPU initiated memory mapped peripheral read operations.
-When CPU2 tries to read the peripheral device (device hereafter), a read message with address $P$ is sent to Interconnect, which then sends this message to CPU1 and the device simultaneously.
-When CPU1 sees this message, it responds with a cache miss message as the address $P$ points to the device.
-At the meantime, the device responds with a data valid message indicating the availability of the requested data.
-Finally, Interconnect synchronizes the both responses, and sends a data valid message back to CPU2.
-This specification is also symmetric for CPU1.
-
-\begin{figure}[tb]
-\begin{center}
-\resizebox{3.2in}{!}{
-\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
-\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
-\tikzset{circ/.style={circle, draw,thick}}
-\tikzset{terminal/.style={circle, draw,line width=.8mm}}
-\tikzset{weight3/.style={line width=.3mm}}
-	% Nodes:
-	\node[circ]		(p1) 		at (0,0) {$p_1$};
-	\node[rectbox]	(msg1)	at	($(p1) + (0,-1.1)$) {$msg_1$};
-	\node[circ] 	(p2) 		at ($(msg1) + (-1.5,0)$) {$p_2$};
-	\node[circ] 	(p3) 		at ($(msg1) + (1.5,0)$) {$p_3$};
-	\node[rectbox]	(msg2)	at	($(p2) + (-1.5,0)$) {$msg_2$};
-	\node[rectbox]	(msg4)	at	($(p3) + (1.5,0)$) {$msg_4$};
-	\node[circ] 	(p4) 		at ($(msg2) + (-1.5,0)$) {$p_4$};
-	\node[rectbox]	(msg3)	at	($(p4) + (0,-1.2)$) {$msg_3$};
-	\node[circ] 	(p6) 		at ($(msg3) + (0,-1.2)$) {$p_6$};
-	\node[circ] 	(p5) 		at ($(msg4) + (1.5,0)$) {$p_5$};
-	\node[rectbox]	(msg5)	at	($(p5) + (0,-1.2)$) {$msg_5$};
-	\node[circ]	(p7)	at	($(msg5) + (0,-1.2)$) {$p_7$};
-	\node[rectbox]	(msg6)	at	($(p6) + (4.5,0)$) {$msg_6$};
-	\node[terminal]	(p8)	at	($(msg6) + (0,-1.2)$) {$p_8$};
-
-	% Edges
-	\draw[->,weight3]	(p1) -- (msg1);
-	\draw[->,weight3]	(msg1) -- (p2);
-	\draw[->,weight3]	(msg1) -- (p3);
-	\draw[->,weight3]	(p2) -- (msg2);
-	\draw[->,weight3]	(msg2) -- (p4);
-	\draw[->,weight3]	(p4) -- (msg3);
-	\draw[->,weight3]	(msg3) -- (p6);
-	\draw[->,weight3]	(p6) -- (msg6);
-	\draw[->,weight3]	(msg6) -- (p8);
-	\draw[->,weight3]	(p3) -- (msg4);
-	\draw[->,weight3]	(msg4) -- (p5);
-	\draw[->,weight3]	(p5) -- (msg5);
-	\draw[->,weight3]	(msg5) -- (p7);
-	\draw[->,weight3]	(p7) -- (msg6);
-	\draw[->,weight3]	(msg6) -- (p8);
-
-		\node[]	(msg-def)	at	($(p8) + (0,-2.5)$) {
-	\begin{minipage}{3in}
-	Definition of the messages:
-\[
-\begin{array}{clllll}
-msg_1: & (\mbox{CPU2},	& \mbox{Icon},		& Rd, 	& P) \\
-msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& Rd,		& P) \\
-msg_3: & (\mbox{CPU1},	& \mbox{Icon}, 	& Miss,			& -) \\
-msg_4: & (\mbox{Icon},	& \mbox{Dev}, 	& Rd,	& P) \\
-msg_5: & (\mbox{Dev},	& \mbox{Icon}, 	& Dval,	& -) \\
-msg_6: & (\mbox{Icon},	& \mbox{CPU2}, 	& Dval,	& -) \\\end{array}
-\]
-\end{minipage}
-		};
-\end{tikzpicture}
-}
-\caption{Flow specification ($F_3$) of a read access to peripheral device from CPU2.}
-\label{fig-peri-read}
-\end{center}
-\end{figure}
-
-The last LPN specification as shown in Fig.~\ref{fig-msi} captures how interrupts from the device are handled.
-In this case study, all interrupts are directed to CPU1.
-When the device triggers an interrupts, it sends a message with $\mathit{Intr}$ in the command field.
-Then, Interconnect notifies CPU1 by sending a message with $\mathit{MSI}$ and $I$ in the command and address fields, respectively, where $I$ is the symbol referring to the entry points to interrupt service routines.
-CPU1 responds with a cache miss message as the receipt of the interrupt.
-
-\begin{figure}[htbp]
-\begin{center}
-\resizebox{3.5in}{!}{
-\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
-\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
-\tikzset{circ/.style={circle, draw,thick}}
-\tikzset{terminal/.style={circle, draw,line width=.8mm}}
-\tikzset{weight3/.style={line width=.3mm}}
-	% Nodes:
-	\node[circ]		(p1) 		at (0,0) {$p_1$};
-	\node[rectbox]	(msg1)	at	($(p1) + (1.5, 0)$) {$msg_1$};
-	\node[circ] 	(p2) 		at ($(msg1) + (1.5, 0)$) {$p_2$};
-	\node[rectbox]	(msg2)	at	($(p2) + (1.5,0)$) {$msg_2$};
-	\node[circ] 	(p3) 		at ($(msg2) + (1.5, 0)$) {$p_3$};
-	\node[rectbox]	(msg3)	at	($(p3) + (1.5, 0)$) {$msg_3$};
-	\node[terminal] 	(p4) 		at ($(msg3) + (1.5,0)$) {$p_4$};
-
-	% Edges
-	\draw[->,weight3]	(p1) -- (msg1);
-	\draw[->,weight3]	(msg1) -- (p2);
-	\draw[->,weight3]	(p2) -- (msg2);
-	\draw[->,weight3]	(msg2) -- (p3);
-	\draw[->,weight3]	(p3) -- (msg3);
-	\draw[->,weight3]	(msg3) -- (p4);
-
-\node[]	(msg-def)	at	($(msg2) + (0,-2)$) {
-	\begin{minipage}{3in}
-	Definition of the messages:
-\[
-\begin{array}{clllll}
-msg_1: & (\mbox{Dev},	& \mbox{Icon},		& \mathit{Intr}, 	& -) \\
-msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& \mathit{MSI},	& I) \\
-msg_3: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Miss},	& -) \\
-\end{array}
-\]
-\end{minipage}
-		};
-\end{tikzpicture}
-}
-\caption{Flow specification ($F_4$) of handling interrupts from the peripheral device.}
-\label{fig-msi}
-\end{center}
-\end{figure}
-
-%
-%
-\subsection{Results and Discussions}
-%
-%
-The transaction level model of the simple system implementing the four flow specifications shown in the previous section is described in SystemC.
-Each component is described as a SystemC module, which may include a number of threads to model concurrency.
-The entire model is concurrent and operates in asynchronous mode.
-
-When testing this model, both CPUs and the peripheral device are set up as flow instance generators.
-They randomly generate the first message in a corresponding flow specification to start a flow instance, and react to incoming messages by generating new messages as defined in the flow specification.
-In the model, monitors are embedded to observe the messages generated by each component.
-When a message is observed, it is written to an output trace file for analysis.
-
-Even though this example is conceptually simple, getting the model to correctly implement the flow specifications is not straightforward.
-On the other hand, results from the trace analysis greatly help the debugging process by providing information to locate problems quickly.
-For example, in an early version of the model, the trace shown in Table~\ref{table-trace} is observed.
-\begin{table*}
-\caption{An observed trace of messages for trace analysis.}
-\label{table-trace}
-\begin{center}
-\begin{tabular}{llll}
-1~~(CPU2, Icon, Rd, M) & 2~~(Icon, CPU1, Rd, M) & 3~~(Icon, Mem, Rd, M) & 4~~(Mem,  Icon, DVal, -) \\
-5~~(CPU1, Icon, Hit, -) & 6~~(CPU1, Icon, DVal, -) & 7~~(Icon, CPU2, Wr, M) & 8~~(Icon, Mem, Wr, M) \\
-9~~(CPU2, Icon, Rd, M) & 10~(Icon, CPU1, Rd, M) & 11~(Icon, Mem, Rd, M) & 12~(CPU2, Icon, Rd, P) \\
-13~(Icon, CPU1, Rd, P) & 14~(CPU1, Icon, Miss, -) & 15~(Icon, Dev, Rd, P) & 16~(CPU1, Icon, DVal, -) \\
-\end{tabular}
-\end{center}
-\end{table*}
-The trace analysis finds out that messages~$1-8$ in the trace are the results of execution of an instance of flow $F_2$ as shown, and the flow execution scenario is $\{(F_{2,1}, \{p_{11}, p_{12}\})\}$.
-Messages~$9-11$ are analyzed as the results of executing another instance of flow $F_2$, message~$12-13$ as the results of executing an instance of flow $F_3$ as shown in Fig.~\ref{fig-peri-read}.
-The flow execution scenario after the first thirteen messages in the trace is
-\[
-\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_4, p_5\}), (F_{3,1}, \{p_4, p_3\})\}.
-\]  
-Message~$14$ can be  the result from executing $F_{2,2}$ or $F_{3,1}$, therefore it leads to two following flow execution scenarios:
-\begin{equation}
-\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_7, p_5\}), (F_{3,1}, \{p_4, p_3\})\}
-\end{equation}
-\begin{equation}
-\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_4, p_5\}), (F_{3,1}, \{p_6, p_3\})\}
-\end{equation}
-In either scenario, after mapping message~$15$ to flow $F_{3,1}$, message~$16$ cannot be mapped to any existing flow instance or to a new flow instance.
-From this inconsistent message, we know that it is generated by CPU1.
-In scenario~(1), CPU1 is in the state after generating the message reporting a cache miss.
-In this case, the DataValid message should not generated.
-In scenario~(2), CPU1 is in the state before generating the message reporting either a cache hit or miss, and again the DataValid message should not generated.
-This inconsistent message helps to locate a bug in the CPU1 model where a DataValid message is generated after either a cache hit or miss message is generated.
-After fixing this bug,  in a few more iterations of analysis and debugging, the trace analysis can eventually extract all initiated flow instances in the model, all in their terminal states, thus showing that the model implements the four flows correctly.
-
-In the second experiment, partial observability is taken into account during the trace analysis with the assumption that the command $\mathit{Wr}$ and $\mathit{Rd}$ and addresses $M$ and $P$ are indistinguishable due to the lack of observability.
-This partial observability is simulated with a modification to the monitors such that in each observed message, command $\mathit{Wr}$ or $\mathit{Rd}$ is replaced with $(\mathit{Wr},\mathit{Rd})$ and address $M$ or $P$ is replaced with $(M,P)$.
-The version of the model generating the trace shown in Table~\ref{table-trace} is reused with the modified the monitors, and the generated trace with the simulated partial observability is shown in Table~\ref{table-trace-2}.
-Similarly, only the first sixteen messages are shown.
-\begin{table*}
-\caption{A trace of messages with partial observability for trace analysis.}
-\label{table-trace-2}
-\begin{center}
-\resizebox{7.3in}{!}{
-\begin{tabular}{llll}
-1~~(CPU2, Icon, (Wr, Rd), (M, P)) & 2~~(Icon, CPU1, (Wr, Rd), (M, P)) & 3~~(Icon, Mem, (Wr, Rd), (M, P)) & 4~~(Mem,  Icon, DVal, -) \\
-5~~(CPU1, Icon, (Hit, Miss), -) & 6~~(CPU1, Icon, DVal, -) & 7~~(Icon, CPU2, (Wr, Rd), (M, P)) & 8~~(Icon, Mem, (Wr, Rd), (M, P)) \\
-9~~(CPU2, Icon, (Wr, Rd), (M, P)) & 10~(Icon, CPU1, (Wr, Rd), (M, P)) & 11~(Icon, Mem, (Wr, Rd), (M, P)) & 12~(CPU2, Icon, (Wr, Rd), (M, P)) \\
-13~(Icon, CPU1, (Wr, Rd), (M, P)) & 14~(CPU1, Icon, (Hit, Miss), -) & 15~(Icon, Dev, (Wr, Rd), (M, P)) & 16~(CPU1, Icon, DVal, -) \\
-\end{tabular}
-}
-\end{center}
-\end{table*}
-
-Each message in the trace with partial observability is referred to as an {\em super} message to distinguish it from the messages of full observability.
-The traces of super messages are referred to as {\em super} traces.
-For example, the first super message in the trace from Table~\ref{table-trace-2}, $\mbox{(CPU2, Icon, (Wr, Rd), (M, P))}$, corresponds to four distinct messages:
-$\mbox{(CPU2, Icon, Wr, M)}$, $\mbox{(CPU2, Icon, Wr, P)}$, $\mbox{(CPU2, Icon, Rd, M)}$, and $\mbox{(CPU2, Icon, Rd, P)}$.
-Some of these messages do not exist in the flow specification, and are ignored during the trace analysis.
-In the above example, message $\mbox{(CPU2, Icon, Wr, P)}$ is ignored.
-
-%Each super trace represents a set of traces, each of which is interpreted to derive a set of flow execution scenarios.
-These flow execution scenarios can show different inconsistent messages.
-For example, the particular trace of the super trace shown in Table~\ref{table-trace-2} below
-%\[
-%\mbox{1 (CPU2, Icon, Wr, M), 2~(Icon, CPU1, Rd, M)}, \ldots
-%\]
-%leads to the following flow execution scenario $\{(F_{1,1}, \langle p_2,p_3\rangle)\}$ showing that the second message is inconsistent.
-In addition, this super trace also represents a trace as shown in Table~\ref{table-trace} that leads to a real bug in the model as discussed above.
-It is also possible in some situation that a super trace may represent a trace with no inconsistent messages.
-Therefore, the trace analysis on a super trace can result in a number of flow execution scenarios revealing different problems or no problem.
-This ambiguity is caused by the partial observability.
-
-Each super trace represents a set of traces, each of which is interpreted to derive a set of flow execution scenarios.
-Due to the partial observability, the number of traces represented by a super trace can become very large.
-For example, the super trace shown in Table~\ref{table-trace-2} represents about twelve thousand possible traces for that short sequence of messages.
-The trace analysis algorithm returns the set of flow execution scenarios for each trace for examination, and a very large number of possible flow execution scenarios can be generated.
-The large number of possible flow execution scenarios not only produces too much information that can overwhelm system validators, but also degrades the performance of the trace analysis algorithm by consuming too much memory.
-As indicated above,  the validators' insights on the SUD can be utilized to trim the possibilities.
-For example, if the validator knows that no flow $F_1$ in Fig.~\ref{fig-cpu-write} is activated in the testing environment, this insight helps to eliminate all flow execution scenarios that include instances of $F_1$ by interpreting message~$\#1$ as either $\mbox{(CPU2, Icon, Rd, M)}$ or $\mbox{(CPU2, Icon, Rd, P)}$.
-Consider another insight such that a new instance of flow $F_2$ as in Fig.~\ref{fig-cpu-read} can be initiated only after the completion of the previous instance of $F_2$.
-If an instant of $F_2$ is assumed to be initiated by the super message $\#9~\mbox{(CPU2, Icon, (Wr, Rd), (M, P))}$ by interpreting it to $\mbox{(CPU2, Icon, Rd, M)}$ during the trace analysis, the super message~$\#12$ can only be interpreted to $\mbox{(CPU2, Icon, Wr, M)}$ or $\mbox{(CPU2, Icon, Rd, P)}$ as the instance of $F_2$ initiated by the super message $\#9$ has not been completed yet at this point.
-According to the above discussion, the validators' insights help restrict how super messages are interpreted, thus reducing the number of flow execution scenarios that can be generated.
-At the end of the analysis, all possible flow execution scenarios are returned to system validators for examination.
-
-%
-%--
-%
-\section{Case Study II}
 
@@
 As there are two CPUs in this model, there are actually six such flows considered for this model.
-More detailed information about these protocols and the flow specifications can be found online\footnote{\url{https://github.com/cao2/paper/blob/master/Flow\%20Specification.pdf}} to meet the page limit.
+These flow specifications are derived from the cache coherence protocols supported in GEM5\footnote{The GEM5 cache coherence protocols can be found at \url{http://www.gem5.org/docs/html/gem5MemorySystem.html}.}.
+More detailed information about the flow specifications used in this case study will be provided as a supplement document if this paper is accepted for publication.
+%More detailed information about these protocols and the flow specifications can be found online\footnote{\url{https://github.com/cao2/paper/blob/master/Flow\%20Specification.pdf}} to meet the page limit.
 
@@
 In this experiment, only the five links involving the memory bus are still considered.
-However, an assumption is made that all messages passing the same link are not distinguishable due to the limitation of the observability.
-More specifically, the monitors are modified such that whenever a message is captured on one of the links, it dumps a set of messages passing through the same link into the trace file.
-Therefore, each line of the trace file corresponds to a set of messages.
+However, an assumption is made that all events passing the same link are not distinguishable due to the limitation of the observability.
+More specifically, the monitors are modified such that whenever an event is captured on one of the links, it dumps a set of events passing through the same link into the trace file.
+Therefore, each line of the trace file corresponds to a set of events.
 After applying the trace analysis to this trace,  a total of 13944 flow execution scenarios are extracted.
-This large number, compared to the number of extracted execution scenarios shown in the Table~\ref{table-case-2} and \ref{table-par-obs}, is due to the ambiguous interpretation of the messages with limited observability.
-
-The whole experiment takes about $15$ minutes and $420$~MB to finish, significantly higher than the numbers for analyzing traces where there is no ambiguity in the observed messages.
-This is due to the fact that a trace of ambiguous events is in fact a set of traces of original messages, which lead to large numbers of execution scenarios either during or at the end of the analysis.
-In this experiment, the peak number of executions during the analysis process is $70384$.
-Compared to the final number, many of the intermediate scenarios are invalid, and removed eventually.
-However, controlling the number of intermediate scenarios during the trace analysis is critical in order for the analysis to be tractable.
+This large number, compared to the number of extracted execution scenarios shown in the Table~\ref{table-case-2} and \ref{table-par-obs}, is due to the ambiguous interpretation of the events with limited observability.
+
+The whole experiment takes about $15$ minutes and $420$~MB to finish, significantly higher than the numbers for analyzing traces where there is no ambiguity in the observed events.
+This is due to the fact that a trace of ambiguous events is in fact a set of traces of original events, which lead to large numbers of execution scenarios either during or at the end of the analysis.
+In this experiment, the peak number of execution scenarios during the analysis process is $70384$.
+Compared to the final number, many of the intermediate execution scenarios are invalid, and removed eventually.
+However, controlling the number of intermediate execution scenarios during the trace analysis is critical in order for the analysis to be tractable.
 Here, insights from validators can help.
+
+As shown above, the ambiguous interpretation of events can lead to large numbers of intermediate and final execution scenarios, which not only make the trace analysis more time consuming but also make it difficult to gain insightful understanding from the derived execution scenarios.
+Careful selection of what to observe may have big impact on results from the trace analysis.
+In this last experiment, we relax the assumption made in the previous experiment such that the events passing each link are partitioned into two groups, one for read operations and one for write operations.
+Similar to the assumption made in the previous experiment, events in the same group are assumed to be non-distinguishable.
+The monitors are modified accordingly such that they output all events in the same group into the trace file if an event from that group is captured.
+After the trace analysis on this new partially observed trace is finished, only one execution scenario is derived where the distribution of the numbers of flow instances is the same as those shown in Table~\ref{table-par-obs}.
+The peak number of execution scenarios encountered during the trace analysis is $4$.
+The total runtime is about $1$~second.
+Compared to the results from the previous experiment, the precision and the performance of the trace analysis are improved dramatically as a result of careful selection of observable signals.
 

### 2015-09-29 paper.tex
updated abstract
the end of introduction is also updated.

Added a paragraph at the beginning of section V to mention the previous
case study.

@@
 post-silicon validation, and currently depends primarily on
-human creativity and insight.  We provide algorithms to
+human creativity and insights.  In this paper, we provide an algorithm to
 automatically infer system-level transactions from
-incomplete, ambiguous, and noisy trace data, together with a
-measure of confidence.  We demonstrate the approach on
-illustrative system-level protocols for SoC models developed
-in SystemC as well as in an FPGA environment.
+incomplete, ambiguous, and noisy trace data.
+% together with a measure of confidence.  
+We demonstrate the approach on a multicore virtual platoform
+developed within the GEM5 environment.
 \end{abstract}
@@
 % together with an estimate of confidence on the inference.  
-We demonstrate our approach on
-two SoC case studies: a pedagogical SystemC prototype and a
-more realistic implementation constructed within the GEM5
-environment~\cite{Binkert2011}.
+We demonstrate our approach on a multicore virtual platform
+constructed within the GEM5 environment~\cite{Binkert2011}.
 
@@
 
+The proposed method is demonstrated on a transaction level model of a simple SoC design with CPU core, memory, and a peripheral device connected by a highly concurrent interconnect in SystemC.
+Since the proposed method is communication centric, the detailed computations of these blocks are not modeled.
+Instead, the modeling is focused on how they participate in system level protocols.
+This simple case study shows that the proposed method is capable of providing useful information to help reveal and locate various bugs in the SystemC model, thus making the debugging more efficient.
+Once a bug-free model is available, the trace analysis method can correctly derive the instances of system flows activated by various components in the design from the observed traces.
+More details about the design model and system flows implemented by the design model will be provided in a supplement document stored online.
+
 \begin{figure} 
@@
 
-In order to find out the efficiency of the trace analysis method for more realistic examples, in this case study, a more detailed transaction level model of a SoC is constructed within the GEM5 environment~\cite{Binkert2011}.
+In order to find out the efficiency of the trace analysis method for more realistic examples, in the case study described in this section, a more detailed transaction level model of a SoC is constructed within the GEM5 environment~\cite{Binkert2011}.
 This SoC model consists of two ARM Cortex-A9 cores, each of which contains two separate 16KB data and instruction caches.

### 2015-09-30 paper.tex
My pass

@@
 post-silicon validation, and currently depends primarily on
-human creativity and insights.  In this paper, we provide an algorithm to
-automatically infer system-level transactions from
-incomplete, ambiguous, and noisy trace data.
-% together with a measure of confidence.  
-We demonstrate the approach on a multicore virtual platoform
+human creativity and insights.  In this paper, we provide an
+algorithm to automatically infer system-level transactions
+from incomplete, ambiguous, and noisy trace data.  We
+demonstrate the approach on a multicore virtual platoform
 developed within the GEM5 environment.
@@
 system-level transactions from traces with missing, noisy,
-and ambiguous signal values.
-% together with an estimate of confidence on the inference.  
-We demonstrate our approach on a multicore virtual platform
-constructed within the GEM5 environment~\cite{Binkert2011}.
+and ambiguous signal values.  We demonstrate our approach on
+a multicore virtual platform constructed within the GEM5
+environment~\cite{Binkert2011}.
 
@@
 protocols are typically specified in architecture documents
-as message flow diagrams.  As illustration,
-Fig~\ref{flow-spec-ex}(a) shows one such diagram for a
-protcol to authenticate and load a firmware during system
-boot for firmware upgrade.  During validation, the system
-under debug (SUD) exercises some complex system-level
-use-case which involves interleaved execution of possibly a
-large number of such protocols.  A trace of a small number
-of hardware signals is then shipped off-chip analysis.  The
-off-chip analysis includes two broad phases: (1)~trace
-abstraction, and (2)~trace interpretation.  Trace
-abstraction maps the hardware trace into higher-level
-architectural constructs, \eg, messages, operations, etc.: a
-message such as {\tt Authorization request} may be
-implemented in hardware through a Boolean or temporal
-combination of specific hardware signals in the NoC fabric
-between {\tt Device} and {\tt CE}, \eg, as a sequence
-containing a header, a specific value of a sequence of data
-words, etc.  We will refer to such architectural constructs
-as {\em protocol events} or {\em flow events}.  Note that
-due to limited observability, it may not be possible to map
-a given set of (observed) hardware signals uniquely to a
-flow event.  Finally, the trace may a result from several
-instances of the same protocol executing concurrently, \eg,
-a firmware authentication protocol may be invoked when
-another instance of the protocol has not completed.
+as message flow diagrams.  For this paper, we use the words
+``protocol'' and ``flow'' interchangeably.
+Fig~\ref{flow-spec-ex}(a) shows one diagram for a protocol
+to authenticate and load a firmware during system boot for
+firmware upgrade.  During validation, the system under debug
+(SUD) exercises some complex system-level use-case which
+involves interleaved execution of possibly a large number of
+such flows.  A trace of a small number of hardware signals
+is then shipped off-chip analysis.  The off-chip analysis
+includes two broad phases: (1)~trace abstraction, and
+(2)~trace interpretation.  Trace abstraction maps the
+hardware trace into higher-level architectural constructs,
+\eg, messages, operations, etc.: a message such as {\tt
+  Authorization request} may be implemented in hardware
+through a Boolean or temporal combination of specific
+hardware signals in the NoC fabric between {\tt Device} and
+{\tt CE}, \eg, as a sequence containing a header, a specific
+value of a sequence of data words, etc.  We will refer to
+such architectural constructs as {\em protocol events} or
+{\em flow events}.  Note that due to limited observability,
+it may not be possible to map a given set of (observed)
+hardware signals uniquely to a flow event.  Finally, the
+trace may a result from several instances of the same
+protocol executing concurrently, \eg, a firmware
+authentication protocol may be invoked when another instance
+of the protocol has not completed.
 
@@
 
-To illustrate the basic idea of the trace analysis method, 
-consider the system flow shown in Figure~Fig.~\ref{flow-spec-ex}(b).
-Let $F_1$ denote such flow.  Suppose that a hardware system 
-implements flow $F_1$, and the following flow trace is abstracted 
-from an observed signal trace as a result of executing such system.
+To illustrate the basic idea, consider the system flow in
+Fig.~\ref{flow-spec-ex}(b), which we will call $F_1$.
+Suppose that the following flow trace is abstracted from an
+observed signal trace of the system.
 \[
@@
 \]  
-where transition names in the LPN are used to represent the 
+Here transition names in the LPN are used to represent the 
 flow events in the trace.  The first four events results in
@@
 \]
-By following Algorithm~\ref{algo:compliance} to handle the remaining 
-four events, the following execution scenario can be derived.
+Using Algorithm~\ref{algo:compliance} to handle the
+remaining four events, the following execution scenario is
+derived.
 \[
@@
 
-The trace interpretation discussion in the preceding section
-assumed full observability.  In this section, we discuss how
-to adapt the method to deal with silicon traces under
-limited (partial) obsrvability.
-
 %% In hardware that implements a given system flow
@@
 
-In general, a signal trace of partial observability corresponding a set of traces of flow events due to the ambiguous interpretation of signal events.
-In the following, we discuss two cases for trace abstraction on partial observability: mapping a single signal event to a flow event or mapping a sequence of signal events to a flow event.
-A signal event is defined as a state on or an assignment to a set of signals.
-
-Hereafter, the term {\em flow traces} is used to refer to traces of flow events. 
-Consider the following example for the first case.
-Suppose that there are three flow events: $e_1$, $e_2$, and $e_3$, which are implemented in hardware by the signal events shown in the list below.
-We use Boolean expressions to represent signal events for the discussion.
+In general, a signal trace of partial observability
+corresponding a set of traces of flow events due to the
+ambiguous interpretation of signal events.  In the
+following, we discuss two cases for trace abstraction on
+partial observability: mapping a single signal event to a
+flow event or mapping a sequence of signal events to a flow
+event.  A signal event is defined as a state on or an
+assignment to a set of signals.
+
+Hereafter, the term {\em flow traces} is used to refer to
+traces of flow events.  Consider the following example for
+the first case.  Suppose that there are three flow events:
+$e_1$, $e_2$, and $e_3$, which are implemented in hardware
+by the signal events shown in the list below.  We use
+Boolean expressions to represent signal events for the
+discussion.
 \[
@@
 \] 
-Now suppose that only signals $b$ and $c$ are observable, and a signal trace of this partial observability is obtained below.
+Suppose that only signals $b$ and $c$ are observable, and we
+obtain the following trace:
 \[
@@
 \]
-During the trace abstraction step, the first two signal events $bc$ can be mapped to $\{e_1, e_2\}$ since $a$ is not observable, and the last one $b'c$ is mapped to $\{e_3\}$.
-Therefore, this signal trace is abstracted to four flow traces, $\{e_1, e_2\} \times \{e_1, e_2\} \times \{e_3\}$.
-
-Next, we consider the case where a flow event is mapped from a sequence of signal events.
-Now suppose that two other flow events are implemented by sequences of signal events as defined in the list below.
+During trace abstraction, the first two signal events $bc$
+can be mapped to $\{e_1, e_2\}$ since $a$ is not observable,
+and the last one $b'c$ is mapped to $\{e_3\}$.  Therefore,
+this signal trace is abstracted to four flow traces, $\{e_1,
+e_2\} \times \{e_1, e_2\} \times \{e_3\}$.
+
+Next, we consider the case where a flow event is mapped from
+a sequence of signal events.  Now suppose that two other
+flow events are implemented by sequences of signal events as
+defined in the list below.
 \[
@@
 \]
-where $\textvisiblespace$ denotes signal events that are not mapped to any flow events.
-Note that the above abstraction leads to three distinct flow traces as the middle three correspond to the same trace of flow events.
-
-From the above discussions, it can be seen that a signal trace of partial observability is generally mapped to a set of flow traces.
-As the the number of signals available for observation becomes smaller, the number of flow event traces corresponding to a signal trace as a result of the abstraction can be enormous.
-This phenomenon can increase the complexity of the trace interpretation as it can cause the number of possible flow execution scenarios generated during the analysis to explode.
-Possible solutions to address this issue include better trace signal selection for observation and assistance from debuggers' insights.
-Trace signal selection itself is an important and difficult subject, and a detailed discussion of it is out of scope of this paper.
-Next, we briefly describe how the debuggers' insights of a system's architecture can help to address the complexity issue in the trace analysis.
-
-\subsection{Inputs from Validators}
-
-During the trace interpretation, the number of intermediate flow execution scenarios may become too large due to ambiguous interpretation from signal events to flow events or from flow events to flow execution scenarios.
-The explosion of the intermediate results can significantly slow down the performance of the trace analysis.
-To address this problem, the debuggers can use their insights and understanding of the SUD to trim the total number of possible system executions derived from the trace analysis.
-When a SUD is validated, a debugger is assumed to have deep knowledge about the system's architecture and microarchitecture, and how the test environment affects the system executions.
-For instance, he/she may have knowledge that in a test environment, the maximal number of instances of a flow can be activated by the SUD, or a flow can be activated after certain other flows have terminated, etc.
-These debugging insights can be encoded as constraints into the trace analysis method, which can then be used to eliminate a large number of flow execution scenarios that violate these constraints during the trace interpretation step.
-
-This approach can be flexible in that it allows a debugger to analyze the observed traces in a trail-and-error manner if the precise knowledge of the system (micro-)architecture is hard to come by.
-For instance, the debugger might initially make a very restricted assumption on how the SUD executes a flow specification, and these assumptions can potentially lead to an empty set of flow execution scenarios.
-Depending on which of these assumptions triggered during the trace interpretation step, the debugger can study these assumptions more carefully, and relax some or all of them for the next run of analysis.
-This iteration can be repeated as many times as necessary until some results deemed meaningful are produced.
-
-Alternatively, if all derived execution scenarios seem to be plausible, the implication that a debugger may draw from this result is that the failure may be independent of the flows being observed.
-Therefore, the testing environment can be adjusted in order for a different part or different behavior of the SUD to be observed.
-This idea, closely related to trace signal selection, is critical for post-silicon validation, and a detailed discussion can only be presented in a separate paper.
+
+where $\textvisiblespace$ denotes signal events that are not
+mapped to any flow events.  Note that the above abstraction
+leads to three distinct flow traces as the middle three
+correspond to the same trace of flow events.
+
+It is clear from above that a partial trace is viewed as a
+set of flow traces, and Algorithm~\ref{algo:compliance} can
+be suitably extended to work with flow traces to obtain the
+set of candidate flows.  However, applicability of the
+algorithm in practice can be gated because the number of
+potential flow execution scenarios generated under partial
+observability may be enormous.  Note that this is not a
+limitation of the algorithm; if the observability of
+critical events is poor there simply {\em are} too many flow
+execution scenarios compliant with the observed trace.
+Nevertheless, we need to address the issue to make trace
+interpretation (whether automatic or not) practicable.
+There are two potential approaches: (1)~better selection of
+post-silicon trace observability, and (2)~use of system
+insights during validation.  Trace signal selection itself
+is an important and orthogonal topic~\cite{nicolici,basu},
+and a detailed discussion of it is out of scope of this
+paper.  However, we briefly describe how the debuggers'
+insights of a system's architecture can help to address the
+complexity issue in the trace trace interpretation.
+
+\subsection{Interactive Trace Interpretation}
+
+Post-silicon validation is performed by debuggers with deep
+knowledge about the system's architecture and
+microarchitecture, and the test environment.  Two key
+insights are (1) the maximal number of instances of a flow
+activated in the test environment, and (2)~the mutual
+relationship between two flows.  For example, the test
+environment may not permit multiple instances of firmware
+authentication to operate concurrently, or a flow involving
+audio and Web browsing to initiate until the flows
+participating in boot are completed.  Our framework permits
+incorporating such insights as constraints in trace
+analysis; flow execution scenarios that violate these
+constraints are ignored.  Apart from the explosion problem,
+this permits the debugger to quickly filter out
+uninteresting combination of flows and focus on interesting
+interleavings.
+
+This approach can be flexible in that it allows a debugger
+to analyze the observed traces in a trail-and-error manner
+if the precise knowledge of the system (micro-)architecture
+is hard to come by.  For instance, the debugger might
+initially make a very restricted assumption on how the SUD
+executes a flow specification, and these assumptions can
+potentially lead to an empty set of flow execution
+scenarios.  Depending on which of these assumptions
+triggered during the trace interpretation step, the debugger
+can study these assumptions more carefully, and relax some
+or all of them for the next run of analysis.  This iteration
+can be repeated as many times as necessary until some
+results deemed meaningful are produced.
+
+%% Alternatively, if all derived execution scenarios seem to be
+%% plausible, the implication that a debugger may draw from
+%% this result is that the failure may be independent of the
+%% flows being observed.  Therefore, the testing environment
+%% can be adjusted in order for a different part or different
+%% behavior of the SUD to be observed.  This idea, closely
+%% related to trace signal selection, is critical for
+%% post-silicon validation, and a detailed discussion can only
+%% be presented in a separate paper.
 
@@
 
-The proposed method is demonstrated on a transaction level model of a simple SoC design with CPU core, memory, and a peripheral device connected by a highly concurrent interconnect in SystemC.
-Since the proposed method is communication centric, the detailed computations of these blocks are not modeled.
-Instead, the modeling is focused on how they participate in system level protocols.
-This simple case study shows that the proposed method is capable of providing useful information to help reveal and locate various bugs in the SystemC model, thus making the debugging more efficient.
-Once a bug-free model is available, the trace analysis method can correctly derive the instances of system flows activated by various components in the design from the observed traces.
-More details about the design model and system flows implemented by the design model will be provided in a supplement document stored online.
+%% The proposed method is demonstrated on a transaction level
+%% model of a simple SoC design with CPU core, memory, and a
+%% peripheral device connected by a highly concurrent
+%% interconnect in SystemC.  Since the proposed method is
+%% communication centric, the detailed computations of these
+%% blocks are not modeled.  Instead, the modeling is focused on
+%% how they participate in system level protocols.  This simple
+%% case study shows that the proposed method is capable of
+%% providing useful information to help reveal and locate
+%% various bugs in the SystemC model, thus making the debugging
+%% more efficient.  Once a bug-free model is available, the
+%% trace analysis method can correctly derive the instances of
+%% system flows activated by various components in the design
+%% from the observed traces.  More details about the design
+%% model and system flows implemented by the design model will
+%% be provided in a supplement document stored online.
 
@@
 
-In order to find out the efficiency of the trace analysis method for more realistic examples, in the case study described in this section, a more detailed transaction level model of a SoC is constructed within the GEM5 environment~\cite{Binkert2011}.
-This SoC model consists of two ARM Cortex-A9 cores, each of which contains two separate 16KB data and instruction caches.
-The caches are connected to a 1GB memory through a memory bus model.
-The architecture of this SoC model is shown in Fig.~\ref{SoC}.
-
-In this model, components communicate with each other by sending and receiving various request and response messages.
-In order to observe and trace communications occurring inside this model during execution, monitors are attached to links connecting the components.
-These monitors record the messages flowing through the links they are attached to, and store them into output trace files.
+In order to find out the efficiency of the trace analysis
+method for more realistic examples, in the case study
+described in this section, a more detailed transaction level
+model of a SoC is constructed within the GEM5
+environment~\cite{Binkert2011}.  This SoC model consists of
+two ARM Cortex-A9 cores, each of which contains two separate
+16KB data and instruction caches.  The caches are connected
+to a 1GB memory through a memory bus model.  The
+architecture of this SoC model is shown in Fig.~\ref{SoC}.
+
+In this model, components communicate with each other by
+sending and receiving various request and response messages.
+In order to observe and trace communications occurring
+inside this model during execution, monitors are attached to
+links connecting the components. These monitors record the
+messages flowing through the links they are attached to, and
+store them into output trace files.
 
@@
 
-For this model, we consider the flow specifications describing the cache coherence protocols supported in GEM5 that is used to build the model in Fig~\ref{SoC}.
-These flow specifications describe data/instruction read operations and data write operations initiated from CPUs.
-Three such flows describe the cache coherent protocols for each CPU.
-As there are two CPUs in this model, there are actually six such flows considered for this model.
-These flow specifications are derived from the cache coherence protocols supported in GEM5\footnote{The GEM5 cache coherence protocols can be found at \url{http://www.gem5.org/docs/html/gem5MemorySystem.html}.}.
-More detailed information about the flow specifications used in this case study will be provided as a supplement document if this paper is accepted for publication.
+For this model, we consider the flow specifications
+describing the cache coherence protocols supported in GEM5
+that is used to build the model in Fig~\ref{SoC}.  These
+flow specifications describe data/instruction read
+operations and data write operations initiated from CPUs.
+Three such flows describe the cache coherent protocols for
+each CPU.  As there are two CPUs in this model, there are
+actually six such flows considered for this model.  These
+flow specifications are derived from the cache coherence
+protocols supported in GEM5.\footnote{The GEM5 cache
+  coherence protocols can be found at
+  \url{http://www.gem5.org/docs/html/gem5MemorySystem.html}.}
+
+%% More detailed information about the flow specifications used in this case study will be provided as a supplement document if this paper is accepted for publication.   
 %More detailed information about these protocols and the flow specifications can be found online\footnote{\url{https://github.com/cao2/paper/blob/master/Flow\%20Specification.pdf}} to meet the page limit.
@@
 
-Two simple programs are written, one for each CPU.
-These simple programs read numbers from a file, perform some operations on these numbers, and store the results back to the file.
-How GEM5 supports shared memory multi-threaded program execution is unclear.
-Therefore, there is no data shared in both caches in this test.
-Furthermore, GEM5 does not support true concurrency.
-When there are two programs running on the CPUs, GEM5 alternates the executions between the two CPUs.
-To simulate asynchronous concurrency with the interleaving semantics, those two simple programs are instrumented with pseudo-blocking commands, one placed before each statement.
-A pseudo blocking command includes a random number generator that returns either $0$ or $1$ and a loop that only exits when the returned random number is $0$.
+We wrote two simple concurrent programs --- one for each CPU
+--- to exercise the flows.  They read numbers from a file,
+perform some operations on these numbers, and store the
+results back to the file.  How GEM5 supports shared memory
+multi-threaded program execution is unclear.  Therefore,
+there is no data shared in both caches in this test.
+Furthermore, GEM5 does not support true concurrency.  When
+there are two programs running on the CPUs, GEM5 alternates
+the executions between the two CPUs.  To simulate
+asynchronous concurrency with the interleaving semantics,
+those two simple programs are instrumented with
+pseudo-blocking commands, one placed before each statement.
+A pseudo blocking command includes a random number generator
+that returns either $0$ or $1$ and a loop that only exits
+when the returned random number is $0$.
 
@@
 
-After the SoC model finishes executing the program, there are totally $343581$ messages collected in the trace file.
-Not all of the messages are relevant to the flow specification as many are used by GEM5 to initialize its simulation environment.
-After removing those irrelevant messages, the number of messages in the trace file is to reduced to $121138$.
-
-The time taken to remove the irrelevant messages from the trace is negligible.
-The total runtime and the peak memory usage for the trace analysis algorithm to finish the reduced trace are 3 seconds and 12MB, respectively.
-From the trace, Table~\ref{table-case-2} shows the number of instances extracted for the six flows describing cache coherent read/write operations initiated from both CPUs.
+After the SoC model finishes executing the program, there
+are totally $343581$ messages collected in the trace file.
+Not all of the messages are relevant to the flow
+specification as many are used by GEM5 to initialize its
+simulation environment.  After removing those irrelevant
+messages, the number of messages in the trace file is to
+reduced to $121138$.
+
+The time taken to remove the irrelevant messages from the
+trace is negligible.  The total runtime and the peak memory
+usage for the trace analysis algorithm to finish the reduced
+trace are 3 seconds and 12MB, respectively.  From the trace,
+Table~\ref{table-case-2} shows the number of instances
+extracted for the six flows describing cache coherent
+read/write operations initiated from both CPUs.
 \begin{table}[tb]
@@
 
-To take the partial observability into account, the four monitors attached to the links between two CPUs and their caches are disabled.
-Then, the trace is generated by the remaining five monitors from the SoC model executing the same program.
-After removing all non-observable messages, the trace only contains 15089 messages.
-The numbers of the flow instances extracted by the trace analysis method are shown in Table~\ref{table-par-obs}.
-From these results, the numbers of the flow instances are dropped significantly compared to the results extracted from the trace with the full observability as shown in Table~\ref{table-case-2}.
-This difference is due to that some communications occurred in the system when executing the program involve the CPUs and their corresponding caches only, and the traffic on the links between the CPUs and their corresponding caches is not observable.
-Therefore, the instances of the flow specifications characterizing these communications do not exist in the trace.
-In other words, all extracted flow instances in Table~\ref{table-par-obs} characterize the communications that pass through the memory bus in the system model.
-The runtime and memory usage is similar to that for analyzing the trace of the full observability as shown above.
+To take the partial observability into account, the four
+monitors attached to the links between two CPUs and their
+caches are disabled. Then, the trace is generated by the
+remaining five monitors from the SoC model executing the
+same program. After removing all non-observable messages,
+the trace only contains 15089 messages. The numbers of the
+flow instances extracted by the trace analysis method are
+shown in Table~\ref{table-par-obs}.  From these results, the
+numbers of the flow instances are dropped significantly
+compared to the results extracted from the trace with the
+full observability as shown in
+Table~\ref{table-case-2}. This difference is due to that
+some communications occurred in the system when executing
+the program involve the CPUs and their corresponding caches
+only, and the traffic on the links between the CPUs and
+their corresponding caches is not observable. Therefore, the
+instances of the flow specifications characterizing these
+communications do not exist in the trace. In other words,
+all extracted flow instances in Table~\ref{table-par-obs}
+characterize the communications that pass through the memory
+bus in the system model.  The runtime and memory usage is
+similar to that for analyzing the trace of the full
+observability as shown above.
 

### 2015-10-01 paper.tex
Final version fro ISQED
Added a results table in section VI.

Fixed missing references issues.

@@
 }
-\Return $\mathit{scen}$ \;
+\Return $\mathit{Scen}$ \;
 \caption{$\textsc{Check-Compliance}(\vec{F}, \, \rho)$}
@@
 
+\begin{table}[tb]
+\caption{Runtime Results of Trace analysis. Time is in seconds and memory usage is in MB.}
+\begin{center}
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+	&	F-Obs. & \begin{tabular}{c}P-Obs. \\ No Amb.\end{tabular} & \begin{tabular}{c}P-Obs. \\ Amb.~1\end{tabular} & \begin{tabular}{c}P-Obs. \\ Amb.~2\end{tabular} \\
+\hline
+\hline
+Time & $3$ & $2.78$ & $896$ & $<1$\\
+\hline
+Mem & $12$ & $10$ & $420$ & $9$\\
+\hline
+
+\end{tabular}
+\end{center}
+\label{table-results}
+\end{table}%
+
+After this model is executed with the simple concurrent programs, 
+the trace analysis is applied to traces with different observabilities
+collected from this model.  The runtime results are shown in Table~\ref{table-results}.
+The first column shows the results from analyzing the trace with the full
+observability, while the next three show the result from 
+analyzing traces with different partial observability assumptions.
+
+In the first experiment, full observability is assumed.
 After the SoC model finishes executing the program, there
@@
 
-To take the partial observability into account, the four
-monitors attached to the links between two CPUs and their
+In the second experiment, partial observability is taken into account
+with the four monitors attached to the links between two CPUs and their
 caches are disabled. Then, the trace is generated by the
@@
 characterize the communications that pass through the memory
-bus in the system model.  The runtime and memory usage is
-similar to that for analyzing the trace of the full
-observability as shown above.
+bus in the system model.  The runtime and memory usage as shown in
+the third column in Table~\ref{table-results} are
+similar to those for analyzing the trace of the full
+observability.
 
@@
 
-The whole experiment takes about $15$ minutes and $420$~MB to finish, significantly higher than the numbers for analyzing traces where there is no ambiguity in the observed events.
+The whole experiment takes about $15$ minutes and $420$~MB to finish as shown in column~4
+in Table~\ref{table-results}, significantly higher than the numbers for analyzing traces where there is no ambiguity in the observed events.
 This is due to the fact that a trace of ambiguous events is in fact a set of traces of original events, which lead to large numbers of execution scenarios either during or at the end of the analysis.
@@
 The peak number of execution scenarios encountered during the trace analysis is $4$.
-The total runtime is about $1$~second.
-Compared to the results from the previous experiment, the precision and the performance of the trace analysis are improved dramatically as a result of careful selection of observable signals.
+The total runtime and memory usage are negligible as shown in the last column in Table~\ref{table-results}.
+Compared to the results from the previous experiment, the precision and the performance of the trace analysis are improved dramatically as a result of careful selection of observable events.
 

### 2015-12-21 paper.tex
added one more reference

@@
 
-%% \author{\large Hao Zheng$^1$, Yuting Cao$^1$, Sandip Ray$^2$, Jin Yang$^2$ \\
-%% $^1$Dept. of Computer Science and Eng., University of South Florida, Tampa, Fl 33620. USA. \\
-%% $^2$Strategic CAD Labs, Intel Corporation, Hillsboro, OR 97124.  USA.}
+ \author{\large Hao Zheng$^1$, Yuting Cao$^1$, Sandip Ray$^2$, Jin Yang$^2$ \\
+ $^1$Dept. of Computer Science and Eng., University of South Florida, Tampa, Fl 33620. USA. \\
+ $^2$Strategic CAD Labs, Intel Corporation, Hillsboro, OR 97124.  USA.}
 
@@
 schedules and accounting for more than $50\%$ of the overall
-design validation cost.  Consequently, it is crucial to
+design validation cost~\cite{Patra2007}.  Consequently, it is crucial to
 develop techniques for streamlining and automating
@@
 \begin{minipage}{3in}
-\resizebox{2.8in}{!}{
+\resizebox{2.4in}{!}{
 \begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
@@
 hardware signals uniquely to a flow event.  Finally, the
-trace may a result from several instances of the same
+trace may be a result from several instances of the same
 protocol executing concurrently, \eg, a firmware
@@
 \noindent 
-{\bf Notations and formalization.}  We assume that the set
-of system flows is provided as a collection ${\vec{F}}$ of
-LPNs.  We define a {\em flow execution scenario} to be a
+{\bf Notations and formalization.}  The set
+of system flows is a collection ${\vec{F}}$ of
+LPNs.  A {\em flow execution scenario} is defined as a
 set $\{(F_{i,j}, s_{i,j})\}$ where $F_{i,j}$ is the
@@
 current states.  Since we assume full observability, we view
-an {\em observed trace} $\rho = e_1e_2\ldots e_n$ is a
+an {\em observed trace} $\rho = e_1e_2\ldots e_n$ as a
 sequence of events.  Given an observed trace $\rho$, the
-goal of of trace interpretation is to construct a set of
+goal of trace interpretation is to construct a set of
 candidate flow execution scenarios whose execution can
@@
 $t$ in $F_i$ such that $L(t) = e$ and $\bullet t \subseteq
-s_{i,j}$.  It returns $\emptyset$ if no such $t$
-exists in $F_i$.
-
-\medskip
-
-Given an obsered trace $\rho$ and the set $\vec{F}$ of LPNs,
+s_{i,j}$.  It returns $\emptyset$ otherwise.
+
+%\medskip
+
+Given an observed trace $\rho$ and the set $\vec{F}$ of LPNs,
 Algorithm~\ref{algo:compliance} provides a basic procedure for
 computing a set of compliant flow execution scenarios. 
+The algorithm operates by keeping track (in variable
+$\mathit{Scen}$) of a set of candidate flow execution scenarios
+compliant with each prefix of $\rho$.  At each iteration,
+for each event $e_h$ in the observed trace, we
+update $Scen$ by either extending a member of
+$\mathit{scen}$ or initiating a new protocol instance for 
+each $scen \in Scen$ with respect to $e_h$ in every possible way.
+If $e_h$ cannot be emitted by any existing or new flow instances, 
+then we report that the trace is {\tt  inconsistent}, \ie, 
+there is no possible interleaving of
+the protocol instances from $\vec{F}$ that is compliant with
+$\rho$.
+
 \begin{algorithm}[ht]
@@
 \end{algorithm}
-The algorithm operates by keeping track (in variable
-$\mathit{Scen}$) of a set of candidate flow execution scenarios
-compliant with each prefix of $\rho$.  At each iteration,
-for each event $e_h$ in the observed trace, we
-update $Scen$ by either extending a member of
-$\mathit{scen}$ or initiating a new protocol instance for 
-each $scen \in Scen$ with respect to $e_h$ in every possible way.
-If $e_h$ cannot be emitted by any existing or new flow instances, 
-then we report that the trace is {\tt  inconsistent}, \ie, 
-there is no possible interleaving of
-the protocol instances from $\vec{F}$ that is compliant with
-$\rho$.
 
@@
 Suppose that the following flow trace is abstracted from an
-observed signal trace of the system.
+observed signal trace.
 \[
@@
 In general, a signal trace of partial observability
-corresponding a set of traces of flow events due to the
+corresponds a set of traces of flow events due to the
 ambiguous interpretation of signal events.  In the
@@
 leads to three distinct flow traces as the middle three
-correspond to the same trace of flow events.
+correspond to the same flow trace.
 
@@
 insights of a system's architecture can help to address the
-complexity issue in the trace trace interpretation.
+complexity issue in the trace interpretation.
 
@@
 analysis; flow execution scenarios that violate these
-constraints are ignored.  Apart from the explosion problem,
-this permits the debugger to quickly filter out
-uninteresting combination of flows and focus on interesting
-interleavings.
+constraints are ignored.  \textcolor{red}{These insights can lead to two advantages.
+First, they help to reduce the potentially large number of partial
+scenarios generated during the trace interpretation step, thus making 
+the analysis more efficient.  Second, they permit the debugger
+to quickly filter out uninteresting combinations of flows and focus on interesting
+interleavings.}
 
@@
 \centerline{
-\includegraphics[width=3.3in]{figures/Fig4.png}}
+\includegraphics[width=3in]{figures/Fig4.png}}
 \caption{SoC platform structure.}
@@
 model of a SoC is constructed within the GEM5
-environment~\cite{Binkert2011}.  This SoC model consists of
+environment~\cite{Binkert2011}.  This SoC model, as shown in Fig.~\ref{SoC},
+consists of
 two ARM Cortex-A9 cores, each of which contains two separate
 16KB data and instruction caches.  The caches are connected
-to a 1GB memory through a memory bus model.  The
-architecture of this SoC model is shown in Fig.~\ref{SoC}.
+to a 1GB memory through a memory bus model.
 
@@
 For this model, we consider the flow specifications
-describing the cache coherence protocols supported in GEM5
+describing the cache coherence protocols supported in GEM5\footnote{The GEM5 cache
+  coherence protocols can be found at
+  \url{http://www.gem5.org/docs/html/gem5MemorySystem.html}.}
 that is used to build the model in Fig~\ref{SoC}.  These
@@
 each CPU.  As there are two CPUs in this model, there are
-actually six such flows considered for this model.  These
-flow specifications are derived from the cache coherence
-protocols supported in GEM5.\footnote{The GEM5 cache
-  coherence protocols can be found at
-  \url{http://www.gem5.org/docs/html/gem5MemorySystem.html}.}
+actually six such flows considered for this model.
 
@@
 
-We wrote two simple concurrent programs --- one for each CPU
---- to exercise the flows.  They read numbers from a file,
+We wrote two simple concurrent programs, one for each CPU,
+ to exercise the flows.  They read numbers from a file,
 perform some operations on these numbers, and store the
@@
 multi-threaded program execution is unclear.  Therefore,
-there is no data shared in both caches in this test.
+no data are shared in both caches in this test.
 Furthermore, GEM5 does not support true concurrency.  When
@@
 trace is negligible.  The total runtime and the peak memory
-usage for the trace analysis algorithm to finish the reduced
-trace are 3 seconds and 12MB, respectively.  From the trace,
-Table~\ref{table-case-2} shows the number of instances
-extracted for the six flows describing cache coherent
-read/write operations initiated from both CPUs.
+taken by the trace analysis algorithm on the reduced
+trace are 3 seconds and 12MB, respectively.  Only one flow execution
+scenario is extracted, and 
+Table~\ref{table-case-2} shows the number of flow instances contained in that
+scenario for the six 
+flows describing cache coherent operations initiated from both CPUs.
 \begin{table}[tb]
@@
 remaining five monitors from the SoC model executing the
-same program. After removing all non-observable messages,
-the trace only contains 15089 messages. The numbers of the
-flow instances extracted by the trace analysis method are
+same program.  The new trace contains 15089 messages.  
+Similarly, only one flow execution
+scenario is extracted, and the numbers of the
+flow instances contained in that execution scenario are
 shown in Table~\ref{table-par-obs}.  From these results, the
@@
 In this experiment, only the five links involving the memory bus are still considered.
-However, an assumption is made that all events passing the same link are not distinguishable due to the limitation of the observability.
-More specifically, the monitors are modified such that whenever an event is captured on one of the links, it dumps a set of events passing through the same link into the trace file.
+However, an assumption is made that all events passing the same link are not distinguishable due to the limited observability.
+The monitors are modified such that whenever an event is captured on one of the links, it dumps a set of events passing through the same link into the trace file.
 Therefore, each line of the trace file corresponds to a set of events.
 After applying the trace analysis to this trace,  a total of 13944 flow execution scenarios are extracted.
-This large number, compared to the number of extracted execution scenarios shown in the Table~\ref{table-case-2} and \ref{table-par-obs}, is due to the ambiguous interpretation of the events with limited observability.
+This large number, compared to the results from the first two experiments, is due to the ambiguous interpretation of the events with limited observability.
 
@@
 This is due to the fact that a trace of ambiguous events is in fact a set of traces of original events, which lead to large numbers of execution scenarios either during or at the end of the analysis.
-In this experiment, the peak number of execution scenarios during the analysis process is $70384$.
-Compared to the final number, many of the intermediate execution scenarios are invalid, and removed eventually.
+In this experiment, the peak number of execution scenarios during the analysis process is $70384$, many of which are invalid and removed eventually.
 However, controlling the number of intermediate execution scenarios during the trace analysis is critical in order for the analysis to be tractable.
-Here, insights from validators can help.
+Here, insights from validators could help, but are not used in this experiment.
 
@@
 
-This paper presents a trace analysis based method for post-silicon validation by interpreting observed raw signal traces at the level of system flow specifications.
+This paper presents a method for post-silicon validation by interpreting observed raw signal traces at the level of system flow specifications.
 The derived flow execution scenarios provide more structured information on system operations, which is more understandable to system validators.

### 2015-12-28 paper.tex
Done minor edits

@@
 schedules and accounting for more than $50\%$ of the overall
-design validation cost~\cite{Patra2007}.  Consequently, it is crucial to
-develop techniques for streamlining and automating
-post-silicon validation activities.
+design validation cost~\cite{Patra2007}.  Consequently, it
+is crucial to develop techniques for streamlining and
+automating post-silicon validation activities.
 
@@
 analysis; flow execution scenarios that violate these
-constraints are ignored.  \textcolor{red}{These insights can lead to two advantages.
-First, they help to reduce the potentially large number of partial
-scenarios generated during the trace interpretation step, thus making 
-the analysis more efficient.  Second, they permit the debugger
-to quickly filter out uninteresting combinations of flows and focus on interesting
-interleavings.}
+constraints are ignored.  These insights can lead to two
+advantages.  First, they help to reduce the potentially
+large number of partial scenarios generated during the trace
+interpretation step, thus making the analysis more
+efficient.  Second, they permit the debugger to quickly
+filter out uninteresting combinations of flows and focus on
+interesting interleavings.
 
 This approach can be flexible in that it allows a debugger
-to analyze the observed traces in a trail-and-error manner
+to analyze the observed traces in a trial-and-error manner
 if the precise knowledge of the system (micro-)architecture
@@
 
-In order to find out the efficiency of the trace analysis
-method for more realistic examples, in the case study
-described in this section, a more detailed transaction level
-model of a SoC is constructed within the GEM5
-environment~\cite{Binkert2011}.  This SoC model, as shown in Fig.~\ref{SoC},
-consists of
-two ARM Cortex-A9 cores, each of which contains two separate
+To determine the efficiency of the trace analysis method for
+a realistic example, a transaction level model of a SoC is
+constructed within the GEM5 environment~\cite{Binkert2011}.
+This SoC model, as shown in Fig.~\ref{SoC}, consists of two
+ARM Cortex-A9 cores, each of which contains two separate
 16KB data and instruction caches.  The caches are connected
-to a 1GB memory through a memory bus model.
-
-In this model, components communicate with each other by
-sending and receiving various request and response messages.
-In order to observe and trace communications occurring
-inside this model during execution, monitors are attached to
-links connecting the components. These monitors record the
-messages flowing through the links they are attached to, and
-store them into output trace files.
+to a 1GB memory through a memory bus model.  Components
+communicate with each other by sending and receiving various
+request and response messages.  In order to observe and
+trace communications occurring inside this model during
+execution, monitors are attached to links connecting the
+components.  These monitors record the messages flowing
+through the links they are attached to, and store them into
+output trace files.
 
@@
 For this model, we consider the flow specifications
-describing the cache coherence protocols supported in GEM5\footnote{The GEM5 cache
-  coherence protocols can be found at
+describing the cache coherence protocols supported in
+GEM5\footnote{The GEM5 cache coherence protocols can be
+  found at
   \url{http://www.gem5.org/docs/html/gem5MemorySystem.html}.}
@@
 Three such flows describe the cache coherent protocols for
-each CPU.  As there are two CPUs in this model, there are
-actually six such flows considered for this model.
+each CPU.  Since there are two CPUs, there are six flows in
+the model.
 
@@
 
-Our work is closely related to communication-centric and transaction based debug.
-An early pioneering work is described in~\cite{Goossens2007NOCS}, which advocates the focus on observing activities on the interconnect network among IP blocks, and mapping these activities to transactions for better correlation between computations and communications.
-Therefore, the communication transactions, as a result of software execution, provide an interface between computation and communication, and facilitate  system-level debug.
-This work is extended in~\cite{Vermeulen2009VLSI-DAT,Goossens2009DATE}.
-However, this line of work is focused on the network-on-chip (NoC) architecture for interconnect using the run/stop debug control method.
-
-A similar transaction-based debug approach is presented in~\cite{Gharehbaghi2012ISQED}.
-Furthermore, it proposes an automated extraction of state machines at transaction level from  high level design models.
-From an observed failure trace, it performs backtracking on this transaction level state machine to derive a set of transaction traces that lead to the observed failure state.
-In the subsequent step, bounded model checking with the constraints on the internal variables is used to refine the set of transaction traces to remove the infeasible traces.
-This approach requires user inputs to identify impossible transaction sequences, and may not find the states causing the failure if the transaction traces leading to the observed failure state is long.
-Backtracking from the observed failure state requires pre-image computation, which can be computationally expensive.
-A transaction-based online debug approach is proposed in~\cite{Dehbash2014} to address these issues.
-This approach utilizes a transaction debug pattern specification language~\cite{Gharehbaghi2009ICCD} to define properties that transactions should meet.
-These transaction properties are checked at runtime by programming debug units in the on-chip debug infrastructure, and the system can be stopped shortly after a violation is detected for any one of those properties.
-In this sense, it can be viewed as the hardware assertion approaches in~\cite{Boule2007ISQED} elevated to the transaction level.
-
-In~\cite{Singerman2011DAC}, a coherent workflow is described where the result from the pre-silicon validation stage can be carried over to the post-silicon stage to improve efficiency and productivity of post-silicon debug.
-This workflow is centered on a repository of system events and simple transactions defined by architects and IP designers.
-It spans across a wide spectrum of the post-silicon validation including DFx instrumentation, test generation, coverage, and debug.
-The DFx instruments are automatically inserted into the design RTL code driven by the defined transactions.
-This instrumentation is optimized for making a large set of events and transactions observable.
-Test generation is also optimized to generate only the necessary but sufficient tests to allow all defined transactions to be exercised.
-Moreover, coverage for post-silicon validation is now defined at the abstract level of events and transactions rather than the raw signals, and thus can be evaluated more efficiently.
-In~\cite{Abarbanel2014DAC}, a model at an even higher-level of abstraction, {\em flows}, is proposed.
-Flows are used to specify more sophisticated cross-IP transactions such as power management, security, etc, and to facilitate reuse of the efforts of the architectural analysis to check HW/SW implementations.
+Our work is closely related to communication-centric and
+transaction based debug.  An early pioneering work is
+described in~\cite{Goossens2007NOCS}, which advocates the
+focus on observing activities on the interconnect network
+among IP blocks, and mapping these activities to
+transactions for better correlation between computations and
+communications.  Therefore, the communication transactions,
+as a result of software execution, provide an interface
+between computation and communication, and facilitate
+system-level debug.  This work is extended
+in~\cite{Vermeulen2009VLSI-DAT,Goossens2009DATE}.  However,
+this line of work is focused on the network-on-chip (NoC)
+architecture for interconnect using the run/stop debug
+control method.
+
+A similar transaction-based debug approach is presented
+in~\cite{Gharehbaghi2012ISQED}.  Furthermore, it proposes an
+automated extraction of state machines at transaction level
+from high level design models.  From an observed failure
+trace, it performs backtracking on this transaction level
+state machine to derive a set of transaction traces that
+lead to the observed failure state.  In the subsequent step,
+bounded model checking with the constraints on the internal
+variables is used to refine the set of transaction traces to
+remove the infeasible traces.  This approach requires user
+inputs to identify impossible transaction sequences, and may
+not find the states causing the failure if the transaction
+traces leading to the observed failure state is long.
+Backtracking from the observed failure state requires
+pre-image computation, which can be computationally
+expensive.  A transaction-based online debug approach is
+proposed in~\cite{Dehbash2014} to address these issues.
+This approach utilizes a transaction debug pattern
+specification language~\cite{Gharehbaghi2009ICCD} to define
+properties that transactions should meet.  These transaction
+properties are checked at runtime by programming debug units
+in the on-chip debug infrastructure, and the system can be
+stopped shortly after a violation is detected for any one of
+those properties.  In this sense, it can be viewed as the
+hardware assertion approaches in~\cite{Boule2007ISQED}
+elevated to the transaction level.
+
+In~\cite{Singerman2011DAC}, a coherent workflow is described
+where the result from the pre-silicon validation stage can
+be carried over to the post-silicon stage to improve
+efficiency and productivity of post-silicon debug.  This
+workflow is centered on a repository of system events and
+simple transactions defined by architects and IP designers.
+It spans across a wide spectrum of the post-silicon
+validation including DFx instrumentation, test generation,
+coverage, and debug.  The DFx instruments are automatically
+inserted into the design RTL code driven by the defined
+transactions.  This instrumentation is optimized for making
+a large set of events and transactions observable.  Test
+generation is also optimized to generate only the necessary
+but sufficient tests to allow all defined transactions to be
+exercised.  Moreover, coverage for post-silicon validation
+is now defined at the abstract level of events and
+transactions rather than the raw signals, and thus can be
+evaluated more efficiently.  In~\cite{Abarbanel2014DAC}, a
+model at an even higher-level of abstraction, {\em flows},
+is proposed.  Flows are used to specify more sophisticated
+cross-IP transactions such as power management, security,
+etc, and to facilitate reuse of the efforts of the
+architectural analysis to check HW/SW implementations.
 
@@
 
-This paper presents a method for post-silicon validation by interpreting observed raw signal traces at the level of system flow specifications.
-The derived flow execution scenarios provide more structured information on system operations, which is more understandable to system validators.
-This information can help to locate design defects more easily, and also provides a measurement of validation coverage.
-
-Due to partial observability, this approach may derive a large number of different flow execution scenarios for a given signal trace.
-Insights from system validators can help to eliminate some false scenarios due to the partial observability.
-An interesting future direction is formalization of the validators' insights using temporal logic on flows so that the validators can express their intents more precisely and concisely.
-
-The trace analysis approach presented in this paper needs to be iterated with different observations selected in different iterations in order to eliminate the false scenarios and to root cause system failures as quickly as possible.
-The observation selection and stitching signal traces of different observations together for the above goal will also be pursued in the future.
+This paper presents a method for post-silicon validation by
+interpreting observed raw signal traces at the level of
+system flow specifications.  The derived flow execution
+scenarios provide more structured information on system
+operations, which is more understandable to system
+validators.  This information can help to locate design
+defects more easily, and also provides a measurement of
+validation coverage.
+
+Due to partial observability, this approach may derive a
+large number of different flow execution scenarios for a
+given signal trace.  Insights from system validators can
+help to eliminate some false scenarios due to the partial
+observability.  An interesting future direction is
+formalization of the validators' insights using temporal
+logic on flows so that the validators can express their
+intents more precisely and concisely.
+
+The trace analysis approach presented in this paper needs to
+be iterated with different observations selected in
+different iterations in order to eliminate the false
+scenarios and to root cause system failures as quickly as
+possible.  The observation selection and stitching signal
+traces of different observations together for the above goal
+will also be pursued in the future.
 

### 2015-12-28 paper2.tex
Done minor edits

%% bare_conf.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compSoC.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference]{IEEEtran}
% Add the compSoC option for Computer SoCiety conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.
\usepackage{pifont}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{mybeamer}
\usepackage{draftwatermark}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

\newcommand{\eg}{\mbox{{\em e.g.}}}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical SoCiety that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{\Large {\bf Flow-guided Analysis of System-level Post-silicon
  Traces Under Limited Observability}}

\author{\large Hao Zheng$^1$, Yuting Cao$^1$, Sandip Ray$^2$, Jin Yang$^2$ \\
$^1$Dept. of Computer Science and Engg., University of South Florida, Tampa, Fl 33620. USA. \\
$^2$Strategic CAD Labs, Intel Corporation, Hillsboro, OR 97124.  USA.}
\maketitle

%% % author names and affiliations
%% % use a multiple column layout for up to three different
%% % affiliations
%% \author{\IEEEauthorblockN{Hao Zheng, Yuting Cao}
%% \IEEEauthorblockA{Computer Science and Engineering\\
%% University of South Florida\\
%% Tampa, Florida 33620}
%% \and
%% \IEEEauthorblockN{Sandip Ray, Jin Yang}
%% \IEEEauthorblockA{Strategic CAD Labs\\
%% Intel Corporation\\
%% Portland, Oregon
%% }}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
%\maketitle

\begin{abstract}
Validation of multicore systems-on-chip (SoC) is very challenging, and main reasons include the limited internal operation observability and non-determinism due to multiple clock domains.   Very often, the observed signal traces offer limited values for debuggers to understand system internal behavior.  This paper presents a trace analysis approach for post-silicon validation.  This approach takes system flow specifications describing system behavior at a high level and an observed signal trace, and interprets the trace with respect to the system flow specification.  The result from the trace interpretation is a set of flow instances such that their execution during a test explains the observed signal trace.   This approach transforms partially observed and non-deterministic signal traces to a representation that is more understandable, and offers more helpful information to locate design defects and a measure to evaluate validation coverage.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
% \IEEEpeermaketitle

\section{Introduction}
% no \IEEEPARstart
%
%Post silicon debug is getting more challenging for current complex system-on-chip design.
%
%Using system level flows in silicon trace analysis helps to reduce the volume of data to be stored and analyzed, difficulty caused by inherent non-determinism.  
%
%Also, abstracts signal traces to system use cases, which helps validation engineering better understand the behavior displayed by the signal traces.

Post-silicon validation makes use of pre-production silicon
integrated circuit (IC) to ensure that the fabricated system
works as desired under actual operating conditions with real
software.  Since the silicon executes at target clock speed,
post-silicon executions are billions of times faster than
RTL simulations, and even provide speed-up of several orders
of magnitude over other pre-silicon platforms (\eg, FPGA,
system-level emulation, etc.).  This makes it possible to
explore deep design states which cannot be exercised in
pre-silicon, and identify errors missed during pre-silicon
validation and debug.  Post-silicon validation is a critical
component of the design validation life-cycle for modern
microprocessors and SoC designs.  Unfortunately, it is also
a highly complex component, performed under aggressive
schedules and accounting for more than $50\%$ of the overall
design validation cost.  Consequently, it is crucial to
develop techniques for streamlining and automating
post-silicon validation activities.

A central component of post-silicon validation of SoC
designs is to correlate trace from silicon execution with
the intended system-level transactions.  An SoC design is
typically composed of a large number of pre-designed
hardware or software blocks (often referred to as
``intellectual properties'' or ``IPs'') that coordinate
through complex protocols to implement system-level
behavior.  Any execution trace of the system involves a
large number of interleaved instances of these protocols.
For example, consider a smartphone executing a usage
scenerio where the end-user browses the Web while listening
to music and sending and receiving occasional text messages.
Typical post-silicon validation use-case involve exercising
such scenarios.  An execution trace for this scenario would
involve activities from the CPU, audio controller, display
controller, wireless radio antenna, etc., reflecting the
interleaved execution of several communication protocols.
On the other hand, due to observability limitations, only a
small small number of participating signals can be actually
traced during silicon execution.  Furthermore, due to
electrical perturbations, silicon data can be noisy, lossy,
and ambiguous.  Consequently, it is non-trivial to identify
all participating protocols and pinpoint the ``right''
interleaving that results in an observed trace.

In this paper, we consider the problem of reconstructing
protocol-level behavior from silicon traces in SoC designs.
More specifically, given a collection of system-level
communication protocols and a trace of (partially observed)
hardware signals, our approach infers, with a certain
measure of confidence, the protocol instances (and their
interleavings) being exercised by the trace.  We show how to
use symbolic simulation and

Systems-on-chips (SoCs) have become more complex as more functionalities are integrated on chips.  To reduce design time, more external IPs are integrated into current SoC designs. At the high level, a SoC can be viewed as a network of IP blocks communicating by a large number of protocols.  A typical SoC design consists of multiple programmable processor cores where a number of software threads can be executed simultaneously.  The resulting designs usually operate in multiple clock domains, and show a high degree of concurrency.  Due to the sophisticated communication mechanisms employed in  current SoC designs to support complex functional behaviors, traditional processor-centric validation has to be enhanced to focus inter-core communications.

To make validation even more challenging, the high degree of integration in current SoC designs causes traditional chip observation points to become inaccessible.  A SoC design usually contains millions of signals.  In post-silicon validation, thousands of those signals are tapped for observation.  Due to the limited availability of the chip pins, only about a hundred of the tapped signals can be traced.   This severely limits observability hinders post-silicon debuggers' capability to root cause an observed failure.  To make the situation even worse, the observed traces often display non-deterministic behavior due to many concurrent operations occurred in multiple clock domains in a design.  The above factors make debugging highly difficult in failure root-causing, and need to be addressed.

In a typical design flow, a design starts with some high-level specification, particularly for communication protocols.   The abstract protocol specification usually captures information exchanged among IP blocks and how it is exchanged to support functional behavior. 
Graphical representations of protocols such as Message Sequence Charts~\cite{Harel03} or Live Sequence Charts~\cite{Damm99} describe protocols intuitively.  However, its success in hardware industry is limited.  Recently, message flows in BPMN~\cite{Krstic14HOST} have been being advocated as a formalism for specifying architecture requirements
 or system-level use cases such as power management, security, etc.  In general, a flow describes a set of message sequences across multiple IP blocks following a system communication protocol.  Specification of system flows BPMN is intuitive and easier for human to understand and to reason about system behavior.  An example of flow specification in BPMN is shown in Figure~\ref{flow-spec-ex}.

\begin{figure}[tb]
\begin{center}
\includegraphics[height=2.5in]{figures/bpmn-flow-ex}
\caption{A graphical representation of a SoC firmware load protocol in BPMN~\cite{Krstic14HOST}.}
\label{flow-spec-ex}
\end{center}
\end{figure}

As indicated above, interpreting and understanding the raw signal traces observed from executing the chip is extremely difficult due to the limited observability and non-determinism.  This paper proposes a trace analysis method which analyzes and interprets the observed signal traces at the level of system flow specifications.  During the step of analysis and interpretation, this method tries to find out possible executions of system flows such that the observed signal trace is a result of such executions.  The resulting scenarios of flow executions often give debuggers more clear and better structured system-level views of the internal operations of the system under debug.  With a better understanding of system operations in a test environment, therefore, it would make easier for debuggers to pinpoint and root cause system failures.  In addition to debugging, another important benefit is better coverage evaluation as a scenario of flow executions captures the information about the types and number of instances of flows executed by the chip in a test environment.  That information can help to determine the quality and coverage of validation.

%---
\section{Labeled Petri-Nets}

Although system flows depicted in BPMN as in Fig.~\ref{flow-spec-ex} are intuitive for system architects to understand communications across multiple components, they are not suitable for algorithmic analysis.  In this section, we present an alternative formalism, \emph{labeled Petri Nets}, to represent communication aspects of system flows with the intuitive descriptions of internal operations for IP blocks abstracted away whenever possible.  Labeled Petri nets are capable of describing sequencing, concurrency, and choices.   This formalism has well defined operational semantics, efficient analysis analysis algorithms and and has been used widely in modeling and analyzing communication protocols, concurrent programs, etc.

%--- LPN definition with levels
%Specifically, a labeled Petri-net (LPN) is a tuple $(P, T, M_0, V, E, L)$ where
%\begin{itemize}
%\item $P$ is a finite set of places,
%\item $T$ is a finite set of transitions,
%\item $M_0$ is the set of initially marked places, also referred to as the initial marking.
%\item $V$ is a finite set of auxiliary variables of integer type,
%\item $E$ is a finite set of events,
%\item $L$ is a labeling function that maps each transition $t \in T$ to a triple $(g, e, A)$ where
%	\begin{itemize}
%	\item $g$ is a predicate over $V$,
%	\item $e \in E$ is an event,
%	\item $A$ is a set of assignments to some variables in $V$.  
%	\end{itemize}
%\end{itemize}
%
%For each transition $t \in T$, its preset, denoted as $\bullet{t} \subseteq P$, is the set of places connected to $t$, and its postset, denoted as $t\bullet \subseteq P$, is the set of places that $t$ is connected to.  A marking of a LPN is a set of places marked with tokens, and a state of a LPN is $(M, \alpha)$ where $M$ is a marking and $\alpha$ is an assignment to all auxiliary variables in $V$.  
%
%The operational semantics of a LPN is defined by transition executions.  A transition can be executed after it is {\em enabled}.  A transition $t = (g, e, A)$ is enabled in a state $(M, \alpha)$ if every place in its preset is included in the marking, i.e. $\bullet{t} \subseteq M$, and the values of the auxiliary variables make the predicate $g$ be true, i.e. $\alpha \models g$.  Execution of $t$ results in a new state $(M', \alpha')$ such that 
%\[
%M' = (M - \bullet{t}) \cup t\bullet,
%\]
%and $\alpha'$ is a new assignment where the values of the auxiliary variables are updated with respect to the assignments $A$.  Furthermore, when $t$ is executed, the event $e$ is emitted.  SoC

Specifically, a labeled Petri-net (LPN) is a tuple $(P, T, s_0, E, L)$ where
\begin{itemize}
\item $P$ is a finite set of places,
\item $T$ is a finite set of transitions,
\item $\mathit{init}$ is the set of initially marked places, also referred to as the initial marking.
\item $E$ is a finite set of events,
\item $L: T \rightarrow E$ is a labeling function that maps each transition $t \in T$ to an event $e \in E$.
\end{itemize}

For each transition $t \in T$, its preset, denoted as $\bullet{t} \subseteq P$, is the set of places connected to $t$, and its postset, denoted as $t\bullet \subseteq P$, is the set of places that $t$ is connected to.  A marking $s \subseteq P$ of a LPN is a subset of places marked with tokens, and it is also referred to as a state of a LPN.  The initial marking $\mathit{init}$ is also the initial state of the LPN.

The operational semantics of a LPN is defined by transition executions.  A transition can be executed after it is {\em enabled}.  A transition $t \in T$ is enabled in a state $s$ if every place in its preset is included in the marking, i.e. $\bullet{t} \subseteq s$.  Execution of $t$ results in a new state $s'$ such that 
\[
s' = (s - \bullet{t}) \cup t\bullet,
\]
and the emission of event $e$ labeled for $t$.

A simple communication protocol specification between two components represented by a LPN is shown in~Figure~\ref{fig-flow-ex-1}.  In this and the following figures for LPNs, the labeled circles denote places, and the labeled boxes denote transitions.  Each transition is labeled with its name and the asSoCiated event.  The solid black places without outgoing edges are {\em terminals}, which indicate termination of protocols represented by the LPNs.  In this LPN, the initial state $\mathit{init} = \{p_1\}$.  In this protocol, component~1 sends two consecutive messages $\mathit{msg}_1$ to component~2, which responds with either $\mathit{msg}_2$ or $\mathit{msg}_3$ non-deterministically.  If $\mathit{msg}_2$ is responded, component~1 repeats the protocol by sending two consecutive $\mathit{msg}_1$ again.  Otherwise, component~1 stops the protocol upon receiving $\mathit{msg}_3$.  This LPN represents two possible communication sequences as shown below.
\[
\begin{array}{l}
\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_1\;\ldots \\
\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\ldots, \mathit{msg}_3\;
\end{array}
\]

\begin{figure}[tb]
\begin{center}
\resizebox{2.5in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(t1)	at	($(p1) + (0,-1.1)$) {$t_1:msg_1$};
	\node[circ] 	(p2) 		at ($(t1) + (0,-1.1)$) {$p_2$};
	\node[rectbox]	(t2)	at	($(p2) + (0,-1.1)$) {$t_2:msg_2$};
	\node[circ] 	(p3) 		at ($(t2) + (0,-1.1)$) {$p_3$};
	\node[rectbox]	(t3)	at	($(p3) + (3,0)$) {$t_3:msg_3$};
	\node[rectbox]	(t4)	at	($(p3) + (-3,0)$) {$t_4:msg_4$};
	\node[circ,fill=black] 	(term) 		at ($(t4) + (0,-1.1)$) {$~~~$};

	% Edges
	\draw[->,weight3]	(p1) -- (t1);
	\draw[->,weight3]	(t1) -- (p2);
	\draw[->,weight3]	(p2) -- (t2);
	\draw[->,weight3]	(t2) -- (p3);
	\draw[->,weight3]	(p3) -- (t3);
	\draw[->,weight3]	(p3) -- (t4);
	\draw[->,weight3]	(t4) -- (term);	
	\draw[->,weight3]	(t3) -- ($(p1)+(3,0)$) -- (p1);

\end{tikzpicture}
}
\caption{An example of a LPN representing a system flow.}
\label{fig-flow-ex-1}
\end{center}
\end{figure}

% @COMMENT This example is for LPNs with guards.
%The above protocol specification can be made more precise by describing how component~2 would respond without using non-determinism.  For example, the system architect may wish to specify that component~2  responds with $\mathit{msg_2}$ to every sequence of two $\mathit{msg_1}$ for five such sequences in a row.  On the sixth such sequence, component~2 responds with $\mathit{msg_3}$.  The LPN modeling such protocol is shown in Figure~\ref{fig-flow-ex-2}.  An alternative without using auxiliary variables in the transition predicates is by unrolling the loops for the specified number of times; however this would result in a larger and more complex LPN.  
%
%\begin{figure}[htbp]
%\begin{center}
%\includegraphics[height=1.6in]{figures/flow-ex-2}
%\caption{Another example of a LPN representing the system flow in Fig.~\ref{fig-flow-ex-1} with a fixed number of message exchanges.}
%\label{fig-flow-ex-2}
%\end{center}
%\end{figure}

%=====
\section{Flow-Directed Trace Analysis}

In a typical validation setting, the system under debug (SUD) is executed in a test environment until it is terminated by the test environment or the system crashes due to a failure.  During the execution, a trace on a small number of observable signals is streamed off the chip for debugging.  Due to the limited observability and inherent non-determinism in today's SoC designs, the observed signal trace is difficult to understand, thus providing limited values for debugging.  In this section, we describe a trace analysis method where the observed signal traces are interpreted at the level of system flows.  In general, the trace analysis can offer debuggers a structured view of communications among the IP blocks during the SUD execution by deriving the types and numbers of system flows activated during SUD executions from the observed signal traces.

The trace analysis method consists of two steps: \emph{trace abstraction} and \emph{interpretation}.  The trace abstraction takes a signal trace observed during the execution of the SUD, and abstracts it with respect to information captured in the system flows.  This requires recognizing signal events and mapping the signal events or sequences of the signal events to flow events appeared in system flow specifications.  For example, a data write message may be a single event in a system flow, however, such a flow event maybe implemented in hardware as a sequence of events including a header, a number of data word transfers and a tail.  This abstraction requires a mapping relation from flow events to sequences of signal events, and it is reasonable to assume that this relation is available.

The trace interpretation takes a finite trace of flow events resulting from the trace abstraction and a set of system flows in LPNs $\vec{F}$, and generates a set of possible system flow execution scenarios.  A \emph{flow execution scenario} is defined as $\{(F_{i,j}, s_{i,j})\}$ where in each element $(F_{i,j}, s_{i,j})$, $F_{i,j}$ is the $j$th activated instance of flow $F_i \in \vec{F}$, and $s_{i,j}$ is a state of $F_{i,j}$.  A flow execution scenario indicates that at a certain point of SUD execution, what types of flows and the number of instances of a particular flow are activated and their corresponding current states.

An observed trace of flow events $\rho = e_1e_2\ldots e_n$ is a result of SUD executing the flow instances of some flow execution scenario.  The goal of the trace interpretation is to derive such scenario.  Let $\mathit{accept(F_{i,j}, s_{i,j}, e)}$ be a function that determines if event $e$ can be emitted by $F_{i,j}$ in state $s_{i,j}$.  This function is used during the trace interpretation for checking whether an event of an observed trace is a result of executing some flow instance.  Formally,  $\mathit{accept(F_{i,j}, s_{i,j}, e)}$ returns $(F_{i,j}, s'_{i,j})$ if there exists a transition $t$ in $F_i$ such that $L(t) = e$ and $\bullet t \subseteq s_{i,j}$.  In this case, $s'_{i,j} = (s_{i,j} - \bullet t) \cup t\bullet$.  It returns $\emptyset$ if no such $t$ exists in $F_i$.

Given a trace of flow events $\rho = e_1e_2\ldots e_n$, the trace interpretation algorithm starts with an empty set of of flow execution scenario $\mathit{Scen} = \emptyset$.  Then, for each $e_h$ where $1 \leq h \leq n$ starting $h=1$, and for each $\mathit{scen} \in \mathit{Scen}$, the following two steps are performed. 
\begin{description}
\item[Step 1]~~For each $(F_{i,j}, s_{i,j}) \in \mathit{scen}$, if $\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = (F_{i,j}, s'_{i,j})$, create a new scenario $\mathit{scen}' = (\mathit{scen} - (F_{i,j}, s_{i,j})) \cup (F_{i,j}, s'_{i,j})$, which is added into $\mathit{Scen}'$.

\item[Step 2]~~For each $F_i \in \vec{F}$, create a new instance $F_{i, j+1}$.  If $\mathit{accept}(F_{i,j+1}, \mathit{init}_{i,j+1}, e_h) = (F_{i,j+1}, s'_{i,j+1})$,  create a new scenario $\mathit{scen}' = \mathit{scen} \cup (F_{i,j+1}, s'_{i,j+1})$, which is added into $\mathit{Scen}'$.
\end{description}
After $e_h$ is processed, $Scen = Scen'$, and the above two steps repeat for the next event $e_{h+1}$.

If every events in $\rho$ is successfully mapped to some flow instance, this algorithm returns a set of flow execution scenarios such that every flow instance is in its terminal state.  On the other hand, inconsistent events can also be encountered.  An event $e$ is \emph{inconsistent} if for each flow execution scenario $\mathit{scen} \in \mathit{Scen}$, the following two conditions hold.
\begin{enumerate}
\item For each $(F_{i,j}, s_{i,j}) \in \mathit{scen}$, $\mathit{accept}(F_{i,j}, s_{i,j}, e_h) = \emptyset$,
\item For each $F_i \in \vec{F}$, $\mathit{accept}(F_{i}, \mathit{init}_{i}, e_h) = \emptyset$.
\end{enumerate}
An inconsistent event is the one produced by SUD execution but cannot be mapped to any flow instances no matter how the trace prior to event $e$ is interpreted.  Inconsistent events indicates possible causes of system failures.

Based on the above discussion, the trace interpretation algorithm returns two pieces of information: 1) a set $G$ of flow execution scenarios where every flow instance in every scenario is in its terminal state, 2) a set $B$ of pairs, each of which includes a set of flow execution scenarios and an inconsistent event.  The set $B$ provides valuable information for debuggers to root cause system failures.  One of these two sets can empty.   With the full observability, the set $G$ includes a single flow execution scenario derived for a trace.  In reality, it is always the case that the SUD is only partially observable.  Therefore, due to the lack of information for precise interpretation, a set of flow execution scenarios is typically derived for a given trace as the result of the trace analysis.

% @Note  \red{\bf Another assumption:  the observed signal trace captures the values on the observed signals in every cycle from the beginning.  Too strong an assumption??}

To illustrate the basic idea of the trace analysis method, consider the system flow shown in Figure~\ref{fig-flow-ex-1}.  Let $F_1$ denote such flow.  Suppose that a hardware system implements flow $F_1$, and the following trace of flow events is abstracted from an observed signal trace as a result of executing such system.     
\[
\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_1\;\mathit{msg}_2\;\mathit{msg}_3\;\ldots
\]  
This trace is interpreted from the first event to the last in order to derive all possible flow execution scenarios.  At the beginning, the first event $\mathit{msg}_1$ is processed.  According to the flow specification $F_1$, we know that one instance of such flow $F_1$, $F_{1,1}$,  is activated by the SUD as $\mathit{accept}(F_{1,1}, init_1, \mathit{msg}_1) = (F_{1,1}, \{p_2\})$ where $\mathit{init} = \{p_1\}$ is the initial state of $F_1$.  As the result, the flow execution scenario after interpreting the first event $\mathit{msg_1}$ is $\{(F_{1,1}, \{p_2\})\}$.

Next, the second $\mathit{msg}_1$ is interpreted on both scenarios   This event could be the result of two possible cases.  In the first case, this event is the result of the continuing execution of $F_{1,1}$ as $\mathit{accept}(F_{1,1}, \{p_2\}, \mathit{msg}_1) = (F_{1,1}, \{p_3\})$.  In the second case, the system may activate another instance of $F_1$, $F_{1,2}$ such that the second event $\mathit{msg_1}$ is a result of executing this new instance.  Therefore, the interpretation of the first two events $\mathit{msg}_1$ leads to two flow execution scenarios as shown below.
\[
\label{tr-analy-s1}\tag{1}
\begin{array}{cl}
1. & \{(F_{1,1}, \{p_3\})\},\\
2. & \{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_2\})\} 
\end{array}
\]

Now, consider the third $\mathit{msg}_1$ for each of the two scenario derived in~(\ref{tr-analy-s1}).    For the execution scenario~1, $F_{1,1}$ is not able to accept $\mathit{msg}_1$ as it is in state $\{p_3\}$.  On the other hand, this event could be the result of activation of a new flow instance.  Therefore, this execution scenario can be revised accordingly as 
\[
\label{tr-analy-s2}\tag{2}
\{(F_{1,1}, \{p_3\}),(F_{1,2},\{p_2\})\}.
\]   
For the execution scenario~2, event $\mathit{msg}_1$ could be the result of continuing execution of $F_{1,1}$ or $F_{1,2}$, or it could be a result of activation of a new flow instance.  Therefore,  three new execution scenarios can be derived as shown below for this event.
\[
\label{tr-analy-s3}\tag{3}
\begin{array}{l}
\{(F_{1,1}, \{p_3\}), (F_{1,2}, \{p_2\})\},\\
\{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_3\})\}, \\
\{(F_{1,1}, \{p_2\}), (F_{1,2}, \{p_2\}), (F_{1,3}, \{p_2\})\}
\end{array}
\]
Since the flow execution scenario in~(\ref{tr-analy-s2}) already exists in~(\ref{tr-analy-s3}), the three flow execution scenarios shown in~(\ref{tr-analy-s3}) is the result from the interpretation of the first three events $\mathit{msg}_1$.

The next flow event in the trace $\mathit{msg_2}$ is analyzed for the flow execution scenarios as shown in~(\ref{tr-analy-s3}).   For the first scenario $\{(F_{1,1}, \{p_3\}), (F_{1,2}, \{p_2\})\}$, event $\mathit{msg_2}$ can only be the result from executing $F_{1,1}$ as $\mathit{accept}(F_{1,1}, \{p_3\}, \mathit{msg_2}) = (F_{1,1}, \{p_1\})$.   Based on the same reasoning, this event can only be the result from executing $F_{1,2}$ in the second scenario, and it moves $F_{1,2}$ to a new state $\{p_3\}$ too.  The interesting case is the third scenario where none of the flow instances can allow $\mathit{msg_2}$ to happen.  This is due to the fact that the flow must be in $\{p_3\}$ for $\mathit{msg_2}$ to happen.  This indicates the third system execution scenario is impossible for the prefix of the flow event trace upto $\mathit{msg_2}$, therefore this scenario is ignored from further analysis.  After analyzing event $\mathit{msg_2}$, the updated system execution scenarios are shown below.
\[
\begin{array}{l}
\{(F_{1,1}, \{p_1\}), (F_{1,2}, \{p_2\})\},\ \{(F_{1,1}, \{p2\}), (F_{1,2}, \{p_1\})\}.
\end{array}
\]

The last flow event in the trace is $\mathit{msg}_3$.  Considering the above two possible system executions, neither can allow this event to be produced as none of the flow instances in both system executions is in state $\{p_3\}$.  What this means is that the system does {\em not} implement the flow specification correctly as it produces something not allowed by the specification.  By examining the system executions right before the ``buggy'' event, debuggers may gain more information on when and where the problem might be.  The trace interpretation algorithm adds these two scenarios along with the fifth event $\mathit{msg_3}$ into $B$, and returns it for debuggers to analyze.

%---
\subsection{Trace Analysis with Partial Observability}

In hardware that implements a given system flow specification, a flow event is defined as an event or a sequence of events on a set of signals.  Due to the limited number of pins on the boundary of chips available for observation, only a small fraction of system signals can be observed during debug.  In this section, we discuss how the trace analysis method presented above can be adapted to deal with signal traces of partial observability.

In general, a signal trace of partial observability corresponding a set of traces of flow events due to the ambiguous interpretation of signal events.  In the following, we discuss two cases for trace abstraction on partial observability: mapping a single signal event to a flow event or mapping a sequence of signal events to a flow event.  A signal event is defined as a state on or an assignment to a set of signals.

Hereafter, the term {\em flow traces} is used to refer to traces of flow events. 
Consider the following example for the first case.  Suppose that there are three flow events: $e_1$, $e_2$, and $e_3$, which are implemented in hardware by the signal events shown in the list below.  We use Boolean expressions to represent signal events for the discussion.
\[
\begin{array}{cl}
e_1: & abc\\
e_2: & \bar{a}bc\\
e_3: & a\bar{b}c
\end{array}
\] 
Now suppose that only signals $b$ and $c$ are observable, and a signal trace of this partial observability is obtained below.
\[
bc\ bc \ \bar{b}c
\]
During the trace abstraction step, the first two signal events $bc$ can be mapped to $\{e_1, e_2\}$ since $a$ is not observable, and the last one $b'c$ is mapped to $\{e_3\}$.  Therefore, this signal trace is abstracted to four flow traces, $\{e_1, e_2\} \times \{e_1, e_2\} \times \{e_3\}$.

Next, we consider the case where a flow event is mapped from a sequence of signal events.  Now suppose that two other flow events are implemented by sequences of signal events as defined in the list below.
\[
\begin{array}{cl}
e_4: & abc\ \bar{a}bc\\
e_5: & abc\ abc\ abc\ \bar{a}bc\\
\end{array}
\] 
Given an observed trace of the same observability shown below
\[
bc\ bc\ bc \ bc,
\]
it is abstracted to the following flow traces.
\[
e_4e_4,\; \textvisiblespace e_4 \textvisiblespace,\;  e_4 \textvisiblespace \textvisiblespace,\;  \textvisiblespace \textvisiblespace e_4, \; e_5
\] 
where $\textvisiblespace$ denotes signal events that are not mapped to any flow events.  Note that the above abstraction leads to three distinct flow traces as the middle three correspond to the same trace of flow events.

From the above discussions, it can be seen that a signal trace of partial observability is generally mapped to a set of flow traces.  As the the number of signals available for observation becomes smaller, the number of flow event traces corresponding to a signal trace as a result of the abstraction can be enormous.  This phenomenon can increase the complexity of the trace interpretation as it can cause the number of possible flow execution scenarios generated during the analysis to explode.   Possible solutions to address this issue include better trace signal selection for observation and assistance from debuggers' insights.  Trace signal selection itself is an important and difficult subject, and a detailed discussion of it is out of scope of this paper.   Next, we briefly describe how the debuggers' insights of a system's architecture can help to address the complexity issue in the trace analysis.

\subsection{Inputs from Validators}

During the trace interpretation, the number of intermediate flow execution scenarios may become too large due to ambiguous interpretation from signal events to flow events or from flow events to flow execution scenarios.  The explosion of the intermediate results can significantly slow down the performance of the trace analysis.  To address this problem, the debuggers can use their insights and understanding of the SUD to trim the total number of possible system executions derived from the trace analysis.  When a SUD is validated, a debugger is assumed to have deep knowledge about the system's architecture and microarchitecture, and how the test environment affects the system executions.  For instance, he/she may have knowledge that in a test environment, the maximal number of instances of a flow can be activated by the SUD, or a flow can be activated after certain other flows have terminated, etc.  These debugging insights can be encoded as constraints into the trace analysis method, which can then be used to eliminate a large number of flow execution scenarios that violate these constraints during the trace interpretation step.

This approach can be flexible in that it allows a debugger to analyze the observed traces in a trail-and-error manner if the precise knowledge of the system (micro-)architecture is hard to come by.  For instance, the debugger might initially make a very restricted assumption on how the SUD executes a flow specification, and these assumptions can potentially lead to an empty set of flow execution scenarios.  Depending on which of these assumptions triggered during the trace interpretation step, the debugger can study these assumptions more carefully, and relax some or all of them for the next run of analysis.  This iteration can be repeated as many times as necessary until some results deemed meaningful are produced.

Alternatively, if all derived execution scenarios seem to be plausible, the implication that a debugger may draw from this result is that the failure may be independent of the flows being observed.  Therefore, the testing environment can be adjusted in order for a different part or different behavior of the SUD to be observed.  This idea, closely related to trace signal selection, is critical for post-silicon validation, and a detailed discussion can only be presented in a separate paper.

%=====
\section{Case Study}

The proof concept of the proposed trace analysis method is demonstrated on a transaction level model of a simple SoC design with two CPU cores, memory, and a peripheral device connected by an interconnect as shown in Fig.~\ref{fig-SoC-ex}.  Since the analysis method presented in this paper is communication centric, the detailed computations of these blocks are not modeled.  Instead, the modeling is focused on how they participate in flows for different system level use cases.

\begin{figure}[tb]
\begin{center}
\includegraphics[width=2in]{figures/simple-ex}
\caption{A simple SoC example.}
\label{fig-SoC-ex}
\end{center}
\end{figure}

\subsection{Flow Specification}

In this case study, four system flows are implemented in the simple SoC model. They include cache coherent memory access operations and a memory-mapped peripheral read operation initiated from the CPUs, a message signaled interrupt operation initiated from the peripheral device.  These flow specifications capture how messages are exchanged for different use cases.  In this model, a message is defined with the following format, $(\mbox{Src}, \mbox{Dest}, \mathit{Cmd}, \mathit{Addr})$, where Src and Dest refer to the source and destination components of messages, $\mathit{Cmd}$ refers to the operations that the destination component should perform, and $\mathit{Addr}$ refers to memory addresses where $\mathit{Cmd}$ applies.  $\mathit{Cmd}$ can be memory accesses or not.  If it is not, then the $\mathit{Addr}$ field of messages is ignored.  In this case study, memory mapped IO mechanism is used.  Furthermore, detailed memory addresses are not modeled. Instead, the address space is partitioned to main memory addresses and peripheral addresses.  In messages, the $\mathit{Addr}$ field is replaced with either $M$ representing a memory address or $P$ representing an address to a peripheral device.

The LPN as shown in Fig.~\ref{fig-cpu-write} specifies a system flow where {CPU2} initiates a memory write operation.   In this flow, CPU2 initiates a memory write request followed by a data valid message to the interconnect.  The data valid messages are used to model availability or validity of data for transfer.  Concurrently, the interconnect inquiries CPU1 if it holds a more updated version of the data by sending a memory write message $\mathit{msg_2}$.   CPU1 generates one of two possible responses.  If CPU1 holds the more updated data in its cache in the same memory space that CPU2 intends to write, a cache hit message $\mathit{msg_4}$ followed by $\mathit{msg_6}$ are sent to the interconnect.  Otherwise, CPU1 sends a cache miss message $\mathit{msg_5}$.  After getting the response from CPU1, Interconnect sends a write request to the memory unit.  This flow is symmetric for CPU1.

\begin{figure}[tb]
\begin{center}
\resizebox{2.8in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(msg1)	at	($(p1) + (0,-1.1)$) {$msg_1$};
	\node[circ] 	(p2) 		at ($(msg1) + (-3,0)$) {$p_2$};
	\node[circ] 	(p3) 		at ($(msg1) + (3,0)$) {$p_3$};
	\node[rectbox]	(msg2)	at	($(p2) + (0,-1.1)$) {$msg_2$};
	\node[rectbox]	(msg3)	at	($(p3) + (0,-1.5)$) {$msg_3$};
	\node[circ] 	(p4) 		at ($(msg2) + (0,-1.1)$) {$p_4$};
	\node[rectbox]	(msg4)	at	($(p4) + (0,-1.1)$) {$msg_4$};
	\node[rectbox]	(msg5)	at	($(p4) + (3,0)$) {$msg_5$};
	\node[circ] 	(p6) 		at ($(msg4) + (0,-1.1)$) {$p_6$};
	\node[circ] 	(p5) 		at ($(msg3) + (0,-1.4)$) {$p_5$};
	\node[rectbox,rotate=90]	(msg6)	at	($(p6) + (1.5,0)$) {$msg_6$};
	\node[circ] 	(p7) 		at ($(msg6) + (1.5,0)$) {$p_7$};
	\node[rectbox]	(msg7)	at	($(p5) + (0,-1.5)$) {$msg_7$};
	\node[terminal]	(p8)	at	($(msg7) + (2,0)$) {$p_8$};
	% Edges
	\draw[->,weight3]	(p1) -- (msg1);
	\draw[->,weight3]	(msg1) -- (p2);
	\draw[->,weight3]	(msg1) -- (p3);
	\draw[->,weight3]	(p2) -- (msg2);
	\draw[->,weight3]	(msg2) -- (p4);
	\draw[->,weight3]	(p4) -- (msg4);
	\draw[->,weight3]	(p4) -- (msg5);
	\draw[->,weight3]	(msg4) -- (p6);
	\draw[->,weight3]	(p6) -- (msg6);
	\draw[->,weight3]	(msg6) -- (p7);
	\draw[->,weight3]	(p7) -- (msg7);
	\draw[->,weight3]	(msg7) -- (p8);
	\draw[->,weight3]	(msg5) -- (p7);
	\draw[->,weight3]	(p3) -- (msg3);
	\draw[->,weight3]	(msg3) -- (p5);
	\draw[->,weight3]	(p5) -- (msg7);

	\node[]	(msg-def)	at	($(p1) + (0,-8.5)$) {
	\begin{minipage}{3in}
	Definition of the messages:
\[
\begin{array}{clllll}
msg_1: & (\mbox{CPU2},	& \mbox{Icon},		& \mathit{Wr}, 	& M) \\
msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& \mathit{Wr},			& M) \\
msg_3: & (\mbox{CPU2},	& \mbox{Icon}, 	& \mathit{DVal},	& -) \\
msg_4: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Hit},	& -) \\
msg_5: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Miss},	& -) \\
msg_6: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{DVal},	& -) \\
msg_7: & (\mbox{Icon},	& \mbox{Mem}, 		& \mathit{Wr},			& M) \\
\end{array}
\]
\end{minipage}
		};
\end{tikzpicture}
}
\caption{Flow specification ($F_1$) of a cache coherent write operation initiated from \texttt{CPU2}.}
\label{fig-cpu-write}
\end{center}
\end{figure}

The LPN specification as shown in Fig.~\ref{fig-cpu-read} captures the system flow where CPU2 initiates a memory read operation.  Basically, CPU2 sends a memory read message to Interconnect, which then generates two concurrent threads, one checks if CPU1 has the more updated data in the memory space for the read operation, and the other thread to get data from the memory.  Once Interconnect gets both responses from CPU1 and memory, it synchronizes the responses, and writes the correct data to the CPU2's cache and memory in parallel.  Again, this specification is symmetric for CPU1.

\begin{figure}[htbp]
\begin{center}
\resizebox{3.2in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(msg1)	at	($(p1) + (0,-1.1)$) {$msg_1$};
	\node[circ] 	(p2) 		at ($(msg1) + (-3,0)$) {$p_2$};
	\node[circ] 	(p3) 		at ($(msg1) + (2,0)$) {$p_3$};
	\node[rectbox]	(msg2)	at	($(p2) + (0,-1.1)$) {$msg_2$};
	\node[rectbox]	(msg3)	at	($(p3) + (2,0)$) {$msg_3$};
	\node[circ] 	(p4) 		at ($(msg2) + (0,-1.1)$) {$p_4$};
	\node[rectbox]	(msg4)	at	($(p4) + (0,-1.1)$) {$msg_4$};
	\node[rectbox]	(msg5)	at	($(p4) + (3,0)$) {$msg_5$};
	\node[circ] 	(p6) 		at ($(msg4) + (0,-1.1)$) {$p_6$};
	\node[circ] 	(p5) 		at ($(msg3) + (0,-1.1)$) {$p_5$};
	\node[rectbox,rotate=90]	(msg6)	at	($(p6) + (1.5,0)$) {$msg_6$};
	\node[circ] 	(p7) 		at ($(msg6) + (1.5,0)$) {$p_7$};
	\node[rectbox]	(msg7)	at	($(p5) + (0,-1.1)$) {$msg_7$};
	\node[circ]	(p8)	at	($(msg7) + (0,-1.1)$) {$p_8$};
	\node[rectbox]	(msg8)	at	($(p8) + (0,-1.1)$) {$msg_8$};
	\node[circ]	(p9)	at	($(msg8) + (-1.5,-1.1)$) {$p_9$};
	\node[circ]	(p10)	at	($(msg8) + (1.5,-1.1)$) {$p_{10}$};
	\node[rectbox]	(msg9)	at	($(p9) + (0,-1.1)$) {$msg_9$};
	\node[rectbox]	(msg10)	at	($(p10) + (0,-1.1)$) {$msg_{10}$};
	\node[terminal]	(p11)	at	($(msg9) + (0,-1.1)$) {$p_{11}$};
	\node[terminal]	(p12)	at	($(msg10) + (0,-1.1)$) {$p_{12}$};

	% Edges
	\draw[->,weight3]	(p1) -- (msg1);
	\draw[->,weight3]	(msg1) -- (p2);
	\draw[->,weight3]	(msg1) -- (p3);
	\draw[->,weight3]	(p2) -- (msg2);
	\draw[->,weight3]	(msg2) -- (p4);
	\draw[->,weight3]	(p4) -- (msg4);
	\draw[->,weight3]	(p4) -- (msg5);
	\draw[->,weight3]	(msg4) -- (p6);
	\draw[->,weight3]	(p6) -- (msg6);
	\draw[->,weight3]	(msg6) -- (p7);
	\draw[->,weight3]	(p7) -- (msg8);
	\draw[->,weight3]	(msg7) -- (p8);
	\draw[->,weight3]	(msg5) -- (p7);
	\draw[->,weight3]	(p3) -- (msg3);
	\draw[->,weight3]	(msg3) -- (p5);
	\draw[->,weight3]	(p5) -- (msg7);
	\draw[->,weight3]	(p8) -- (msg8);
	\draw[->,weight3]	(msg8) -- (p9);
	\draw[->,weight3]	(msg8) -- (p10);
	\draw[->,weight3]	(p9)  -- (msg9);
	\draw[->,weight3]	(p10)  -- (msg10);
	\draw[->,weight3]	(msg9) -- (p11);
	\draw[->,weight3]	(msg10) -- (p12);

	\node[]	(msg-def)	at	($(p1) + (-2.7,-10)$) {
	\begin{minipage}{2in}
	Definition of the messages:
\[
\begin{array}{clllll}
msg_1: & (\mbox{CPU2},	& \mbox{Icon},		& \mathit{Rd}, 		& M) \\
msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& \mathit{Rd},	& M) \\
msg_3: & (\mbox{Icon},	& \mbox{Mem}, 		& \mathit{Rd},			& M) \\
msg_4: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Hit},	& -) \\
msg_5: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Miss},& -) \\
msg_6: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{DVal},	& -) \\
msg_7: & (\mbox{Mem},		& \mbox{Icon}, & \mathit{DVal},			& -) \\
msg_8: & (-,		& -, 		& -,				& -) \\
msg_9: & (\mbox{Icon},	& \mbox{Mem}, 		& \mathit{Wr},			& M) \\
msg_{10}: & (\mbox{Icon},	& \mbox{CPU2}, 		& \mathit{Wr},			& M) \\
\end{array}
\]
\end{minipage}
		};

\end{tikzpicture}
}
\caption{Flow specification ($F_2$) of a cache coherent read operation from CPU2.}
\label{fig-cpu-read}
\end{center}
\end{figure}

The LPN specification as shown in Fig.~\ref{fig-peri-read} captures a system flow for CPU initiated memory mapped peripheral read operations.  When CPU2 tries to read the peripheral device (device hereafter), a read message with address $P$ is sent to Interconnect, which then sends this message to CPU1 and the device simultaneously.  When CPU1 sees this message, it responds with a cache miss message as the address $P$ points to the device.  At the meantime, the device responds with a data valid message indicating the availability of the requested data.  Finally, Interconnect synchronizes the both responses, and sends a data valid message back to CPU2.   This specification is also symmetric for CPU1.

\begin{figure}[tb]
\begin{center}
\resizebox{3.2in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(msg1)	at	($(p1) + (0,-1.1)$) {$msg_1$};
	\node[circ] 	(p2) 		at ($(msg1) + (-1.5,0)$) {$p_2$};
	\node[circ] 	(p3) 		at ($(msg1) + (1.5,0)$) {$p_3$};
	\node[rectbox]	(msg2)	at	($(p2) + (-1.5,0)$) {$msg_2$};
	\node[rectbox]	(msg4)	at	($(p3) + (1.5,0)$) {$msg_4$};
	\node[circ] 	(p4) 		at ($(msg2) + (-1.5,0)$) {$p_4$};
	\node[rectbox]	(msg3)	at	($(p4) + (0,-1.2)$) {$msg_3$};
	\node[circ] 	(p6) 		at ($(msg3) + (0,-1.2)$) {$p_6$};
	\node[circ] 	(p5) 		at ($(msg4) + (1.5,0)$) {$p_5$};
	\node[rectbox]	(msg5)	at	($(p5) + (0,-1.2)$) {$msg_5$};
	\node[circ]	(p7)	at	($(msg5) + (0,-1.2)$) {$p_7$};
	\node[rectbox]	(msg6)	at	($(p6) + (4.5,0)$) {$msg_6$};
	\node[terminal]	(p8)	at	($(msg6) + (0,-1.2)$) {$p_8$};

	% Edges
	\draw[->,weight3]	(p1) -- (msg1);
	\draw[->,weight3]	(msg1) -- (p2);
	\draw[->,weight3]	(msg1) -- (p3);
	\draw[->,weight3]	(p2) -- (msg2);
	\draw[->,weight3]	(msg2) -- (p4);
	\draw[->,weight3]	(p4) -- (msg3);
	\draw[->,weight3]	(msg3) -- (p6);
	\draw[->,weight3]	(p6) -- (msg6);
	\draw[->,weight3]	(msg6) -- (p8);
	\draw[->,weight3]	(p3) -- (msg4);
	\draw[->,weight3]	(msg4) -- (p5);
	\draw[->,weight3]	(p5) -- (msg5);
	\draw[->,weight3]	(msg5) -- (p7);
	\draw[->,weight3]	(p7) -- (msg6);
	\draw[->,weight3]	(msg6) -- (p8);

		\node[]	(msg-def)	at	($(p8) + (0,-2.5)$) {
	\begin{minipage}{3in}
	Definition of the messages:
\[
\begin{array}{clllll}
msg_1: & (\mbox{CPU2},	& \mbox{Icon},		& Rd, 	& P) \\
msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& Rd,		& P) \\
msg_3: & (\mbox{CPU1},	& \mbox{Icon}, 	& Miss,			& -) \\
msg_4: & (\mbox{Icon},	& \mbox{Dev}, 	& Rd,	& P) \\
msg_5: & (\mbox{Dev},	& \mbox{Icon}, 	& Dval,	& -) \\
msg_6: & (\mbox{Icon},	& \mbox{CPU2}, 	& Dval,	& -) \\\end{array}
\]
\end{minipage}
		};
\end{tikzpicture}
}
\caption{Flow specification ($F_3$) of a read access to peripheral device from CPU2.}
\label{fig-peri-read}
\end{center}
\end{figure}

The last LPN specification as shown in Fig.~\ref{fig-msi} captures how interrupts from the device are handled.  In this case study, all interrupts are directed to CPU1.  When the device triggers an interrupts, it sends a message with $\mathit{Intr}$ in the command field.  Then, Interconnect notifies CPU1 by sending a message with $\mathit{MSI}$ and $I$ in the command and address fields, respectively, where $I$ is the symbol referring to the entry points to interrupt service routines.   CPU1 responds with a cache miss message as the receipt of the interrupt.

\begin{figure}[htbp]
\begin{center}
\resizebox{3.5in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(msg1)	at	($(p1) + (1.5, 0)$) {$msg_1$};
	\node[circ] 	(p2) 		at ($(msg1) + (1.5, 0)$) {$p_2$};
	\node[rectbox]	(msg2)	at	($(p2) + (1.5,0)$) {$msg_2$};
	\node[circ] 	(p3) 		at ($(msg2) + (1.5, 0)$) {$p_3$};
	\node[rectbox]	(msg3)	at	($(p3) + (1.5, 0)$) {$msg_3$};
	\node[terminal] 	(p4) 		at ($(msg3) + (1.5,0)$) {$p_4$};

	% Edges
	\draw[->,weight3]	(p1) -- (msg1);
	\draw[->,weight3]	(msg1) -- (p2);
	\draw[->,weight3]	(p2) -- (msg2);
	\draw[->,weight3]	(msg2) -- (p3);
	\draw[->,weight3]	(p3) -- (msg3);
	\draw[->,weight3]	(msg3) -- (p4);

\node[]	(msg-def)	at	($(msg2) + (0,-2)$) {
	\begin{minipage}{3in}
	Definition of the messages:
\[
\begin{array}{clllll}
msg_1: & (\mbox{Dev},	& \mbox{Icon},		& \mathit{Intr}, 	& -) \\
msg_2: & (\mbox{Icon},	& \mbox{CPU1}, 	& \mathit{MSI},	& I) \\
msg_3: & (\mbox{CPU1},	& \mbox{Icon}, 	& \mathit{Miss},	& -) \\
\end{array}
\]
\end{minipage}
		};
\end{tikzpicture}
}
\caption{Flow specification ($F_4$) of handling interrupts from the peripheral device.}
\label{fig-msi}
\end{center}
\end{figure}

%
%
\subsection{Results and Discussions}
%
%
The transaction level model of the simple system implementing the four flow specifications shown in the previous section is described in SystemC.  Each component is described as a SystemC module, which may include a number of threads to model concurrency.  The entire model is concurrent and operates in asynchronous mode.

When testing this model, both CPUs and the peripheral device are set up as flow instance generators.  They randomly generate the first message in a corresponding flow specification to start a flow instance, and react to incoming messages by generating new messages as defined in the flow specification.  In the model, monitors are embedded to observe the messages generated by each component.  When a message is observed, it is written to an output trace file for analysis.

Even though this example is conceptually simple, getting the model to correctly implement the flow specifications is not straightforward.  On the other hand, results from the trace analysis greatly help the debugging process by providing information to locate problems quickly.  For example, in an early version of the model, the trace shown in Table~\ref{table-trace} is observed.  
\begin{table*}
\caption{An observed trace of messages for trace analysis.}
\label{table-trace}
\begin{center}
\begin{tabular}{llll}
1~~(CPU2, Icon, Rd, M) & 2~~(Icon, CPU1, Rd, M) & 3~~(Icon, Mem, Rd, M) & 4~~(Mem,  Icon, DVal, -) \\
5~~(CPU1, Icon, Hit, -) & 6~~(CPU1, Icon, DVal, -) & 7~~(Icon, CPU2, Wr, M) & 8~~(Icon, Mem, Wr, M) \\
9~~(CPU2, Icon, Rd, M) & 10~(Icon, CPU1, Rd, M) & 11~(Icon, Mem, Rd, M) & 12~(CPU2, Icon, Rd, P) \\
13~(Icon, CPU1, Rd, P) & 14~(CPU1, Icon, Miss, -) & 15~(Icon, Dev, Rd, P) & 16~(CPU1, Icon, DVal, -) \\
\end{tabular}
\end{center}
\end{table*}
The trace analysis finds out that messages~$1-8$ in the trace are the results of execution of an instance of flow $F_2$ as shown, and the flow execution scenario is $\{(F_{2,1}, \{p_{11}, p_{12}\})\}$.  Messages~$9-11$ are analyzed as the results of executing another instance of flow $F_2$, message~$12-13$ as the results of executing an instance of flow $F_3$ as shown in Fig.~\ref{fig-peri-read}.  The flow execution scenario after the first thirteen messages in the trace is 
\[
\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_4, p_5\}), (F_{3,1}, \{p_4, p_3\})\}.
\]  
Message~$14$ can be  the result from executing $F_{2,2}$ or $F_{3,1}$, therefore it leads to two following flow execution scenarios:
\begin{equation}
\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_7, p_5\}), (F_{3,1}, \{p_4, p_3\})\}
\end{equation}
\begin{equation}
\{(F_{2,1}, \{p_{11}, p_{12}\}), (F_{2,2}, \{p_4, p_5\}), (F_{3,1}, \{p_6, p_3\})\}
\end{equation}
In either scenario, after mapping message~$15$ to flow $F_{3,1}$, message~$16$ cannot be mapped to any existing flow instance or to a new flow instance.  From this inconsistent message, we know that it is generated by CPU1.  In scenario~(1), CPU1 is in the state after generating the message reporting a cache miss.  In this case, the DataValid message should not generated.  In scenario~(2), CPU1 is in the state before generating the message reporting either a cache hit or miss, and again the DataValid message should not generated.   This inconsistent message helps to locate a bug in the CPU1 model where a DataValid message is generated after either a cache hit or miss message is generated.  After fixing this bug,  in a few more iterations of analysis and debugging, the trace analysis can eventually extract all initiated flow instances in the model, all in their terminal states, thus showing that the model implements the four flows correctly.

In the second experiment, partial observability is taken into account during the trace analysis with the assumption that the command $\mathit{Wr}$ and $\mathit{Rd}$ and addresses $M$ and $P$ are indistinguishable due to the lack of observability.  This partial observability is simulated with a modification to the monitors such that in each observed message, command $\mathit{Wr}$ or $\mathit{Rd}$ is replaced with $(\mathit{Wr},\mathit{Rd})$ and address $M$ or $P$ is replaced with $(M,P)$.  The version of the model generating the trace shown in Table~\ref{table-trace} is reused with the modified the monitors, and the generated trace with the simulated partial observability is shown in Table~\ref{table-trace-2}.  Similarly, only the first sixteen messages are shown. 
\begin{table*}
\caption{A trace of messages with partial observability for trace analysis.}
\label{table-trace-2}
\begin{center}
\resizebox{7.3in}{!}{
\begin{tabular}{llll}
1~~(CPU2, Icon, (Wr, Rd), (M, P)) & 2~~(Icon, CPU1, (Wr, Rd), (M, P)) & 3~~(Icon, Mem, (Wr, Rd), (M, P)) & 4~~(Mem,  Icon, DVal, -) \\
5~~(CPU1, Icon, (Hit, Miss), -) & 6~~(CPU1, Icon, DVal, -) & 7~~(Icon, CPU2, (Wr, Rd), (M, P)) & 8~~(Icon, Mem, (Wr, Rd), (M, P)) \\
9~~(CPU2, Icon, (Wr, Rd), (M, P)) & 10~(Icon, CPU1, (Wr, Rd), (M, P)) & 11~(Icon, Mem, (Wr, Rd), (M, P)) & 12~(CPU2, Icon, (Wr, Rd), (M, P)) \\
13~(Icon, CPU1, (Wr, Rd), (M, P)) & 14~(CPU1, Icon, (Hit, Miss), -) & 15~(Icon, Dev, (Wr, Rd), (M, P)) & 16~(CPU1, Icon, DVal, -) \\
\end{tabular}
}
\end{center}
\end{table*}

Each message in the trace with partial observability is referred to as an {\em super} message to distinguish it from the messages of full observability.  The traces of super messages are referred to as {\em super} traces.  For example, the first super message in the trace from Table~\ref{table-trace-2}, $\mbox{(CPU2, Icon, (Wr, Rd), (M, P))}$, corresponds to four distinct messages:
$\mbox{(CPU2, Icon, Wr, M)}$, $\mbox{(CPU2, Icon, Wr, P)}$, $\mbox{(CPU2, Icon, Rd, M)}$, and $\mbox{(CPU2, Icon, Rd, P)}$.  Some of these messages do not exist in the flow specification, and are ignored during the trace analysis.   In the above example, message $\mbox{(CPU2, Icon, Wr, P)}$ is ignored.

%Each super trace represents a set of traces, each of which is interpreted to derive a set of flow execution scenarios.  These flow execution scenarios can show different inconsistent messages.  For example, the particular trace of the super trace shown in Table~\ref{table-trace-2} below
%\[
%\mbox{1 (CPU2, Icon, Wr, M), 2~(Icon, CPU1, Rd, M)}, \ldots
%\]
%leads to the following flow execution scenario $\{(F_{1,1}, \langle p_2,p_3\rangle)\}$ showing that the second message is inconsistent.  In addition, this super trace also represents a trace as shown in Table~\ref{table-trace} that leads to a real bug in the model as discussed above.  It is also possible in some situation that a super trace may represent a trace with no inconsistent messages.  Therefore, the trace analysis on a super trace can result in a number of flow execution scenarios revealing different problems or no problem.  This ambiguity is caused by the partial observability.

Each super trace represents a set of traces, each of which is interpreted to derive a set of flow execution scenarios.
Due to the partial observability, the number of traces represented by a super trace can become very large. For example, the super trace shown in Table~\ref{table-trace-2} represents about twelve thousand possible traces for that short sequence of messages.  The trace analysis algorithm returns the set of flow execution scenarios for each trace for examination, and a very large number of possible flow execution scenarios can be generated.  The large number of possible flow execution scenarios not only produces too much information that can overwhelm system validators, but also degrades the performance of the trace analysis algorithm by consuming too much memory.  As indicated above,  the validators' insights on the SUD can be utilized to trim the possibilities.  For example, if the validator knows that no flow $F_1$ in Fig.~\ref{fig-cpu-write} is activated in the testing environment, this insight helps to eliminate all flow execution scenarios that include instances of $F_1$ by interpreting message~$\#1$ as either $\mbox{(CPU2, Icon, Rd, M)}$ or $\mbox{(CPU2, Icon, Rd, P)}$.  Consider another insight such that a new instance of flow $F_2$ as in Fig.~\ref{fig-cpu-read} can be initiated only after the completion of the previous instance of $F_2$.  If an instant of $F_2$ is assumed to be initiated by the super message $\#9~\mbox{(CPU2, Icon, (Wr, Rd), (M, P))}$ by interpreting it to $\mbox{(CPU2, Icon, Rd, M)}$ during the trace analysis, the super message~$\#12$ can only be interpreted to $\mbox{(CPU2, Icon, Wr, M)}$ or $\mbox{(CPU2, Icon, Rd, P)}$ as the instance of $F_2$ initiated by the super message $\#9$ has not been completed yet at this point.   According to the above discussion, the validators' insights help restrict how super messages are interpreted, thus reducing the number of flow execution scenarios that can be generated.
At the end of the analysis, all possible flow execution scenarios are returned to system validators for examination.

%
%--
%
\section{Case Study II}

\begin{figure} 
\centerline{
\includegraphics[width=3.3in]{figures/Fig4.png}}
\caption{SoC platform structure.}
\label{SoC}
\end{figure}

In order to find out the efficiency of the trace analysis method for more realistic examples, in this case study, a more detailed transaction level model of a SoC is constructed within the GEM5 environment~\cite{Binkert2011}.  This SoC model consists of two ARM Cortex-A9 cores, each of which contains two separate 16KB data and instruction caches.  The caches are connected to a 1GB memory through a memory bus model.  The architecture of this SoC model is shown in Fig.~\ref{SoC}.

In this model, components communicate with each other by sending and receiving various request and response messages.  In order to observe and trace communications occurring inside this model during execution, monitors are attached to links connecting the components. These monitors record the messages flowing through the links they are attached to, and store them into output trace files.

%By combining the informations from all of the nine communication monitors, required communication traces are obtained. As a virtualized SoC platform, GEM5 has three types of request: timing, atomic and functional. Timing request include the modeling of queuing delay and resource contention. Atomic request is a faster than timing request with no delay.  Functional request is used for loading binaries, examining/changing variables in the simulated system, and so on. In our case, we specify all our request to be timed as it's the most detailed one.However, some of the system initiation is still atomic and it's not related to our research, so we only took the messages that are timing request from our collected data.

%\subsection{Flow Specification}

For this model, we consider the flow specification describing the cache coherence protocol supported in GEM5 that is used to build the model in Fig~\ref{SoC}.  This specification consists of two flows, one for cache coherent read requests and one for  cache coherent write requests, from CPUs.  Two such flows describe the cache coherent protocol for each CPU.  As there are two CPUs in this model, there are actually four such flows considered for this model.  More detailed information about the flow specification can be found at \url{https://github.com/cao2/paper/blob/master/Flow\%20Specification.pdf}

%In the following section, a brief overview for the implemented instances of these flows are given. These flow specifications capture how messages are exchanged for different use cases. In this model, a message is defined with the following format, (Src,Dest,Cmd), where Src and Dest refer to the source and destination components of messages, Cmd refers to the operations that the destination component should perform. 
%\begin{figure} 
%\centerline{
%\includegraphics[width=3.7In]{figures/Fig5.png}}
%
%{\footnotesize
%\[
%\begin{array}{llll}
%msg_0: (&\mbox{ CPU1},&\mbox{writeReq},&\mbox{cache1   })\\       
%msg_1: (&\mbox{ cache1},&\mbox{ readExreq },&\mbox{Bus     })\\        
%msg_2: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2 })\\  
%msg_3: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2         })\\   
%msg_4: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce2           })\\  
%msg_5: (&\mbox{ cache2},&\mbox{readExreq},&\mbox{cpu2 })\\  
%msg_6: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ cahce1       })\\     
%msg_7: (&\mbox{ cache1},&\mbox{readExreq},&\mbox{cpu1           })\\  
%msg_8: (&\mbox{ Bus},&\mbox{ readExreq},&\mbox{ Memory })\\  
%msg_9: (&\mbox{ true                                          })\\  
%msg_{10}: (&\mbox{ Memory},&\mbox{ readExres},&\mbox{ Bus        })\\  
%msg_{11}: (&\mbox{ cache2},&\mbox{ readExres},&\mbox{ Bus })\\  
%msg_{12}: (&\mbox{ Bus},&\mbox{ readExres},&\mbox{ cache1        })\\  
%msg_{13}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1         })\\  
%msg_{14}: (&\mbox{ cache1},&\mbox{ writeRes},&\mbox{ CPU1 })\\  
%msg_{15}: (&\mbox{ cache1},&\mbox{UpgradeReq             })\\  
%msg_{16}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ cahce2      })\\   
%msg_{17}: (&\mbox{ Bus},&\mbox{ UpgradeReq},&\mbox{ Memory })\\  
%msg_{18}: (&\mbox{ cache2},&\mbox{UpgradeRes},&\mbox{ Bus     })\\  
%msg_{19}: (&\mbox{ Bus},&\mbox{ UpgradeRes},&\mbox{ cache1      })\\  
%msg_{20}: (&\mbox{ cache1},&\mbox{ WriteRes},&\mbox{ CPU1 })\\  
%\end{array}
%\]}
%\caption{Flow specification ($F_1$) of a cache coherent write operation initiated from CPU1}
%\label{write-flow}
%\end{figure}
%
%The LPN as shown in Fig.~\ref{write-flow} specifies a system flow where CPU1 initiates a memory write operation. In this flow, CPU1 initiates a memory write request to L1 Cache. Next, depends on whether the required data is included in cache1, the cache1 will generate three possible responses. First, the cache1 will send read exclusive request message msg2 to interconnect if data is not present in cache.; Or if the data is shared by CPU0, it will send upgrade message msg16 to interconnect to disable CPU0's ownership of this block of data, else if CPU1 has exclusive right of required data, cache1 will perform the write operation and sent an response message msg15 to CPU1. Afterwards, interconnect will sent request to all connected component (data cache and instruction cache of each CPU and memory). Once the interconnect obtains the response from either CPU0 or memory, it will generate a response message to cache1, then cache1 will generate a write response message msg14 (or msg 21 ) to CPU1. The flow is symmetric for CPU2.
%
%
%\begin{figure} 
%\centerline{
%\includegraphics[width=4in]{figures/Fih6.png}}
%
%{\footnotesize
%\[
%\begin{array}{llll}
%t0: (&\mbox{ CPU1},&\mbox{ReadReq},&\mbox{cache1  })\\                   
%t1: (&\mbox{ cache1},&\mbox{ StoreCondreq },&\mbox{Bus })\\           
%t2: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2 })\\
%t3: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2       })\\      
%t4: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce2           })\\ 
%t5: (&\mbox{ cache2},&\mbox{StoreCondreq},&\mbox{cpu2 })\\
%t6: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ cahce1     })\\       
%t7: (&\mbox{ cache1},&\mbox{StoreCondreq},&\mbox{cpu1           })\\ 
%t8: (&\mbox{ Bus},&\mbox{ StoreCondreq},&\mbox{ Memory })\\
%t9: (&\mbox{ true                                        })\\
%t10: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus            })\\    
%t11: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus })\\
%t12: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1      })\\            
%t13: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1          })\\  
%t14: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
%t21: (&\mbox{ cache1},&\mbox{ LoadLockedreq },&\mbox{Bus })\\     
%t22: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\
%t23: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2 })\\
%t24: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce2     })\\ 
%t25: (&\mbox{ cache2},&\mbox{LoadLockedreq},&\mbox{cpu2     })\\
%t26: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ cahce1 })\\
%t27: (&\mbox{ cache1},&\mbox{LoadLockedreq},&\mbox{cpu1       })\\
%t28: (&\mbox{ Bus},&\mbox{ LoadLockedreq},&\mbox{ Memory     })\\
%t29: (&\mbox{ true })\\
%t30: (&\mbox{ Memory},&\mbox{ ReadRes},&\mbox{ Bus       })\\     
%t31: (&\mbox{ cache2},&\mbox{ ReadRes},&\mbox{ Bus    })\\
%t32: (&\mbox{ Bus},&\mbox{ ReadRes},&\mbox{ cache1 })\\
%t33: (&\mbox{ cache1},&\mbox{ ReadRes},&\mbox{ CPU1 })\\
%\end{array}
%\]}
%\caption{Flow specification ($F_2$) of a cache coherent read operation initiated from CPU1}
%\label{read-flow}
%
%\end{figure}
%The LPN specification as shown in Fig.~\ref{read-flow} captures the system flow where CPU1 initiate a memory read operation. Similar to memory write, CPU1 will first initiate a memory read request message msg1 to cache1. Cache1 can generate three possible responses. First, if the requested data is not in cache1, cache1 will generate an storeCondReq message msg2 to interconnect asking for data. Second, if the requested data is inside of cache1 but it's shared, cache1 will generate a upgrade request message msg22 to make sure it has the newest data; Last,  if cache1 has and data and it's not shared, it will generate a read response message msg15 to CPU1. Afterwards, interconnect will send corresponding request to all of its connected components. When interconnect receives the response, it will generate response message to cache1. Then cache1 will generate the read response message msg14 or msg34 to CPU1. This specification is also symmetric for CPU2.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Results and Discussions}

In order to trigger cache coherence protocols, a simple two-threaded program including one producer and one consumer is created and executed on the CPUs, each of which executes one thread.  \red{How is shared memory is simulated???} . The producer will write to a specified memory address, and the consumer will read from the same address. GEM5 does not support true concurrency.  When there are two programs running on the CPUs, GEM5 alternates the executions between the two  CPUs.

%We produce a result file including all of the communication messages between every components from two simple programs running simultaneously on each CPU.  The program assigned to CPU1 read one file three times for one letter and writing one letter for three times to the same file. CPU2's program will do the same read and write functionalities to the same file, only difference is that CPU2 will write first and then read. One thing we should know about GEM5 is that even when we run these two programs concurrently, it will attempt to produce a concurrent result. The data we collected shows that it's not the real nondeterministic concurrency. What GEM5 did was it allowed two CPUs to execute its own instruction in turn, therefore the order was deterministic. Therefore, no matter how many times we ran the program, it produced the same result. We tried other virtual SoC platform softwares, and this was the best nondeterministic concurrency we can get. Even if this is slightly off with what the real chip should work, it still surve our purpose of testing the correctness of our algorithm.

After the SoC model finishes executing the program, there are totally $343581$ messages collected in the trace file.  Not all of the messages are relevant to the flow specification as many are used by GEM5 to initialize its simulation environment.  After removing those irrelevant messages, the number of messages in the trace file is to reduced to $121138$.

The time taken to remove the irrelevant messages from the trace is negligible.  The total runtime and the peak memory usage for the trace analysis algorithm to finish the reduced trace are 3 seconds and 12MB, respectively.  From the trace, Table~\ref{table-case-2} shows the number of instances extracted for the four flows describing cache coherent read/write operations. 
\begin{table}[td]
\caption{The number of flow instances derived by the trace analysis}
\begin{center}
\begin{tabular}{|c|c|}
\hline
Flows & $\#$Instances \\
\hline
\hline
CPU1 Read		&  $4002$\\
CPU1 Write	&  $3370$\\
CPU1 Failed Read  &$95$\\
CPU2 Read		&  $3955$\\
CPU2 Write	&  $3308$\\
CPU2 Failed Read  &$93$\\
\hline
\end{tabular}
\end{center}
\label{table-case-2}
\end{table}%

\red{Also need to consider the case where some monitors are removed?}

%=====
\section{Related Work}

Our work is closely related to communication-centric and transaction based debug.  An early pioneering work is described in~\cite{Goossens2007NOCS}, which advocates the focus on observing activities on the interconnect network among IP blocks, and mapping these activities to transactions for better correlation between computations and communications.  Therefore, the communication transactions, as a result of software execution, provide an interface between computation and communication, and facilitate  system-level debug.  This work is extended in~\cite{Vermeulen2009VLSI-DAT,Goossens2009DATE}.  However, this line of work is focused on the network-on-chip (NoC) architecture for interconnect using the run/stop debug control method.

A similar transaction-based debug approach is presented in~\cite{Gharehbaghi2012ISQED}.  Furthermore, it proposes an automated extraction of state machines at transaction level from  high level design models.  From an observed failure trace, it performs backtracking on this transaction level state machine to derive a set of transaction traces that lead to the observed failure state.  In the subsequent step, bounded model checking with the constraints on the internal variables is used to refine the set of transaction traces to remove the infeasible traces.  This approach requires user inputs to identify impossible transaction sequences, and may not find the states causing the failure if the transaction traces leading to the observed failure state is long.  Backtracking from the observed failure state requires pre-image computation, which can be computationally expensive.  A transaction-based online debug approach is proposed in~\cite{Dehbash2014} to address these issues.  This approach utilizes a transaction debug pattern specification language~\cite{Gharehbaghi2009ICCD} to define properties that transactions should meet.  These transaction properties are checked at runtime by programming debug units in the on-chip debug infrastructure, and the system can be stopped shortly after a violation is detected for any one of those properties.  In this sense, it can be viewed as the hardware assertion approaches in~\cite{Boule2007ISQED} elevated to the transaction level.

In~\cite{Singerman2011DAC}, a coherent workflow is described where the result from the pre-silicon validation stage can be carried over to the post-silicon stage to improve efficiency and productivity of post-silicon debug.  This workflow is centered on a repository of system events and simple transactions defined by architects and IP designers.  It spans across a wide spectrum of the post-silicon validation including DFx instrumentation, test generation, coverage, and debug.  The DFx instruments are automatically inserted into the design RTL code driven by the defined transactions.  This instrumentation is optimized for making a large set of events and transactions observable.   Test generation is also optimized to generate only the necessary but sufficient tests to allow all defined transactions to be exercised.   Moreover, coverage for post-silicon validation is now defined at the abstract level of events and transactions rather than the raw signals, and thus can be evaluated more efficiently.  In~\cite{Abarbanel2014DAC}, a model at an even higher-level of abstraction, {\em flows}, is proposed.  Flows are used to specify more sophisticated cross-IP transactions such as power management, security, etc, and to facilitate reuse of the efforts of the architectural analysis to check HW/SW implementations.

%
%--
\section{Conclusion}

This paper presents a trace analysis based method for post-silicon validation by interpreting observed raw signal traces at the level of system flow specifications.  The derived flow execution scenarios provide more structured information on system operations, which is more understandable to system validators.   This information can help to locate design defects more easily, and also provides a measurement of validation coverage.

Due to partial observability, this approach may derive a large number of different flow execution scenarios for a given signal trace.  Insights from system validators can help to eliminate some false scenarios due to the partial observability.  An interesting future direction is formalization of the validators' insights using temporal logic on flows so that the validators can express their intents more precisely and concisely.

The trace analysis approach presented in this paper needs to be iterated with different observations selected in different iterations in order to eliminate the false scenarios and to root cause system failures as quickly as possible.  The observation selection and stitching signal traces of different observations together for the above goal will also be pursued in the future.

% conference papers do not normally have an appendix

% use section* for acknowledgement
%\section*{Acknowledgment}
%
%
%The authors would like to thank...
%

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{SoC}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}
%
%\bibitem{IEEEhowto:kopka}
%H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
%
%\end{thebibliography}

% that's all folks
\end{document}

### 2016-01-03 paper.tex
removed bookmarks
Added an option in order to disable bookmarks in the compiled PDF.

@@
 \documentclass[conference]{IEEEtran}
+
+\PassOptionsToPackage{bookmarks={false}}{hyperref}
+
 \usepackage{pifont}
@@
 %\usepackage{draftwatermark}
-\usepackage{hyperref}
-\hypersetup{
-    colorlinks=true,
-    linkcolor=blue,
-    filecolor=magenta,      
-    urlcolor=cyan,
-}
+%\usepackage[]{hyperref}
+%\hypersetup{
+%    bookmarks    = false,            % show bookmarks bar?
+%    colorlinks=true,
+%    linkcolor=blue,
+%    filecolor=magenta,      
+%    urlcolor=cyan,
+%}
+%
 
@@
 describing the cache coherence protocols supported in
-GEM5\footnote{The GEM5 cache coherence protocols can be
-  found at
-  \url{http://www.gem5.org/docs/html/gem5MemorySystem.html}.}
-that is used to build the model in Fig~\ref{SoC}.  These
-flow specifications describe data/instruction read
+GEM5 that is used to build the model in Fig~\ref{SoC}.
+The GEM5 cache coherence protocols can be found at~\cite{gem5}.
+These flow specifications describe data/instruction read
 operations and data write operations initiated from CPUs.
